### 2013-09-04 background.tex
converted background and methods from Word to LaTeX

Clinical Background and Significance 
Prostate cancer (PCa) is the most common noncutaneous malignancy among men in the United States. Approximately 1 in every 6 men will develop PCa during their lifetime, with the median age of diagnosis at 67. i It is also the second leading cause of cancer related death, with 1 in 36 men dying from the disease. The National Cancer Institute estimates that 238,590 men will be diagnosed with PCa in 2013 and 29,720 will die from the disease.ii

PCa diagnosis usually begins by screening with prostate specific antigen (PSA) and digital rectal examination (DRE).  Definitive diagnosis is made by random ultrasonography-guided transrectal (TRUS) biopsies, which are then used to provide the clinician with the proper Gleason score. The combination of these factors—as well as staging—determines the appropriate therapy and prognosis.

PCa screening has led to earlier diagnosis of smaller tumors and more localized disease.  However, it is well known that the sensitivity and specificity of PSA and DRE are not optimal. In addition, DRE has a low predictive value at lower PSA ranges, and PSA yields many false positives. .iii iv  As such, a theoretical risk of over-diagnosis and treatment of low-grade—and possibly clinically insignificant—disease exists. Moreover, due to the random nature of biopsies, cancer located outside the routine sampling site can be missed and the extent of the cancer might be underestimated.v vi For example, in a study by Mufarrij et al 45.9%-47.2% of patients who were candidates for active surveillance but underwent radical prostatectomy had a higher Gleason score on final histopathology than after TRUS biopsy.vii These inaccuracies may lead to inappropriate diagnosis, imprecise risk assessment and potentially avoidable morbidity.

The Use of Magnetic Resonance Imaging in Prostate Diagnostics
Magnetic resonance (MR) imaging has been available for use in the workup of patients with PCa since the early 1980s but the early work on its diagnostic accuracy is heterogeneous. viii Earlier MR techniques relied mostly on morphology via T1 and T2- weighted imaging (T2WI). The more recent ability to include not only anatomic but also biologic and functional dynamic parameters into MR analysis—via diffusion-weighted imaging (DWI), dynamic contrast-enhanced (DCE) imaging or MR spectroscopic imaging (MRSI)—is promising in the future diagnosis and management of PCA.

 Currently prostate MR focuses on a multiparametric approach, where 2 or more imaging sequences—including anatomic and functional data—are used together to try to arrive to a diagnosis.ix As MR technology continues to evolve and improve, its role in PCA diagnosis, staging, treatment planning and follow-up has gained much attention.

 T2-Weighted Imaging
 T2WI sequences are crucial components of prostate MR imaging.  T2WI is particularly useful in prostate analysis due to its excellent soft tissue contrast resolution, which can be maximized by using thin sections of 3-4mm and a small field of view of approximately 14cm. x xi xii T2 sequences are the most helpful for tumor localization as they can clearly show overall prostate morphology, internal structures and prostatic margins.xiii Seventy five percent of prostatic tumors are found in the peripheral zone (PZ) and normally show hypointense T2 signal when compared to the higher intensity PZ.xiv xvHowever tumors can sometimes seem of similar intensity as the surrounding tissue and false positives can occur secondary to post biopsy changes/hemorrhage, hyperplasia or prostatitis, making diagnosis more challenging.xvi

 Functional MR Sequences
 Even though T2WI is the mainstay of prostate MR, its overall performance in prostate cancer diagnosis is not optimal. The incorporation of two or more functional sequences in multiparametric MR imaging (mpMRI) has been shown to significantly improve the performance of MR in cancer diagnosis.xvii In fact, the European Society of Urogenital Radiology’s (ESUR) prostate MR guidelines recommend at least 2 functional imaging techniques in addition to T2WI in order to better characterize prostate tumors.xviii  In a study by Turkbey et al, researchers found that mpMRI had a positive predictive value of 98% in prostate cancer detection.xix These functional sequences include diffusion- weighted imaging (DWI), dynamic contrast enhanced imaging (DCE-MRI) and MR spectroscopic imaging (MRSI).

 Diffusion-weighted imaging
 DWI is based on the free movement of water particles in tissue and measures the degree of motion restriction.xx Normal prostatic tissue is very glandular with plenty of water molecule movement. On the other hand tumors have high cellular density and restricted water movement, which leads to decreased diffusion. Due to this restricted diffusion, tumors seem to be of higher intensity on DWI.xxi

 DWI sequences can be processed to obtain apparent diffusion coefficient (ADC) maps. These serve as a more objective measure of diffusivity since they are independent of magnetic field and thus overcome the effects of T2 shine-through.xxii xxiii ADC maps are also used to visually assess the tumor, which appears as a focus of decreased signal intensity when compared to normal prostate tissue.xxiv

 In addition to knowing the location of the cancer within the prostate, being able to identify which cancers will exhibit more aggressive behavior is important when making clinical management decisions. A number of studies have found that lower ADC values correlate with tumors that have higher Gleason scores. xxv xxvi xxviiThis could be explained by the fact that higher-grade lesions are usually of higher cellular density and thus have even more marked restricted diffusion. This inverse correlation suggests that ADC maps might be helpful when characterizing lesions and assessing tumor aggressiveness. xxviii

 Dynamic contrast enhanced imaging
 DCE-MRI has been shown to have a high sensitivity in cancer detection, especially when combined with other MR imaging modalities.xxix xxx xxxi xxxii In a study by Hara et al, this functional modality identified 93% of clinically significant cancers.xxxiii DCE-MRI provides a measure of tumor vascularity and takes advantage of tumor-induced angiogenesis in PCa. xxxiv Increased vascularity in prostate cancer results in earlier and higher peak contrast enhancement and a faster washout when compared to normal prostatic tissue.xxxv  These known characteristics of PCa are helpful when trying to localize lesions.

 A variety of different approaches can be used to analyze DCE-MRI and no report has, up to date, described the superiority of one method over the others.xxxvi The approaches include qualitative analysis by visual assessment of the images, semi-quantitative methods that look at kinetic parameters on a voxel-by-voxel basis, or quantitative methods that that convert signal intensity into kinetic parameters to determine how much contrast is being exchanged between the vascular, extravascular and extracellular spaces.xxxvii

    MR spectroscopy
    Like DWI, MRSI can be helpful in lesion characterization as it provides information on the presence of certain metabolites in tissue. xxxviii Choline and citrate are especially useful in the setting of PCa. Choline is critical in cell membrane synthesis and is thus usually elevated in cells with cancerous behavior.xxxix Healthy prostate epithelium synthesizes and secretes large quantities of citrate but levels are decreased in PCa.xl xliTherefore, an increase in the choline-to-citrate ratio on MRSI can be used as an indicator of malignancy.xlii

    Although some data suggests that MRSI increases the accuracy of tumor volume detection and staging when used in addition to anatomic imaging, it is unclear whether it is superior to other modalities.xliii ESUR lists MSRI as an optional modality for the diagnosis of PCa since it significantly lengthens exam time and requires specific technical expertise.xliv Many academic centers in the United States do not include MSRI in their mpMRI prostate cancer protocols.xlv

### 2013-09-04 methods.tex
converted background and methods from Word to LaTeX

Methods & Materials
Study Inclusion Criteria & MR Imaging
Prostates removed by radical prostatectomy were used in this IRB-approved (Duke IRB# Pro00006458), HIPAA compliant study from men ranging in age from XX--XX diagnosed with biopsy-proven prostate cancer destined for surgical removal. Between DATES, a total of XX patients were recruited and enrolled in this study. Inclusion criteria were undergoing complete pelvic MRI with endorectal coil for detection of prostate cancer, including multiplanar T2-weighted anatomic imaging, diffusion-weighted imaging (DWI), and dynamic contrast enhanced MRI (DCE-MRI) as well as radical prostatectomy and whole mount histology. Patients with previous treatments of prostate cancer or benign prostatic hyperplasia (BPH), or anatomic anomalies of the rectum, were excluded.  All patients enrolled in this study provided written informed consent. Our final cohort included XX subjects. Table XX has a summary of the patient demographics for the cases shown in this manuscript.
All imaging was performed on one of two 3.0 Tesla MR scanners (General Electric HDx, GE Healthcare, Waukesha, WI;  Siemens Skyra, Siemens Healthcare, Erlangan, Germany) using a single channel Medrad eCoil endorectal coil (Medrad, Indianola, PA) as well as multichannel surface coils.  Imaging sequences included thin-section (3 mm section thickness) fast spin echo T2-weighted images in the coronal, axial and sagittal planes.  Diffusion weighted images were obtained using multiple b-values and calculation of ADC maps was also performed.  Dynamic contrast enhanced MR sequences were obtained after administration of a weight-based dose of extracellular MR contrast agent with 4-5 second temporal resolution for 5-6 minutes. (If we need it, can put in table with full MR parameters)
Prostates were radically removed using a da Vinci Surgical System.  After exision, the prostates were formalin fixed for at least 24 hours without being cut, and then processed for whole mount histology.
Whole Mount Histology
Fixed in formalin; hand-cut ~3 mm spacing
H&E whole-mount images
Clinically read, with the following markings (ELN)
Digitally-scanned @ 600 dpi
Manually segmented using itk-SNAP; saved as nifti stack
Rigid Histology Image Registration
MR T2WI Data
Diffeomorphic MR:Histology Image Registration
Rigid T2WI:ADC Image Registration

### 2013-09-09 background.tex
updated background from Kirema

@@
 Approximately 1 in every 6 men will develop PCa during their lifetime, with the median age of diagnosis at 67.
-i It is also the second leading cause of cancer related death, with 1 in 36 men dying from the disease.
-The National Cancer Institute estimates that 238,590 men will be diagnosed with PCa in 2013 and 29,720 will die from the disease.ii
+1It is also the second leading cause of cancer related death, with 1 in 36 men dying from the disease.
+The National Cancer Institute estimates that 238,590 men will be diagnosed with PCa in 2013 and 29,720 will die from the disease.1
 
 PCa diagnosis usually begins by screening with prostate specific antigen (PSA) and digital rectal examination (DRE).
-Definitive diagnosis is made by random ultrasonography-guided transrectal (TRUS) biopsies, which are then used to provide the clinician with the proper Gleason score.
+Definitive diagnosis is made by random transrectal ultrasonography-guided (TRUS) biopsies, which are then used to provide the clinician with the proper Gleason score.
 The combination of these factors—as well as staging—determines the appropriate therapy and prognosis.
@@
 However, it is well known that the sensitivity and specificity of PSA and DRE are not optimal.
-In addition, DRE has a low predictive value at lower PSA ranges, and PSA yields many false positives.
-.iii iv  As such, a theoretical risk of over-diagnosis and treatment of low-grade—and possibly clinically insignificant—disease exists.
-Moreover, due to the random nature of biopsies, cancer located outside the routine sampling site can be missed and the extent of the cancer might be underestimated.v vi For example, in a study by Mufarrij et al 45.9%-47.2% of patients who were candidates for active surveillance but underwent radical prostatectomy had a higher Gleason score on final histopathology than after TRUS biopsy.vii These inaccuracies may lead to inappropriate diagnosis, imprecise risk assessment and potentially avoidable morbidity.
+In addition, DRE has a low predictive value at lower PSA ranges, and PSA yields many false positives.2-4 As such, a theoretical risk of over-diagnosis and treatment of low-grade—and possibly clinically insignificant—disease exists.
+Moreover, due to the random nature of biopsies, cancer located outside the routine sampling site can be missed and the extent of the cancer might be underestimated.3 5 For example, in a study by Mufarrij et al 45.9\%-47.2\% of patients who were candidates for active surveillance but underwent radical prostatectomy had a higher Gleason score on final histopathology than after TRUS biopsy.6 These inaccuracies may lead to inappropriate diagnosis, imprecise risk assessment and potentially avoidable morbidity.
 
 The Use of Magnetic Resonance Imaging in Prostate Diagnostics
-Magnetic resonance (MR) imaging has been available for use in the workup of patients with PCa since the early 1980s but the early work on its diagnostic accuracy is heterogeneous.
-viii Earlier MR techniques relied mostly on morphology via T1 and T2- weighted imaging (T2WI).
+Magnetic resonance (MR) imaging has been available for use in the workup of patients with PCa since the early 1980s but the early work on its diagnostic accuracy is heterogeneous.7 Earlier MR techniques relied mostly on morphology via T1 and T2- weighted imaging (T2WI).
 The more recent ability to include not only anatomic but also biologic and functional dynamic parameters into MR analysis—via diffusion-weighted imaging (DWI), dynamic contrast-enhanced (DCE) imaging or MR spectroscopic imaging (MRSI)—is promising in the future diagnosis and management of PCA.
 
- Currently prostate MR focuses on a multiparametric approach, where 2 or more imaging sequences—including anatomic and functional data—are used together to try to arrive to a diagnosis.ix As MR technology continues to evolve and improve, its role in PCA diagnosis, staging, treatment planning and follow-up has gained much attention.
+ Currently prostate MR focuses on a multiparametric approach, where 2 or more imaging sequences—including anatomic and functional data—are used together to try to arrive to a diagnosis.8 As MR technology continues to evolve and improve, its role in PCA diagnosis, staging, treatment planning and follow-up has gained much attention.
 
- T2-Weighted Imaging
+ T2-Weighted Imaging and Prostate Anatomy
  T2WI sequences are crucial components of prostate MR imaging.
-T2WI is particularly useful in prostate analysis due to its excellent soft tissue contrast resolution, which can be maximized by using thin sections of 3-4mm and a small field of view of approximately 14cm.
-x xi xii T2 sequences are the most helpful for tumor localization as they can clearly show overall prostate morphology, internal structures and prostatic margins.xiii Seventy five percent of prostatic tumors are found in the peripheral zone (PZ) and normally show hypointense T2 signal when compared to the higher intensity PZ.xiv xvHowever tumors can sometimes seem of similar intensity as the surrounding tissue and false positives can occur secondary to post biopsy changes/hemorrhage, hyperplasia or prostatitis, making diagnosis more challenging.xvi
+T2WI is particularly useful in prostate analysis due to its excellent soft tissue contrast resolution, which can be maximized by using thin sections of 3-4mm and a small field of view of approximately 14cm.3,9 T2 sequences are the most helpful for tumor localization as they can clearly show overall prostate morphology, internal structures and prostatic margins.3
+
+ The prostate can be divided into glandular and nonglandular components.
+The glandular components include the peripheral zone (PZ) and the central gland, which are typically easily distinguishable on T2WI.
+The central gland includes the central zone, transition zone and the periurethral glandular tissue.10 Other anatomical markers such as the urethra, verumontanum and ejaculatory are also often seen on T2WI.
+
+ Approximately 70\% of the prostatic tissue is found in the peripheral zone, which is high in water content and thus of higher signal intensity in T2WI.10 Seventy five percent of prostatic tumors are found in the PZ and normally show hypointense T2 signal when compared to the higher intensity PZ.4 11 However tumors can sometimes seem of similar intensity as the surrounding tissue and false positives can occur secondary to post biopsy changes/hemorrhage, hyperplasia or prostatitis, making diagnosis more challenging.11
 
@@
  Even though T2WI is the mainstay of prostate MR, its overall performance in prostate cancer diagnosis is not optimal.
-The incorporation of two or more functional sequences in multiparametric MR imaging (mpMRI) has been shown to significantly improve the performance of MR in cancer diagnosis.xvii In fact, the European Society of Urogenital Radiology’s (ESUR) prostate MR guidelines recommend at least 2 functional imaging techniques in addition to T2WI in order to better characterize prostate tumors.xviii  In a study by Turkbey et al, researchers found that mpMRI had a positive predictive value of 98% in prostate cancer detection.xix These functional sequences include diffusion- weighted imaging (DWI), dynamic contrast enhanced imaging (DCE-MRI) and MR spectroscopic imaging (MRSI).
+The incorporation of two or more functional sequences in multiparametric MR imaging (mpMRI) has been shown to significantly improve the performance of MRI in cancer diagnosis.12 In fact, the European Society of Urogenital Radiology’s (ESUR) prostate MR guidelines recommend at least 2 functional imaging techniques in addition to T2WI in order to better characterize prostate tumors.8 In a study by Turkbey et al, researchers found that mpMRI had a positive predictive value of 98\% in prostate cancer detection.12 Functional sequences include diffusion- weighted imaging (DWI), dynamic contrast enhanced imaging (DCE-MRI) and MR spectroscopic imaging (MRSI).
 
  Diffusion-weighted imaging
- DWI is based on the free movement of water particles in tissue and measures the degree of motion restriction.xx Normal prostatic tissue is very glandular with plenty of water molecule movement.
-On the other hand tumors have high cellular density and restricted water movement, which leads to decreased diffusion.
-Due to this restricted diffusion, tumors seem to be of higher intensity on DWI.xxi
+ DWI is based on the free movement of water particles in tissue and measures the degree of motion restriction.13 Normal prostatic tissue is very glandular with plenty of water molecule movement.
+On the other hand, tumors have high cellular density and restricted water movement, which leads to decreased diffusion.
+Due to this restricted diffusion, tumors seem to be of higher intensity on DWI.13
 
  DWI sequences can be processed to obtain apparent diffusion coefficient (ADC) maps.
-These serve as a more objective measure of diffusivity since they are independent of magnetic field and thus overcome the effects of T2 shine-through.xxii xxiii ADC maps are also used to visually assess the tumor, which appears as a focus of decreased signal intensity when compared to normal prostate tissue.xxiv
+These serve as a more objective measure of diffusivity since they are independent of magnetic field and thus overcome the effects of T2 shine-through.13,14  ADC maps are also used to visually assess the tumor, which appears as a focus of decreased signal intensity when compared to normal prostate tissue.14
 
@@
 A number of studies have found that lower ADC values correlate with tumors that have higher Gleason scores.
-xxv xxvi xxviiThis could be explained by the fact that higher-grade lesions are usually of higher cellular density and thus have even more marked restricted diffusion.
-This inverse correlation suggests that ADC maps might be helpful when characterizing lesions and assessing tumor aggressiveness.
-xxviii
+15-17 This could be explained by the fact that higher-grade lesions are usually of higher cellular density and thus have even more marked restricted diffusion.
+This inverse correlation suggests that ADC maps might be helpful when characterizing lesions and assessing tumor aggressiveness.18
 
  Dynamic contrast enhanced imaging
- DCE-MRI has been shown to have a high sensitivity in cancer detection, especially when combined with other MR imaging modalities.xxix xxx xxxi xxxii In a study by Hara et al, this functional modality identified 93% of clinically significant cancers.xxxiii DCE-MRI provides a measure of tumor vascularity and takes advantage of tumor-induced angiogenesis in PCa.
-xxxiv Increased vascularity in prostate cancer results in earlier and higher peak contrast enhancement and a faster washout when compared to normal prostatic tissue.xxxv  These known characteristics of PCa are helpful when trying to localize lesions.
+ DCE-MRI has been shown to have a high sensitivity in cancer detection, especially when combined with other MR imaging modalities.19-22 In a study by Hara et al, this functional modality identified 93\% of clinically significant cancers.23 DCE-MRI provides a measure of tumor vascularity and takes advantage of tumor-induced angiogenesis in PCa.
+24Increased vascularity in PCa results in earlier and higher peak contrast enhancement and a faster washout when compared to normal prostatic tissue.
+These known characteristics of PCa are helpful when trying to localize lesions.
 
- A variety of different approaches can be used to analyze DCE-MRI and no report has, up to date, described the superiority of one method over the others.xxxvi The approaches include qualitative analysis by visual assessment of the images, semi-quantitative methods that look at kinetic parameters on a voxel-by-voxel basis, or quantitative methods that that convert signal intensity into kinetic parameters to determine how much contrast is being exchanged between the vascular, extravascular and extracellular spaces.xxxvii
+ A variety of different approaches can be used to analyze DCE-MRI and no report has, up to date, described the superiority of one method over the others.3 The approaches include qualitative analysis by visual assessment of the images, semi-quantitative methods that look at kinetic parameters on a voxel-by-voxel basis, or quantitative methods that that convert signal intensity into kinetic parameters to determine how much contrast is being exchanged between the vascular, extravascular and extracellular spaces.3
 
@@
     Like DWI, MRSI can be helpful in lesion characterization as it provides information on the presence of certain metabolites in tissue.
-xxxviii Choline and citrate are especially useful in the setting of PCa.
-Choline is critical in cell membrane synthesis and is thus usually elevated in cells with cancerous behavior.xxxix Healthy prostate epithelium synthesizes and secretes large quantities of citrate but levels are decreased in PCa.xl xliTherefore, an increase in the choline-to-citrate ratio on MRSI can be used as an indicator of malignancy.xlii
+8Choline and citrate are especially useful in the setting of PCa.
+Choline is critical in cell membrane synthesis and is thus usually elevated in cells with cancerous behavior.3 Healthy prostate epithelium synthesizes and secretes large quantities of citrate but levels are decreased in PCa.25,26 Therefore, an increase in the choline-to-citrate ratio on MRSI can be used as an indicator of malignancy.9
 
-    Although some data suggests that MRSI increases the accuracy of tumor volume detection and staging when used in addition to anatomic imaging, it is unclear whether it is superior to other modalities.xliii ESUR lists MSRI as an optional modality for the diagnosis of PCa since it significantly lengthens exam time and requires specific technical expertise.xliv Many academic centers in the United States do not include MSRI in their mpMRI prostate cancer protocols.xlv
+    Although some data suggests that MRSI increases the accuracy of tumor volume detection and staging when used in addition to anatomic imaging, it is unclear whether it is superior to other modalities.9 ESUR lists MSRI as an optional modality for the diagnosis of PCa since it significantly lengthens exam time and requires specific technical expertise.8 Thus, it is not surprising that many academic centers in the United States do not include MSRI in their mpMRI prostate cancer protocols.3

### 2013-09-09 background.tex
placeholders

@@
-Clinical Background and Significance 
+\subsection{Clinical Background and Significance}
 Prostate cancer (PCa) is the most common noncutaneous malignancy among men in the United States.
@@
 
-The Use of Magnetic Resonance Imaging in Prostate Diagnostics
+\subsection{The Use of Magnetic Resonance Imaging in Prostate Diagnostics}
 Magnetic resonance (MR) imaging has been available for use in the workup of patients with PCa since the early 1980s but the early work on its diagnostic accuracy is heterogeneous.7 Earlier MR techniques relied mostly on morphology via T1 and T2- weighted imaging (T2WI).
@@
 
- T2-Weighted Imaging and Prostate Anatomy
+\subsubsection{T2-Weighted Imaging and Prostate Anatomy}
  T2WI sequences are crucial components of prostate MR imaging.
@@
 
- Functional MR Sequences
+\subsubsection{Functional MR Sequences}
  Even though T2WI is the mainstay of prostate MR, its overall performance in prostate cancer diagnosis is not optimal.
@@
 
- Diffusion-weighted imaging
+\subsubsection{Diffusion-weighted imaging}
  DWI is based on the free movement of water particles in tissue and measures the degree of motion restriction.13 Normal prostatic tissue is very glandular with plenty of water molecule movement.
@@
 
- Dynamic contrast enhanced imaging
+\subsubsection{Dynamic contrast enhanced imaging}
  DCE-MRI has been shown to have a high sensitivity in cancer detection, especially when combined with other MR imaging modalities.19-22 In a study by Hara et al, this functional modality identified 93\% of clinically significant cancers.23 DCE-MRI provides a measure of tumor vascularity and takes advantage of tumor-induced angiogenesis in PCa.
@@
 
-    MR spectroscopy
+\subsubsection{MR spectroscopy}
 Like DWI, MRSI can be helpful in lesion characterization as it provides information on the presence of certain metabolites in tissue.

### 2013-09-09 compile.sh

@@ -1,0 +1,12 @@
+#!/bin/bash
+texfile='mr_arfi_anatomy'
+pdflatex $texfile.tex
+pdflatex $texfile.tex
+bibtex $texfile
+pdflatex $texfile.tex
+pdflatex $texfile.tex
+rm $texfile.aux $texfile.bbl $texfile.out $texfile.blg $texfile.log
+
+#response='ResponseToReviewsR2'
+#pdflatex $response.tex
+#rm $response.aux $response.log

### 2013-09-09 methods.tex
syntax fixed

@@
-Methods & Materials
-Study Inclusion Criteria & MR Imaging
-Prostates removed by radical prostatectomy were used in this IRB-approved (Duke IRB# Pro00006458), HIPAA compliant study from men ranging in age from XX--XX diagnosed with biopsy-proven prostate cancer destined for surgical removal.
-Between DATES, a total of XX patients were recruited and enrolled in this study.
-Inclusion criteria were undergoing complete pelvic MRI with endorectal coil for detection of prostate cancer, including multiplanar T2-weighted anatomic imaging, diffusion-weighted imaging (DWI), and dynamic contrast enhanced MRI (DCE-MRI) as well as radical prostatectomy and whole mount histology.
-Patients with previous treatments of prostate cancer or benign prostatic hyperplasia (BPH), or anatomic anomalies of the rectum, were excluded.
-All patients enrolled in this study provided written informed consent.
-Our final cohort included XX subjects.
-Table XX has a summary of the patient demographics for the cases shown in this manuscript.
-All imaging was performed on one of two 3.0 Tesla MR scanners (General Electric HDx, GE Healthcare, Waukesha, WI;  Siemens Skyra, Siemens Healthcare, Erlangan, Germany) using a single channel Medrad eCoil endorectal coil (Medrad, Indianola, PA) as well as multichannel surface coils.
-Imaging sequences included thin-section (3 mm section thickness) fast spin echo T2-weighted images in the coronal, axial and sagittal planes.
-Diffusion weighted images were obtained using multiple b-values and calculation of ADC maps was also performed.
-Dynamic contrast enhanced MR sequences were obtained after administration of a weight-based dose of extracellular MR contrast agent with 4-5 second temporal resolution for 5-6 minutes.
-(If we need it, can put in table with full MR parameters)
-Prostates were radically removed using a da Vinci Surgical System.
-After exision, the prostates were formalin fixed for at least 24 hours without being cut, and then processed for whole mount histology.
+\subsection{Study Inclusion Criteria \& MR Imaging}
+Prostates removed by radical prostatectomy were used in this IRB-approved (Duke
+IRB# Pro00006458), HIPAA compliant study from men ranging in age from XX--XX
+diagnosed with biopsy-proven prostate cancer destined for surgical removal.
+Between DATES, a total of XX patients were recruited and enrolled in this
+study. Inclusion criteria were undergoing complete pelvic MRI with endorectal
+coil for detection of prostate cancer, including multiplanar T2-weighted
+anatomic imaging, diffusion-weighted imaging (DWI), and dynamic contrast
+enhanced MRI (DCE-MRI) as well as radical prostatectomy and whole mount
+histology. Patients with previous treatments of prostate cancer or benign
+prostatic hyperplasia (BPH), or anatomic anomalies of the rectum, were
+excluded.  All patients enrolled in this study provided written informed
+consent. Our final cohort included XX subjects. Table XX has a summary of the
+patient demographics for the cases shown in this manuscript.
+
+All imaging was performed on one of two 3.0 Tesla MR scanners (General Electric
+HDx, GE Healthcare, Waukesha, WI;  Siemens Skyra, Siemens Healthcare, Erlangan,
+Germany) using a single channel Medrad eCoil endorectal coil (Medrad,
+Indianola, PA) as well as multichannel surface coils.  Imaging sequences
+included thin-section (3 mm section thickness) fast spin echo T2-weighted
+images in the coronal, axial and sagittal planes.  Diffusion weighted images
+were obtained using multiple b-values and calculation of ADC maps was also
+performed.  Dynamic contrast enhanced MR sequences were obtained after
+administration of a weight-based dose of extracellular MR contrast agent with
+4-5 second temporal resolution for 5-6 minutes. (If we need it, can put in
+table with full MR parameters) Prostates were radically removed using a da
+Vinci Surgical System.  After exision, the prostates were formalin fixed for at
+least 24 hours without being cut, and then processed for whole mount histology.
+
 Whole Mount Histology

### 2013-09-09 methods.tex
more sytnax fixed

@@
 Prostates removed by radical prostatectomy were used in this IRB-approved (Duke
-IRB# Pro00006458), HIPAA compliant study from men ranging in age from XX--XX
+IRB\# Pro00006458), HIPAA compliant study from men ranging in age from XX--XX
 diagnosed with biopsy-proven prostate cancer destined for surgical removal.
@@
 Fixed in formalin; hand-cut ~3 mm spacing
-H&E whole-mount images
+H\&E whole-mount images
 Clinically read, with the following markings (ELN)

### 2013-09-10 methods.tex
new subsection headings

--- 2013-09-09 methods.tex
@@
 Prostates removed by radical prostatectomy were used in this IRB-approved (Duke
-IRB# Pro00006458), HIPAA compliant study from men ranging in age from XX--XX
+IRB\# Pro00006458), HIPAA-compliant study from men ranging in age from XX--XX
 diagnosed with biopsy-proven prostate cancer destined for surgical removal.
@@
 table with full MR parameters) Prostates were radically removed using a da
-Vinci Surgical System.  After exision, the prostates were formalin fixed for at
-least 24 hours without being cut, and then processed for whole mount histology.
+Vinci Surgical System (INSERT COMPANY INFORMATION).  After exision, the
+prostates were formalin fixed for at least 24 hours without being cut, and then
+processed for whole mount histology.
 
-Whole Mount Histology
-Fixed in formalin; hand-cut ~3 mm spacing
-H&E whole-mount images
-Clinically read, with the following markings (ELN)
-Digitally-scanned @ 600 dpi
-Manually segmented using itk-SNAP; saved as nifti stack
-Rigid Histology Image Registration
-MR T2WI Data
-Diffeomorphic MR:Histology Image Registration
-Rigid T2WI:ADC Image Registration
+\subsection{ARFI Imaging}
+
+\subsection{Image Zonal Anatomy Segmentation and 3D Model Rendering}

### 2013-09-13 arfi_methods.tex
Stephen ARFI methods

\subsection{ARFI Imaging Methods}
Experimental ARFI imaging data were acquired using a modified Siemens Acuson
SC2000 ultrasound scanner (Siemens Healthcare, Ultrasound Business Unit,
Mountain View, CA, USA) and the longitudinal array of an Acuson ER7B
transducer~\cite{Rosenzweig2012}.  The ARFI imaging sequence was comprised of
standard B-mode ultrasonic imaging, or tracking beams, and pushing beams. For
each lateral location, two pre-push reference images were acquired, then three
300 cycle pushing pulses were transmitted in rapid succession, focused at 30
mm, 22.5 mm, and 15 mm, respectively, and finally the response of the tissue
was tracked for up to 6ms at a PRF of 8kHz. This pushing strategy is similar to
what has been published by Bercoff et al~\cite{BercoffSSI}. The 30 mm and 22.5
mm foci pushing pulses were transmitted at 4.6 MHz with a F/2 geometry and the
15 mm focus pushing pulse was transmitted at 5.4 MHz with a F/2.35 geometry to
maintain the same beamwidth (0.67 mm) throughout the region of excitation. A
total of 82 lateral locations were interrogated to cover the 55 mm field of
view, translating 0.67 mm laterally per location.

For the tracking pulses, 16 parallel receive lines at 5.0 MHz were spaced to
observe both the on and off-axis response of the tissue to the pushing pulses.
Specifically, four lines were dedicated to tracking the on-axis displacement,
with all 4 beams located inside the beamwidth of the pushing pulses such that
the beam spacing was 0.17 mm. The other twelve lines were separated into two
groups to observe both the left and right propagating waves. For each sub-group
of 6, the beams were located 1.89 mm laterally offset from the push and had an
inter-beam spacing of 0.76 mm to cover a total field of view of 11.3 mm
laterally. For brevity, the 4 on-axis beams will be referred to as the ARFI
data and the 12 off-axis beams will be referred to as the SWEI data.

Displacement estimation was performed using a phase shift estimator on the
beamformed IQ data~\cite{Loupas1,Pinton1}. The ARFI data were then normalized
as a function of depth to account for attenuation and focal gain effects by
taking the mean displacement across the entire data set at each time step then
low-pass filtering with a cutoff frequency of 0.8 mm$^{-1}$.

### 2013-09-13 methods.tex
Stephen ARFI methods

@@
-\subsection{Study Inclusion Criteria \& MR Imaging}
-Prostates removed by radical prostatectomy were used in this IRB-approved (Duke
-IRB\# Pro00006458), HIPAA-compliant study from men ranging in age from XX--XX
-diagnosed with biopsy-proven prostate cancer destined for surgical removal.
-Between DATES, a total of XX patients were recruited and enrolled in this
-study. Inclusion criteria were undergoing complete pelvic MRI with endorectal
-coil for detection of prostate cancer, including multiplanar T2-weighted
-anatomic imaging, diffusion-weighted imaging (DWI), and dynamic contrast
-enhanced MRI (DCE-MRI) as well as radical prostatectomy and whole mount
-histology. Patients with previous treatments of prostate cancer or benign
-prostatic hyperplasia (BPH), or anatomic anomalies of the rectum, were
-excluded.  All patients enrolled in this study provided written informed
-consent. Our final cohort included XX subjects. Table XX has a summary of the
-patient demographics for the cases shown in this manuscript.
+\input{mr_methods}
 
-All imaging was performed on one of two 3.0 Tesla MR scanners (General Electric
-HDx, GE Healthcare, Waukesha, WI;  Siemens Skyra, Siemens Healthcare, Erlangan,
-Germany) using a single channel Medrad eCoil endorectal coil (Medrad,
-Indianola, PA) as well as multichannel surface coils.  Imaging sequences
-included thin-section (3 mm section thickness) fast spin echo T2-weighted
-images in the coronal, axial and sagittal planes.  Diffusion weighted images
-were obtained using multiple b-values and calculation of ADC maps was also
-performed.  Dynamic contrast enhanced MR sequences were obtained after
-administration of a weight-based dose of extracellular MR contrast agent with
-4-5 second temporal resolution for 5-6 minutes. (If we need it, can put in
-table with full MR parameters) Prostates were radically removed using a da
-Vinci Surgical System (INSERT COMPANY INFORMATION).  After exision, the
-prostates were formalin fixed for at least 24 hours without being cut, and then
-processed for whole mount histology.
+\input{arfi_methods}
 
-\subsection{ARFI Imaging}
-
-\subsection{Image Zonal Anatomy Segmentation and 3D Model Rendering}
+\input{image_proc_methods}

### 2013-09-13 mr_methods.tex
MR methods split to separate file

\subsection{Study Inclusion Criteria \& MR Imaging}
Prostates removed by radical prostatectomy were used in this IRB-approved (Duke
IRB\# Pro00006458), HIPAA-compliant study from men ranging in age from XX--XX
diagnosed with biopsy-proven prostate cancer destined for surgical removal.
Between DATES, a total of XX patients were recruited and enrolled in this
study. Inclusion criteria were undergoing complete pelvic MRI with endorectal
coil for detection of prostate cancer, including multiplanar T2-weighted
anatomic imaging, diffusion-weighted imaging (DWI), and dynamic contrast
enhanced MRI (DCE-MRI) as well as radical prostatectomy and whole mount
histology. Patients with previous treatments of prostate cancer or benign
prostatic hyperplasia (BPH), or anatomic anomalies of the rectum, were
excluded.  All patients enrolled in this study provided written informed
consent. Our final cohort included XX subjects. Table XX has a summary of the
patient demographics for the cases shown in this manuscript.

All imaging was performed on one of two 3.0 Tesla MR scanners (General Electric
HDx, GE Healthcare, Waukesha, WI;  Siemens Skyra, Siemens Healthcare, Erlangan,
Germany) using a single channel Medrad eCoil endorectal coil (Medrad,
Indianola, PA) as well as multichannel surface coils.  Imaging sequences
included thin-section (3 mm section thickness) fast spin echo T2-weighted
images in the coronal, axial and sagittal planes.  Diffusion weighted images
were obtained using multiple b-values and calculation of ADC maps was also
performed.  Dynamic contrast enhanced MR sequences were obtained after
administration of a weight-based dose of extracellular MR contrast agent with
4-5 second temporal resolution for 5-6 minutes. (If we need it, can put in
table with full MR parameters) Prostates were radically removed using a da
Vinci Surgical System (INSERT COMPANY INFORMATION).  After exision, the
prostates were formalin fixed for at least 24 hours without being cut, and then
processed for whole mount histology.

### 2013-09-13 background.tex
Kirema seg methods included, and converting background refs, but question about numbers

@@
 \subsection{Clinical Background and Significance}
-Prostate cancer (PCa) is the most common noncutaneous malignancy among men in the United States.
-Approximately 1 in every 6 men will develop PCa during their lifetime, with the median age of diagnosis at 67.
-1It is also the second leading cause of cancer related death, with 1 in 36 men dying from the disease.
-The National Cancer Institute estimates that 238,590 men will be diagnosed with PCa in 2013 and 29,720 will die from the disease.1
+Prostate cancer (PCa) is the most common noncutaneous malignancy among men in
+the United States. Approximately 1 in every 6 men will develop PCa during their
+lifetime, with the median age of diagnosis at 67. 1It is also the second
+leading cause of cancer related death, with 1 in 36 men dying from the disease.
+The National Cancer Institute estimates that 238,590 men will be diagnosed with
+PCa in 2013 and 29,720 will die from the disease.~\cite{Howlader2010}
 
-PCa diagnosis usually begins by screening with prostate specific antigen (PSA) and digital rectal examination (DRE).
-Definitive diagnosis is made by random transrectal ultrasonography-guided (TRUS) biopsies, which are then used to provide the clinician with the proper Gleason score.
-The combination of these factors—as well as staging—determines the appropriate therapy and prognosis.
+PCa diagnosis usually begins by screening with prostate specific antigen (PSA)
+and digital rectal examination (DRE).  Definitive diagnosis is made by random
+transrectal ultrasonography-guided (TRUS) biopsies, which are then used to
+provide the clinician with the proper Gleason score. The combination of these
+factors—as well as staging—determines the appropriate therapy and prognosis.
 
-PCa screening has led to earlier diagnosis of smaller tumors and more localized disease.
-However, it is well known that the sensitivity and specificity of PSA and DRE are not optimal.
-In addition, DRE has a low predictive value at lower PSA ranges, and PSA yields many false positives.2-4 As such, a theoretical risk of over-diagnosis and treatment of low-grade—and possibly clinically insignificant—disease exists.
-Moreover, due to the random nature of biopsies, cancer located outside the routine sampling site can be missed and the extent of the cancer might be underestimated.3 5 For example, in a study by Mufarrij et al 45.9\%-47.2\% of patients who were candidates for active surveillance but underwent radical prostatectomy had a higher Gleason score on final histopathology than after TRUS biopsy.6 These inaccuracies may lead to inappropriate diagnosis, imprecise risk assessment and potentially avoidable morbidity.
+PCa screening has led to earlier diagnosis of smaller tumors and more localized
+disease.  However, it is well known that the sensitivity and specificity of PSA
+and DRE are not optimal. In addition, DRE has a low predictive value at lower
+PSA ranges, and PSA yields many false
+positives.~\cite{Howlader2010,Schroder2000,Gosselaar2006} As such, a
+theoretical risk of over-diagnosis and treatment of low-grade—and possibly
+clinically insignificant—disease exists. Moreover, due to the random nature of
+biopsies, cancer located outside the routine sampling site can be missed and
+the extent of the cancer might be underestimated.~\cite{Schroder2000,Gupta2013}
+For example, in a study by Mufarrij et al. 45.9--47.2\% of patients who were
+candidates for active surveillance but underwent radical prostatectomy had a
+higher Gleason score on final histopathology than after TRUS
+biopsy.~\cite{Mufarrih2010} These inaccuracies may lead to inappropriate
+diagnosis, imprecise risk assessment and potentially avoidable morbidity.
 
 \subsection{The Use of Magnetic Resonance Imaging in Prostate Diagnostics}
-Magnetic resonance (MR) imaging has been available for use in the workup of patients with PCa since the early 1980s but the early work on its diagnostic accuracy is heterogeneous.7 Earlier MR techniques relied mostly on morphology via T1 and T2- weighted imaging (T2WI).
-The more recent ability to include not only anatomic but also biologic and functional dynamic parameters into MR analysis—via diffusion-weighted imaging (DWI), dynamic contrast-enhanced (DCE) imaging or MR spectroscopic imaging (MRSI)—is promising in the future diagnosis and management of PCA.
+Magnetic resonance (MR) imaging has been available for use in the workup of
+patients with PCa since the early 1980s but the early work on its diagnostic
+accuracy is heterogeneous.7 Earlier MR techniques relied mostly on morphology
+via T1 and T2- weighted imaging (T2WI). The more recent ability to include not
+only anatomic but also biologic and functional dynamic parameters into MR
+analysis—via diffusion-weighted imaging (DWI), dynamic contrast-enhanced (DCE)
+imaging or MR spectroscopic imaging (MRSI)—is promising in the future diagnosis
+and management of PCA.
 
- Currently prostate MR focuses on a multiparametric approach, where 2 or more imaging sequences—including anatomic and functional data—are used together to try to arrive to a diagnosis.8 As MR technology continues to evolve and improve, its role in PCA diagnosis, staging, treatment planning and follow-up has gained much attention.
+Currently prostate MR focuses on a multiparametric approach, where 2 or more
+imaging sequences—including anatomic and functional data—are used together to
+try to arrive to a diagnosis.8 As MR technology continues to evolve and
+improve, its role in PCA diagnosis, staging, treatment planning and follow-up
+has gained much attention.
 
 \subsubsection{T2-Weighted Imaging and Prostate Anatomy}
- T2WI sequences are crucial components of prostate MR imaging.
-T2WI is particularly useful in prostate analysis due to its excellent soft tissue contrast resolution, which can be maximized by using thin sections of 3-4mm and a small field of view of approximately 14cm.3,9 T2 sequences are the most helpful for tumor localization as they can clearly show overall prostate morphology, internal structures and prostatic margins.3
+T2WI sequences are crucial components of prostate MR imaging.  T2WI is
+particularly useful in prostate analysis due to its excellent soft tissue
+contrast resolution, which can be maximized by using thin sections of 3-4mm and
+a small field of view of approximately 14cm.3,9 T2 sequences are the most
+helpful for tumor localization as they can clearly show overall prostate
+morphology, internal structures and prostatic margins.3
 
- The prostate can be divided into glandular and nonglandular components.
-The glandular components include the peripheral zone (PZ) and the central gland, which are typically easily distinguishable on T2WI.
-The central gland includes the central zone, transition zone and the periurethral glandular tissue.10 Other anatomical markers such as the urethra, verumontanum and ejaculatory are also often seen on T2WI.
+The prostate can be divided into glandular and nonglandular components. The
+glandular components include the peripheral zone (PZ) and the central gland,
+which are typically easily distinguishable on T2WI. The central gland includes
+the central zone, transition zone and the periurethral glandular tissue.10
+Other anatomical markers such as the urethra, verumontanum and ejaculatory are
+also often seen on T2WI.
 
- Approximately 70\% of the prostatic tissue is found in the peripheral zone, which is high in water content and thus of higher signal intensity in T2WI.10 Seventy five percent of prostatic tumors are found in the PZ and normally show hypointense T2 signal when compared to the higher intensity PZ.4 11 However tumors can sometimes seem of similar intensity as the surrounding tissue and false positives can occur secondary to post biopsy changes/hemorrhage, hyperplasia or prostatitis, making diagnosis more challenging.11
+Approximately 70\% of the prostatic tissue is found in the peripheral zone,
+which is high in water content and thus of higher signal intensity in T2WI.10
+Seventy five percent of prostatic tumors are found in the PZ and normally show
+hypointense T2 signal when compared to the higher intensity PZ.4 11 However
+tumors can sometimes seem of similar intensity as the surrounding tissue and
+false positives can occur secondary to post biopsy changes/hemorrhage,
+hyperplasia or prostatitis, making diagnosis more challenging.11
 
 \subsubsection{Functional MR Sequences}
- Even though T2WI is the mainstay of prostate MR, its overall performance in prostate cancer diagnosis is not optimal.
-The incorporation of two or more functional sequences in multiparametric MR imaging (mpMRI) has been shown to significantly improve the performance of MRI in cancer diagnosis.12 In fact, the European Society of Urogenital Radiology’s (ESUR) prostate MR guidelines recommend at least 2 functional imaging techniques in addition to T2WI in order to better characterize prostate tumors.8 In a study by Turkbey et al, researchers found that mpMRI had a positive predictive value of 98\% in prostate cancer detection.12 Functional sequences include diffusion- weighted imaging (DWI), dynamic contrast enhanced imaging (DCE-MRI) and MR spectroscopic imaging (MRSI).
+Even though T2WI is the mainstay of prostate MR, its overall performance in
+prostate cancer diagnosis is not optimal. The incorporation of two or more
+functional sequences in multiparametric MR imaging (mpMRI) has been shown to
+significantly improve the performance of MRI in cancer diagnosis.12 In fact,
+the European Society of Urogenital Radiology’s (ESUR) prostate MR guidelines
+recommend at least 2 functional imaging techniques in addition to T2WI in order
+to better characterize prostate tumors.8 In a study by Turkbey et al,
+researchers found that mpMRI had a positive predictive value of 98\% in
+prostate cancer detection.12 Functional sequences include diffusion- weighted
+imaging (DWI), dynamic contrast enhanced imaging (DCE-MRI) and MR spectroscopic
+imaging (MRSI).
 
 \subsubsection{Diffusion-weighted imaging}
- DWI is based on the free movement of water particles in tissue and measures the degree of motion restriction.13 Normal prostatic tissue is very glandular with plenty of water molecule movement.
-On the other hand, tumors have high cellular density and restricted water movement, which leads to decreased diffusion.
-Due to this restricted diffusion, tumors seem to be of higher intensity on DWI.13
+DWI is based on the free movement of water particles in tissue and measures the
+degree of motion restriction.13 Normal prostatic tissue is very glandular with
+plenty of water molecule movement. On the other hand, tumors have high cellular
+density and restricted water movement, which leads to decreased diffusion. Due
+to this restricted diffusion, tumors seem to be of higher intensity on DWI.13
 
- DWI sequences can be processed to obtain apparent diffusion coefficient (ADC) maps.
-These serve as a more objective measure of diffusivity since they are independent of magnetic field and thus overcome the effects of T2 shine-through.13,14  ADC maps are also used to visually assess the tumor, which appears as a focus of decreased signal intensity when compared to normal prostate tissue.14
+DWI sequences can be processed to obtain apparent diffusion coefficient (ADC)
+maps. These serve as a more objective measure of diffusivity since they are
+independent of magnetic field and thus overcome the effects of T2
+shine-through.13,14  ADC maps are also used to visually assess the tumor, which
+appears as a focus of decreased signal intensity when compared to normal
+prostate tissue.14
 
- In addition to knowing the location of the cancer within the prostate, being able to identify which cancers will exhibit more aggressive behavior is important when making clinical management decisions.
-A number of studies have found that lower ADC values correlate with tumors that have higher Gleason scores.
-15-17 This could be explained by the fact that higher-grade lesions are usually of higher cellular density and thus have even more marked restricted diffusion.
-This inverse correlation suggests that ADC maps might be helpful when characterizing lesions and assessing tumor aggressiveness.18
+In addition to knowing the location of the cancer within the prostate, being
+able to identify which cancers will exhibit more aggressive behavior is
+important when making clinical management decisions. A number of studies have
+found that lower ADC values correlate with tumors that have higher Gleason
+scores. 15-17 This could be explained by the fact that higher-grade lesions are
+usually of higher cellular density and thus have even more marked restricted
+diffusion. This inverse correlation suggests that ADC maps might be helpful
+when characterizing lesions and assessing tumor aggressiveness.18
 
 \subsubsection{Dynamic contrast enhanced imaging}
- DCE-MRI has been shown to have a high sensitivity in cancer detection, especially when combined with other MR imaging modalities.19-22 In a study by Hara et al, this functional modality identified 93\% of clinically significant cancers.23 DCE-MRI provides a measure of tumor vascularity and takes advantage of tumor-induced angiogenesis in PCa.
-24Increased vascularity in PCa results in earlier and higher peak contrast enhancement and a faster washout when compared to normal prostatic tissue.
-These known characteristics of PCa are helpful when trying to localize lesions.
+DCE-MRI has been shown to have a high sensitivity in cancer detection,
+especially when combined with other MR imaging modalities.19-22 In a study by
+Hara et al, this functional modality identified 93\% of clinically significant
+cancers.23 DCE-MRI provides a measure of tumor vascularity and takes advantage
+of tumor-induced angiogenesis in PCa. 24Increased vascularity in PCa results in
+earlier and higher peak contrast enhancement and a faster washout when compared
+to normal prostatic tissue. These known characteristics of PCa are helpful when
+trying to localize lesions.
 
- A variety of different approaches can be used to analyze DCE-MRI and no report has, up to date, described the superiority of one method over the others.3 The approaches include qualitative analysis by visual assessment of the images, semi-quantitative methods that look at kinetic parameters on a voxel-by-voxel basis, or quantitative methods that that convert signal intensity into kinetic parameters to determine how much contrast is being exchanged between the vascular, extravascular and extracellular spaces.3
+A variety of different approaches can be used to analyze DCE-MRI and no report
+has, up to date, described the superiority of one method over the others.3 The
+approaches include qualitative analysis by visual assessment of the images,
+semi-quantitative methods that look at kinetic parameters on a voxel-by-voxel
+basis, or quantitative methods that that convert signal intensity into kinetic
+parameters to determine how much contrast is being exchanged between the
+vascular, extravascular and extracellular spaces.3
 
 \subsubsection{MR spectroscopy}
-Like DWI, MRSI can be helpful in lesion characterization as it provides information on the presence of certain metabolites in tissue.
-8Choline and citrate are especially useful in the setting of PCa.
-Choline is critical in cell membrane synthesis and is thus usually elevated in cells with cancerous behavior.3 Healthy prostate epithelium synthesizes and secretes large quantities of citrate but levels are decreased in PCa.25,26 Therefore, an increase in the choline-to-citrate ratio on MRSI can be used as an indicator of malignancy.9
+Like DWI, MRSI can be helpful in lesion characterization as it provides
+information on the presence of certain metabolites in tissue. 8Choline and
+citrate are especially useful in the setting of PCa. Choline is critical in
+cell membrane synthesis and is thus usually elevated in cells with cancerous
+behavior.3 Healthy prostate epithelium synthesizes and secretes large
+quantities of citrate but levels are decreased in PCa.25,26 Therefore, an
+increase in the choline-to-citrate ratio on MRSI can be used as an indicator of
+malignancy.9
 
-Although some data suggests that MRSI increases the accuracy of tumor volume detection and staging when used in addition to anatomic imaging, it is unclear whether it is superior to other modalities.9 ESUR lists MSRI as an optional modality for the diagnosis of PCa since it significantly lengthens exam time and requires specific technical expertise.8 Thus, it is not surprising that many academic centers in the United States do not include MSRI in their mpMRI prostate cancer protocols.3
+Although some data suggests that MRSI increases the accuracy of tumor volume
+detection and staging when used in addition to anatomic imaging, it is unclear
+whether it is superior to other modalities.9 ESUR lists MSRI as an optional
+modality for the diagnosis of PCa since it significantly lengthens exam time
+and requires specific technical expertise.8 Thus, it is not surprising that
+many academic centers in the United States do not include MSRI in their mpMRI
+prostate cancer protocols.3

### 2013-09-13 image_proc_methods.tex
Kirema seg methods included, and converting background refs, but question about numbers

\subsection{Image Zonal Anatomy Segmentation and 3D Model Rendering}
Axial T2W MR slices were manually segmented using the polygon tool on ITK SNAP
using separate labels for the peripheral zone (PZ), central gland (CG) and
anterior fibromuscular stroma (AFS). The gland was segmented from base to apex.
The base was identified below the bladder and subsequent images were segmented
until the last slice with visible prostatic tissue was identified caudally. The
CG, PZ and AFS were segmented independently according to their well-established
anatomical characteristics on
T2WI.~\cite{Verma2011,Jung2012,Poon1985,Hricak2007,Bonekamp2011} The PZ was
identified by its homogenous high signal intensity on T2WI, which is usually
similar to that of the nearby periprostatic fat. The CG was visualized and
delineated based on its heterogeneous and lower signal intensity as well as its
location. Although not readily visible on every case, the AFS was identified by
its low T2 signal intensity and its location anterior to the central gland.

### 2013-09-13 acknowledgements.tex
updated formatting

\section*{ACKNOWLEDGEMENTS}
The authors would like to thank Siemens Medical Solution USA, Ultrasound
Division for their in-kind technical support and Ned Danieley for system administration support.
This work was supported by NIH R01CA142824 and the Duke Coulter Translational Grant Program.

### 2013-09-13 background.tex
updated formatting

--- 2013-09-09 background.tex
@@
-Clinical Background and Significance 
-Prostate cancer (PCa) is the most common noncutaneous malignancy among men in the United States.
-Approximately 1 in every 6 men will develop PCa during their lifetime, with the median age of diagnosis at 67.
-1It is also the second leading cause of cancer related death, with 1 in 36 men dying from the disease.
-The National Cancer Institute estimates that 238,590 men will be diagnosed with PCa in 2013 and 29,720 will die from the disease.1
+\section{BACKGROUND}\label{sect:background}
 
-PCa diagnosis usually begins by screening with prostate specific antigen (PSA) and digital rectal examination (DRE).
-Definitive diagnosis is made by random transrectal ultrasonography-guided (TRUS) biopsies, which are then used to provide the clinician with the proper Gleason score.
-The combination of these factors—as well as staging—determines the appropriate therapy and prognosis.
-
-PCa screening has led to earlier diagnosis of smaller tumors and more localized disease.
-However, it is well known that the sensitivity and specificity of PSA and DRE are not optimal.
-In addition, DRE has a low predictive value at lower PSA ranges, and PSA yields many false positives.2-4 As such, a theoretical risk of over-diagnosis and treatment of low-grade—and possibly clinically insignificant—disease exists.
-Moreover, due to the random nature of biopsies, cancer located outside the routine sampling site can be missed and the extent of the cancer might be underestimated.3 5 For example, in a study by Mufarrij et al 45.9\%-47.2\% of patients who were candidates for active surveillance but underwent radical prostatectomy had a higher Gleason score on final histopathology than after TRUS biopsy.6 These inaccuracies may lead to inappropriate diagnosis, imprecise risk assessment and potentially avoidable morbidity.
-
-The Use of Magnetic Resonance Imaging in Prostate Diagnostics
-Magnetic resonance (MR) imaging has been available for use in the workup of patients with PCa since the early 1980s but the early work on its diagnostic accuracy is heterogeneous.7 Earlier MR techniques relied mostly on morphology via T1 and T2- weighted imaging (T2WI).
-The more recent ability to include not only anatomic but also biologic and functional dynamic parameters into MR analysis—via diffusion-weighted imaging (DWI), dynamic contrast-enhanced (DCE) imaging or MR spectroscopic imaging (MRSI)—is promising in the future diagnosis and management of PCA.
-
- Currently prostate MR focuses on a multiparametric approach, where 2 or more imaging sequences—including anatomic and functional data—are used together to try to arrive to a diagnosis.8 As MR technology continues to evolve and improve, its role in PCA diagnosis, staging, treatment planning and follow-up has gained much attention.
-
- T2-Weighted Imaging and Prostate Anatomy
- T2WI sequences are crucial components of prostate MR imaging.
-T2WI is particularly useful in prostate analysis due to its excellent soft tissue contrast resolution, which can be maximized by using thin sections of 3-4mm and a small field of view of approximately 14cm.3,9 T2 sequences are the most helpful for tumor localization as they can clearly show overall prostate morphology, internal structures and prostatic margins.3
-
- The prostate can be divided into glandular and nonglandular components.
-The glandular components include the peripheral zone (PZ) and the central gland, which are typically easily distinguishable on T2WI.
-The central gland includes the central zone, transition zone and the periurethral glandular tissue.10 Other anatomical markers such as the urethra, verumontanum and ejaculatory are also often seen on T2WI.
-
- Approximately 70\% of the prostatic tissue is found in the peripheral zone, which is high in water content and thus of higher signal intensity in T2WI.10 Seventy five percent of prostatic tumors are found in the PZ and normally show hypointense T2 signal when compared to the higher intensity PZ.4 11 However tumors can sometimes seem of similar intensity as the surrounding tissue and false positives can occur secondary to post biopsy changes/hemorrhage, hyperplasia or prostatitis, making diagnosis more challenging.11
-
- Functional MR Sequences
- Even though T2WI is the mainstay of prostate MR, its overall performance in prostate cancer diagnosis is not optimal.
-The incorporation of two or more functional sequences in multiparametric MR imaging (mpMRI) has been shown to significantly improve the performance of MRI in cancer diagnosis.12 In fact, the European Society of Urogenital Radiology’s (ESUR) prostate MR guidelines recommend at least 2 functional imaging techniques in addition to T2WI in order to better characterize prostate tumors.8 In a study by Turkbey et al, researchers found that mpMRI had a positive predictive value of 98\% in prostate cancer detection.12 Functional sequences include diffusion- weighted imaging (DWI), dynamic contrast enhanced imaging (DCE-MRI) and MR spectroscopic imaging (MRSI).
-
- Diffusion-weighted imaging
- DWI is based on the free movement of water particles in tissue and measures the degree of motion restriction.13 Normal prostatic tissue is very glandular with plenty of water molecule movement.
-On the other hand, tumors have high cellular density and restricted water movement, which leads to decreased diffusion.
-Due to this restricted diffusion, tumors seem to be of higher intensity on DWI.13
-
- DWI sequences can be processed to obtain apparent diffusion coefficient (ADC) maps.
-These serve as a more objective measure of diffusivity since they are independent of magnetic field and thus overcome the effects of T2 shine-through.13,14  ADC maps are also used to visually assess the tumor, which appears as a focus of decreased signal intensity when compared to normal prostate tissue.14
-
- In addition to knowing the location of the cancer within the prostate, being able to identify which cancers will exhibit more aggressive behavior is important when making clinical management decisions.
-A number of studies have found that lower ADC values correlate with tumors that have higher Gleason scores.
-15-17 This could be explained by the fact that higher-grade lesions are usually of higher cellular density and thus have even more marked restricted diffusion.
-This inverse correlation suggests that ADC maps might be helpful when characterizing lesions and assessing tumor aggressiveness.18
-
- Dynamic contrast enhanced imaging
- DCE-MRI has been shown to have a high sensitivity in cancer detection, especially when combined with other MR imaging modalities.19-22 In a study by Hara et al, this functional modality identified 93\% of clinically significant cancers.23 DCE-MRI provides a measure of tumor vascularity and takes advantage of tumor-induced angiogenesis in PCa.
-24Increased vascularity in PCa results in earlier and higher peak contrast enhancement and a faster washout when compared to normal prostatic tissue.
-These known characteristics of PCa are helpful when trying to localize lesions.
-
- A variety of different approaches can be used to analyze DCE-MRI and no report has, up to date, described the superiority of one method over the others.3 The approaches include qualitative analysis by visual assessment of the images, semi-quantitative methods that look at kinetic parameters on a voxel-by-voxel basis, or quantitative methods that that convert signal intensity into kinetic parameters to determine how much contrast is being exchanged between the vascular, extravascular and extracellular spaces.3
-
-    MR spectroscopy
-    Like DWI, MRSI can be helpful in lesion characterization as it provides information on the presence of certain metabolites in tissue.
-8Choline and citrate are especially useful in the setting of PCa.
-Choline is critical in cell membrane synthesis and is thus usually elevated in cells with cancerous behavior.3 Healthy prostate epithelium synthesizes and secretes large quantities of citrate but levels are decreased in PCa.25,26 Therefore, an increase in the choline-to-citrate ratio on MRSI can be used as an indicator of malignancy.9
-
-    Although some data suggests that MRSI increases the accuracy of tumor volume detection and staging when used in addition to anatomic imaging, it is unclear whether it is superior to other modalities.9 ESUR lists MSRI as an optional modality for the diagnosis of PCa since it significantly lengthens exam time and requires specific technical expertise.8 Thus, it is not surprising that many academic centers in the United States do not include MSRI in their mpMRI prostate cancer protocols.3
+\input{mr_background}

### 2013-09-13 mr_background.tex
original Kirema background

Clinical Background and Significance 
Prostate cancer (PCa) is the most common noncutaneous malignancy among men in the United States. Approximately 1 in every 6 men will develop PCa during their lifetime, with the median age of diagnosis at 67. 1It is also the second leading cause of cancer related death, with 1 in 36 men dying from the disease. The National Cancer Institute estimates that 238,590 men will be diagnosed with PCa in 2013 and 29,720 will die from the disease.1

PCa diagnosis usually begins by screening with prostate specific antigen (PSA) and digital rectal examination (DRE).  Definitive diagnosis is made by random transrectal ultrasonography-guided (TRUS) biopsies, which are then used to provide the clinician with the proper Gleason score. The combination of these factors—as well as staging—determines the appropriate therapy and prognosis.

PCa screening has led to earlier diagnosis of smaller tumors and more localized disease.  However, it is well known that the sensitivity and specificity of PSA and DRE are not optimal. In addition, DRE has a low predictive value at lower PSA ranges, and PSA yields many false positives.2-4 As such, a theoretical risk of over-diagnosis and treatment of low-grade—and possibly clinically insignificant—disease exists. Moreover, due to the random nature of biopsies, cancer located outside the routine sampling site can be missed and the extent of the cancer might be underestimated.3 5 For example, in a study by Mufarrij et al 45.9%-47.2% of patients who were candidates for active surveillance but underwent radical prostatectomy had a higher Gleason score on final histopathology than after TRUS biopsy.6 These inaccuracies may lead to inappropriate diagnosis, imprecise risk assessment and potentially avoidable morbidity.

The Use of Magnetic Resonance Imaging in Prostate Diagnostics
Magnetic resonance (MR) imaging has been available for use in the workup of patients with PCa since the early 1980s but the early work on its diagnostic accuracy is heterogeneous.7 Earlier MR techniques relied mostly on morphology via T1 and T2- weighted imaging (T2WI). The more recent ability to include not only anatomic but also biologic and functional dynamic parameters into MR analysis—via diffusion-weighted imaging (DWI), dynamic contrast-enhanced (DCE) imaging or MR spectroscopic imaging (MRSI)—is promising in the future diagnosis and management of PCA.

 Currently prostate MR focuses on a multiparametric approach, where 2 or more imaging sequences—including anatomic and functional data—are used together to try to arrive to a diagnosis.8 As MR technology continues to evolve and improve, its role in PCA diagnosis, staging, treatment planning and follow-up has gained much attention.

T2-Weighted Imaging and Prostate Anatomy
T2WI sequences are crucial components of prostate MR imaging.  T2WI is particularly useful in prostate analysis due to its excellent soft tissue contrast resolution, which can be maximized by using thin sections of 3-4mm and a small field of view of approximately 14cm.3,9 T2 sequences are the most helpful for tumor localization as they can clearly show overall prostate morphology, internal structures and prostatic margins.3

The prostate can be divided into glandular and nonglandular components. The glandular components include the peripheral zone (PZ) and the central gland, which are typically easily distinguishable on T2WI. The central gland includes the central zone, transition zone and the periurethral glandular tissue.10 Other anatomical markers such as the urethra, verumontanum and ejaculatory are also often seen on T2WI.

Approximately 70% of the prostatic tissue is found in the peripheral zone, which is high in water content and thus of higher signal intensity in T2WI.10 Seventy five percent of prostatic tumors are found in the PZ and normally show hypointense T2 signal when compared to the higher intensity PZ.4 11 However tumors can sometimes seem of similar intensity as the surrounding tissue and false positives can occur secondary to post biopsy changes/hemorrhage, hyperplasia or prostatitis, making diagnosis more challenging.11

Functional MR Sequences
Even though T2WI is the mainstay of prostate MR, its overall performance in prostate cancer diagnosis is not optimal. The incorporation of two or more functional sequences in multiparametric MR imaging (mpMRI) has been shown to significantly improve the performance of MRI in cancer diagnosis.12 In fact, the European Society of Urogenital Radiology’s (ESUR) prostate MR guidelines recommend at least 2 functional imaging techniques in addition to T2WI in order to better characterize prostate tumors.8 In a study by Turkbey et al, researchers found that mpMRI had a positive predictive value of 98% in prostate cancer detection.12 Functional sequences include diffusion- weighted imaging (DWI), dynamic contrast enhanced imaging (DCE-MRI) and MR spectroscopic imaging (MRSI).

Diffusion-weighted imaging
DWI is based on the free movement of water particles in tissue and measures the degree of motion restriction.13 Normal prostatic tissue is very glandular with plenty of water molecule movement. On the other hand, tumors have high cellular density and restricted water movement, which leads to decreased diffusion. Due to this restricted diffusion, tumors seem to be of higher intensity on DWI.13

DWI sequences can be processed to obtain apparent diffusion coefficient (ADC) maps. These serve as a more objective measure of diffusivity since they are independent of magnetic field and thus overcome the effects of T2 shine-through.13,14  ADC maps are also used to visually assess the tumor, which appears as a focus of decreased signal intensity when compared to normal prostate tissue.14

In addition to knowing the location of the cancer within the prostate, being able to identify which cancers will exhibit more aggressive behavior is important when making clinical management decisions. A number of studies have found that lower ADC values correlate with tumors that have higher Gleason scores. 15-17 This could be explained by the fact that higher-grade lesions are usually of higher cellular density and thus have even more marked restricted diffusion. This inverse correlation suggests that ADC maps might be helpful when characterizing lesions and assessing tumor aggressiveness.18

Dynamic contrast enhanced imaging
DCE-MRI has been shown to have a high sensitivity in cancer detection, especially when combined with other MR imaging modalities.19-22 In a study by Hara et al, this functional modality identified 93% of clinically significant cancers.23 DCE-MRI provides a measure of tumor vascularity and takes advantage of tumor-induced angiogenesis in PCa. 24Increased vascularity in PCa results in earlier and higher peak contrast enhancement and a faster washout when compared to normal prostatic tissue. These known characteristics of PCa are helpful when trying to localize lesions.

A variety of different approaches can be used to analyze DCE-MRI and no report has, up to date, described the superiority of one method over the others.3 The approaches include qualitative analysis by visual assessment of the images, semi-quantitative methods that look at kinetic parameters on a voxel-by-voxel basis, or quantitative methods that that convert signal intensity into kinetic parameters to determine how much contrast is being exchanged between the vascular, extravascular and extracellular spaces.3

MR spectroscopy
Like DWI, MRSI can be helpful in lesion characterization as it provides information on the presence of certain metabolites in tissue. 8Choline and citrate are especially useful in the setting of PCa. Choline is critical in cell membrane synthesis and is thus usually elevated in cells with cancerous behavior.3 Healthy prostate epithelium synthesizes and secretes large quantities of citrate but levels are decreased in PCa.25,26 Therefore, an increase in the choline-to-citrate ratio on MRSI can be used as an indicator of malignancy.9

Although some data suggests that MRSI increases the accuracy of tumor volume detection and staging when used in addition to anatomic imaging, it is unclear whether it is superior to other modalities.9 ESUR lists MSRI as an optional modality for the diagnosis of PCa since it significantly lengthens exam time and requires specific technical expertise.8 Thus, it is not surprising that many academic centers in the United States do not include MSRI in their mpMRI prostate cancer protocols.3

1.	Howlader N, Noone AM, Krapcho M, Neyman N, Aminou R, Waldron W. SEER Cancer Statistics Review, 1975–2010, National Cancer Institute. Bethesda, MD, based on November 2012 SEER data submission, posted to the SEER web site, 2013. http://seer.cancer.gov/csr/1975_2010 (Accessed on June 08, 2013). 2011.
2.	Gosselaar C, Roobol MJ, Roemeling S, van der Kwast TH, Schröder FH. Screening for prostate cancer at low PSA range: the impact of digital rectal examination on tumor incidence and tumor characteristics. The Prostate. 2007;67(2):154-161 %@ 1097-0045.
3.	Gupta RT, Kauffman CR, Polascik TJ, Taneja SS, Rosenkrantz AB. The state of prostate MRI in 2013. Oncology. Apr 2013;27(4):262-270.
4.	Hricak H, Choyke PL, Eberhardt SC, Leibel SA, Scardino PT. Imaging prostate cancer: A multidisciplinary perspective1. Radiology. 2007;243(1):28-53 %@ 0033-8419.
5.	Cornud F, Delongchamps NB, Mozer P, et al. Value of Multiparametric MRI in the Work-up of Prostate Cancer. Current urology reports. 2012;13(1):82-92 %@ 1527-2737.
6.	Mufarrij P, Sankin A, Godoy G, Lepor H. Pathologic outcomes of candidates for active surveillance undergoing radical prostatectomy. Urology. 2010;76(3):689-692 %@ 0090-4295.
7.	Poon PY, McCallum RW, Henkelman MM, et al. Magnetic resonance imaging of the prostate. Radiology. 1985;154(1):143-149 %@ 0033-8419.
8.	Barentsz JO, Richenberg J, Clements R, et al. ESUR prostate MR guidelines 2012. European radiology. 2012;22(4):746-757 %@ 0938-7994.
9.	Bonekamp D, Jacobs MA, El-Khouli R, Stoianovici D, Macura KJ. Advancements in MR imaging of the prostate: from diagnosis to interventions. Radiographics. 2011;31(3):677-703 %@ 0271-5333.
10.	Jung AJ, Westphalen AC. Imaging Prostate Cancer. Radiologic Clinics of North America. 2012;50(6):1043-1059 %@ 0033-8389.
11.	Hegde JV, Mulkern RV, Panych LP, et al. Multiparametric MRI of prostate cancer: An update on state‐of‐the‐art techniques and their performance in detecting and localizing prostate cancer. Journal of Magnetic Resonance Imaging. 2013;37(5):1035-1054 %@ 1522-2586.
12.	Turkbey B, Choyke PL. Multiparametric MRI and prostate cancer diagnosis and risk stratification. Current opinion in urology. 2012;22(4):310-315 %@ 0963-0643.
13.	Koh D-M, Collins DJ. Diffusion-weighted MRI in the body: applications and challenges in oncology. American Journal of Roentgenology. 2007;188(6):1622-1635 %@ 0361-1803X.
14.	Tan CH, Wang J, Kundra V. Diffusion weighted imaging in prostate cancer. European radiology. 2011;21(3):593-603 %@ 0938-7994.
15.	Kobus T, Vos PC, Hambrock T, et al. Prostate cancer aggressiveness: in vivo assessment of MR spectroscopy and diffusion-weighted imaging at 3 T. Radiology. 2012;265(2):457-467 %@ 0033-8419.
16.	Woodfield CA, Tung GA, Grand DJ, Pezzullo JA, Machan JT, Renzulli JF. Diffusion-weighted MRI of peripheral zone prostate cancer: comparison of tumor apparent diffusion coefficient with Gleason score and percentage of tumor on core biopsy. American Journal of Roentgenology. 2010;194(4):W316-W322 %@ 0361-0803X.
17.	Hambrock T, Somford DM, Huisman HJ, et al. Relationship between apparent diffusion coefficients at 3.0-T MR imaging and Gleason grade in peripheral zone prostate cancer. Radiology. 2011;259(2):453-461 %@ 0033-8419.
18.	Vargas HA, Akin O, Franiel T, et al. Diffusion-weighted endorectal MR imaging at 3 T for prostate cancer: tumor detection and assessment of aggressiveness. Radiology. 2011;259(3):775-784 %@ 0033-8419.
19.	Tanimoto A, Nakashima J, Kohno H, Shinmoto H, Kuribayashi S. Prostate cancer screening: The clinical value of diffusion‐weighted imaging and dynamic MR imaging in combination with T2‐weighted imaging. Journal of Magnetic Resonance Imaging. 2007;25(1):146-152 %@ 1522-2586.
20.	Girouin N, Mège-Lechevallier F, Senes AT, et al. Prostate dynamic contrast-enhanced MRI with simple visual diagnostic criteria: is it reasonable? European radiology. 2007;17(6):1498-1509 %@ 0938-7994.
21.	Ocak I, Bernardo M, Metzger G, et al. Dynamic contrast-enhanced MRI of prostate cancer at 3 T: a study of pharmacokinetic parameters. American Journal of Roentgenology. 2007;189(4):W192-W201 %@ 0361-0803X.
22.	Kim JK, Hong SS, Choi YJ, et al. Wash‐in rate on the basis of dynamic contrast‐enhanced MRI: Usefulness for prostate cancer detection and localization. Journal of Magnetic Resonance Imaging. 2005;22(5):639-646 %@ 1522-2586.
23.	Hara N, Okuizumi M, Koike H, Kawaguchi M, Bilim V. Dynamic contrast‐enhanced magnetic resonance imaging (DCE‐MRI) is a useful modality for the precise detection and staging of early prostate cancer. The Prostate. 2005;62(2):140-147 %@ 1097-0045.
24.	Noworolski SM, Henry RG, Vigneron DB, Kurhanewicz J. Dynamic contrast‐enhanced MRI in normal and abnormal prostate tissues as defined by biopsy, MRI, and 3D MRSI. Magnetic resonance in medicine. 2005;53(2):249-255 %@ 1522-2594.
25.	Costello LC, Franklin RBa, Narayan P. Citrate in the diagnosis of prostate cancer. The prostate. 1999;38(3):237-245 %@ 1097-0045.
26.	Kurhanewicz J, Swanson MG, Nelson SJ, Vigneron DB. Combined magnetic resonance imaging and spectroscopic imaging approach to molecular imaging of prostate cancer. Journal of Magnetic Resonance Imaging. 2002;16(4):451-463 %@ 1522-2586.

### 2013-09-13 mr_background.tex
updated refs from Word -> LaTeX

@@
-Clinical Background and Significance 
-Prostate cancer (PCa) is the most common noncutaneous malignancy among men in the United States.
-Approximately 1 in every 6 men will develop PCa during their lifetime, with the median age of diagnosis at 67.
-1It is also the second leading cause of cancer related death, with 1 in 36 men dying from the disease.
-The National Cancer Institute estimates that 238,590 men will be diagnosed with PCa in 2013 and 29,720 will die from the disease.1
+\subsection{Clinical Background and Significance}
+Prostate cancer (PCa) is the most common noncutaneous malignancy among men in
+the United States. Approximately 1 in every 6 men will develop PCa during their
+lifetime, with the median age of diagnosis at 67. 1It is also the second
+leading cause of cancer related death, with 1 in 36 men dying from the disease.
+The National Cancer Institute estimates that 238,590 men will be diagnosed with
+PCa in 2013 and 29,720 will die from the disease.~\cite{Howlader2011}
 
-PCa diagnosis usually begins by screening with prostate specific antigen (PSA) and digital rectal examination (DRE).
-Definitive diagnosis is made by random transrectal ultrasonography-guided (TRUS) biopsies, which are then used to provide the clinician with the proper Gleason score.
-The combination of these factors—as well as staging—determines the appropriate therapy and prognosis.
+PCa diagnosis usually begins by screening with prostate specific antigen (PSA)
+and digital rectal examination (DRE).  Definitive diagnosis is made by random
+transrectal ultrasonography-guided (TRUS) biopsies, which are then used to
+provide the clinician with the proper Gleason score. The combination of these
+factors—as well as staging—determines the appropriate therapy and prognosis.
 
-PCa screening has led to earlier diagnosis of smaller tumors and more localized disease.
-However, it is well known that the sensitivity and specificity of PSA and DRE are not optimal.
-In addition, DRE has a low predictive value at lower PSA ranges, and PSA yields many false positives.2-4 As such, a theoretical risk of over-diagnosis and treatment of low-grade—and possibly clinically insignificant—disease exists.
-Moreover, due to the random nature of biopsies, cancer located outside the routine sampling site can be missed and the extent of the cancer might be underestimated.3 5 For example, in a study by Mufarrij et al 45.9%-47.2% of patients who were candidates for active surveillance but underwent radical prostatectomy had a higher Gleason score on final histopathology than after TRUS biopsy.6 These inaccuracies may lead to inappropriate diagnosis, imprecise risk assessment and potentially avoidable morbidity.
+PCa screening has led to earlier diagnosis of smaller tumors and more localized
+disease.  However, it is well known that the sensitivity and specificity of PSA
+and DRE are not optimal. In addition, DRE has a low predictive value at lower
+PSA ranges, and PSA yields many false
+positives.~\cite{Gosselaar2007,Gupta2013,Hricak2007} As such, a theoretical
+risk of over-diagnosis and treatment of low-grade—and possibly clinically
+insignificant—disease exists. Moreover, due to the random nature of biopsies,
+cancer located outside the routine sampling site can be missed and the extent
+of the cancer might be underestimated.~\cite{Gupta2013,Cornud1012} For example,
+in a study by Mufarrij et al 45.9--47.2\% of patients who were candidates for
+active surveillance but underwent radical prostatectomy had a higher Gleason
+score on final histopathology than after TRUS biopsy.~\cite{Mufarrij2010} These
+inaccuracies may lead to inappropriate diagnosis, imprecise risk assessment and
+potentially avoidable morbidity.
 
-The Use of Magnetic Resonance Imaging in Prostate Diagnostics
-Magnetic resonance (MR) imaging has been available for use in the workup of patients with PCa since the early 1980s but the early work on its diagnostic accuracy is heterogeneous.7 Earlier MR techniques relied mostly on morphology via T1 and T2- weighted imaging (T2WI).
-The more recent ability to include not only anatomic but also biologic and functional dynamic parameters into MR analysis—via diffusion-weighted imaging (DWI), dynamic contrast-enhanced (DCE) imaging or MR spectroscopic imaging (MRSI)—is promising in the future diagnosis and management of PCA.
+\subsection*{The Use of Magnetic Resonance Imaging in Prostate Diagnostics}
+Magnetic resonance (MR) imaging has been available for use in the workup of
+patients with PCa since the early 1980s but the early work on its diagnostic
+accuracy is heterogeneous.7 Earlier MR techniques relied mostly on morphology
+via T1 and T2- weighted imaging (T2WI). The more recent ability to include not
+only anatomic but also biologic and functional dynamic parameters into MR
+analysis—via diffusion-weighted imaging (DWI), dynamic contrast-enhanced (DCE)
+imaging or MR spectroscopic imaging (MRSI)—is promising in the future diagnosis
+and management of PCA.
 
- Currently prostate MR focuses on a multiparametric approach, where 2 or more imaging sequences—including anatomic and functional data—are used together to try to arrive to a diagnosis.8 As MR technology continues to evolve and improve, its role in PCA diagnosis, staging, treatment planning and follow-up has gained much attention.
+Currently prostate MR focuses on a multiparametric approach, where 2 or more
+imaging sequences—including anatomic and functional data—are used together to
+try to arrive to a diagnosis.~\cite{Barentsz2012} As MR technology continues to evolve and
+improve, its role in PCA diagnosis, staging, treatment planning and follow-up
+has gained much attention.
 
-T2-Weighted Imaging and Prostate Anatomy
-T2WI sequences are crucial components of prostate MR imaging.
-T2WI is particularly useful in prostate analysis due to its excellent soft tissue contrast resolution, which can be maximized by using thin sections of 3-4mm and a small field of view of approximately 14cm.3,9 T2 sequences are the most helpful for tumor localization as they can clearly show overall prostate morphology, internal structures and prostatic margins.3
+\subsection{T2-Weighted Imaging and Prostate Anatomy}
+T2WI sequences are crucial components of prostate MR imaging.  T2WI is
+particularly useful in prostate analysis due to its excellent soft tissue
+contrast resolution, which can be maximized by using thin sections of 3--4 mm
+and a small field of view of approximately 14 cm.~\cite{Gupta2013,Bonekamp2011}
+T2 sequences are the most helpful for tumor localization as they can clearly
+show overall prostate morphology, internal structures and prostatic
+margins.~\cite{Gupta2013}
 
-The prostate can be divided into glandular and nonglandular components.
-The glandular components include the peripheral zone (PZ) and the central gland, which are typically easily distinguishable on T2WI.
-The central gland includes the central zone, transition zone and the periurethral glandular tissue.10 Other anatomical markers such as the urethra, verumontanum and ejaculatory are also often seen on T2WI.
+The prostate can be divided into glandular and nonglandular components. The
+glandular components include the peripheral zone (PZ) and the central gland,
+which are typically easily distinguishable on T2WI. The central gland includes
+the central zone, transition zone and the periurethral glandular
+tissue.~\cite{Jung2012} Other anatomical markers such as the urethra,
+verumontanum and ejaculatory are also often seen on T2WI.
 
-Approximately 70% of the prostatic tissue is found in the peripheral zone, which is high in water content and thus of higher signal intensity in T2WI.10 Seventy five percent of prostatic tumors are found in the PZ and normally show hypointense T2 signal when compared to the higher intensity PZ.4 11 However tumors can sometimes seem of similar intensity as the surrounding tissue and false positives can occur secondary to post biopsy changes/hemorrhage, hyperplasia or prostatitis, making diagnosis more challenging.11
+Approximately 70\% of the prostatic tissue is found in the peripheral zone,
+which is high in water content and thus of higher signal intensity in
+T2WI.~\cite{Jung2012} Seventy five percent of prostatic tumors are found in the
+PZ and normally show hypointense T2 signal when compared to the higher
+intensity PZ.~\cite{Hricak2007,Hegde2013} However, tumors can sometimes seem of
+similar intensity as the surrounding tissue and false positives can occur
+secondary to post biopsy changes/hemorrhage, hyperplasia or prostatitis, making
+diagnosis more challenging.~\cite{Hegde2013}
 
-Functional MR Sequences
-Even though T2WI is the mainstay of prostate MR, its overall performance in prostate cancer diagnosis is not optimal.
-The incorporation of two or more functional sequences in multiparametric MR imaging (mpMRI) has been shown to significantly improve the performance of MRI in cancer diagnosis.12 In fact, the European Society of Urogenital Radiology’s (ESUR) prostate MR guidelines recommend at least 2 functional imaging techniques in addition to T2WI in order to better characterize prostate tumors.8 In a study by Turkbey et al, researchers found that mpMRI had a positive predictive value of 98% in prostate cancer detection.12 Functional sequences include diffusion- weighted imaging (DWI), dynamic contrast enhanced imaging (DCE-MRI) and MR spectroscopic imaging (MRSI).
+\subsection{Functional MR Sequences}
+Even though T2WI is the mainstay of prostate MR, its overall performance in
+prostate cancer diagnosis is not optimal. The incorporation of two or more
+functional sequences in multiparametric MR imaging (mpMRI) has been shown to
+significantly improve the performance of MRI in cancer
+diagnosis.~\cite{Turkbey2012} In fact, the European Society of Urogenital
+Radiology’s (ESUR) prostate MR guidelines recommend at least 2 functional
+imaging techniques in addition to T2WI in order to better characterize prostate
+tumors.8 In a study by Turkbey et al, researchers found that mpMRI had a
+positive predictive value of 98\% in prostate cancer
+detection.~\cite{Turkbey2012} Functional sequences include diffusion- weighted
+imaging (DWI), dynamic contrast enhanced imaging (DCE-MRI) and MR spectroscopic
+imaging (MRSI).
 
-Diffusion-weighted imaging
-DWI is based on the free movement of water particles in tissue and measures the degree of motion restriction.13 Normal prostatic tissue is very glandular with plenty of water molecule movement.
-On the other hand, tumors have high cellular density and restricted water movement, which leads to decreased diffusion.
-Due to this restricted diffusion, tumors seem to be of higher intensity on DWI.13
+\subsection{Diffusion-weighted imaging}
+DWI is based on the free movement of water particles in tissue and measures the
+degree of motion restriction.13 Normal prostatic tissue is very glandular with
+plenty of water molecule movement. On the other hand, tumors have high cellular
+density and restricted water movement, which leads to decreased diffusion. Due
+to this restricted diffusion, tumors seem to be of higher intensity on
+DWI.~\cite{Koh2007}
 
-DWI sequences can be processed to obtain apparent diffusion coefficient (ADC) maps.
-These serve as a more objective measure of diffusivity since they are independent of magnetic field and thus overcome the effects of T2 shine-through.13,14  ADC maps are also used to visually assess the tumor, which appears as a focus of decreased signal intensity when compared to normal prostate tissue.14
+DWI sequences can be processed to obtain apparent diffusion coefficient (ADC)
+maps. These serve as a more objective measure of diffusivity since they are
+independent of magnetic field and thus overcome the effects of T2
+shine-through.~\cite{Turkbey2012,Koh2007}  ADC maps are also used to visually
+assess the tumor, which appears as a focus of decreased signal intensity when
+compared to normal prostate tissue.~\cite{Tan2011}
 
-In addition to knowing the location of the cancer within the prostate, being able to identify which cancers will exhibit more aggressive behavior is important when making clinical management decisions.
-A number of studies have found that lower ADC values correlate with tumors that have higher Gleason scores.
-15-17 This could be explained by the fact that higher-grade lesions are usually of higher cellular density and thus have even more marked restricted diffusion.
-This inverse correlation suggests that ADC maps might be helpful when characterizing lesions and assessing tumor aggressiveness.18
+In addition to knowing the location of the cancer within the prostate, being
+able to identify which cancers will exhibit more aggressive behavior is
+important when making clinical management decisions. A number of studies have
+found that lower ADC values correlate with tumors that have higher Gleason
+scores.~\cite{Kobus2012,Woodfield2010,Hambrock2011} This could be explained by
+the fact that higher-grade lesions are usually of higher cellular density and
+thus have even more marked restricted diffusion. This inverse correlation
+suggests that ADC maps might be helpful when characterizing lesions and
+assessing tumor aggressiveness.~\cite{Vargas2011}
 
-Dynamic contrast enhanced imaging
-DCE-MRI has been shown to have a high sensitivity in cancer detection, especially when combined with other MR imaging modalities.19-22 In a study by Hara et al, this functional modality identified 93% of clinically significant cancers.23 DCE-MRI provides a measure of tumor vascularity and takes advantage of tumor-induced angiogenesis in PCa.
-24Increased vascularity in PCa results in earlier and higher peak contrast enhancement and a faster washout when compared to normal prostatic tissue.
+\subsection{Dynamic contrast enhanced imaging}
+DCE-MRI has been shown to have a high sensitivity in cancer detection,
+especially when combined with other MR imaging
+modalities.~\cite{Tanimoto2007,Girouin2007,Ocak2007,Kim2005} In a study by Hara
+et al, this functional modality identified 93\% of clinically significant
+cancers.~\cite{Hara2005} DCE-MRI provides a measure of tumor vascularity and
+takes advantage of tumor-induced angiogenesis in PCa.~\cite{Noworolski2005}
+Increased vascularity in PCa results in earlier and higher peak contrast
+enhancement and a faster washout when compared to normal prostatic tissue.
 These known characteristics of PCa are helpful when trying to localize lesions.
 
-A variety of different approaches can be used to analyze DCE-MRI and no report has, up to date, described the superiority of one method over the others.3 The approaches include qualitative analysis by visual assessment of the images, semi-quantitative methods that look at kinetic parameters on a voxel-by-voxel basis, or quantitative methods that that convert signal intensity into kinetic parameters to determine how much contrast is being exchanged between the vascular, extravascular and extracellular spaces.3
+A variety of different approaches can be used to analyze DCE-MRI and no report
+has, up to date, described the superiority of one method over the others.3 The
+approaches include qualitative analysis by visual assessment of the images,
+semi-quantitative methods that look at kinetic parameters on a voxel-by-voxel
+basis, or quantitative methods that that convert signal intensity into kinetic
+parameters to determine how much contrast is being exchanged between the
+vascular, extravascular and extracellular spaces.~\cite{Gupta2013}
 
-MR spectroscopy
-Like DWI, MRSI can be helpful in lesion characterization as it provides information on the presence of certain metabolites in tissue.
-8Choline and citrate are especially useful in the setting of PCa.
-Choline is critical in cell membrane synthesis and is thus usually elevated in cells with cancerous behavior.3 Healthy prostate epithelium synthesizes and secretes large quantities of citrate but levels are decreased in PCa.25,26 Therefore, an increase in the choline-to-citrate ratio on MRSI can be used as an indicator of malignancy.9
+\subsection{MR spectroscopy}
+Like DWI, MRSI can be helpful in lesion characterization as it provides
+information on the presence of certain metabolites in
+tissue.~\cite{Barentsz2012} Choline and citrate are especially useful in the
+setting of PCa. Choline is critical in cell membrane synthesis and is thus
+usually elevated in cells with cancerous behavior.3 Healthy prostate epithelium
+synthesizes and secretes large quantities of citrate but levels are decreased
+in PCa.~\cite{Costello1999,Kurhanewicz2002} Therefore, an increase in the
+choline-to-citrate ratio on MRSI can be used as an indicator of
+malignancy.~\cite{Bonekamp2011}
 
-Although some data suggests that MRSI increases the accuracy of tumor volume detection and staging when used in addition to anatomic imaging, it is unclear whether it is superior to other modalities.9 ESUR lists MSRI as an optional modality for the diagnosis of PCa since it significantly lengthens exam time and requires specific technical expertise.8 Thus, it is not surprising that many academic centers in the United States do not include MSRI in their mpMRI prostate cancer protocols.3
-
-1.	Howlader N, Noone AM, Krapcho M, Neyman N, Aminou R, Waldron W.
-SEER Cancer Statistics Review, 1975–2010, National Cancer Institute.
-Bethesda, MD, based on November 2012 SEER data submission, posted to the SEER web site, 2013.
-http://seer.cancer.gov/csr/1975_2010 (Accessed on June 08, 2013).
-2011.
-2.	Gosselaar C, Roobol MJ, Roemeling S, van der Kwast TH, Schröder FH.
-Screening for prostate cancer at low PSA range: the impact of digital rectal examination on tumor incidence and tumor characteristics.
-The Prostate.
-2007;67(2):154-161 %@ 1097-0045.
-3.	Gupta RT, Kauffman CR, Polascik TJ, Taneja SS, Rosenkrantz AB. The state of prostate MRI in 2013. Oncology. Apr 2013;27(4):262-270.
-4.	Hricak H, Choyke PL, Eberhardt SC, Leibel SA, Scardino PT. Imaging prostate cancer: A multidisciplinary perspective1. Radiology. 2007;243(1):28-53 %@ 0033-8419.
-5.	Cornud F, Delongchamps NB, Mozer P, et al. Value of Multiparametric MRI in the Work-up of Prostate Cancer. Current urology reports. 2012;13(1):82-92 %@ 1527-2737.
-6.	Mufarrij P, Sankin A, Godoy G, Lepor H. Pathologic outcomes of candidates for active surveillance undergoing radical prostatectomy. Urology. 2010;76(3):689-692 %@ 0090-4295.
-7.	Poon PY, McCallum RW, Henkelman MM, et al. Magnetic resonance imaging of the prostate. Radiology. 1985;154(1):143-149 %@ 0033-8419.
-8.	Barentsz JO, Richenberg J, Clements R, et al. ESUR prostate MR guidelines 2012. European radiology. 2012;22(4):746-757 %@ 0938-7994.
-9.	Bonekamp D, Jacobs MA, El-Khouli R, Stoianovici D, Macura KJ. Advancements in MR imaging of the prostate: from diagnosis to interventions. Radiographics. 2011;31(3):677-703 %@ 0271-5333.
-10.	Jung AJ, Westphalen AC. Imaging Prostate Cancer. Radiologic Clinics of North America. 2012;50(6):1043-1059 %@ 0033-8389.
-11.	Hegde JV, Mulkern RV, Panych LP, et al.
-Multiparametric MRI of prostate cancer: An update on state‐of‐the‐art techniques and their performance in detecting and localizing prostate cancer.
-Journal of Magnetic Resonance Imaging.
-2013;37(5):1035-1054 %@ 1522-2586.
-12.	Turkbey B, Choyke PL. Multiparametric MRI and prostate cancer diagnosis and risk stratification. Current opinion in urology. 2012;22(4):310-315 %@ 0963-0643.
-13.	Koh D-M, Collins DJ. Diffusion-weighted MRI in the body: applications and challenges in oncology. American Journal of Roentgenology. 2007;188(6):1622-1635 %@ 0361-1803X.
-14.	Tan CH, Wang J, Kundra V. Diffusion weighted imaging in prostate cancer. European radiology. 2011;21(3):593-603 %@ 0938-7994.
-15.	Kobus T, Vos PC, Hambrock T, et al. Prostate cancer aggressiveness: in vivo assessment of MR spectroscopy and diffusion-weighted imaging at 3 T. Radiology. 2012;265(2):457-467 %@ 0033-8419.
-16.	Woodfield CA, Tung GA, Grand DJ, Pezzullo JA, Machan JT, Renzulli JF.
-Diffusion-weighted MRI of peripheral zone prostate cancer: comparison of tumor apparent diffusion coefficient with Gleason score and percentage of tumor on core biopsy.
-American Journal of Roentgenology.
-2010;194(4):W316-W322 %@ 0361-0803X.
-17.	Hambrock T, Somford DM, Huisman HJ, et al.
-Relationship between apparent diffusion coefficients at 3.0-T MR imaging and Gleason grade in peripheral zone prostate cancer.
-Radiology.
-2011;259(2):453-461 %@ 0033-8419.
-18.	Vargas HA, Akin O, Franiel T, et al.
-Diffusion-weighted endorectal MR imaging at 3 T for prostate cancer: tumor detection and assessment of aggressiveness.
-Radiology.
-2011;259(3):775-784 %@ 0033-8419.
-19.	Tanimoto A, Nakashima J, Kohno H, Shinmoto H, Kuribayashi S.
-Prostate cancer screening: The clinical value of diffusion‐weighted imaging and dynamic MR imaging in combination with T2‐weighted imaging.
-Journal of Magnetic Resonance Imaging.
-2007;25(1):146-152 %@ 1522-2586.
-20.	Girouin N, Mège-Lechevallier F, Senes AT, et al.
-Prostate dynamic contrast-enhanced MRI with simple visual diagnostic criteria: is it reasonable?
-European radiology.
-2007;17(6):1498-1509 %@ 0938-7994.
-21.	Ocak I, Bernardo M, Metzger G, et al.
-Dynamic contrast-enhanced MRI of prostate cancer at 3 T: a study of pharmacokinetic parameters.
-American Journal of Roentgenology.
-2007;189(4):W192-W201 %@ 0361-0803X.
-22.	Kim JK, Hong SS, Choi YJ, et al.
-Wash‐in rate on the basis of dynamic contrast‐enhanced MRI: Usefulness for prostate cancer detection and localization.
-Journal of Magnetic Resonance Imaging.
-2005;22(5):639-646 %@ 1522-2586.
-23.	Hara N, Okuizumi M, Koike H, Kawaguchi M, Bilim V.
-Dynamic contrast‐enhanced magnetic resonance imaging (DCE‐MRI) is a useful modality for the precise detection and staging of early prostate cancer.
-The Prostate.
-2005;62(2):140-147 %@ 1097-0045.
-24.	Noworolski SM, Henry RG, Vigneron DB, Kurhanewicz J.
-Dynamic contrast‐enhanced MRI in normal and abnormal prostate tissues as defined by biopsy, MRI, and 3D MRSI.
-Magnetic resonance in medicine.
-2005;53(2):249-255 %@ 1522-2594.
-25.	Costello LC, Franklin RBa, Narayan P. Citrate in the diagnosis of prostate cancer. The prostate. 1999;38(3):237-245 %@ 1097-0045.
-26.	Kurhanewicz J, Swanson MG, Nelson SJ, Vigneron DB.
-Combined magnetic resonance imaging and spectroscopic imaging approach to molecular imaging of prostate cancer.
-Journal of Magnetic Resonance Imaging.
-2002;16(4):451-463 %@ 1522-2586.
+Although some data suggests that MRSI increases the accuracy of tumor volume
+detection and staging when used in addition to anatomic imaging, it is unclear
+whether it is superior to other modalities.~\cite{Bonekamp2011} ESUR lists MSRI
+as an optional modality for the diagnosis of PCa since it significantly
+lengthens exam time and requires specific technical
+expertise.~\cite{Barentsz2012} Thus, it is not surprising that many academic
+centers in the United States do not include MSRI in their mpMRI prostate cancer
+protocols.~\cite{Gupta2013}

### 2013-09-13 compile.sh
updated keywords and title / author lists

@@ -1,12 +1,0 @@
-#!/bin/bash
-texfile='mr_arfi_anatomy'
-pdflatex $texfile.tex
-pdflatex $texfile.tex
-bibtex $texfile
-pdflatex $texfile.tex
-pdflatex $texfile.tex
-rm $texfile.aux $texfile.bbl $texfile.out $texfile.blg $texfile.log
-
-#response='ResponseToReviewsR2'
-#pdflatex $response.tex
-#rm $response.aux $response.log

### 2013-09-13 compile.sh
latexmk compilation

@@ -1,0 +1,1 @@
+latexmk -bibtex -pdf mr_arfi_anatomy.tex

### 2013-09-13 intro.tex
using continual compilation

\section{Introduction}\label{sect:intro}

The introduction text will go here.

### 2013-09-13 compile.sh

@@ -1,1 +1,1 @@
-latexmk -bibtex -pdf mr_arfi_anatomy.tex
+latexmk -bibtex -pdf -pvc mr_arfi_anatomy.tex

### 2013-09-13 mr_background.tex
updated new refs and anatomy background

@@
 imaging techniques in addition to T2WI in order to better characterize prostate
-tumors.8 In a study by Turkbey et al, researchers found that mpMRI had a
+tumors.~\cite{Barentsz2012} In a study by Turkbey et al., researchers found that mpMRI had a
 positive predictive value of 98\% in prostate cancer

### 2013-09-13 prostate_anatomy.tex
prostate anatomy background

\subsection{Prostate Anatomy}

The prostate gland sits below the urinary bladder and surrounds the urethra.
Its superior borders include the bladder and seminal vesicles and the
urogenital diaphragm delineates its inferior boundary. The gland is bordered
anteriorly by the pubic symphysis and posteriorly by the rectum.  The prostate
is separated from the rectum by 2--3 mm fascial layer,~\cite{Jung2012} and can
be easily palpated on rectal examination.

The gland can be divided from superior to inferior into the base, midgland and
apex. The urethra enters the prostate proximally at the base and extends to the
midgland at which point the ejaculatory ducts open into the urethra at the
verumontanum.~\cite{Jung2012} The urethra then continues past the apex and
travels through the penis. The prostate can be divided into glandular and
non-glandular components.  The glandular components include the transitional
zone, central zone and peripheral zone. Each zone contains approximately 5\%,
20\% and 70--80\% of glandular tissue, respectively.~\cite{Bonekamp2011} The
non-glandular components include the anterior fibromuscular stroma and the
urethra.

Although not a true capsule, an outer band of fibromuscular tissue surrounds
the prostate.~\cite{Bonekamp2011} This ``capsule'' is important when assessing
the extraprostatic extension of cancer as tumor can spread by disrupting this
tissue. Two neurovascular bundles course posterior and lateral to the prostate,
which can also be invaded by malignant cells.

### 2013-09-13 discussion.tex
notes

\section{Discussion}\label{sect:discussion}

Discussion text will go here.

### 2013-09-13 results.tex
notes

\section{Results}\label{sect:results}

\begin{itemize}
    \item Images of select cases, demonstrating segmentation process
    \item Table of volumes in MR and ARFI images
    \item Table comparing measurements
\end{itemize}

### 2013-09-13 conclusions.tex
notes added

\section{Conclusions}
The delineation of prostate zonal anatomy in ARFI images has been compared with
the established methods for identifying zonal anatomy using T2 MR images.  In
XX cases of prostates containing varying degrees of PCa and BPH, perpipheral
zone volumes\ldots XXXXX and central gland volumes\ldots XXXXX.  Aspect ratios
of the central glad agreed to within XX\% between MR and ARFI imaging datasets.
Appreciable amounts of BPH made determining the transition between the central
gland and peripheral zone difficult to discern and, these cases were not
included in this analysis.  Additionally, large PCa lesions can also distort
prostate zonal anatomy appreciably, making the distinction between peripheral
zone and central gland challenging.  Overall, ARFI imaging is able to delineate
central gland from peripheral zone in the prostate in the absence of excessive
BPH or PCa tumor burden.

### 2013-09-13 discussion.tex
notes added

@@
 
-Discussion text will go here.
+\begin{itemize}
+    \item BPH and PCa can heavily compromise ability to see zones.
+No healthy prostates were studied, which would have made this analysis easier; however, the prostate of the older man, which would be in our screening population, is not that of the healthier, younger male, making analysis in that population less meaningful.
+\end{itemize}

### 2013-09-13 tab_mr_arfi_volumes.tex
volume comparison data and results table

\begin{table}
\centering
\caption{Comparison of Peripheral Zone and Central Gland Volumes in MR T2WI and ARFI Imaging}
\begin{tabular}{|l|l|l|l|l|} \hline
Study Subject ID & MR Central Gland & MR Total & ARFI Central Zone & ARFI Total \\ 
& Volume (mm$^3$) & Volume (mm$^3$) & Volume (mm$^3$) & Volume (mm$^3$) \\ \hline
P86 & 12738.12 & 24567.06 & 14296. & 40031. \\
P79 & 14263.86 & 28514.79 & 8370 & 33978 \\
P78 & 23471.81 & 32484.36 & 13290.47 & 37420 \\
P77 & 17317.03 & 32492.83 & 10829.7 & 31819 \\
P75 & 57561.69 & 70950.98 & 30369 & 73684.64 \\
P74 & 12010.3 & 27841.21 & 15679 & 43298.34 \\
P72 & 8824.44 & 19587.02 & 11776.44 & 28545.79 \\
P71 & 10969.65 & 21276.98 & 16020.73 & 31424.61 \\
P70 & 13633.36 & 20749.91 & 19282 & 36876.3\\
P69 & 23580.12 & 36112.58 & 33500 & 70654 \\
P67 & 16569.17 & 27332.39 & 25219 & 34656.2 \\
P61 & 25381.36 & 49213.93 & 18137 & 54044.3 \\
P59 & 9247.34 & 26360.37 & 6144 & 35137 \\
P58 & 14787.88 & 23361.47 & 19654.6 & 35211 \\
P57 & 17872.15 & 35371.12 & 10153 & 38028.47 \\
P56 & 33315.41 & 48495.39 & 39351 & 66425 \\ \hline
\end{tabular}
\label{tab:mr_arfi_volumes}
\end{table}

### 2013-09-13 mr_background.tex
fixed citations; more to be fixed

@@
 DWI is based on the free movement of water particles in tissue and measures the
-degree of motion restriction.13 Normal prostatic tissue is very glandular with
+degree of motion restriction.~\cite{Koh2007} Normal prostatic tissue is very glandular with
 plenty of water molecule movement. On the other hand, tumors have high cellular
@@
 A variety of different approaches can be used to analyze DCE-MRI and no report
-has, up to date, described the superiority of one method over the others.3 The
+has, up to date, described the superiority of one method over the others.~\cite{Gupta2013} The
 approaches include qualitative analysis by visual assessment of the images,

### 2013-09-13 figs/mr_arfi_volumes.m
stacked bar plots of MR and ARFI volumens

@@ -1,0 +1,61 @@
+% bar_stacked_volumes.m
+%
+% Generate stacked bar graphs comparing the central and total prostate volumes
+% in MR and ARFI images.  Data are loaded in from the CSV file exported from
+% Tyler's Google spreadsheet.
+%
+% Mark Palmeri
+% mlp6@duke.edu
+% 2013-09-13
+
+vols = csvread('../data/Prostate_CZ_PZ_Volume_Axis_Measurements.csv');
+
+mr_cg_vol = vols(:,2);
+mr_total_vol = vols(:,3);
+mr_pz_vol = mr_total_vol - mr_cg_vol;
+arfi_cz_vol = vols(:,10);
+arfi_total_vol = vols(:,11);
+arfi_pz_vol = arfi_total_vol - arfi_cz_vol;
+
+MRdataStack = [mr_cg_vol, mr_pz_vol];
+ARFIdataStack = [arfi_cz_vol, arfi_pz_vol];
+% convert from mm^3 -> cm^3
+MRdataStack = MRdataStack/1e3;
+ARFIdataStack = ARFIdataStack/1e3;
+
+% fontsize
+fs = 18;
+bar_width = 0.3;
+
+figure;
+hold on;
+h=bar([1:length(MRdataStack)],MRdataStack,bar_width,'stacked');
+g=bar([1:length(ARFIdataStack)]+bar_width,ARFIdataStack,bar_width,'stacked');
+set(h(1),'FaceColor',[0 0 1]);
+set(h(2),'FaceColor',[1 0 0]);
+set(g(1),'FaceColor',[0 1 0]);
+set(g(2),'FaceColor',[1 1 0]);
+colormap(winter)
+ylabel('Zone Volumes (cm^3)','FontSize',fs)
+xlabel('Study Subject','FontSize',fs);
+title('MR (Red/Blue) and ARFI (Yellow/Green) Zone Volumes','FontSize',fs);
+a=axis;
+a(2) = 17;
+axis(a);
+legend('Central Gland','Peripheral Zone','Location','NorthWest');
+legend boxoff;
+
+set(gca, ...
+    'Box'         , 'off'     , ...
+    'TickDir'     , 'out'     , ...
+    'TickLength'  , [.01 .01] , ...
+    'XMinorTick'  , 'off'      , ...
+    'YMinorTick'  , 'off'      , ...
+    'XGrid'       , 'off'      , ...
[…]

### 2013-09-13 fig_mr_arfi_volumes.tex
text surrounding the MR / ARFI volume figs

\begin{figure}[htb!]
\centering
\includegraphics[width=0.5\linewidth]{figs/mr_arfi_volumes.pdf}
\caption{MR and ARFI volumes - need more text here}
\label{fig:mr_arfi_volumes} 
\end{figure}

### 2013-09-13 figs/mr_arfi_volumes.m
MR:ARFI total volume comparisons

@@ -59,3 +59,21 @@ set(gca, ...
     'FontSize'    , fs);
 
 print('-depsc2','mr_arfi_volumes.eps');
+
+% lets do a little more processing to figure out what is going on... first the
+% relative volume in ARFI relative to MR
+
+arfi_mr_vol_diff = ((arfi_total_vol./mr_total_vol) - 1)*100;
+figure;
+bar([1:length(MRdataStack)],arfi_mr_vol_diff,0.5);
+title('ARFI:MR Total Volume Comparison','FontSize',fs);
+xlabel('Study Subject','FontSize',fs);
+ylabel('Percent Difference','FontSize',fs);
+set(gca,'FontSize',fs);
+a=axis;
+a(2) = 17;
+a(3) = -5;
+a(4) = 110;
+axis(a);
+text(0.5,100,sprintf('Mean Diff = %.1f +/- %.1f',mean(arfi_mr_vol_diff),std(arfi_mr_vol_diff)),'FontSize',18);
+print('-depsc2','mr_arfi_volume_diff.eps');

### 2013-09-13 figs/mr_arfi_volumes.m
comparison of central : total volumes

@@ -35,7 +35,6 @@ set(h(1),'FaceColor',[0 0 1]);
 set(h(2),'FaceColor',[1 0 0]);
 set(g(1),'FaceColor',[0 1 0]);
 set(g(2),'FaceColor',[1 1 0]);
-colormap(winter)
 ylabel('Zone Volumes (cm^3)','FontSize',fs)
 xlabel('Study Subject','FontSize',fs);
 title('MR (Red/Blue) and ARFI (Yellow/Green) Zone Volumes','FontSize',fs);
@@ -75,5 +74,29 @@ a(2) = 17;
 a(3) = -5;
 a(4) = 110;
 axis(a);
-text(0.5,100,sprintf('Mean Diff = %.1f +/- %.1f',mean(arfi_mr_vol_diff),std(arfi_mr_vol_diff)),'FontSize',18);
+text(0.5,100,sprintf('Mean Diff = %.1f +/- %.1f',mean(arfi_mr_vol_diff),std(arfi_mr_vol_diff)),'FontSize',fs);
 print('-depsc2','mr_arfi_volume_diff.eps');
+
+% now lets look at the relative ratios of CG:total volume for each modality
+mr_cg_total = (mr_cg_vol./mr_total_vol)*100;
+arfi_cz_total = (arfi_cz_vol./arfi_total_vol)*100;
+
+figure;
+hold on;
+h=bar([1:length(MRdataStack)],mr_cg_total,bar_width);
+set(h(1),'FaceColor',[0 0 1]);
+g=bar([1:length(ARFIdataStack)]+bar_width,arfi_cz_total,bar_width);
+set(g(1),'FaceColor',[0 1 0]);
+xlabel('Study Subject','FontSize',fs);
+ylabel('Ratio of Central : Total Volume','FontSize',fs);
+title('MR and ARFI Central : Total Ratios','FontSize',fs);
+set(gca,'FontSize',fs);
+a = axis;
+a(2) = 17;
+a(4) = 100;
+axis(a);
+
+text(1,90,sprintf('Mean Diff = %.1f +/- %.1f',mean(arfi_cz_total-mr_cg_total),std(arfi_cz_total-mr_cg_total)),'FontSize',fs);
+
+print('-depsc2','mr_arfi_central_total_diff.eps');
+

### 2013-09-13 fig_mr_arfi_volumes.tex
updated caption and figures for mr:arfi volumes

@@
 \centering
-\includegraphics[width=0.5\linewidth]{figs/mr_arfi_volumes.pdf}
-\caption{MR and ARFI volumes - need more text here}
+\begin{tabular}{ccc}
+\includegraphics[width=0.3\linewidth]{figs/mr_arfi_volumes.pdf} &
+\includegraphics[width=0.3\linewidth]{figs/mr_arfi_volume_diff.pdf} &
+\includegraphics[width=0.3\linewidth]{figs/mr_arfi_central_total_diff.pdf} \\
+(a) Central:Peripheral Volume & (b) Total Volume Difference & (c) Central:Total Volume Difference\\
+\end{tabular}
+\caption{Comparison of MR and ARFI zonal anatomy volume estimates from manually-segmented images.
+Total prostate volumes ranged from 19.6--71.0 cm$^3$ based on MR image models (a), with ARFI image models overestimating total prostate volume by 36.7 $\pm$ 27.9\% (b).
+ARFI image delineation of the central zone volume relative to total prostate volume showed a mean underestimation of -15.3 $\pm$ 13.0\% (c, green) relative to the MR central gland : total volume ratios (c, blue).}
 \label{fig:mr_arfi_volumes} 

### 2013-09-13 figs/convert_eps_to_pdf.sh
updated figures; need EPS intermediate to get correct bounding box

@@ -1,0 +1,1 @@
+for i in `ls *.eps`; do epspdf $i; done;

### 2013-09-13 figs/mr_arfi_volumes.m

@@ -58,6 +58,7 @@ set(gca, ...
     'FontSize'    , fs);
 
 print('-depsc2','mr_arfi_volumes.eps');
+close;
 
 % lets do a little more processing to figure out what is going on... first the
 % relative volume in ARFI relative to MR
@@ -76,6 +77,7 @@ a(4) = 110;
 axis(a);
 text(0.5,100,sprintf('Mean Diff = %.1f +/- %.1f',mean(arfi_mr_vol_diff),std(arfi_mr_vol_diff)),'FontSize',fs);
 print('-depsc2','mr_arfi_volume_diff.eps');
+close;
 
 % now lets look at the relative ratios of CG:total volume for each modality
 mr_cg_total = (mr_cg_vol./mr_total_vol)*100;
@@ -99,4 +101,6 @@ axis(a);
 text(1,90,sprintf('Mean Diff = %.1f +/- %.1f',mean(arfi_cz_total-mr_cg_total),std(arfi_cz_total-mr_cg_total)),'FontSize',fs);
 
 print('-depsc2','mr_arfi_central_total_diff.eps');
+close;
 
+system('convert_eps_to_pdf.sh');

### 2013-09-13 tab_mr_arfi_volumes.tex
column headings in bold

@@
 \centering
-\caption{Comparison of Peripheral Zone and Central Gland Volumes in MR T2WI and ARFI Imaging}
+\caption{Comparison of Central Gland / Zone and Total Prostate Volumes in MR T2WI and ARFI Imaging}
 \begin{tabular}{|l|l|l|l|l|} \hline
-Study Subject ID & MR Central Gland & MR Total & ARFI Central Zone & ARFI Total \\ 
-& Volume (mm$^3$) & Volume (mm$^3$) & Volume (mm$^3$) & Volume (mm$^3$) \\ \hline
+{\bf Study Subject} & {\bf MR Central Gland} & {\bf MR Total} & {\bf ARFI Central Zone} & {\bf ARFI Total} \\ 
+& {\bf Volume (mm$^3$)} & {\bf Volume (mm$^3$)} & {\bf Volume (mm$^3$)} & {\bf Volume (mm$^3$)} \\ \hline
 P86 & 12738.12 & 24567.06 & 14296. & 40031. \\

### 2013-09-13 data/gen_latex_table.m
generated script to render LaTeX table data directly from CSV file instead of manually entering it

@@ -1,0 +1,17 @@
+% gen_latex_tab.m
+%
+% generate LaTeX table from CSV file
+% 
+% Mark Palmeri
+% mlp6@duke.edu
+% 2013-09-13
+
+data = csvread('Prostate_CZ_PZ_Volume_Axis_Measurements.csv');
+
+fid=fopen('vol_data.tex','w');
+
+for i=1:size(data,1),
+    fprintf(fid,'%i & %.2f & %.2f & %.2f & %.2f \\\\ \n',i,data(i,2)/1e3,data(i,3)/1e3,data(i,10)/1e3,data(i,11)/1e3);
+end;
+
+fclose(fid);

### 2013-09-13 figs/mr_arfi_volumes.m
updated to just use 1:16, not P#

@@ -10,6 +10,7 @@
 
 vols = csvread('../data/Prostate_CZ_PZ_Volume_Axis_Measurements.csv');
 
+Pnum = num2cell(vols(:,1));
 mr_cg_vol = vols(:,2);
 mr_total_vol = vols(:,3);
 mr_pz_vol = mr_total_vol - mr_cg_vol;
@@ -43,6 +44,7 @@ a(2) = 17;
 axis(a);
 legend('Central Gland','Peripheral Zone','Location','NorthWest');
 legend boxoff;
+%set(gca,'XTickLabel',Pcell)
 
 set(gca, ...
     'Box'         , 'off'     , ...

### 2013-09-13 tab_mr_arfi_volumes.tex
updated formatting in table and figs

@@
 {\bf Study Subject} & {\bf MR Central Gland} & {\bf MR Total} & {\bf ARFI Central Zone} & {\bf ARFI Total} \\ 
-& {\bf Volume (mm$^3$)} & {\bf Volume (mm$^3$)} & {\bf Volume (mm$^3$)} & {\bf Volume (mm$^3$)} \\ \hline
-P86 & 12738.12 & 24567.06 & 14296. & 40031. \\
-P79 & 14263.86 & 28514.79 & 8370 & 33978 \\
-P78 & 23471.81 & 32484.36 & 13290.47 & 37420 \\
-P77 & 17317.03 & 32492.83 & 10829.7 & 31819 \\
-P75 & 57561.69 & 70950.98 & 30369 & 73684.64 \\
-P74 & 12010.3 & 27841.21 & 15679 & 43298.34 \\
-P72 & 8824.44 & 19587.02 & 11776.44 & 28545.79 \\
-P71 & 10969.65 & 21276.98 & 16020.73 & 31424.61 \\
-P70 & 13633.36 & 20749.91 & 19282 & 36876.3\\
-P69 & 23580.12 & 36112.58 & 33500 & 70654 \\
-P67 & 16569.17 & 27332.39 & 25219 & 34656.2 \\
-P61 & 25381.36 & 49213.93 & 18137 & 54044.3 \\
-P59 & 9247.34 & 26360.37 & 6144 & 35137 \\
-P58 & 14787.88 & 23361.47 & 19654.6 & 35211 \\
-P57 & 17872.15 & 35371.12 & 10153 & 38028.47 \\
-P56 & 33315.41 & 48495.39 & 39351 & 66425 \\ \hline
+& {\bf Volume (cm$^3$)} & {\bf Volume (cm$^3$)} & {\bf Volume (cm$^3$)} & {\bf Volume (cm$^3$)} \\ \hline
+\input{data/vol_data}
+\hline
 \end{tabular}

### 2013-09-13 tab_mr_arfi_axes.tex
added axis data table

\begin{table}
\centering
\caption{Comparison of Central Gland / Zone and Total Prostate Major / Minos Axes in MR T2WI and ARFI Imaging}
\begin{tabular}{|l|l|l|l|l|l|l|l|l|l|l|l|l|} \hline
{\bf Study} & {\bf MR} & {\bf MR} & {\bf MR} & {\bf MR} & {\bf MR} & {\bf MR} & {\bf ARFI} & {\bf ARFI} & {\bf ARFI} & {\bf ARFI} & {\bf ARFI} & {\bf ARFI} \\ 
{\bf Subject} & {\bf CA} & {\bf CL} & {\bf CE} & {\bf TA} & {\bf TL} & {\bf TE} & {\bf CA} & {\bf CL} & {\bf CE} & {\bf TA} & {\bf TL} & {\bf TE} \\
 & {\bf (cm)} & {\bf (cm)} & {\bf (cm)} & {\bf (cm)} & {\bf (cm)} & {\bf (cm)} & {\bf (cm)} & {\bf (cm)} & {\bf (cm)} & {\bf (cm)} & {\bf (cm)} & {\bf (cm)} \\ \hline
\input{data/axis_data}
\hline
\end{tabular}
\label{tab:mr_arfi_axes}
\end{table}

### 2013-09-13 data/gen_latex_table.m

@@ -15,3 +15,11 @@ for i=1:size(data,1),
 end;
 
 fclose(fid);
+
+fid = fopen('axis_data.tex','w');
+
+for i=1:size(data,1),
+    fprintf(fid,'%i & %.2f & %.2f & %.2f & %.2f & %.2f & %.2f & %.2f & %.2f & %.2f & %.2f & %.2f & %.2f \\\\ \n',i,data(i,4)/10,data(i,5)/10,data(i,6)/10,data(i,7)/10,data(i,8)/10,data(i,9)/10,data(i,12)/10,data(i,13)/10,data(i,14)/10,data(i,15)/10,data(i,16)/10,data(i,17)/10);
+end;
+
+fclose(fid);

### 2013-09-14 figs/corr_weight_vol.m
figure showing no correlation b/w prostate weight and imaging volume

@@ -1,0 +1,67 @@
+% corr_weight_vol.m
+%
+% Correlate excised path weights (g) with total volume estimates with each
+% modality.
+%
+% Mark Palmeri
+% mlp6@duke.edu
+% 2013-09-13
+
+vols = csvread('../data/Prostate_CZ_PZ_Volume_Axis_Measurements.csv');
+
+Pnum = num2cell(vols(:,1));
+mr_total_vol = vols(:,3);
+arfi_total_vol = vols(:,11);
+w = vols(:,12);
+
+mr_total_vol = mr_total_vol/1e3;
+arfi_total_vol = arfi_total_vol/1e3;
+
+% fontsize
+fs = 18;
+
+figure;
+hold on;
+plot(w,mr_total_vol,'bx','MarkerSize',14,'LineWidth',3);
+plot(w,arfi_total_vol,'go','MarkerSize',14,'LineWidth',3);
+ylabel('Total Volume (cm^3)','FontSize',fs)
+xlabel('Prostate Weight (g)','FontSize',fs);
+
+set(gca, ...
+    'Box'         , 'off'     , ...
+    'TickDir'     , 'out'     , ...
+    'TickLength'  , [.01 .01] , ...
+    'XMinorTick'  , 'off'      , ...
+    'YMinorTick'  , 'off'      , ...
+    'XGrid'       , 'off'      , ...
+    'XColor'      , [0 0 0], ...
+    'YColor'      , [0 0 0], ...
+    'XMinorGrid'  , 'off'      , ...
+    'LineWidth'   , 2, ...
+    'FontSize'    , fs);
+
+% lets do a linear regression and see what we get
+mr_p = polyfit(w,mr_total_vol,1);
+mr_fit = polyval(mr_p,w);
+mr_resid = mr_total_vol - mr_fit;
+mr_ss_resid = sum(mr_resid.^2);
+mr_ss_total = (length(mr_total_vol)- 1) * var(mr_total_vol);
+mr_Rsq = 1 - mr_ss_resid / mr_ss_total
+
+plot(w,mr_fit,'-b','LineWidth',3);
+
+arfi_p = polyfit(w,arfi_total_vol,1);
+arfi_fit = polyval(arfi_p,w);
+arfi_resid = arfi_total_vol - arfi_fit;
+arfi_ss_resid = sum(arfi_resid.^2);
[…]

### 2013-09-14 fig_mr_arfi_weight.tex
figure correlating weight with imaging volumens

\begin{figure}[htb!]
\centering
\includegraphics[width=0.5\linewidth]{figs/corr_weight_vol.pdf}
\caption{The prostates analyzed in this study showed no significant correlation
    between their measured excised pathology weight and the reconstructed
    volumes in T2WI MR (blue) and ARFI imaging (green).  The coefficients of
    determination (R$^2$) for the linear regressions for each imaging modality were
    0.05 and 0.13, respectively, indicating no significant correlation.}
\label{fig:mr_arfi_weight} 
\end{figure}

### 2013-09-14 results.tex
fixed citations; more to be fixed
text surrounding the MR / ARFI volume figs
added axis data table
text additions for figures

@@
     \item Images of select cases, demonstrating segmentation process
-    \item Table of volumes in MR and ARFI images
-    \item Table comparing measurements
 \end{itemize}
+
+The computed volumes of the prostates from the MR and ARFI imaging data are
+presented in Table~\ref{tab:fig_mr_arfi_volumes}.  The subvolumes associated
+with the zonal anatomy in each imaging modality was measured
+(Figure~\ref{fig:fig_mr_arfi_volumes}(a)), showing an mean overestimation of
+total prostate volume of 36.7 $\pm$ 27.9\% by ARFI imaging compared to MR
+volumes (Figure~\ref{fig:fig_mr_arfi_volumes}(b)), and a relative
+underestimation of the central zone volume relative to the central gland in MR
+imaging of -15.3 $\pm$ 13.0 \% (Figure~\ref{fig:fig_mr_arfi_volumes}(c)).
+
+\input{tab_mr_arfi_volumes}
+
+\input{fig_mr_arfi_volumes}
+
+MR and ARFI imaging total prostate volumes were also compared with the weights
+and prostate volumes, approximated to be tri-axial ellipsoids, and calculated
+from measurements of the prostate semi-principal axes.  That was no significant
+correlation between the prostate weight and the image-reconstructed prostate
+volumes (Figure~\ref{fig:fig_mr_arfi_weight}, R$^2$ = 0.05 (MR) and 0.13
+(ARFI)), and there was XXXX correlation with the ellipsoidal approximation of
+the measurement prostate volume and the image-recontructed volumens (Figure
+XX).
+
+\input{fig_mr_arfi_weight}
+
+\input{tab_mr_arfi_axes}

### 2013-09-14 figs/corr_weight_vol.m
correlations b/w imaging volumes and path ellipsoidal volumes

@@ -1,67 +1,76 @@
-% corr_weight_vol.m
-%
-% Correlate excised path weights (g) with total volume estimates with each
-% modality.
-%
-% Mark Palmeri
-% mlp6@duke.edu
-% 2013-09-13
+function corr_weight_vol()
+    % function corr_weight_vol
+    %
+    % Correlate excised path weights (g) with total volume estimates with each
+    % modality.
+    %
+    % Mark Palmeri
+    % mlp6@duke.edu
+    % 2013-09-13
 
-vols = csvread('../data/Prostate_CZ_PZ_Volume_Axis_Measurements.csv');
+    vols = csvread('../data/Prostate_CZ_PZ_Volume_Axis_Measurements.csv');
 
-Pnum = num2cell(vols(:,1));
-mr_total_vol = vols(:,3);
-arfi_total_vol = vols(:,11);
-w = vols(:,12);
+    Pnum = num2cell(vols(:,1));
+    mr_total_vol = vols(:,3);
+    arfi_total_vol = vols(:,11);
+    w = vols(:,12);
+    path_vols = vols(:,16);
 
-mr_total_vol = mr_total_vol/1e3;
-arfi_total_vol = arfi_total_vol/1e3;
+    mr_total_vol = mr_total_vol/1e3;
+    arfi_total_vol = arfi_total_vol/1e3;
 
-% fontsize
-fs = 18;
+    gen_fig_n_fits(w,mr_total_vol,arfi_total_vol,'weight');
+    gen_fig_n_fits(path_vols,mr_total_vol,arfi_total_vol,'pathVol');
 
-figure;
-hold on;
-plot(w,mr_total_vol,'bx','MarkerSize',14,'LineWidth',3);
-plot(w,arfi_total_vol,'go','MarkerSize',14,'LineWidth',3);
-ylabel('Total Volume (cm^3)','FontSize',fs)
-xlabel('Prostate Weight (g)','FontSize',fs);
+    system('convert_eps_to_pdf.sh');
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+function gen_fig_n_fits(w,mr_total_vol,arfi_total_vol,metric)
+    % fontsize
+    fs = 18;
+    figure;
+    hold on;
+    plot(w,mr_total_vol,'bx','MarkerSize',14,'LineWidth',3);
[…]

### 2013-09-14 fig_mr_arfi_weight.tex
added fig with correlation b/w imaging volumes and path tri-axial volumes

@@
 \centering
-\includegraphics[width=0.5\linewidth]{figs/corr_weight_vol.pdf}
+\begin{tabular}{ll}
+\includegraphics[width=0.33\linewidth]{figs/corr_weight_vol.pdf} &
+\includegraphics[width=0.33\linewidth]{figs/corr_pathVol_vol.pdf} \\
+(a) Image Volume : Prostate Weight & (b) Image Volume : Pathology Volume \\
+\end{tabular}
 \caption{The prostates analyzed in this study showed no significant correlation
-    between their measured excised pathology weight and the reconstructed
+    between their measured excised pathology weight (a) and the reconstructed
     volumes in T2WI MR (blue) and ARFI imaging (green).  The coefficients of
-    determination (R$^2$) for the linear regressions for each imaging modality were
-    0.05 and 0.13, respectively, indicating no significant correlation.}
+    determination (R$^2$) for the linear regressions for each imaging modality
+    were 0.05 and 0.13, respectively, indicating no significant correlation.
+    Tri-axial pathology measurements were used to make an ellipsoidal prostate
+    volume approximation (b), which showed mild correlation with T2WI MR (R$^2$
+    = 0.32) and moderate correlation with ARFI imaging (R$^2$ = 0.71).}
 \label{fig:mr_arfi_weight} 

### 2013-09-14 results.tex
added image:path vol in text

@@
 correlation between the prostate weight and the image-reconstructed prostate
-volumes (Figure~\ref{fig:fig_mr_arfi_weight}, R$^2$ = 0.05 (MR) and 0.13
-(ARFI)), and there was XXXX correlation with the ellipsoidal approximation of
-the measurement prostate volume and the image-recontructed volumens (Figure
-XX).
+volumes (Figure~\ref{fig:mr_arfi_weight}(a), R$^2$ = 0.05 (MR) and 0.13
+(ARFI)), and there was mild-to-moderate correlation with the ellipsoidal
+approximation of the measurement prostate volume and the image-recontructed
+volumens (Figure~\ref{fig:mr_arfi_weight}(b), R$^2$ = 0.32 (MR) and 0.71
+(ARFI)).
 

### 2013-09-15 figs/corr_weight_vol.m
added figure of weight vs ellipsoidal volume

@@ -21,6 +21,7 @@ function corr_weight_vol()
 
     gen_fig_n_fits(w,mr_total_vol,arfi_total_vol,'weight');
     gen_fig_n_fits(path_vols,mr_total_vol,arfi_total_vol,'pathVol');
+    gen_fig_n_fits(w,path_vols,path_vols,'path_vol_weight');
 
     system('convert_eps_to_pdf.sh');
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@ -29,10 +30,18 @@ function gen_fig_n_fits(w,mr_total_vol,arfi_total_vol,metric)
     fs = 18;
     figure;
     hold on;
-    plot(w,mr_total_vol,'bx','MarkerSize',14,'LineWidth',3);
-    plot(w,arfi_total_vol,'go','MarkerSize',14,'LineWidth',3);
-    ylabel('Total Volume (cm^3)','FontSize',fs)
-    if(strcmp(metric,'weight')),
+    if(strcmp(metric,'path_vol_weight')),
+        plot(w,mr_total_vol,'bx','MarkerSize',14,'LineWidth',3);
+    else,
+        plot(w,mr_total_vol,'bx','MarkerSize',14,'LineWidth',3);
+        plot(w,arfi_total_vol,'go','MarkerSize',14,'LineWidth',3);
+    end;
+    if(strcmp(metric,'path_vol_weight')),
+        ylabel('Pathology Ellipsoidal Volume (cm^3)','FontSize',fs)
+    else,
+        ylabel('Total Volume (cm^3)','FontSize',fs)
+    end;
+    if(strcmp(metric,'weight') || strcmp(metric,'path_vol_weight')),
         xlabel('Prostate Weight (g)','FontSize',fs);
     elseif(strcmp(metric,'pathVol')),
         xlabel('Prostate Pathology Volume (cm^3)','FontSize',fs);
@@ -56,12 +65,20 @@ function gen_fig_n_fits(w,mr_total_vol,arfi_total_vol,metric)
 
     plot(w,mr_fit,'-b','LineWidth',3);
 
-    [arfi_fit,arfi_Rsq]=compute_linreg_Rsq(w,arfi_total_vol);
+    if(~strcmp(metric,'path_vol_weight')),
+        [arfi_fit,arfi_Rsq]=compute_linreg_Rsq(w,arfi_total_vol);
 
[…]

### 2013-09-15 fig_mr_arfi_weight.tex
added weight : ellipsoidal volume to text and figs

@@
 \centering
-\begin{tabular}{ll}
-\includegraphics[width=0.33\linewidth]{figs/corr_weight_vol.pdf} &
-\includegraphics[width=0.33\linewidth]{figs/corr_pathVol_vol.pdf} \\
-(a) Image Volume : Prostate Weight & (b) Image Volume : Pathology Volume \\
+\begin{tabular}{lll}
+\includegraphics[width=0.3\linewidth]{figs/corr_weight_vol.pdf} &
+\includegraphics[width=0.3\linewidth]{figs/corr_pathVol_vol.pdf} &
+\includegraphics[width=0.3\linewidth]{figs/corr_path_vol_weight_vol.pdf} \\
+(a) Image Volume : Prostate Weight & (b) Image Volume : Path Volume & (c) Path Weight : Path Volume \\
 \end{tabular}
@@
     volume approximation (b), which showed mild correlation with T2WI MR (R$^2$
-    = 0.32) and moderate correlation with ARFI imaging (R$^2$ = 0.71).}
+    = 0.32) and moderate correlation with ARFI imaging (R$^2$ = 0.71).  The
+    lack of correlationed between imaging volumes and prostate weight is not
+    surprising since the weights were also not correlated with the estimated
+    ellipsoidal prostate volumes (c, R$^2$ = 0.04).} 
 \label{fig:mr_arfi_weight}

### 2013-09-15 tab_mr_arfi_axes.tex
redid presentation of mr/arfi axis table data

@@
 \begin{tabular}{|l|l|l|l|l|l|l|l|l|l|l|l|l|} \hline
-{\bf Study} & {\bf MR} & {\bf MR} & {\bf MR} & {\bf MR} & {\bf MR} & {\bf MR} & {\bf ARFI} & {\bf ARFI} & {\bf ARFI} & {\bf ARFI} & {\bf ARFI} & {\bf ARFI} \\ 
-{\bf Subject} & {\bf CA} & {\bf CL} & {\bf CE} & {\bf TA} & {\bf TL} & {\bf TE} & {\bf CA} & {\bf CL} & {\bf CE} & {\bf TA} & {\bf TL} & {\bf TE} \\
+{\bf Study} & {\bf MR} & {\bf ARFI} & {\bf MR} & {\bf ARFI} & {\bf MR} & {\bf ARFI} & {\bf MR} & {\bf ARFI} & {\bf MR} & {\bf ARFI} & {\bf MR} & {\bf ARFI} \\ 
+{\bf Subject} & {\bf CA} & {\bf CA} & {\bf CL} & {\bf CL} & {\bf CE} & {\bf CE} & {\bf TA} & {\bf TA} & {\bf TL} & {\bf TL} & {\bf TE} & {\bf TE} \\
  & {\bf (cm)} & {\bf (cm)} & {\bf (cm)} & {\bf (cm)} & {\bf (cm)} & {\bf (cm)} & {\bf (cm)} & {\bf (cm)} & {\bf (cm)} & {\bf (cm)} & {\bf (cm)} & {\bf (cm)} \\ \hline

### 2013-09-15 data/gen_latex_table.m

@@ -19,7 +19,7 @@ fclose(fid);
 fid = fopen('axis_data.tex','w');
 
 for i=1:size(data,1),
-    fprintf(fid,'%i & %.2f & %.2f & %.2f & %.2f & %.2f & %.2f & %.2f & %.2f & %.2f & %.2f & %.2f & %.2f \\\\ \n',i,data(i,4)/10,data(i,5)/10,data(i,6)/10,data(i,7)/10,data(i,8)/10,data(i,9)/10,data(i,12)/10,data(i,13)/10,data(i,14)/10,data(i,15)/10,data(i,16)/10,data(i,17)/10);
+    fprintf(fid,'%i & %.2f & %.2f & %.2f & %.2f & %.2f & %.2f & %.2f & %.2f & %.2f & %.2f & %.2f & %.2f \\\\ \n',i,data(i,4)/10,data(i,12)/10,data(i,5)/10,data(i,13)/10,data(i,6)/10,data(i,14)/10,data(i,7)/10,data(i,15)/10,data(i,8)/10,data(i,16)/10,data(i,9)/10,data(i,17)/10);
 end;
 
 fclose(fid);

### 2013-09-15 fig_mr_arfi_volumes.tex
updated figures for CORRECT weights and just using EPS with epstopdf LaTeX package

@@
 \begin{tabular}{ccc}
-\includegraphics[width=0.3\linewidth]{figs/mr_arfi_volumes.pdf} &
-\includegraphics[width=0.3\linewidth]{figs/mr_arfi_volume_diff.pdf} &
-\includegraphics[width=0.3\linewidth]{figs/mr_arfi_central_total_diff.pdf} \\
+\includegraphics[width=0.3\linewidth]{figs/mr_arfi_volumes} &
+\includegraphics[width=0.3\linewidth]{figs/mr_arfi_volume_diff} &
+\includegraphics[width=0.3\linewidth]{figs/mr_arfi_central_total_diff} \\
 (a) Central:Peripheral Volume & (b) Total Volume Difference & (c) Central:Total Volume Difference\\

### 2013-09-15 fig_mr_arfi_weight.tex
updated figures for CORRECT weights and just using EPS with epstopdf LaTeX package

@@
 \begin{tabular}{lll}
-\includegraphics[width=0.3\linewidth]{figs/corr_weight_vol.pdf} &
-\includegraphics[width=0.3\linewidth]{figs/corr_pathVol_vol.pdf} &
-\includegraphics[width=0.3\linewidth]{figs/corr_path_vol_weight_vol.pdf} \\
+\includegraphics[width=0.3\linewidth]{figs/corr_weight_vol} &
+\includegraphics[width=0.3\linewidth]{figs/corr_pathVol_vol} &
+\includegraphics[width=0.3\linewidth]{figs/corr_path_vol_weight_vol} \\
 (a) Image Volume : Prostate Weight & (b) Image Volume : Path Volume & (c) Path Weight : Path Volume \\

### 2013-09-15 figs/convert_eps_to_pdf.sh

@@ -1,1 +1,0 @@
-for i in `ls *.eps`; do epspdf $i; done;

### 2013-09-15 figs/corr_weight_vol.m

@@ -13,8 +13,8 @@ function corr_weight_vol()
     Pnum = num2cell(vols(:,1));
     mr_total_vol = vols(:,3);
     arfi_total_vol = vols(:,11);
-    w = vols(:,12);
-    path_vols = vols(:,16);
+    w = vols(:,18);
+    path_vols = vols(:,22);
 
     mr_total_vol = mr_total_vol/1e3;
     arfi_total_vol = arfi_total_vol/1e3;
@@ -23,7 +23,6 @@ function corr_weight_vol()
     gen_fig_n_fits(path_vols,mr_total_vol,arfi_total_vol,'pathVol');
     gen_fig_n_fits(w,path_vols,path_vols,'path_vol_weight');
 
-    system('convert_eps_to_pdf.sh');
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 function gen_fig_n_fits(w,mr_total_vol,arfi_total_vol,metric)
     % fontsize

### 2013-09-15 figs/mr_arfi_volumes.m

@@ -104,5 +104,3 @@ text(1,90,sprintf('Mean Diff = %.1f +/- %.1f',mean(arfi_cz_total-mr_cg_total),st
 
 print('-depsc2','mr_arfi_central_total_diff.eps');
 close;
-
-system('convert_eps_to_pdf.sh');

### 2013-09-15 compile-n-view.sh
changed to latex from pdflatex for compilation

@@ -1,1 +1,1 @@
-latexmk -bibtex -pdf -pvc mr_arfi_anatomy.tex
+latexmk -bibtex -pdfdvi -latex="latex --shell-escape %O %S" -pvc mr_arfi_anatomy.tex

### 2013-09-15 compile-n-view.sh
removed explicit latex specification

@@ -1,1 +1,1 @@
-latexmk -bibtex -pdfdvi -latex="latex --shell-escape %O %S" -pvc mr_arfi_anatomy.tex
+latexmk -bibtex -pdfdvi -pvc mr_arfi_anatomy.tex

### 2013-09-15 fig_mr_arfi_weight.tex
updated text for corrected prostate weights and ellipsoidal volumes

--- 2013-09-14 fig_mr_arfi_weight.tex
@@
 \centering
-\begin{tabular}{ll}
-\includegraphics[width=0.33\linewidth]{figs/corr_weight_vol.pdf} &
-\includegraphics[width=0.33\linewidth]{figs/corr_pathVol_vol.pdf} \\
-(a) Image Volume : Prostate Weight & (b) Image Volume : Pathology Volume \\
+\begin{tabular}{lll}
+\includegraphics[width=0.3\linewidth]{figs/corr_path_vol_weight_vol} &
+\includegraphics[width=0.3\linewidth]{figs/corr_weight_vol} &
+\includegraphics[width=0.3\linewidth]{figs/corr_pathVol_vol} \\
+(a) Path Weight : Path Volume & (b) Image Volume : Prostate Weight & (c) Image Volume : Path Volume \\
 \end{tabular}
-\caption{The prostates analyzed in this study showed no significant correlation
-    between their measured excised pathology weight (a) and the reconstructed
-    volumes in T2WI MR (blue) and ARFI imaging (green).  The coefficients of
-    determination (R$^2$) for the linear regressions for each imaging modality
-    were 0.05 and 0.13, respectively, indicating no significant correlation.
-    Tri-axial pathology measurements were used to make an ellipsoidal prostate
-    volume approximation (b), which showed mild correlation with T2WI MR (R$^2$
-    = 0.32) and moderate correlation with ARFI imaging (R$^2$ = 0.71).}
+\caption{Tri-axial pathology measurements were used to make an ellipsoidal
+    prostate volume approximation based on gross pathology axis measurements,
+    which was moderately well-correlated with the excised prostated weights (a,
+    R$^2$ = 0.68).  T2WI MR (blue, X) showed a moderate correlation between the
+    reconstructed volumes and prostate weight (R$^2$ = 0.44), while volumes
+    reconstructed from ARFI images (green, O) showed weaker correlation (R$^2$
+    = 0.21) (b).  Even weaker correlations existed between both T2WI MR and
+    ARFI image volumens and approaximated ellipsoidal prostate pathology
+    volumes (R$^2$ = 0.08 and 0.01, respectively) (c).}
 \label{fig:mr_arfi_weight}

### 2013-09-15 results.tex
added weight : ellipsoidal volume to text and figs
updated text for corrected prostate weights and ellipsoidal volumes

--- 2013-09-14 results.tex
@@
 and prostate volumes, approximated to be tri-axial ellipsoids, and calculated
-from measurements of the prostate semi-principal axes.  That was no significant
+from measurements of the prostate semi-principal axes.  Prostate weights were
+moderately correlated with estimated pathology ellipsoidal prostate volumes
+(Figure~\ref{fig:mr_arfi_weight}(a), R$^2$ = 0.68).  There was moderate
 correlation between the prostate weight and the image-reconstructed prostate
-volumes (Figure~\ref{fig:fig_mr_arfi_weight}, R$^2$ = 0.05 (MR) and 0.13
-(ARFI)), and there was XXXX correlation with the ellipsoidal approximation of
-the measurement prostate volume and the image-recontructed volumens (Figure
-XX).
+volumes (Figure~\ref{fig:mr_arfi_weight}(b), R$^2$ = 0.44 (MR) and 0.21
+(ARFI)), though there was weaker correlation with the ellipsoidal approximation
+of the measurement prostate volume and the image-recontructed volumens
+(Figure~\ref{fig:mr_arfi_weight}(c), R$^2$ = 0.08 (MR) and 0.01 (ARFI)).
 

### 2013-09-15 data/gen_latex_table.m
extract path data

@@ -23,3 +23,11 @@ for i=1:size(data,1),
 end;
 
 fclose(fid);
+
+fid = fopen('path_data.tex','w');
+
+for i=1:size(data,1),
+    fprintf(fid,'%i & %.2f & %.2f & %.2f & %.2f & %.2f \\\\ \n',i,data(i,18),data(i,19),data(i,20),data(i,21),data(i,22)); 
+end;
+
+fclose(fid);

### 2013-09-15 results.tex
path metrics tabulated

@@
 
+Weights and axis measurements from the gross pathology processing of the
+excised prostates were collected (Table~\ref{tab:path_data}), and using the
+axis measurements (lateral-to-lateral, anterior-to-posterior, and
+apex-to-base), the prostate volume was approximated as a tri-axial ellipsoid,
+and its volume was estimated.
+
+\input{tab_path_data}
+
 MR and ARFI imaging total prostate volumes were also compared with the weights

### 2013-09-15 data/gen_latex_table.m

@@ -27,7 +27,7 @@ fclose(fid);
 fid = fopen('path_data.tex','w');
 
 for i=1:size(data,1),
-    fprintf(fid,'%i & %.2f & %.2f & %.2f & %.2f & %.2f \\\\ \n',i,data(i,18),data(i,19),data(i,20),data(i,21),data(i,22)); 
+    fprintf(fid,'%i & %.f. & %.1f & %.1f & %.1f & %.2f \\\\ \n',i,data(i,18),data(i,19),data(i,20),data(i,21),data(i,22)); 
 end;
 
 fclose(fid);

### 2013-09-15 tab_mr_arfi_axes.tex
now consistent with path table axis labeling

@@
 \centering
-\caption{Comparison of Central Gland / Zone and Total Prostate Major / Minos Axes in MR T2WI and ARFI Imaging}
+\caption{Comparison of Central Gland / Zone (C) and Total (T) Prostate Axes in MR T2WI and ARFI Imaging.
+Axes are approximated in orientation to match those specified in gross pathology: lateral-to-lateral (LL), anterior-to-posterior (AP) and apex-to-base (AB).}
 \begin{tabular}{|l|l|l|l|l|l|l|l|l|l|l|l|l|} \hline
 {\bf Study} & {\bf MR} & {\bf ARFI} & {\bf MR} & {\bf ARFI} & {\bf MR} & {\bf ARFI} & {\bf MR} & {\bf ARFI} & {\bf MR} & {\bf ARFI} & {\bf MR} & {\bf ARFI} \\ 
-{\bf Subject} & {\bf CA} & {\bf CA} & {\bf CL} & {\bf CL} & {\bf CE} & {\bf CE} & {\bf TA} & {\bf TA} & {\bf TL} & {\bf TL} & {\bf TE} & {\bf TE} \\
+{\bf Subject} & {\bf C-AB} & {\bf C-AB} & {\bf C-LL} & {\bf C-LL} & {\bf C-AP} & {\bf C-AP} & {\bf T-AB} & {\bf T-AB} & {\bf T-LL} & {\bf T-LL} & {\bf T-AP} & {\bf T-AP} \\
  & {\bf (cm)} & {\bf (cm)} & {\bf (cm)} & {\bf (cm)} & {\bf (cm)} & {\bf (cm)} & {\bf (cm)} & {\bf (cm)} & {\bf (cm)} & {\bf (cm)} & {\bf (cm)} & {\bf (cm)} \\ \hline

### 2013-09-15 acknowledgements.tex
typo

@@
 The authors would like to thank Siemens Medical Solution USA, Ultrasound
-Division for their in-kind technical support and Ned Danieley for system administration support.
-This work was supported by NIH R01CA142824 and the Duke Coulter Translational Grant Program.
+Division for their in-kind technical support and Ned Danieley for computer
+system administration support.  This work was supported by NIH R01CA142824 and
+the Duke Coulter Translational Grant Program.

### 2013-09-15 figs/mr_arfi_path_axes.m
first-stab at axis plots... not at all done!!

@@ -1,0 +1,100 @@
+% mr_arfi_path_axes.m
+%
+% Generate plots of the axis-ratios for MR, ARFI and pathology.
+%
+% Mark Palmeri
+% mlp6@duke.edu
+% 2013-09-14
+
+data = csvread('../data/Prostate_CZ_PZ_Volume_Axis_Measurements.csv');
+
+Pnum = num2cell(data(:,1));
+
+% ordering the axes as:
+% (1) APEX-TO-BASE (axial)
+% (2) LATERAL-TO-LATERAL (lat)
+% (3) ANTERIOR-TO-POSTERIOR (elev)
+MRaxes = [data(:,IAMHERE];
+ARFIaxes = [arfi_cz_vol, arfi_pz_vol];
+PATHaxes
+
+% fontsize
+fs = 18;
+bar_width = 0.3;
+
+figure;
+hold on;
+h=bar([1:length(MRdataStack)],MRdataStack,bar_width,'stacked');
+g=bar([1:length(ARFIdataStack)]+bar_width,ARFIdataStack,bar_width,'stacked');
+set(h(1),'FaceColor',[0 0 1]);
+set(h(2),'FaceColor',[1 0 0]);
+set(g(1),'FaceColor',[0 1 0]);
+set(g(2),'FaceColor',[1 1 0]);
+ylabel('Zone Volumes (cm^3)','FontSize',fs)
+xlabel('Study Subject','FontSize',fs);
+title('MR (Red/Blue) and ARFI (Yellow/Green) Zone Volumes','FontSize',fs);
+a=axis;
+a(2) = 17;
+axis(a);
+legend('Central Gland','Peripheral Zone','Location','NorthWest');
+legend boxoff;
+%set(gca,'XTickLabel',Pcell)
+
+set(gca, ...
+    'Box'         , 'off'     , ...
+    'TickDir'     , 'out'     , ...
+    'TickLength'  , [.01 .01] , ...
+    'XMinorTick'  , 'off'      , ...
+    'YMinorTick'  , 'off'      , ...
+    'XGrid'       , 'off'      , ...
+    'XColor'      , [0 0 0], ...
+    'YColor'      , [0 0 0], ...
+    'XMinorGrid'  , 'off'      , ...
+    'LineWidth'   , 2, ...
+    'FontSize'    , fs);
+
+print('-depsc2','mr_arfi_volumes.eps');
+close;
+
[…]

### 2013-09-15 acknowledgements.tex
removed all caps section title

--- 2013-09-13 acknowledgements.tex
@@
-\section*{ACKNOWLEDGEMENTS}
+\section*{Acknowledgements}
 The authors would like to thank Siemens Medical Solution USA, Ultrasound
-Division for their in-kind technical support and Ned Danieley for system administration support.
-This work was supported by NIH R01CA142824 and the Duke Coulter Translational Grant Program.
+Division for their in-kind technical support and Ned Danieley for computer
+system administration support.  This work was supported by NIH R01CA142824 and
+the Duke Coulter Translational Grant Program.

### 2013-09-15 results.tex
fixed refs to table and figs

@@
 The computed volumes of the prostates from the MR and ARFI imaging data are
-presented in Table~\ref{tab:fig_mr_arfi_volumes}.  The subvolumes associated
+presented in Table~\ref{tab:mr_arfi_volumes}.  The subvolumes associated
 with the zonal anatomy in each imaging modality was measured
-(Figure~\ref{fig:fig_mr_arfi_volumes}(a)), showing an mean overestimation of
+(Figure~\ref{fig:mr_arfi_volumes}(a)), showing an mean overestimation of
 total prostate volume of 36.7 $\pm$ 27.9\% by ARFI imaging compared to MR
-volumes (Figure~\ref{fig:fig_mr_arfi_volumes}(b)), and a relative
+volumes (Figure~\ref{fig:mr_arfi_volumes}(b)), and a relative
 underestimation of the central zone volume relative to the central gland in MR
-imaging of -15.3 $\pm$ 13.0 \% (Figure~\ref{fig:fig_mr_arfi_volumes}(c)).
+imaging of -15.3 $\pm$ 13.0 \% (Figure~\ref{fig:mr_arfi_volumes}(c)).
 

### 2013-09-15 fig_mr_arfi_total_axes.tex
total prostate axis comparisons b/w MR, ARFI and PATH

\begin{figure}[htb!]
\centering
\begin{tabular}{ccc}
\includegraphics[width=0.3\linewidth]{figs/mr_arfi_total_axes1} &
\includegraphics[width=0.3\linewidth]{figs/mr_arfi_total_axes2} &
\includegraphics[width=0.3\linewidth]{figs/mr_arfi_total_axes3} \\
(a) Lat-Lat : Apex-Base & (b) Ant-Post : Apex-Base & (c) Lat-Lat : Ant-Post \\
\end{tabular}
\label{fig:mr_arfi_total_axes} 
\end{figure}

### 2013-09-15 results.tex
total prostate axis comparisons b/w MR, ARFI and PATH

@@
 
+While difficult to compare the absolute lengths of the major and minor
+anatomical axes of the prostates between the imaging modalities and pathology
+due to different geometric distortions during imaging and changes to the organ
+after excision, measurements of the prostate dimensions along the three
+standard anatomic axes (apex-to-base, lateral-to-lateral, and
+anterior-to-posterior) were made (Table~\ref{tab:mr_arfi_axes}), and the
+relative ratios of the axes in imaging and pathology for the total prostate
+volume were compared (Figure~\ref{fig:mr_arfi_total_axes}), and the relative
+ratios of the central gland / zone anatomy were compared for MR and ARFI
+imaging (Figure~\ref{fig:mr_arfi_central_axes}).
+
 \input{tab_mr_arfi_axes}
+
+\input{fig_mr_arfi_total_axes}

### 2013-09-15 figs/mr_arfi_path_axes.m

@@ -1,100 +1,114 @@
-% mr_arfi_path_axes.m
-%
-% Generate plots of the axis-ratios for MR, ARFI and pathology.
-%
-% Mark Palmeri
-% mlp6@duke.edu
-% 2013-09-14
+function mr_arfi_path_axes
+    % function mr_arfi_path_axes
+    %
+    % Generate plots of the axis-ratios for MR, ARFI and pathology.
+    %
+    % Mark Palmeri
+    % mlp6@duke.edu
+    % 2013-09-14
 
-data = csvread('../data/Prostate_CZ_PZ_Volume_Axis_Measurements.csv');
+    data = csvread('../data/Prostate_CZ_PZ_Volume_Axis_Measurements.csv');
 
-Pnum = num2cell(data(:,1));
+    Pnum = num2cell(data(:,1));
 
-% ordering the axes as:
-% (1) APEX-TO-BASE (axial)
-% (2) LATERAL-TO-LATERAL (lat)
-% (3) ANTERIOR-TO-POSTERIOR (elev)
-MRaxes = [data(:,IAMHERE];
-ARFIaxes = [arfi_cz_vol, arfi_pz_vol];
-PATHaxes
+    % ordering the axes as:
+    % (1) APEX-TO-BASE (axial)
+    % (2) LATERAL-TO-LATERAL (lat)
+    % (3) ANTERIOR-TO-POSTERIOR (elev)
+    MR_central_axes = [data(:,4:6)];
+    MR_total_axes = [data(:,7:9)];
+    ARFI_central_axes = [data(:,12:14)];
+    ARFI_total_axes = [data(:,15:18)];
+    PATH_total_axes = [data(:,19:21)];
 
-% fontsize
-fs = 18;
-bar_width = 0.3;
+    % I will now compute ratios of AB:LL (1:2), AB:AP (1:3), and LL:AP (2:3) for
+    % imaging central and total and path total volumes
+    MR_central_ratios = compute_ratios(MR_central_axes);
+    MR_total_ratios = compute_ratios(MR_total_axes);
+    ARFI_central_ratios = compute_ratios(ARFI_central_axes);
+    ARFI_total_ratios = compute_ratios(ARFI_total_axes);
+    PATH_total_ratios = compute_ratios(PATH_total_axes);
 
-figure;
-hold on;
-h=bar([1:length(MRdataStack)],MRdataStack,bar_width,'stacked');
[…]

### 2013-09-15 fig_mr_arfi_central_axes.tex
added central axis mr:arfi fig

\begin{figure}[htb!]
\centering
\begin{tabular}{ccc}
\includegraphics[width=0.3\linewidth]{figs/mr_arfi_central_axes1} &
\includegraphics[width=0.3\linewidth]{figs/mr_arfi_central_axes2} &
\includegraphics[width=0.3\linewidth]{figs/mr_arfi_central_axes3} \\
(a) Lat-Lat : Apex-Base & (b) Ant-Post : Apex-Base & (c) Lat-Lat : Ant-Post \\
\end{tabular}
\label{fig:mr_arfi_central_axes} 
\end{figure}

### 2013-09-16 fig_mr_arfi_total_axes.tex
updated metrics on axis ratio over/underestimation

@@
 \includegraphics[width=0.3\linewidth]{figs/mr_arfi_total_axes3} \\
-(a) Lat-Lat : Apex-Base & (b) Ant-Post : Apex-Base & (c) Lat-Lat : Ant-Post \\
+(a) & (b) & (c) \\
+\includegraphics[width=0.3\linewidth]{figs/mr_arfi_total_over_under1} &
+\includegraphics[width=0.3\linewidth]{figs/mr_arfi_total_over_under2} &
+\includegraphics[width=0.3\linewidth]{figs/mr_arfi_total_over_under3} \\
+(d) & (e) & (f) \\
 \end{tabular}
+\caption{Comparison of the ratios of the three anatomic axis measurement ratios
+    for T2WI MR (top row, blue), ARFI imaging (top row, green) and gross
+    pathology (top row, red).  The over/underestimation of the axis ratios
+    between ARFI imaging : T2WI MR : Pathology are shown in the bottom row
+    (d-f), with mean ratio differences compiled in
+    Table~\ref{tab:axis_ratio_over_under}.}
 \label{fig:mr_arfi_total_axes} 

### 2013-09-16 tab_axis_ratio_over_under.tex
updated metrics on axis ratio over/underestimation

\begin{table}
\centering
\caption{Mean axis ratio differences between ARFI imaging, T2WI MR and
    pathology for the total prostate volume and central gland / zone for the
    imaging modalities.  The three axis ratios analyzed were:
    lateral-to-lateral : apex-to-base (LL:AB), anterior-to-posterio :
    apex-to-base (AP:AB), and lateral-to-lateral : anterior-to-posterior
    (LL:AP).}
\begin{tabular}{|l|l|l|l|l|} \hline
{\bf Image Modality} & {\bf Comparative Measure} & {\bf Total / Central} & {\bf Axes} & {\bf Axis Ratio Difference (\%)} \\ \hline
\input{figs/tab_axis_ratio_over_under_data}
\hline
\end{tabular}
\label{tab:axis_ratio_over_under}
\end{table}

### 2013-09-16 figs/mr_arfi_path_axes.m

@@ -36,6 +36,9 @@ function mr_arfi_path_axes
             '(Anterior-to-Posterior : Apex-to-Base)',...
             '(Lateral-to-Lateral : Anterior-to-Posterior)'};
 
+    % create a file for writing the data for LaTeX table
+    fid = fopen('tab_axis_ratio_over_under_data.tex','w');
+
     for i=1:3,
         figure;
         hold on;
@@ -83,7 +86,7 @@ function mr_arfi_path_axes
         set(g(1),'FaceColor',[0 1 0]);
         ylabel('Axis Ratios','FontSize',fs)
         xlabel('Study Subject','FontSize',fs);
-        title(sprintf('Total Prostate Axes: %s',titles{i}),'FontSize',fs);
+        title(sprintf('Central Gland / Zone Axes: %s',titles{i}),'FontSize',fs);
         a=axis;
         a(2) = 17;
         axis(a);
@@ -108,7 +111,90 @@ function mr_arfi_path_axes
         close;
     end;
 
+    % now lets compute how much over/under estimation of axis ratios ARFI is compared to the MR (and PATH).
+    ARFI_MR_Total_OverUnder = compute_over_under(MR_total_ratios,ARFI_total_ratios);
+    ARFI_PATH_Total_OverUnder = compute_over_under(PATH_total_ratios,ARFI_total_ratios);
+    MR_PATH_Total_OverUnder = compute_over_under(PATH_total_ratios,MR_total_ratios);
+    ARFI_MR_Central_OverUnder = compute_over_under(MR_central_ratios,ARFI_central_ratios);
+
+    axis_combos = {'LL:AB','AP:AB','LL:AP'};
+    % another set of bar plots! 
+    for i=1:3,
+        figure;
+        hold on;
+        bar_width=0.25;
+        h=bar([1:length(ARFI_MR_Total_OverUnder(:,i))],ARFI_MR_Total_OverUnder(:,i),bar_width);
+        g=bar([1:length(ARFI_PATH_Total_OverUnder(:,i))]+bar_width,ARFI_PATH_Total_OverUnder(:,i),bar_width);
+        f=bar([1:length(MR_PATH_Total_OverUnder(:,i))]+2*bar_width,MR_PATH_Total_OverUnder(:,i),bar_width);
+        set(h(1),'FaceColor',[0 0 1]);
+        set(g(1),'FaceColor',[0 1 0]);
+        set(f(1),'FaceColor',[1 0 0]);
[…]

### 2013-09-16 fig_mr_arfi_central_axes.tex
added relative diffs in central axis ratios

@@
 \includegraphics[width=0.3\linewidth]{figs/mr_arfi_central_axes3} \\
-(a) Lat-Lat : Apex-Base & (b) Ant-Post : Apex-Base & (c) Lat-Lat : Ant-Post \\
+(a) & (b) & (c) \\
+\includegraphics[width=0.3\linewidth]{figs/mr_arfi_central_over_under1.eps} &
+\includegraphics[width=0.3\linewidth]{figs/mr_arfi_central_over_under2.eps} &
+\includegraphics[width=0.3\linewidth]{figs/mr_arfi_central_over_under3.eps} \\
+(a) & (b) & (c) \\
 \end{tabular}
+\caption{Comparison of the ratios of the three anatomic axis measurement ratios
+    for T2WI MR (top row, blue) and ARFI imaging (top row, green).  The
+    over/underestimation of the axis ratios between ARFI imaging and T2WI MR
+    are shown in the bottom row (d-f), with mean ratio differences compiled in
+    Table~\ref{tab:axis_ratio_over_under}.}
 \label{fig:mr_arfi_central_axes} 

### 2013-09-16 fig_mr_anatomy.tex
MR annotated anatomy included

\begin{figure}[htb!]
\centering
\begin{tabular}{ccc}
\includegraphics[width=0.29\linewidth]{figs/mr_anatomy/T2_apex} &
\includegraphics[width=0.275\linewidth]{figs/mr_anatomy/T2_midgland} &
\includegraphics[width=0.3\linewidth]{figs/mr_anatomy/T2_base} \\
(a) MR T2WI Apex & (b) MR T2WI Mid-gland & (c) MR T2WI Base \\
\end{tabular}
\caption{Axial T2-weighted MR images of the prostate show the apex (a),
    mid-gland (b), and base (c).  The peripheral zone (PZ) is of higher signal
    intensity than the central gland (CG), the latter which is composed of the
    central zone and the transitional zone. The apex (A) is composed mostly of
    PZ glandular tissue and the urethra is seen at the level of the midgland as
    an inverted ``U'' (b). Note the area of hypointense signal in the
    peripheral zone at the midgland and base (asterisk, b and c), which
    represents a prostatic tumor.  The posterior (P) aspect of the prostate is
    adjacent to the endorectal coil, and the right (R)-to-left (L) extent of
    the prostate is refered to as the lateral-to-lateral axis in the subsequent
    analysis.}
\label{fig:mr_anatomy} 
\end{figure}

### 2013-09-16 figs/compute_linreg_Rsq.m
added figures showing the axis measurement comparisons

@@ -1,0 +1,8 @@
+function [fit,Rsq]=compute_linreg_Rsq(w,total_vol)
+    p = polyfit(w,total_vol,1);
+    fit = polyval(p,w);
+    resid = total_vol - fit;
+    ss_resid = sum(resid.^2);
+    ss_total = (length(total_vol)- 1) * var(total_vol);
+    Rsq = 1 - ss_resid / ss_total
+

### 2013-09-16 figs/corr_weight_vol.m

@@ -82,11 +82,3 @@ function gen_fig_n_fits(w,mr_total_vol,arfi_total_vol,metric)
     print('-depsc2',sprintf('corr_%s_vol.eps',metric));
     close;
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-function [fit,Rsq]=compute_linreg_Rsq(w,total_vol)
-    p = polyfit(w,total_vol,1);
-    fit = polyval(p,w);
-    resid = total_vol - fit;
-    ss_resid = sum(resid.^2);
-    ss_total = (length(total_vol)- 1) * var(total_vol);
-    Rsq = 1 - ss_resid / ss_total
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2013-09-16 figs/mr_arfi_path_axes.m

@@ -21,6 +21,66 @@ function mr_arfi_path_axes
     ARFI_total_axes = [data(:,15:18)];
     PATH_total_axes = [data(:,19:21)];
 
+    titles={'(Lateral-to-Lateral : Apex-to-Base)',...
+            '(Anterior-to-Posterior : Apex-to-Base)',...
+            '(Lateral-to-Lateral : Anterior-to-Posterior)'};
+    axis_titles={'Apex-to-Base',...
+            'Lateral-to-Lateral',...
+            'Anterior-to-Posterior'};
+
+    % fontsize
+    fs = 18;
+
+    % lets plot some direct axis measurements of the total prostate and central gland
+    for i=1:3,
+        figure;
+        hold on;
+        plot(PATH_total_axes(:,i),MR_total_axes(:,i)/10,'bx','MarkerSize',14,'LineWidth',3);
+        plot(PATH_total_axes(:,i),ARFI_total_axes(:,i)/10,'go','MarkerSize',14,'LineWidth',3);
+        xlabel('Pathology Measurement (cm)','FontSize',fs);
+        ylabel('Imaging Measurement (cm)','FontSize',fs);
+
+        set(gca, ...
+            'Box'         , 'off'     , ...
+            'TickDir'     , 'out'     , ...
+            'TickLength'  , [.01 .01] , ...
+            'XMinorTick'  , 'off'      , ...
+            'YMinorTick'  , 'off'      , ...
+            'XGrid'       , 'off'      , ...
+            'XColor'      , [0 0 0], ...
+            'YColor'      , [0 0 0], ...
+            'XMinorGrid'  , 'off'      , ...
+            'LineWidth'   , 2, ...
+            'FontSize'    , fs);
+
+         [mr_fit,mr_Rsq]=compute_linreg_Rsq(PATH_total_axes(:,i),MR_total_axes(:,i)/10);
+         plot(PATH_total_axes(:,i),mr_fit,'-b','LineWidth',3);
+         [arfi_fit,arfi_Rsq]=compute_linreg_Rsq(PATH_total_axes(:,i),ARFI_total_axes(:,i)/10);
+         plot(PATH_total_axes(:,i),arfi_fit,'-g','LineWidth',3);
+
+       axis([2 8 2 8]); 
+
+       title(sprintf('%s',axis_titles{i}),'FontSize',fs);
+
+       switch i
+       case 1
[…]

### 2013-09-16 fig_mr_arfi_path_axes.tex
latest raw axis plots

\begin{figure}[htb!]
\centering
\begin{tabular}{ccc}
\includegraphics[width=0.3\linewidth]{figs/Lateral-to-Lateral} &
\includegraphics[width=0.3\linewidth]{figs/Anterior-to-Posterior} &
\includegraphics[width=0.3\linewidth]{figs/Apex-to-Base} \\
(a) & (b) & (c) \\
\end{tabular}
\caption{Measurements of the prostate dimensions along the three standard
anatomic axes: lateral-to-lateral (a), anterior-to-posterior (b) and
apex-to-base (c).  The correlation between the imaging axis measurements and
pathology was performed for each orientation.  The black dashed-line represents
the projection of perfectly-correlated measurements between imaging and
patholoy.  Notice that there is no correlation between imaging and pathology in
the lateral-to-lateral and anterior-to-posterior axes, with a general
over-estimation of the lateral-to-lateral axis in the imaging.  The MR
estimation of the apex-to-base dimension showed moderate correlation with
pathology, with the ARFI image apex-to-base approximation having an overall
underestimation in this orientation.}
\label{fig:mr_arfi_path_axes} 
\end{figure}

### 2013-09-16 results.tex
added central axis mr:arfi fig
updated metrics on axis ratio over/underestimation
added relative diffs in central axis ratios
MR annotated anatomy included
latest raw axis plots

--- 2013-09-15 results.tex
@@
 \end{itemize}
+
+Figure~\ref{fig:mr_anatomy} shows the anatomic zones, etc. in ARFI imaging, blah blah blah\ldots
+
+\input{fig_mr_anatomy}
 
@@
 
+Measurements of the prostate dimensions along the three standard anatomic axes
+(apex-to-base, lateral-to-lateral, and anterior-to-posterior) were made
+(Table~\ref{tab:mr_arfi_axes}), and the correlation between the imaging axis
+measurements and pathology was performed (Figure~\ref{fig:mr_arfi_path_axes}).
+To evaluate the relative shape of the prostate in each imaging modality, the
+relative ratios of the axes in imaging and pathology for the total prostate
+volume were compared (Figure~\ref{fig:mr_arfi_total_axes}), and the relative
+ratios of the central gland / zone anatomy were compared for MR and ARFI
+imaging (Figure~\ref{fig:mr_arfi_central_axes}).  The mean ratio differences
+between ARFI imaging, T2WI MR and pathology are compiled in
+Table~\ref{tab:axis_ratio_over_under}.
+
 \input{tab_mr_arfi_axes}
+
+\input{fig_mr_arfi_path_axes}
+
+\input{fig_mr_arfi_total_axes}
+
+\input{fig_mr_arfi_central_axes}
+
+\input{tab_axis_ratio_over_under}

### 2013-09-16 fig_mr_arfi_path_axes.tex
added error table for imaging axes

@@
 pathology, with the ARFI image apex-to-base approximation having an overall
-underestimation in this orientation.}
+underestimation in this orientation.  The over-/under-estimation of each imaging modality relative to gross pathology is summarized in Table~\ref{tab:mr_arfi_path_axes}.}
 \label{fig:mr_arfi_path_axes} 

### 2013-09-16 figs/mr_arfi_path_axes.m

@@ -77,9 +77,16 @@ function mr_arfi_path_axes
 
         print('-depsc2',sprintf('%s.eps',axis_titles{i}));
         close;
+
+        % compute over/unders for each modality
+        mr_axis_OverUnder(:,i) = compute_over_under(PATH_total_axes(:,i),MR_total_axes(:,i)/10);
+        arfi_axis_OverUnder(:,i) = compute_over_under(PATH_total_axes(:,i),ARFI_total_axes(:,i)/10);
     end;
 
-
+    mean(mr_axis_OverUnder)
+    std(mr_axis_OverUnder)
+    mean(arfi_axis_OverUnder)
+    std(arfi_axis_OverUnder)
 
     % I will now compute ratios of AB:LL (1:2), AB:AP (1:3), and LL:AP (2:3) for
     % imaging central and total and path total volumes
@@ -88,8 +95,6 @@ function mr_arfi_path_axes
     ARFI_central_ratios = compute_ratios(ARFI_central_axes);
     ARFI_total_ratios = compute_ratios(ARFI_total_axes);
     PATH_total_ratios = compute_ratios(PATH_total_axes);
-
-
 
     % create a file for writing the data for LaTeX table
     fid = fopen('tab_axis_ratio_over_under_data.tex','w');

### 2013-09-17 fig_mr_arfi_weight.tex
updated correlations with weight and volume after excluding pt 4

@@
 (a) Path Weight : Path Volume & (b) Image Volume : Prostate Weight & (c) Image Volume : Path Volume \\
+\end{tabular}
+\begin{tabular}{ll}
+\includegraphics[width=0.3\linewidth]{figs/corr_weight_vol_no4} &
+\includegraphics[width=0.3\linewidth]{figs/corr_pathVol_vol_no4} \\
+(d) Image Volume : Prostate Weight (-4) & (e) Image Volume : Path Volume (-4) \\
 \end{tabular}
@@
     ARFI image volumens and approaximated ellipsoidal prostate pathology
-    volumes (R$^2$ = 0.08 and 0.01, respectively) (c).}
+    volumes (R$^2$ = 0.08 and 0.01, respectively) (c).  It should be noted in
+    these figures that Study Subject 4 had an excessively large prostate that
+    was difficult to fully capture in imaging, and was therefore grossly
+    underestimated in size.  We retrospectively excluded that study subject
+    from the analysis and re-calculated the coefficients of determination,
+    showing that MR and ARFI imaging improved to R$^2$ = 0.74 and 0.60,
+    respectively, for prostate weight (d), and R$^2$ = 0.39 for both imaging
+    modalities relative to pathology gross volume (e).}
 \label{fig:mr_arfi_weight}

### 2013-09-17 figs/corr_weight_vol_no4.m

@@ -1,0 +1,83 @@
+function corr_weight_vol_no4()
+    % function corr_weight_vol
+    %
+    % Correlate excised path weights (g) with total volume estimates with each
+    % modality, but exclude study subject 4 who is a huge outlier.
+    %
+    % Mark Palmeri
+    % mlp6@duke.edu
+    % 2013-09-17
+
+    vols = csvread('../data/Prostate_CZ_PZ_Volume_Axis_Measurements.csv');
+
+    mr_total_vol = vols([1:3,5:end],3);
+    arfi_total_vol = vols([1:3,5:end],11);
+    w = vols([1:3,5:end],18);
+    path_vols = vols([1:3,5:end],22);
+
+    mr_total_vol = mr_total_vol/1e3;
+    arfi_total_vol = arfi_total_vol/1e3;
+
+    gen_fig_n_fits(w,mr_total_vol,arfi_total_vol,'weight');
+    gen_fig_n_fits(path_vols,mr_total_vol,arfi_total_vol,'pathVol');
+    gen_fig_n_fits(w,path_vols,path_vols,'path_vol_weight');
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+function gen_fig_n_fits(w,mr_total_vol,arfi_total_vol,metric)
+    % fontsize
+    fs = 18;
+    figure;
+    hold on;
+    if(strcmp(metric,'path_vol_weight')),
+        plot(w,mr_total_vol,'bx','MarkerSize',14,'LineWidth',3);
+    else,
+        plot(w,mr_total_vol,'bx','MarkerSize',14,'LineWidth',3);
+        plot(w,arfi_total_vol,'go','MarkerSize',14,'LineWidth',3);
+    end;
+    if(strcmp(metric,'path_vol_weight')),
+        ylabel('Pathology Ellipsoidal Volume (cm^3)','FontSize',fs)
+    else,
+        ylabel('Total Volume (cm^3)','FontSize',fs)
+    end;
+    if(strcmp(metric,'weight') || strcmp(metric,'path_vol_weight')),
+        xlabel('Prostate Weight (g)','FontSize',fs);
+    elseif(strcmp(metric,'pathVol')),
+        xlabel('Prostate Pathology Volume (cm^3)','FontSize',fs);
+    end;
+
+    set(gca, ...
+        'Box'         , 'off'     , ...
[…]

### 2013-09-17 results.tex
added error table for imaging axes
latest and greatest results; placeholder for more

@@
 
-Figure~\ref{fig:mr_anatomy} shows the anatomic zones, etc. in ARFI imaging, blah blah blah\ldots
+Figure~\ref{fig:mr_anatomy} shows the anatomic zones and some anatomic
+orientation labels in the prostate of a representative study subject.  The
+zonal anatomy of the prostate was manually segmented (Figure XX) across the
+image stack, and 3D models of the zonal anatomy volumes were rendered (Figure
+XX).
 
 \input{fig_mr_anatomy}
+
+INSERT MR IMAGE SEGMENTS HERE (P79)
+
+INSERT 3D MODEL OF MR HERE
 
@@
 
+\input{tab_mr_arfi_path_axes}
+
 \input{fig_mr_arfi_total_axes}

### 2013-09-18 abstract.tex
added abstract and intro text

\section*{Abstract}
I need a few sentences from the intro here as a lead in...

MR and ARFI imaging datasets were compared with gross pathology measurements
most immediately post radical prostatectomy.  Both imaging modalities showed
moderate correlations (XX $<$ R$^2 < $ XX) between estimated organ volume and
gross pathologic weights and estimated volumes from tri-axial measurements.
ARFI images, on average, over-estimated prostate volumes by XX\% $\pm$ XX\%,
primarily due to over-estimation of the lateral (right-left) dimension of the
prostate.  MR images also over-estimated the prostate volume by XX\% $\pm$
XX\%, again, with a primary over-estimation of the lateral dimension.  MR and
ARFI imaging yielded different estimates of organ eccentricity as
characterized by ratio of the tri-axial measurements, with both imaging
modalities estimating greater degrees of eccentricity, dominated, again, by
lateral axis over-estimation.  Central gland volumes of the prostate were
well-visualized in both imaging modalities INCLUDE MORE QUANTITATIVE DETAILS
HERE.  ARFI imaging of the prostate can accurately delineate the central gland
of the prostate and the boundaries of the capsule, though care must be taken in
delineating the lateral edges of the prostate in the posterior peripheral zone.

### 2013-09-18 intro.tex
texting new compilation
intro content outline
added abstract and intro text

@@
 
-The introduction text will go here.
+Kirema's intro text will go here.
+
+Section~\ref{sect:background} provides an overview of MR and ARFI imaging in
+the prostate and an overall clinical motivation for prostate imaging.
+Section~\ref{sect:methods} describes the methods used to experimentally-acquire
+our imaging data, process of gross pathological specimens post radical
+prostatectomy, and image process our datasets.  The results of our analysis,
+including gross pathology and imaging prostate axis and volume estimates, are
+presented in Section~\ref{sect:results}, along with a statistical analysis of
+the bias and variability associated with each imaging modalities measurements,
+all of which is discussed in Section~\ref{sect:discussion}.

### 2013-09-20 intro.tex
updated / new Kirema content

@@
 
-Kirema's intro text will go here.
+Prostate cancer (PCa) is the most common non-cutaneous malignancy among men in
+the United States and the second leading cause of cancer related
+death.~\cite{Howlander2011} Screening with prostate specific antigen (PSA) and
+digital rectal examination (DRE) has led to earlier PCa detection but
+performance of these measures is not optimal, leading to imprecise risk
+assessment.  In addition, the random nature of transrectal ultrasonography
+(TRUS) guided biopsies can miss or underestimate the burden of
+cancer.~\cite{Gupta2013} Earlier disease diagnosis leads to challenges in
+deciding optimal management strategies for patients presenting with fewer
+disease burden, whereas missed tumors on TRUS-guided biopsies can result in
+inappropriate diagnoses.
+
+The use of non-invasive imaging in the evaluation of PCa could lead to improved
+diagnosis, risk-stratification, and management. Magnetic resonance imaging
+(MRI) has been available for use in the workup of patients with PCa since the
+early 1980s but early studies on its diagnostic accuracy were heterogeneous.
+However, the more recent ability to include functional parameters in PCa MRI
+analysis has yielded promising results.~\cite{Gupta2013,Hricak2007} Among the
+MRI modalities currently used in the study of PCa, it is well established that
+T2-weighted imaging (T2WI) offers the best assessment of prostate anatomy due
+to its ability to delineate prostatic margins, distinguish internal structures
+and differentiate among the glandular zones.
+
+Acoustic radiation force impulse (ARFI) imaging is an ultrasound-based modality
+that evaluates the mechanical properties of tissues.~\cite{Nightingale2002}
+ARFI imaging has the potential to aid in PCa diagnosis and management by
+evaluating the structural composition of prostate zones and tumors base on
+their stiffness.  Zhai et al were able to visualize prostatic anatomy by
+utilizing ARFI imaging \exvivo. In a second study, Zhai et al. demonstrated the
+feasibility of ARFI prostate imaging \inviv\invivo.~\cite{Zhai2012}  However, to the authors’ best
+knowledge, there have been no studies to date that compare \invivo ARFI
+prostate imaging to other imaging modalities.~\cite{Zhai2010} The goal of this
+study was to evaluate the ability of ARFI to distinguish prostate zonal anatomy
+\invivo as compared to endorectal T2WI.
 

### 2013-09-20 arfi_background.tex
updated to Kirema 2nd draft

\subsection{Acoustic Radiation Force Impulse (ARFI) Imaging}
ARFI imaging is an ultrasound technique that evaluates the mechanical
properties of tissues. It generates short-duration acoustic radiation forces
that result in tissue displacement. The response can then be measured for
tissue characterization, with displacement being inversely proportional to
stiffness.~\cite{Nightingale2002} This characterization is of particular
importance in prostate evaluation due to the gland’s complex and heterogeneous
composition.  Previous studies have shown that the central zone and PCA can be
up to 3 times stiffer than PZ tissue.~\cite{Zhai2010a,Zhai2012,Zhai2010b} This
finding can be explained by the PZ’s higher water content and is analogous to
the higher signal intensity of the PZ on T2WI.

The potential of ARFI imaging to distinguish normal prostatic anatomy from
pathologic tissue is promising in the future of PCa diagnosis and image guided
therapies. B-mode US, which is currently used for real-time guidance in
TRUS-guided biopsies, does not have the ability to clearly visualize PCa or to
differentiate it from other normal structures or disease processes such as
benign prostatic hyperplasia (BPH) or prostatitis.~\cite{Zhai2010a} This has
resulted in a systematic yet random approach to prostate core biopsies, with
6-12 cores routinely sampled from different anatomical regions.~\cite{Loch2004}
The ability to clearly distinguish PCa from other tissues could lead to
targeted biopsies with increased sensitivity and resultant decreased morbidity.
It could also aid in the development of targeted therapies for those with low
burden of disease.

### 2013-09-20 mr_background.tex
updated to Kirema 2nd draft

--- 2013-09-13 mr_background.tex
@@
 \subsection{Clinical Background and Significance}
-Prostate cancer (PCa) is the most common noncutaneous malignancy among men in
+Prostate cancer (PCa) is the most common non-cutaneous malignancy among men in
 the United States. Approximately 1 in every 6 men will develop PCa during their
@@
 of the cancer might be underestimated.~\cite{Gupta2013,Cornud1012} For example,
-in a study by Mufarrij et al 45.9--47.2\% of patients who were candidates for
+in a study by Mufarrij et al. 45.9--47.2\% of patients who were candidates for
 active surveillance but underwent radical prostatectomy had a higher Gleason
@@
 
-The prostate can be divided into glandular and nonglandular components. The
+The prostate can be divided into glandular and non-glandular components. The
 glandular components include the peripheral zone (PZ) and the central gland,
@@
 imaging techniques in addition to T2WI in order to better characterize prostate
-tumors.~\cite{Barentsz2012} In a study by Turkbey et al., researchers found that mpMRI had a
-positive predictive value of 98\% in prostate cancer
+tumors.~\cite{Barentsz2012} In a study by Turkbey et al., researchers found
+that mpMRI had a positive predictive value of 98\% in prostate cancer
 detection.~\cite{Turkbey2012} Functional sequences include diffusion- weighted
@@
 
-\subsection{Diffusion-weighted imaging}
 DWI is based on the free movement of water particles in tissue and measures the
-degree of motion restriction.13 Normal prostatic tissue is very glandular with
-plenty of water molecule movement. On the other hand, tumors have high cellular
-density and restricted water movement, which leads to decreased diffusion. Due
-to this restricted diffusion, tumors seem to be of higher intensity on
-DWI.~\cite{Koh2007}
+degree of motion restriction.~\cite{Koh2007} Normal prostatic tissue is very
+glandular with plenty of water molecule movement. On the other hand, tumors
+have high cellular density and restricted water movement, which leads to
+decreased diffusion. Due to this restricted diffusion, tumors seem to be of
+higher intensity on DWI.~\cite{Koh2007}
 
-DWI sequences can be processed to obtain apparent diffusion coefficient (ADC)
-maps. These serve as a more objective measure of diffusivity since they are
-independent of magnetic field and thus overcome the effects of T2
-shine-through.~\cite{Turkbey2012,Koh2007}  ADC maps are also used to visually
-assess the tumor, which appears as a focus of decreased signal intensity when
-compared to normal prostate tissue.~\cite{Tan2011}
+DCE-MRI provides a measure of tumor vascularity and takes advantage of
+tumor-induced angiogenesis in PCa.~\cite{Noworolski2005} Increased vascularity
+in PCa results in earlier and higher peak contrast enhancement and a faster
+washout when compared to normal prostatic tissue. These known characteristics
+of PCa are helpful when trying to localize lesions.
 
-In addition to knowing the location of the cancer within the prostate, being
-able to identify which cancers will exhibit more aggressive behavior is
-important when making clinical management decisions. A number of studies have
-found that lower ADC values correlate with tumors that have higher Gleason
-scores.~\cite{Kobus2012,Woodfield2010,Hambrock2011} This could be explained by
-the fact that higher-grade lesions are usually of higher cellular density and
-thus have even more marked restricted diffusion. This inverse correlation
-suggests that ADC maps might be helpful when characterizing lesions and
-assessing tumor aggressiveness.~\cite{Vargas2011}
-
-\subsection{Dynamic contrast enhanced imaging}
-DCE-MRI has been shown to have a high sensitivity in cancer detection,
-especially when combined with other MR imaging
-modalities.~\cite{Tanimoto2007,Girouin2007,Ocak2007,Kim2005} In a study by Hara
-et al, this functional modality identified 93\% of clinically significant
-cancers.~\cite{Hara2005} DCE-MRI provides a measure of tumor vascularity and
-takes advantage of tumor-induced angiogenesis in PCa.~\cite{Noworolski2005}
-Increased vascularity in PCa results in earlier and higher peak contrast
-enhancement and a faster washout when compared to normal prostatic tissue.
-These known characteristics of PCa are helpful when trying to localize lesions.
-
-A variety of different approaches can be used to analyze DCE-MRI and no report
-has, up to date, described the superiority of one method over the others.3 The
-approaches include qualitative analysis by visual assessment of the images,
-semi-quantitative methods that look at kinetic parameters on a voxel-by-voxel
-basis, or quantitative methods that that convert signal intensity into kinetic
-parameters to determine how much contrast is being exchanged between the
-vascular, extravascular and extracellular spaces.~\cite{Gupta2013}
-
-\subsection{MR spectroscopy}
 Like DWI, MRSI can be helpful in lesion characterization as it provides
@@
 setting of PCa. Choline is critical in cell membrane synthesis and is thus
-usually elevated in cells with cancerous behavior.3 Healthy prostate epithelium
-synthesizes and secretes large quantities of citrate but levels are decreased
-in PCa.~\cite{Costello1999,Kurhanewicz2002} Therefore, an increase in the
-choline-to-citrate ratio on MRSI can be used as an indicator of
+usually elevated in cells with cancerous behavior.~\cite{Gupta2013} Healthy
+prostate epithelium synthesizes and secretes large quantities of citrate but
+levels are decreased in PCa.~\cite{Costello1999,Kurhanewicz2002} Therefore, an
+increase in the choline-to-citrate ratio on MRSI can be used as an indicator of
 malignancy.~\cite{Bonekamp2011}
-
-Although some data suggests that MRSI increases the accuracy of tumor volume
-detection and staging when used in addition to anatomic imaging, it is unclear
-whether it is superior to other modalities.~\cite{Bonekamp2011} ESUR lists MSRI
-as an optional modality for the diagnosis of PCa since it significantly
-lengthens exam time and requires specific technical
-expertise.~\cite{Barentsz2012} Thus, it is not surprising that many academic
-centers in the United States do not include MSRI in their mpMRI prostate cancer
-protocols.~\cite{Gupta2013}

### 2013-09-21 abstract.tex
added central gland volume comparisons

@@
 \section*{Abstract}
-I need a few sentences from the intro here as a lead in...
-
-MR and ARFI imaging datasets were compared with gross pathology measurements
-most immediately post radical prostatectomy.  Both imaging modalities showed
-moderate correlations (XX $<$ R$^2 < $ XX) between estimated organ volume and
-gross pathologic weights and estimated volumes from tri-axial measurements.
-ARFI images, on average, over-estimated prostate volumes by XX\% $\pm$ XX\%,
+Prostate cancer (PCa) is the most common non-cutaneous malignancy among men in
+the United States and the second leading cause of cancer related death.  The
+use of non-invasive imaging in the evaluation of PCa could lead to improved
+diagnosis, risk-stratification, and management.  Magnetic resonance imaging
+(MRI) has been available for use in the workup of patients with PCa since the
+early 1980s, and recent advances with functional parameters has greatly
+improved its clinical diagnostic utility.  Acoustic radiation force impulse
+(ARFI) imaging is an ultrasound-based modality that evaluates the mechanical
+properties of tissues. ARFI imaging has the potential to aid in PCa diagnosis
+and management by evaluating the structural composition of prostate zones and
+tumors base on their stiffness.  In this study, MR and ARFI imaging datasets
+were compared with gross pathology measurements immediately post radical
+prostatectomy.  Both imaging modalities showed moderate correlations (0.39 $<$
+R$^2 < $ 0.74) between estimated organ volume and gross pathologic weights and
+estimated volumes from tri-axial measurements.  ARFI images, on average,
+over-estimated prostate volumes by 36\% $\pm$ 28\% compared to MR images,
 primarily due to over-estimation of the lateral (right-left) dimension of the
-prostate.  MR images also over-estimated the prostate volume by XX\% $\pm$
-XX\%, again, with a primary over-estimation of the lateral dimension.  MR and
-ARFI imaging yielded different estimates of organ eccentricity as
-characterized by ratio of the tri-axial measurements, with both imaging
-modalities estimating greater degrees of eccentricity, dominated, again, by
-lateral axis over-estimation.  Central gland volumes of the prostate were
-well-visualized in both imaging modalities INCLUDE MORE QUANTITATIVE DETAILS
-HERE.  ARFI imaging of the prostate can accurately delineate the central gland
-of the prostate and the boundaries of the capsule, though care must be taken in
-delineating the lateral edges of the prostate in the posterior peripheral zone.
+prostate.  The central zone volumes of the prostate agreed to within 2.1 $\pm$
+39.1\% between ARFI and MR imaging.  MR and ARFI imaging yielded different
+estimates of organ eccentricity as characterized by ratio of the tri-axial
+measurements, with both imaging modalities estimating greater degrees of
+eccentricity, dominated, again, by lateral axis over-estimation.  ARFI imaging
+of the prostate can accurately delineate the central gland of the prostate and
+the boundaries of the capsule, though care must be taken in delineating the
+lateral edges of the prostate in the posterior peripheral zone.

### 2013-09-21 fig_mr_arfi_volumes.tex
added central gland volume comparisons

@@
 \includegraphics[width=0.3\linewidth]{figs/mr_arfi_volume_diff} &
-\includegraphics[width=0.3\linewidth]{figs/mr_arfi_central_total_diff} \\
-(a) Central:Peripheral Volume & (b) Total Volume Difference & (c) Central:Total Volume Difference\\
+\includegraphics[width=0.3\linewidth]{figs/mr_arfi_central_diff} \\
+(a) Central:Peripheral Volume & (b) Total Volume Difference & (c) Central Volume Difference\\
 \end{tabular}
-\caption{Comparison of MR and ARFI zonal anatomy volume estimates from manually-segmented images.
-Total prostate volumes ranged from 19.6--71.0 cm$^3$ based on MR image models (a), with ARFI image models overestimating total prostate volume by 36.7 $\pm$ 27.9\% (b).
-ARFI image delineation of the central zone volume relative to total prostate volume showed a mean underestimation of -15.3 $\pm$ 13.0\% (c, green) relative to the MR central gland : total volume ratios (c, blue).}
+\caption{Comparison of MR and ARFI zonal anatomy volume estimates from
+    manually-segmented images.  Total prostate volumes ranged from 19.6--71.0
+    cm$^3$ based on MR image models (a), with ARFI image models overestimating
+    total prostate volume by 36.7 $\pm$ 27.9\% (b).  ARFI image estimation of
+    the central zone volume relative to the MR central gland volume varied by
+    2.1 $\pm$ 39.0\% (c).  Table~\ref{tab:mr_arfi_volumes} contains the
+    individual volume estimates for the entire prostate and the central
+    glands.}
 \label{fig:mr_arfi_volumes} 

### 2013-09-21 intro.tex
added central gland volume comparisons

@@
 utilizing ARFI imaging \exvivo. In a second study, Zhai et al. demonstrated the
-feasibility of ARFI prostate imaging \inviv\invivo.~\cite{Zhai2012}  However, to the authors’ best
+feasibility of ARFI prostate imaging \invivo.~\cite{Zhai2012}  However, to the authors’ best
 knowledge, there have been no studies to date that compare \invivo ARFI

### 2013-09-21 results.tex
added central gland volume comparisons

@@
 The computed volumes of the prostates from the MR and ARFI imaging data are
-presented in Table~\ref{tab:mr_arfi_volumes}.  The subvolumes associated
-with the zonal anatomy in each imaging modality was measured
-(Figure~\ref{fig:mr_arfi_volumes}(a)), showing an mean overestimation of
-total prostate volume of 36.7 $\pm$ 27.9\% by ARFI imaging compared to MR
-volumes (Figure~\ref{fig:mr_arfi_volumes}(b)), and a relative
-underestimation of the central zone volume relative to the central gland in MR
-imaging of -15.3 $\pm$ 13.0 \% (Figure~\ref{fig:mr_arfi_volumes}(c)).
+presented in Table~\ref{tab:mr_arfi_volumes}.  The subvolumes associated with
+the zonal anatomy in each imaging modality was measured
+(Figure~\ref{fig:mr_arfi_volumes}(a)), showing an mean overestimation of total
+prostate volume of 36.7 $\pm$ 27.9\% by ARFI imaging compared to MR volumes
+(Figure~\ref{fig:mr_arfi_volumes}(b)), and central gland volumes differed
+between ARFI and MR images by 2.1 $\pm$ 39.1\%
+(Figure~\ref{fig:mr_arfi_volumes}(c)).
 

### 2013-09-21 figs/mr_arfi_volumes.m

@@ -65,7 +65,7 @@ close;
 % lets do a little more processing to figure out what is going on... first the
 % relative volume in ARFI relative to MR
 
-arfi_mr_vol_diff = ((arfi_total_vol./mr_total_vol) - 1)*100;
+arfi_mr_vol_diff = ((arfi_total_vol-mr_total_vol)./mr_total_vol)*100;
 figure;
 bar([1:length(MRdataStack)],arfi_mr_vol_diff,0.5);
 title('ARFI:MR Total Volume Comparison','FontSize',fs);
@@ -104,3 +104,21 @@ text(1,90,sprintf('Mean Diff = %.1f +/- %.1f',mean(arfi_cz_total-mr_cg_total),st
 
 print('-depsc2','mr_arfi_central_total_diff.eps');
 close;
+
+% finally, lets just look at central glands % diffs
+arfi_mr_central_diff = ((arfi_cz_vol-mr_cg_vol)./mr_cg_vol)*100;
+figure;                            
+bar([1:length(MRdataStack)],arfi_mr_central_diff,0.5);
+title('ARFI:MR Central Gland Volume Comparison','FontSize',fs);
+xlabel('Study Subject','FontSize',fs);
+ylabel('Percent Difference','FontSize',fs);
+set(gca,'FontSize',fs);            
+a=axis;                            
+a(2) = 17;                         
+%a(3) = -5;                         
+%a(4) = 110;                        
+axis(a);
+text(0.5,70,sprintf('Mean Diff = %.1f +/- %.1f',mean(arfi_mr_central_diff),std(arfi_mr_central_diff)),'FontSize',fs);
+print('-depsc2','mr_arfi_central_diff.eps');
+close;
+

### 2013-09-21 fig_mr_segs_vol.tex
P79 seg vol added

\begin{figure}[htb!]
\centering
\includegraphics[width=1.0\textwidth]{figs/MR_P79_SegsVol.eps}
\caption{Representative MR T2WI segmentations and rendered volume performed in
    3D Slicer, with the peripheral zone being delineated in red, and the
    central gland being delineated in blue.  The native imaging plane for
    segmentation is the axial view, shown in the bottom left image.  The bottom
    middle and right images show the projections of the rendered model segment
    outlines in the sagital and coronal views, respectively.}
\label{fig:mr_segs_vol} 
\end{figure}

### 2013-09-21 results.tex
P79 seg vol added

@@
 orientation labels in the prostate of a representative study subject.  The
-zonal anatomy of the prostate was manually segmented (Figure XX) across the
-image stack, and 3D models of the zonal anatomy volumes were rendered (Figure
-XX).
+zonal anatomy of the prostate was manually segmented
+(Figure~\ref{fig:mr_segs_vol}, bottom left) across the image stack, and 3D
+models of the zonal anatomy volumes were rendered
+(Figure~\ref{fig:mr_segs_vol}, top).
 
@@
 
-INSERT MR IMAGE SEGMENTS HERE (P79)
-
-INSERT 3D MODEL OF MR HERE
+\input{fig_mr_segs_vol}
 

### 2013-09-21 fig_mr_segs_vol.tex
3D slicer methods added

@@
 \caption{Representative MR T2WI segmentations and rendered volume performed in
-    3D Slicer, with the peripheral zone being delineated in red, and the
-    central gland being delineated in blue.  The native imaging plane for
+    3D Slicer, with the peripheral zone (PZ) being delineated in red, the
+    central gland (CG) being delineated in blue, and the anterior fibromuscular
+    stroma (AFS) being shown in gray.  The AFS was combined with the PZ for
+    quantitative analyses shown herein.  The native imaging plane for
     segmentation is the axial view, shown in the bottom left image.  The bottom

### 2013-09-21 image_proc_methods.tex
3D slicer methods added

@@
 \subsection{Image Zonal Anatomy Segmentation and 3D Model Rendering}
-Axial T2W MR slices were manually segmented using the polygon tool on ITK SNAP
-using separate labels for the peripheral zone (PZ), central gland (CG) and
-anterior fibromuscular stroma (AFS). The gland was segmented from base to apex.
-The base was identified below the bladder and subsequent images were segmented
-until the last slice with visible prostatic tissue was identified caudally. The
-CG, PZ and AFS were segmented independently according to their well-established
-anatomical characteristics on
+Axial T2W MR slices were manually segmented using the polygon tool on
+ITK-SNAP~\cite{Yushkevich2006} using separate labels for the peripheral zone
+(PZ), central gland (CG) and anterior fibromuscular stroma (AFS). The gland was
+segmented from base to apex.  The base was identified below the bladder and
+subsequent images were segmented until the last slice with visible prostatic
+tissue was identified caudally. The CG, PZ and AFS were segmented independently
+according to their well-established anatomical characteristics on
 T2WI.~\cite{Verma2011,Jung2012,Poon1985,Hricak2007,Bonekamp2011} The PZ was
@@
 its low T2 signal intensity and its location anterior to the central gland.
+
+ARFI images were segmented using a similar procedure to the MR images, using
+XXXX (INSERT CRITERIA HERE).
+
+Segmented image stacks were imported into 3D Slicer (v4.3.0, CITATION) and 3D
+models were rendered using the following parameters (Table~\ref{tab:3dslicer}:
+
+\begin{table}[htb!]
+\centering
+\caption{3D model volume rendering parameters}
+\begin{tabular}{ll}
+{\bf Parameter} & {\bf Value} \\ \hline
+Decimation & 0.1 \\
+Smoothing Algorithm & Laplacian \\
+Smoothing  & 70.0 \\
+Joint Smoothing & Enabled \\
+\end{tabular}
+\end{table}

### 2013-09-22 figs/mr_arfi_path_axes.m
new plot of MR:ARFI axis comparisons

@@ -88,6 +88,54 @@ function mr_arfi_path_axes
     mean(arfi_axis_OverUnder)
     std(arfi_axis_OverUnder)
 
+    % let's make plots of the correlation between MR:ARFI for total gland and central gland
+    for i=1:3,
+        figure;
+        hold on;
+        plot(MR_total_axes(:,i),ARFI_total_axes(:,i)/10,'kx','MarkerSize',14,'LineWidth',3);
+        plot(MR_central_axes(:,i),ARFI_central_axes(:,i)/10,'ro','MarkerSize',14,'LineWidth',3);
+        xlabel('MR Measurement (cm)','FontSize',fs);
+        ylabel('ARFI Measurement (cm)','FontSize',fs);
+
+        set(gca, ...
+            'Box'         , 'off'     , ...
+            'TickDir'     , 'out'     , ...
+            'TickLength'  , [.01 .01] , ...
+            'XMinorTick'  , 'off'      , ...
+            'YMinorTick'  , 'off'      , ...
+            'XGrid'       , 'off'      , ...
+            'XColor'      , [0 0 0], ...
+            'YColor'      , [0 0 0], ...
+            'XMinorGrid'  , 'off'      , ...
+            'LineWidth'   , 2, ...
+            'FontSize'    , fs);
+
+         [image_total_fit,image_total_Rsq]=compute_linreg_Rsq(MR_total_axes(:,i)/10,ARFI_total_axes(:,i)/10);
+         plot(MR_total_axes(:,i),image_total_fit,'-k','LineWidth',3);
+         [image_central_fit,image_central_Rsq]=compute_linreg_Rsq(MR_central_axes(:,i)/10,ARFI_central_axes(:,i)/10);
+         plot(MR_central_axes(:,i),image_central_fit,'-r','LineWidth',3);
+
+       axis([2 8 2 8]); 
+
+       title(sprintf('%s',axis_titles{i}),'FontSize',fs);
+
+       switch i
+       case 1
+           legend(sprintf('Total Gland (R^2 = %.2f)',image_total_Rsq),sprintf('Central Gland (R^2 = %.2f)',image_central_Rsq),'Location','NorthWest');
+        case 2
+           legend(sprintf('Total Gland (R^2 = %.2f)',image_total_Rsq),sprintf('Central Gland (R^2 = %.2f)',image_central_Rsq),'Location','NorthWest');
+        case 3
[…]

### 2013-09-22 figs/mr_arfi_path_axes.m
added new MR:ARFI axis comparisons

@@ -92,8 +92,8 @@ function mr_arfi_path_axes
     for i=1:3,
         figure;
         hold on;
-        plot(MR_total_axes(:,i),ARFI_total_axes(:,i)/10,'kx','MarkerSize',14,'LineWidth',3);
-        plot(MR_central_axes(:,i),ARFI_central_axes(:,i)/10,'ro','MarkerSize',14,'LineWidth',3);
+        plot(MR_total_axes(:,i)/10,ARFI_total_axes(:,i)/10,'kx','MarkerSize',14,'LineWidth',3);
+        plot(MR_central_axes(:,i)/10,ARFI_central_axes(:,i)/10,'ro','MarkerSize',14,'LineWidth',3);
         xlabel('MR Measurement (cm)','FontSize',fs);
         ylabel('ARFI Measurement (cm)','FontSize',fs);
 
@@ -111,19 +111,19 @@ function mr_arfi_path_axes
             'FontSize'    , fs);
 
          [image_total_fit,image_total_Rsq]=compute_linreg_Rsq(MR_total_axes(:,i)/10,ARFI_total_axes(:,i)/10);
-         plot(MR_total_axes(:,i),image_total_fit,'-k','LineWidth',3);
+         plot(MR_total_axes(:,i)/10,image_total_fit,'-k','LineWidth',3);
          [image_central_fit,image_central_Rsq]=compute_linreg_Rsq(MR_central_axes(:,i)/10,ARFI_central_axes(:,i)/10);
-         plot(MR_central_axes(:,i),image_central_fit,'-r','LineWidth',3);
+         plot(MR_central_axes(:,i)/10,image_central_fit,'-r','LineWidth',3);
 
        axis([2 8 2 8]); 
 
-       title(sprintf('%s',axis_titles{i}),'FontSize',fs);
+       title(sprintf('%s Axis',axis_titles{i}),'FontSize',fs);
 
        switch i
        case 1
            legend(sprintf('Total Gland (R^2 = %.2f)',image_total_Rsq),sprintf('Central Gland (R^2 = %.2f)',image_central_Rsq),'Location','NorthWest');
         case 2
-           legend(sprintf('Total Gland (R^2 = %.2f)',image_total_Rsq),sprintf('Central Gland (R^2 = %.2f)',image_central_Rsq),'Location','NorthWest');
[…]

### 2013-09-22 figs/mr_arfi_path_axes.m
computing mean & std for mr:arfi axes

@@ -134,7 +134,17 @@ function mr_arfi_path_axes
 
         print('-depsc2',sprintf('Imaging_%s.eps',axis_titles{i}));
         close;
+
+        % compute over/unders for each modality
+        mr_arfi_total_OverUnder(:,i) = compute_over_under(MR_total_axes(:,i)/10,ARFI_total_axes(:,i)/10);
+        mr_arfi_central_OverUnder(:,i) = compute_over_under(MR_central_axes(:,i)/10,ARFI_central_axes(:,i)/10);
     end;
+
+    disp('MR:ARFI Total and Central');
+    mean(mr_arfi_total_OverUnder)
+    std(mr_arfi_total_OverUnder)
+    mean(mr_arfi_central_OverUnder)
+    std(mr_arfi_central_OverUnder)
 
     % I will now compute ratios of AB:LL (1:2), AB:AP (1:3), and LL:AP (2:3) for
     % imaging central and total and path total volumes

### 2013-09-22 fig_mr_arfi_path_axes.tex
added error estimates for mr:arfi axes

@@
 (a) & (b) & (c) \\
+\includegraphics[width=0.3\linewidth]{figs/Imaging_Lateral-to-Lateral} &
+\includegraphics[width=0.3\linewidth]{figs/Imaging_Anterior-to-Posterior} &
+\includegraphics[width=0.3\linewidth]{figs/Imaging_Apex-to-Base} \\
+(d) & (e) & (f) \\
 \end{tabular}
@@
     anatomic axes: lateral-to-lateral (a), anterior-to-posterior (b) and
-apex-to-base (c).  The correlation between the imaging axis measurements and
-pathology was performed for each orientation.  The black dashed-line represents
-the projection of perfectly-correlated measurements between imaging and
-patholoy.  Notice that there is no correlation between imaging and pathology in
-the lateral-to-lateral and anterior-to-posterior axes, with a general
-over-estimation of the lateral-to-lateral axis in the imaging.  The MR
-estimation of the apex-to-base dimension showed moderate correlation with
-pathology, with the ARFI image apex-to-base approximation having an overall
-underestimation in this orientation.  The over-/under-estimation of each imaging modality relative to gross pathology is summarized in Table~\ref{tab:mr_arfi_path_axes}.}
+    apex-to-base (c).  The correlation between the MR and ARFI imaging axis
+    measurements was performed for each orientation.  The black dashed-line
+    represents the projection of perfectly-correlated measurements between
+    imaging and patholoy.  Notice that there is no correlation between imaging
+    and pathology in the lateral-to-lateral and anterior-to-posterior axes,
+    with a general over-estimation of the lateral-to-lateral axis in the
+    imaging.  The MR estimation of the apex-to-base dimension showed moderate
+    correlation with pathology, with the ARFI image apex-to-base approximation
+    having an overall underestimation in this orientation.  The axis
+    measurements of the imaging modalities realtive to one another, for both
+    total gland (black) and central gland (red) are shown in the bottom row.
+    The over-/under-estimation of each imaging modality relative to gross
+    pathology and each other is summarized in
+    Table~\ref{tab:mr_arfi_path_axes}.} 
 \label{fig:mr_arfi_path_axes}

### 2013-09-22 figs/mr_arfi_volumes.m
updated volume figures to be lin regs

@@ -46,18 +46,7 @@ legend('Central Gland','Peripheral Zone','Location','NorthWest');
 legend boxoff;
 %set(gca,'XTickLabel',Pcell)
 
-set(gca, ...
-    'Box'         , 'off'     , ...
-    'TickDir'     , 'out'     , ...
-    'TickLength'  , [.01 .01] , ...
-    'XMinorTick'  , 'off'      , ...
-    'YMinorTick'  , 'off'      , ...
-    'XGrid'       , 'off'      , ...
-    'XColor'      , [0 0 0], ...
-    'YColor'      , [0 0 0], ...
-    'XMinorGrid'  , 'off'      , ...
-    'LineWidth'   , 2, ...
-    'FontSize'    , fs);
+set_axes(gca,fs);
 
 print('-depsc2','mr_arfi_volumes.eps');
 close;
@@ -82,43 +71,81 @@ print('-depsc2','mr_arfi_volume_diff.eps');
 close;
 
 % now lets look at the relative ratios of CG:total volume for each modality
-mr_cg_total = (mr_cg_vol./mr_total_vol)*100;
-arfi_cz_total = (arfi_cz_vol./arfi_total_vol)*100;
-
-figure;
-hold on;
-h=bar([1:length(MRdataStack)],mr_cg_total,bar_width);
-set(h(1),'FaceColor',[0 0 1]);
-g=bar([1:length(ARFIdataStack)]+bar_width,arfi_cz_total,bar_width);
-set(g(1),'FaceColor',[0 1 0]);
-xlabel('Study Subject','FontSize',fs);
-ylabel('Ratio of Central : Total Volume','FontSize',fs);
-title('MR and ARFI Central : Total Ratios','FontSize',fs);
-set(gca,'FontSize',fs);
-a = axis;
-a(2) = 17;
-a(4) = 100;
-axis(a);
-
-text(1,90,sprintf('Mean Diff = %.1f +/- %.1f',mean(arfi_cz_total-mr_cg_total),std(arfi_cz_total-mr_cg_total)),'FontSize',fs);
-
-print('-depsc2','mr_arfi_central_total_diff.eps');
-close;
+%mr_cg_total = (mr_cg_vol./mr_total_vol)*100;
+%arfi_cz_total = (arfi_cz_vol./arfi_total_vol)*100;
+%
+%figure;
+%hold on;
+%h=bar([1:length(MRdataStack)],mr_cg_total,bar_width);
+%set(h(1),'FaceColor',[0 0 1]);
[…]

### 2013-09-22 figs/set_axes.m

@@ -1,0 +1,15 @@
+function set_axes(gca,fs)
+
+set(gca, ...
+    'Box'         , 'off'     , ...
+    'TickDir'     , 'out'     , ...
+    'TickLength'  , [.01 .01] , ...
+    'XMinorTick'  , 'off'      , ...
+    'YMinorTick'  , 'off'      , ...
+    'XGrid'       , 'off'      , ...
+    'XColor'      , [0 0 0], ...
+    'YColor'      , [0 0 0], ...
+    'XMinorGrid'  , 'off'      , ...
+    'LineWidth'   , 2, ...
+    'FontSize'    , fs);
+

### 2013-09-22 fig_mr_arfi_volumes.tex
integrated lin reg vol figs

--- 2013-09-15 fig_mr_arfi_volumes.tex
@@
 \includegraphics[width=0.3\linewidth]{figs/mr_arfi_volumes} &
-\includegraphics[width=0.3\linewidth]{figs/mr_arfi_volume_diff} &
-\includegraphics[width=0.3\linewidth]{figs/mr_arfi_central_total_diff} \\
-(a) Central:Peripheral Volume & (b) Total Volume Difference & (c) Central:Total Volume Difference\\
+\includegraphics[width=0.3\linewidth]{figs/mr_arfi_total_linreg} &
+\includegraphics[width=0.3\linewidth]{figs/mr_arfi_central_linreg} \\
+(a) CG:PZ Volume & (b) Total Gland Volume & (c) Central Gland Volume \\
 \end{tabular}
-\caption{Comparison of MR and ARFI zonal anatomy volume estimates from manually-segmented images.
-Total prostate volumes ranged from 19.6--71.0 cm$^3$ based on MR image models (a), with ARFI image models overestimating total prostate volume by 36.7 $\pm$ 27.9\% (b).
-ARFI image delineation of the central zone volume relative to total prostate volume showed a mean underestimation of -15.3 $\pm$ 13.0\% (c, green) relative to the MR central gland : total volume ratios (c, blue).}
+\caption{Comparison of MR and ARFI zonal anatomy volume estimates from
+    manually-segmented images.  Total prostate gland volumes ranged from 19.6--71.0
+    cm$^3$ based on MR image models (a), with a moderate correlation between MR
+    and ARFI imaging for both the total gland volume (R$^2$ = 0.68, (b)) and the
+    central gland (R$^2$ = 0.41, (c)).  The ARFI total gland volumes had a positive bias
+    compared with the MR volumes (36.7 $\pm$ 27.9\% (b)), while the ARFI
+    central zone volumes were on average more similar (2.1 $\pm$ 39.0\%, (c)), but
+    with some significant variability between study subjects.
+    Table~\ref{tab:mr_arfi_volumes} contains the individual volume estimates
+    for the total prostate and central glands.}
 \label{fig:mr_arfi_volumes} 

### 2013-09-22 figs/mr_arfi_volumes.m
updated title and legend

@@ -36,13 +36,13 @@ set(h(1),'FaceColor',[0 0 1]);
 set(h(2),'FaceColor',[1 0 0]);
 set(g(1),'FaceColor',[0 1 0]);
 set(g(2),'FaceColor',[1 1 0]);
-ylabel('Zone Volumes (cm^3)','FontSize',fs)
+ylabel('Gland/Zone Volumes (cm^3)','FontSize',fs)
 xlabel('Study Subject','FontSize',fs);
-title('MR (Red/Blue) and ARFI (Yellow/Green) Zone Volumes','FontSize',fs);
+title('MR (Red/Blue) and ARFI (Yellow/Green) Volumes','FontSize',fs);
 a=axis;
 a(2) = 17;
 axis(a);
-legend('Central Gland','Peripheral Zone','Location','NorthWest');
+legend('CG','PZ','Location','NorthWest');
 legend boxoff;
 %set(gca,'XTickLabel',Pcell)
 
@@ -117,7 +117,7 @@ figure;
     hold on
     plot(mr_total_vol/1e3,arfi_total_vol/1e3,'kx','MarkerSize',14,'LineWidth',3);
     plot(mr_total_vol/1e3,mr_arfi_total_fit,'-k','LineWidth',3);
-    title('ARFI:MR Total Gland Volume','FontSize',fs);
+    title('ARFI:MR Volumes','FontSize',fs);
     xlabel('MR Volume (cm^3)','FontSize',fs);
     ylabel('ARFI Volume (cm^3)','FontSize',fs);
     set(gca,'FontSize',fs);            

### 2013-09-22 fig_mr_arfi_path_axes.tex
updated figures for correlation analyses and removed image comparisons to path axes

--- 2013-09-16 fig_mr_arfi_path_axes.tex
@@
 \begin{tabular}{ccc}
-\includegraphics[width=0.3\linewidth]{figs/Lateral-to-Lateral} &
-\includegraphics[width=0.3\linewidth]{figs/Anterior-to-Posterior} &
-\includegraphics[width=0.3\linewidth]{figs/Apex-to-Base} \\
+%\includegraphics[width=0.3\linewidth]{figs/Lateral-to-Lateral} &
+%\includegraphics[width=0.3\linewidth]{figs/Anterior-to-Posterior} &
+%\includegraphics[width=0.3\linewidth]{figs/Apex-to-Base} \\
+%(a) & (b) & (c) \\
+\includegraphics[width=0.3\linewidth]{figs/Imaging_Lateral-to-Lateral} &
+\includegraphics[width=0.3\linewidth]{figs/Imaging_Anterior-to-Posterior} &
+\includegraphics[width=0.3\linewidth]{figs/Imaging_Apex-to-Base} \\
 (a) & (b) & (c) \\
+%(d) & (e) & (f) \\
 \end{tabular}
@@
     anatomic axes: lateral-to-lateral (a), anterior-to-posterior (b) and
-apex-to-base (c).  The correlation between the imaging axis measurements and
-pathology was performed for each orientation.  The black dashed-line represents
+    apex-to-base (c).  The correlation between the MR and ARFI imaging axis
+    measurements was performed in each orientation for the total gland (black
+    crosses) and central gland (red circles).  The black dashed-line represents
     the projection of perfectly-correlated measurements between imaging and
-patholoy.  Notice that there is no correlation between imaging and pathology in
-the lateral-to-lateral and anterior-to-posterior axes, with a general
-over-estimation of the lateral-to-lateral axis in the imaging.  The MR
-estimation of the apex-to-base dimension showed moderate correlation with
-pathology, with the ARFI image apex-to-base approximation having an overall
-underestimation in this orientation.}
+    patholoy.  The over-/under-estimation of each imaging modality relative to
+    gross pathology and each other is summarized in
+    Table~\ref{tab:mr_arfi_axes_error}.} 
 \label{fig:mr_arfi_path_axes}

### 2013-09-22 abstract.tex
final draft version

@@
 Prostate cancer (PCa) is the most common non-cutaneous malignancy among men in
-the United States and the second leading cause of cancer related death.  The
-use of non-invasive imaging in the evaluation of PCa could lead to improved
-diagnosis, risk-stratification, and management.  Magnetic resonance imaging
-(MRI) has been available for use in the workup of patients with PCa since the
-early 1980s, and recent advances with functional parameters has greatly
-improved its clinical diagnostic utility.  Acoustic radiation force impulse
-(ARFI) imaging is an ultrasound-based modality that evaluates the mechanical
-properties of tissues. ARFI imaging has the potential to aid in PCa diagnosis
-and management by evaluating the structural composition of prostate zones and
-tumors base on their stiffness.  In this study, MR and ARFI imaging datasets
-were compared with gross pathology measurements immediately post radical
-prostatectomy.  Both imaging modalities showed moderate correlations (0.39 $<$
-R$^2 < $ 0.74) between estimated organ volume and gross pathologic weights and
-estimated volumes from tri-axial measurements.  ARFI images, on average,
-over-estimated prostate volumes by 36\% $\pm$ 28\% compared to MR images,
-primarily due to over-estimation of the lateral (right-left) dimension of the
-prostate.  The central zone volumes of the prostate agreed to within 2.1 $\pm$
-39.1\% between ARFI and MR imaging.  MR and ARFI imaging yielded different
-estimates of organ eccentricity as characterized by ratio of the tri-axial
-measurements, with both imaging modalities estimating greater degrees of
-eccentricity, dominated, again, by lateral axis over-estimation.  ARFI imaging
-of the prostate can accurately delineate the central gland of the prostate and
-the boundaries of the capsule, though care must be taken in delineating the
-lateral edges of the prostate in the posterior peripheral zone.
+the United States and the second leading cause of cancer related death.
+Non-invasive imaging could lead to improved diagnosis, risk-stratification, and
+PCa management.  Magnetic resonance imaging (MRI) has been available for use in
+the workup of patients with PCa since the early 1980s, and recent advances with
+functional parameters has greatly improved its clinical diagnostic utility.
+Acoustic Radiation Force Impulse (ARFI) imaging is an ultrasound-based modality
+that evaluates the mechanical properties of soft tissues. ARFI imaging has the
+potential to aid in PCa diagnosis and management by evaluating the structural
+composition of prostate zones and tumors based on their stiffness.  In this
+study, MR and ARFI imaging datasets were compared to one another and with gross
+pathology measurements made immediately post radical prostatectomy.  Imaging
+datasets were manually segmented to delineate the central gland and prostate
+capsule, and 3D models were rendered to evaluate zonal anatomy dimensions and
+volumes.  Both imaging modalities showed moderate correlations (0.39 $<$ R$^2 <
+$ 0.74) between estimated organ volume and gross pathologic weights ARFI and MR
+total prostate gland volumes were well-correlated (R$^2$ = 0.68), but ARFI
+images yielded prostate volumes that were, on average, larger (36\% $\pm$ 28\%)
+than MR images, primarily due to over-estimation of the anterior-to-posterior
+dimension of the prostate total gland (17.0 $\pm$ 12.1\%), while over-estimated
+of the other dimensions were less significant contributors (8.1 $\pm$ 18.4\%
+and 0.58 $\pm$ 12.9\%).  The central zone volumes of ARFI and MR images were
+also moderated correlated (R$^2$ = 0.41), with minimal volume bias between the
+imaging modalities, but significant variability case-to-case (2.1 $\pm$
+39.1\%).  Central zone volume differences were, again, strongly attributed to
+over-estimation of the anterior-to-posterior axis (14.8 $\pm$ 23.1\%), with a
+significant underestimation of the apex-to-base dimension (-10.8 $\pm$ 22.3\%)
+and no mean bias in the lateral-to-lateral measurements (0.006 $\pm$ 17.2\%).
+Strong variability in central gland volumes is believed to be related to the
+extent of benign prostatic hyperplasia (BPH) for select cases.  Overall, ARFI
+imaging of the prostate yielded prostate volumes and dimensions that were
+correlated with MR T2WI estimates, with biases in the anterior-to-posterior
+dimension, most likely related to poor displacement SNR in the anterior region
+of the prostate from greater distance from the rectal wall imaging surface.
+ARFI imaging is a promising low-cost, real-time imaging modality that can
+compliment MR imaging for diagnosis, treatment planning and management of PCa.

### 2013-09-22 intro.tex
final draft version

--- 2013-09-18 intro.tex
@@
 \section{Introduction}\label{sect:intro}
+Prostate cancer (PCa) is the most common non-cutaneous malignancy among men in
+the United States and the second leading cause of cancer related
+death.~\cite{Howlader2011} Screening with prostate specific antigen (PSA) and
+digital rectal examination (DRE) has led to earlier PCa detection, but
+performance of these measures is not optimal, leading to imprecise risk
+assessment.  Additionally, the random nature of transrectal ultrasound
+(TRUS)-guided biopsies can miss or underestimate the burden of
+cancer.~\cite{Gupta2013} Earlier disease diagnosis leads to challenges in
+deciding optimal management strategies for patients presenting with less
+disease burden, whereas missed tumors on TRUS-guided biopsies can result in
+inappropriate diagnoses.
 
-Kirema's intro text will go here.
+The use of non-invasive prostate imaging could lead to improved PCa diagnosis,
+risk-stratification, and management.  Magnetic resonance imaging (MRI) has been
+available for use in the workup of patients with PCa since the early 1980s, but
+early studies on its diagnostic accuracy were heterogeneous.  However, the more
+recent ability to include functional parameters in PCa MRI analysis has yielded
+promising results.~\cite{Gupta2013,Hricak2007} Among the MRI modalities
+currently used in the study of PCa, it is well established that T2-Weighted
+Imaging (T2WI) offers the best assessment of prostate anatomy due to its
+ability to delineate prostatic margins, distinguish internal structures and
+differentiate among the glandular zones.
 
-Section~\ref{sect:background} provides an overview of MR and ARFI imaging in
-the prostate and an overall clinical motivation for prostate imaging.
-Section~\ref{sect:methods} describes the methods used to experimentally-acquire
-our imaging data, process of gross pathological specimens post radical
-prostatectomy, and image process our datasets.  The results of our analysis,
-including gross pathology and imaging prostate axis and volume estimates, are
-presented in Section~\ref{sect:results}, along with a statistical analysis of
-the bias and variability associated with each imaging modalities measurements,
-all of which is discussed in Section~\ref{sect:discussion}.
+Acoustic Radiation Force Impulse (ARFI) imaging is an ultrasound-based modality
+that evaluates the mechanical properties of tissues.~\cite{Nightingale2002b}
+ARFI imaging has the potential to aid in PCa diagnosis and management by
+evaluating the structural composition of prostate zones and tumors based on
+their stiffness contrast.  Zhai \etal were able to visualize prostatic anatomy
+by utilizing ARFI imaging in freshly-excised prostates. In a second study, Zhai
+\etal demonstrated the feasibility of ARFI prostate imaging
+\invivo.~\cite{Zhai2012}; however, to the authors’ best knowledge, there have
+been no studies to date that compare \invivo ARFI prostate imaging to other
+imaging modalities.~\cite{Zhai2010}
+
+The goal of this study is to evaluate the ability of ARFI imaging to delineate
+prostate zonal anatomy, specifically central gland and total prostate gland,
+\invivo as compared to endorectal MR T2WI.  Section~\ref{sect:background}
+provides an overview of MR and ARFI imaging in the prostate and an overall
+clinical motivation for prostate imaging.  Section~\ref{sect:methods} describes
+the methods used to experimentally-acquire our imaging data, process of gross
+pathological specimens post radical prostatectomy, and image process our
+datasets.  The results of our analysis, including gross pathology and imaging
+prostate axis and volume estimates, are presented in
+Section~\ref{sect:results}, along with a statistical analysis of the bias and
+variability associated with each imaging modalities measurements, all of which
+is discussed in Section~\ref{sect:discussion}.

### 2013-09-22 clinical_significance.tex
final draft version

\subsection{Clinical Background and Significance}
PCa is the most common non-cutaneous malignancy among men in the United States.
Approximately 1 in every 6 men will develop PCa during their lifetime, with the
median age of diagnosis at 67 years old.~\cite{Howlader2011} PCa is also the
second leading cause of cancer-related death, with 1 in 36 men dying from the
disease.  The National Cancer Institute (NCI) estimates that 238,590 men will
be diagnosed with PCa in 2013 and 29,720 will die from the
disease.~\cite{Howlader2011}

PCa diagnosis usually begins by screening with prostate specific antigen (PSA)
and digital rectal examination (DRE).  Definitive diagnosis is made by random
transrectal ultrasound (TRUS)-guided biopsies, which are then used to provide
the clinician with the proper Gleason score. The combination of these factors,
as well as staging, determines the appropriate therapy and prognosis.

PCa screening has led to earlier diagnosis of smaller tumors and more localized
disease; however, it is well known that the sensitivity and specificity of PSA
and DRE are not optimal. In addition, DRE has a low predictive value at lower
PSA ranges, and PSA yields many false
positives.~\cite{Gosselaar2007,Gupta2013,Hricak2007} As such, a theoretical
risk of over-diagnosis and treatment of low-grade, and possibly clinically
insignificant, disease exists.  Moreover, due to the random nature of biopsies,
PCa located outside the routine sampling sites can be missed and the extent
of the cancer might be underestimated.~\cite{Gupta2013,Cornud1012} For example,
in a study by Mufarrij \etal, 45.9--47.2\% of patients who were candidates for
active surveillance, but underwent radical prostatectomy, had a higher Gleason
score on final histopathology than after TRUS biopsy.~\cite{Mufarrij2010} These
inaccuracies may lead to inappropriate diagnosis, imprecise risk assessment and
potentially avoidable morbidity.

### 2013-09-22 prostate_anatomy.tex
final draft version

@@
 \subsection{Prostate Anatomy}
-
 The prostate gland sits below the urinary bladder and surrounds the urethra.
-Its superior borders include the bladder and seminal vesicles and the
+Its superior borders include the bladder and seminal vesicles, and the
 urogenital diaphragm delineates its inferior boundary. The gland is bordered
 anteriorly by the pubic symphysis and posteriorly by the rectum.  The prostate
-is separated from the rectum by 2--3 mm fascial layer,~\cite{Jung2012} and can
-be easily palpated on rectal examination.
+is separated from the rectum by a 2--3 mm fascial layer,~\cite{Jung2012} and it
+can be easily palpated on rectal examination.
 
-The gland can be divided from superior to inferior into the base, midgland and
-apex. The urethra enters the prostate proximally at the base and extends to the
-midgland at which point the ejaculatory ducts open into the urethra at the
-verumontanum.~\cite{Jung2012} The urethra then continues past the apex and
-travels through the penis. The prostate can be divided into glandular and
-non-glandular components.  The glandular components include the transitional
-zone, central zone and peripheral zone. Each zone contains approximately 5\%,
-20\% and 70--80\% of glandular tissue, respectively.~\cite{Bonekamp2011} The
-non-glandular components include the anterior fibromuscular stroma and the
-urethra.
+The prostate gland can be divided from superior-to-inferior into the base,
+mid-gland and apex. The urethra enters the prostate proximally at the base and
+extends to the mid-gland, at which point the ejaculatory ducts open into the
+urethra at the verumontanum.~\cite{Jung2012} The urethra then continues past
+the apex and travels through the penis. The prostate can be divided into
+glandular and non-glandular components.  The glandular components include the
+transitional zone (TZ), central zone (CZ) and peripheral zone (PZ). In the
+healthy prostate, each zone contains approximately 5\%, 20\% and 70--80\% of
+glandular tissue, respectively.~\cite{Bonekamp2011} The non-glandular
+components include the anterior fibromuscular stroma (AFS) and the urethra.  In
+some imaging modalities, the central and transition zones cannot be visualized
+as discrete entities, and they are collectively identified as the central gland
+(CG).  This CG was one of the prostate structures manually segmented in MR and
+ARFI images in this study.
 
@@
 tissue. Two neurovascular bundles course posterior and lateral to the prostate,
-which can also be invaded by malignant cells.
+which can also be invaded by malignant cells.  The prostate capsule is another
+structure that was manually segmented in both MR and ARFI images in this study,
+and its overall extent was used to delineate the total prostate gland.

### 2013-09-22 arfi_background.tex
final draft version

@@
 \subsection{Acoustic Radiation Force Impulse (ARFI) Imaging}
-ARFI imaging is an ultrasound technique that evaluates the mechanical
-properties of tissues. It generates short-duration acoustic radiation forces
-that result in tissue displacement. The response can then be measured for
-tissue characterization, with displacement being inversely proportional to
-stiffness.~\cite{Nightingale2002} This characterization is of particular
-importance in prostate evaluation due to the gland’s complex and heterogeneous
-composition.  Previous studies have shown that the central zone and PCA can be
-up to 3 times stiffer than PZ tissue.~\cite{Zhai2010a,Zhai2012,Zhai2010b} This
-finding can be explained by the PZ’s higher water content and is analogous to
-the higher signal intensity of the PZ on T2WI.
-
-The potential of ARFI imaging to distinguish normal prostatic anatomy from
-pathologic tissue is promising in the future of PCa diagnosis and image guided
-therapies. B-mode US, which is currently used for real-time guidance in
-TRUS-guided biopsies, does not have the ability to clearly visualize PCa or to
-differentiate it from other normal structures or disease processes such as
-benign prostatic hyperplasia (BPH) or prostatitis.~\cite{Zhai2010a} This has
-resulted in a systematic yet random approach to prostate core biopsies, with
-6-12 cores routinely sampled from different anatomical regions.~\cite{Loch2004}
-The ability to clearly distinguish PCa from other tissues could lead to
-targeted biopsies with increased sensitivity and resultant decreased morbidity.
-It could also aid in the development of targeted therapies for those with low
-burden of disease.
+Acoustic radiation force-based elasticity imaging has recently been developed
+for investigation of mechanical properties of human soft tissues, including
+liver, arteries, heart, prostate, and
+breast~\cite{Bouchard2009,Dumont2009,Trahey2004,Dahl2009,proc_congdon04,Palmeri2011,Zhai2010a,hsu07,Palmeri2011a}.
+and it provides a unique tool to use to characterize the cervix during
+pregnancy.  The acoustic radiation force $\vec{F}$ generated in soft tissues by
+focused ultrasound can be described by~\cite{nyborg65,torr84}:
+\begin{equation}
+\vec{F} = \frac{2 \alpha \vec{I}}{c},
+\label{eqn:radforce}
+\end{equation}
+where $\alpha$ is the acoustic attenuation coefficient of the tissue, $c$ is
+the tissue's sound speed, and $\vec{I}$ is the acoustic intensity at a given
+point in space.  The acoustic radiation force is generated by a transfer of
+momentum from the propagating acoustic wave to the propagation medium through
+attenuation mechanisms, such as absorption and scattering of the ultrasonic
+wave.  In these studies, the acoustic attenuation and sound speed of the
+prostate tissue were assumed to be constant and not vary as a function of the
+zonal anatomy, meaning that all displacement differences seen in the prostate
+would be assumed to be stiffness-related after normalizing for depth-dependent
+focal gain.

### 2013-09-22 mr_background.tex
final draft version

@@
-\subsection{Clinical Background and Significance}
-Prostate cancer (PCa) is the most common non-cutaneous malignancy among men in
-the United States. Approximately 1 in every 6 men will develop PCa during their
-lifetime, with the median age of diagnosis at 67. 1It is also the second
-leading cause of cancer related death, with 1 in 36 men dying from the disease.
-The National Cancer Institute estimates that 238,590 men will be diagnosed with
-PCa in 2013 and 29,720 will die from the disease.~\cite{Howlader2011}
+\subsection{MR Imaging in Prostate Diagnostics}
+MR imaging has been available for use in the workup of patients with PCa since
+the early 1980s, but the early work on its diagnostic accuracy is heterogeneous
+due to heavy reliance on morphology via T1 and T2-weighted imaging (T2WI). More
+recent abilities to include not only anatomic, but also biologic and functional
+dynamic parameters, into MR analysis via diffusion-weighted imaging (DWI),
+dynamic contrast-enhanced (DCE) imaging or MR spectroscopic imaging (MRSI) is
+promising in the future diagnosis and management of PCa.
 
-PCa diagnosis usually begins by screening with prostate specific antigen (PSA)
-and digital rectal examination (DRE).  Definitive diagnosis is made by random
-transrectal ultrasonography-guided (TRUS) biopsies, which are then used to
-provide the clinician with the proper Gleason score. The combination of these
-factors—as well as staging—determines the appropriate therapy and prognosis.
-
-PCa screening has led to earlier diagnosis of smaller tumors and more localized
-disease.  However, it is well known that the sensitivity and specificity of PSA
-and DRE are not optimal. In addition, DRE has a low predictive value at lower
-PSA ranges, and PSA yields many false
-positives.~\cite{Gosselaar2007,Gupta2013,Hricak2007} As such, a theoretical
-risk of over-diagnosis and treatment of low-grade—and possibly clinically
-insignificant—disease exists. Moreover, due to the random nature of biopsies,
-cancer located outside the routine sampling site can be missed and the extent
-of the cancer might be underestimated.~\cite{Gupta2013,Cornud1012} For example,
-in a study by Mufarrij et al. 45.9--47.2\% of patients who were candidates for
-active surveillance but underwent radical prostatectomy had a higher Gleason
-score on final histopathology than after TRUS biopsy.~\cite{Mufarrij2010} These
-inaccuracies may lead to inappropriate diagnosis, imprecise risk assessment and
-potentially avoidable morbidity.
-
-\subsection*{The Use of Magnetic Resonance Imaging in Prostate Diagnostics}
-Magnetic resonance (MR) imaging has been available for use in the workup of
-patients with PCa since the early 1980s but the early work on its diagnostic
-accuracy is heterogeneous.7 Earlier MR techniques relied mostly on morphology
-via T1 and T2- weighted imaging (T2WI). The more recent ability to include not
-only anatomic but also biologic and functional dynamic parameters into MR
-analysis—via diffusion-weighted imaging (DWI), dynamic contrast-enhanced (DCE)
-imaging or MR spectroscopic imaging (MRSI)—is promising in the future diagnosis
-and management of PCA.
-
-Currently prostate MR focuses on a multiparametric approach, where 2 or more
-imaging sequences—including anatomic and functional data—are used together to
-try to arrive to a diagnosis.~\cite{Barentsz2012} As MR technology continues to evolve and
-improve, its role in PCA diagnosis, staging, treatment planning and follow-up
+Currently prostate MR focuses on a multiparametric approach, where two-or-more
+imaging sequences, including anatomic and functional data, are used together to
+make a diagnosis.~\cite{Barentsz2012} As MR technology continues to evolve and
+improve, its role in PCa diagnosis, staging, treatment planning and follow-up
 has gained much attention.
 
-\subsection{T2-Weighted Imaging and Prostate Anatomy}
+\subsubsection{T2-Weighted Imaging and Prostate Anatomy}
 T2WI sequences are crucial components of prostate MR imaging.  T2WI is
@@
 and a small field of view of approximately 14 cm.~\cite{Gupta2013,Bonekamp2011}
-T2 sequences are the most helpful for tumor localization as they can clearly
+T2 sequences are the most helpful for tumor localization, as they can clearly
 show overall prostate morphology, internal structures and prostatic
@@
 
-Approximately 70\% of the prostatic tissue is found in the peripheral zone,
-which is high in water content and thus of higher signal intensity in
-T2WI.~\cite{Jung2012} Seventy five percent of prostatic tumors are found in the
-PZ and normally show hypointense T2 signal when compared to the higher
-intensity PZ.~\cite{Hricak2007,Hegde2013} However, tumors can sometimes seem of
-similar intensity as the surrounding tissue and false positives can occur
-secondary to post biopsy changes/hemorrhage, hyperplasia or prostatitis, making
-diagnosis more challenging.~\cite{Hegde2013}
+Approximately 70\% of the prostatic tissue is found in the PZ, which is high in
+water content, and thus has higher signal intensity in T2WI.~\cite{Jung2012}
+Seventy-five percent of prostatic tumors are found in the PZ and normally show
+hypointense T2 signal when compared to the higher intensity
+PZ.~\cite{Hricak2007,Hegde2013}; however, tumors can sometimes have similar
+intensity as the surrounding tissue and false positives can occur secondary to
+post-biopsy changes/hemorrhage, hyperplasia or prostatitis, making diagnosis
+more challenging.~\cite{Hegde2013}
 
-\subsection{Functional MR Sequences}
+\subsubsection{Functional MR Sequences}
 Even though T2WI is the mainstay of prostate MR, its overall performance in
@@
 significantly improve the performance of MRI in cancer
-diagnosis.~\cite{Turkbey2012} In fact, the European Society of Urogenital
-Radiology’s (ESUR) prostate MR guidelines recommend at least 2 functional
-imaging techniques in addition to T2WI in order to better characterize prostate
-tumors.~\cite{Barentsz2012} In a study by Turkbey et al., researchers found
-that mpMRI had a positive predictive value of 98\% in prostate cancer
-detection.~\cite{Turkbey2012} Functional sequences include diffusion- weighted
-imaging (DWI), dynamic contrast enhanced imaging (DCE-MRI) and MR spectroscopic
-imaging (MRSI).
+diagnosis.~\cite{Turkbey2012}  The European Society of Urogenital Radiology’s
+(ESUR) prostate MR guidelines recommend at least 2 functional imaging
+techniques, in addition to T2WI, to better characterize prostate
+tumors.~\cite{Barentsz2012} In a study by Turkbey \etal, researchers found that
+mpMRI had a PCa detection positive predictive value of 98\%.~\cite{Turkbey2012}
+Functional sequences include diffusion-weighted imaging (DWI), dynamic contrast
+enhanced imaging (DCE-MRI) and MR spectroscopic imaging (MRSI).  While mpMRI is
+critical in improving PCa detection, T2WI is mainstay of delineating prostate
+anatomy, and for the purposes of comparing prostate anatomy between ARFI and MR
+images, T2WI were used exclusively in this study.
 
-DWI is based on the free movement of water particles in tissue and measures the
-degree of motion restriction.~\cite{Koh2007} Normal prostatic tissue is very
-glandular with plenty of water molecule movement. On the other hand, tumors
-have high cellular density and restricted water movement, which leads to
-decreased diffusion. Due to this restricted diffusion, tumors seem to be of
-higher intensity on DWI.~\cite{Koh2007}
-
-DCE-MRI provides a measure of tumor vascularity and takes advantage of
-tumor-induced angiogenesis in PCa.~\cite{Noworolski2005} Increased vascularity
-in PCa results in earlier and higher peak contrast enhancement and a faster
-washout when compared to normal prostatic tissue. These known characteristics
-of PCa are helpful when trying to localize lesions.
-
-Like DWI, MRSI can be helpful in lesion characterization as it provides
-information on the presence of certain metabolites in
-tissue.~\cite{Barentsz2012} Choline and citrate are especially useful in the
-setting of PCa. Choline is critical in cell membrane synthesis and is thus
-usually elevated in cells with cancerous behavior.~\cite{Gupta2013} Healthy
-prostate epithelium synthesizes and secretes large quantities of citrate but
-levels are decreased in PCa.~\cite{Costello1999,Kurhanewicz2002} Therefore, an
-increase in the choline-to-citrate ratio on MRSI can be used as an indicator of
-malignancy.~\cite{Bonekamp2011}
+%DWI is based on the free movement of water particles in tissue and measures the
+%degree of motion restriction.~\cite{Koh2007} Normal prostatic tissue is very
+%glandular with plenty of water molecule movement. On the other hand, tumors
+%have high cellular density and restricted water movement, which leads to
+%decreased diffusion. Due to this restricted diffusion, tumors seem to be of
+%higher intensity on DWI.~\cite{Koh2007}
+%
+%DCE-MRI provides a measure of tumor vascularity and takes advantage of
+%tumor-induced angiogenesis in PCa.~\cite{Noworolski2005} Increased vascularity
+%in PCa results in earlier and higher peak contrast enhancement and a faster
+%washout when compared to normal prostatic tissue. These known characteristics
+%of PCa are helpful when trying to localize lesions. 
+%
+%Like DWI, MRSI can be helpful in lesion characterization as it provides
+%information on the presence of certain metabolites in
+%tissue.~\cite{Barentsz2012} Choline and citrate are especially useful in the
+%setting of PCa. Choline is critical in cell membrane synthesis and is thus
+%usually elevated in cells with cancerous behavior.~\cite{Gupta2013} Healthy
+%prostate epithelium synthesizes and secretes large quantities of citrate but
+%levels are decreased in PCa.~\cite{Costello1999,Kurhanewicz2002} Therefore, an
+%increase in the choline-to-citrate ratio on MRSI can be used as an indicator of
+%malignancy.~\cite{Bonekamp2011}

### 2013-09-22 arfi_methods.tex
final draft version

@@
-\subsection{ARFI Imaging Methods}
+\subsection{ARFI Imaging}
 Experimental ARFI imaging data were acquired using a modified Siemens Acuson
-SC2000 ultrasound scanner (Siemens Healthcare, Ultrasound Business Unit,
+SC2000\texttrademark~ultrasound scanner (Siemens Healthcare, Ultrasound Business Unit,
 Mountain View, CA, USA) and the longitudinal array of an Acuson ER7B
-transducer~\cite{Rosenzweig2012}.  The ARFI imaging sequence was comprised of
+transducer.  The ARFI imaging sequence was comprised of
 standard B-mode ultrasonic imaging, or tracking beams, and pushing beams. For
@@
 was tracked for up to 6ms at a PRF of 8kHz. This pushing strategy is similar to
-what has been published by Bercoff et al~\cite{BercoffSSI}. The 30 mm and 22.5
+what has been published by Bercoff \etal~\cite{Bercoff2004}. The 30 mm and 22.5
 mm foci pushing pulses were transmitted at 4.6 MHz with a F/2 geometry and the
@@
 with all 4 beams located inside the beamwidth of the pushing pulses such that
-the beam spacing was 0.17 mm. The other twelve lines were separated into two
-groups to observe both the left and right propagating waves. For each sub-group
-of 6, the beams were located 1.89 mm laterally offset from the push and had an
-inter-beam spacing of 0.76 mm to cover a total field of view of 11.3 mm
-laterally. For brevity, the 4 on-axis beams will be referred to as the ARFI
-data and the 12 off-axis beams will be referred to as the SWEI data.
+the beam spacing was 0.17 mm.
 
-Displacement estimation was performed using a phase shift estimator on the
-beamformed IQ data~\cite{Loupas1,Pinton1}. The ARFI data were then normalized
-as a function of depth to account for attenuation and focal gain effects by
-taking the mean displacement across the entire data set at each time step then
-low-pass filtering with a cutoff frequency of 0.8 mm$^{-1}$.
+Displacement estimation was performed using a phase-shift estimator on the
+beamformed in-phase and quadrature (IQ) data~\cite{Loupas95,pinton06}. The ARFI
+data were then normalized as a function of depth to account for attenuation and
+focal gain effects.  This normalization was performed using a displacement
+profile measured in a homogeneous tissue-mimicking phantom, which was applied
+to all displacements in the entire data set at each time step, and then
+low-pass filtered with a cutoff-frequency of 0.8 mm$^{-1}$.

### 2013-09-22 mr_methods.tex
final draft version

@@
-\subsection{Study Inclusion Criteria \& MR Imaging}
-Prostates removed by radical prostatectomy were used in this IRB-approved (Duke
-IRB\# Pro00006458), HIPAA-compliant study from men ranging in age from XX--XX
-diagnosed with biopsy-proven prostate cancer destined for surgical removal.
-Between DATES, a total of XX patients were recruited and enrolled in this
-study. Inclusion criteria were undergoing complete pelvic MRI with endorectal
-coil for detection of prostate cancer, including multiplanar T2-weighted
-anatomic imaging, diffusion-weighted imaging (DWI), and dynamic contrast
-enhanced MRI (DCE-MRI) as well as radical prostatectomy and whole mount
-histology. Patients with previous treatments of prostate cancer or benign
-prostatic hyperplasia (BPH), or anatomic anomalies of the rectum, were
-excluded.  All patients enrolled in this study provided written informed
-consent. Our final cohort included XX subjects. Table XX has a summary of the
-patient demographics for the cases shown in this manuscript.
+\subsection{Study Inclusion Criteria}
+Patients undergoing radical prostatectomy for biopsy-proven PCa treatment were
+enrolled as study subjects in this IRB-approved (Duke IRB\# Pro00006458),
+HIPAA-compliant study.  A total of 16 patients were recruited and enrolled in
+this study.  Inclusion criteria were undergoing complete pelvic MRI with
+endorectal coil for detection of prostate cancer, including multiplanar
+T2-weighted anatomic imaging, as well as pre-operative ARFI imaging and radical
+prostatectomy.  Patients with previous treatments of PCa or benign prostatic
+hyperplasia (BPH), or anatomic anomalies of the rectum, were excluded from this
+study.  All patients enrolled in this study provided written informed consent.
 
-All imaging was performed on one of two 3.0 Tesla MR scanners (General Electric
-HDx, GE Healthcare, Waukesha, WI;  Siemens Skyra, Siemens Healthcare, Erlangan,
-Germany) using a single channel Medrad eCoil endorectal coil (Medrad,
-Indianola, PA) as well as multichannel surface coils.  Imaging sequences
+\emph{CO-AUTHOR QUESTION: DO WE WANT TO INCLUDE A TABLE OF PATIENT DEMOGRAPHIC INFORMATION?}
+
+\subsection{MR Imaging}
+All MR imaging was performed on one of two 3.0 Tesla MR scanners (General
+Electric HDx, GE Healthcare, Waukesha, WI;  Siemens Skyra, Siemens Healthcare,
+Erlangan, Germany) using a single channel Medrad eCoil endorectal coil (Medrad,
+Indianola, PA), as well as multi-channel surface coils.  Imaging sequences
 included thin-section (3 mm section thickness) fast spin echo T2-weighted
@@
 table with full MR parameters) Prostates were radically removed using a da
-Vinci Surgical System (INSERT COMPANY INFORMATION).  After exision, the
-prostates were formalin fixed for at least 24 hours without being cut, and then
-processed for whole mount histology.
+Vinci Surgical System (Intuitive Surgical\textregistered, Inc., Sunnyvale, CA).
+After exision, the prostates were weighed and measured, formalin fixed for at
+least 24 hours without being cut, and then processed for whole mount histology.

### 2013-09-22 image_proc_methods.tex
final draft methods

--- 2013-09-13 image_proc_methods.tex
@@
 \subsection{Image Zonal Anatomy Segmentation and 3D Model Rendering}
-Axial T2W MR slices were manually segmented using the polygon tool on ITK SNAP
-using separate labels for the peripheral zone (PZ), central gland (CG) and
-anterior fibromuscular stroma (AFS). The gland was segmented from base to apex.
-The base was identified below the bladder and subsequent images were segmented
-until the last slice with visible prostatic tissue was identified caudally. The
-CG, PZ and AFS were segmented independently according to their well-established
-anatomical characteristics on
+Axial MR T2WI images were manually segmented using the smooth polygon tool in
+ITK-SNAP~\cite{Yushkevich2006}, using unique labels for the PZ, CG and AFS. The
+gland was segmented from base to apex.  The base was identified below the
+bladder, and subsequent images were segmented until the last slice with visible
+prostatic tissue was identified caudally. The CG, PZ and AFS were segmented
+independently according to their well-established anatomical characteristics on
 T2WI.~\cite{Verma2011,Jung2012,Poon1985,Hricak2007,Bonekamp2011} The PZ was
@@
 its low T2 signal intensity and its location anterior to the central gland.
+
+ARFI images were segmented using a similar procedure to the MR images.
+
+\emph{CO-AUTHOR QUESTION: CAN WE EASILY DELINEATE OUR SEGMENTATION CRITERIA?}
+
+Segmented image stacks were imported into 3D Slicer (v4.3.0) and 3D
+models were rendered using the following parameters (Table~\ref{tab:3dslicer}):
+
+\begin{table}[h!]
+\centering
+\caption{3D model volume rendering parameters}
+\begin{tabular}{ll}
+{\bf Parameter} & {\bf Value} \\ \hline
+Decimation & 0.1 \\
+Smoothing Algorithm & Laplacian \\
+Smoothing  & 70.0 \\
+Joint Smoothing & Enabled \\
+\end{tabular}
+\end{table}
+
+The 3D slicer models were used to render volume estimates of the PZ and CG,
+with the sum of PZ and CG representing the total prostate gland volume.  AFS
+volume estimates in select MR cases were included in the total prostate gland
+volume estimates.  Orthogonal tri-axial measurements in the lateral-to-lateral,
+axex-to-base, and anterior-to-posterior dimensions were made in 3D slicer.

### 2013-09-22 mr_methods.tex
final draft methods

@@
-\subsection{Study Inclusion Criteria}
-Patients undergoing radical prostatectomy for biopsy-proven PCa treatment were
-enrolled as study subjects in this IRB-approved (Duke IRB\# Pro00006458),
-HIPAA-compliant study.  A total of 16 patients were recruited and enrolled in
-this study.  Inclusion criteria were undergoing complete pelvic MRI with
-endorectal coil for detection of prostate cancer, including multiplanar
-T2-weighted anatomic imaging, as well as pre-operative ARFI imaging and radical
-prostatectomy.  Patients with previous treatments of PCa or benign prostatic
-hyperplasia (BPH), or anatomic anomalies of the rectum, were excluded from this
-study.  All patients enrolled in this study provided written informed consent.
-
-\emph{CO-AUTHOR QUESTION: DO WE WANT TO INCLUDE A TABLE OF PATIENT DEMOGRAPHIC INFORMATION?}
-
 \subsection{MR Imaging}
@@
 Vinci Surgical System (Intuitive Surgical\textregistered, Inc., Sunnyvale, CA).
-After exision, the prostates were weighed and measured, formalin fixed for at
-least 24 hours without being cut, and then processed for whole mount histology.

### 2013-09-22 path_methods.tex
final draft methods

\subsection{Pathology Processing}
After excision, the prostates were weighed and tri-axially measured, formalin
fixed for at least 24 hours without being cut, and then processed for whole
mount histology.  While tri-axial measurements of the gross prostate were not necessarily anatomically aligned, rough estimates of the prostate volume were made using a tri-axial, ellipsoidal organ volume assumption, where the volume could be estimated as:

\begin{equation}
V = \frac{4}{3}\pi \sqrt{\det{A^{-1}}},
\end{equation}

where $A$ represents the eigenvector matrix of the tri-axial pathology
measurements.

### 2013-09-22 study_methods.tex
final draft methods

\subsection{Study Inclusion Criteria}
Patients undergoing radical prostatectomy for biopsy-proven PCa treatment were
enrolled as study subjects in this IRB-approved (Duke IRB\# Pro00006458),
HIPAA-compliant study.  A total of 16 patients were recruited and enrolled in
this study.  Inclusion criteria were undergoing complete pelvic MRI with
endorectal coil for detection of prostate cancer, including multiplanar
T2-weighted anatomic imaging, as well as pre-operative ARFI imaging and radical
prostatectomy.  Patients with previous treatments of PCa or benign prostatic
hyperplasia (BPH), or anatomic anomalies of the rectum, were excluded from this
study.  All patients enrolled in this study provided written informed consent.

\emph{CO-AUTHOR QUESTION: DO WE WANT TO INCLUDE A TABLE OF PATIENT DEMOGRAPHIC INFORMATION?}

### 2013-09-22 conclusions.tex
final draft version

@@
 The delineation of prostate zonal anatomy in ARFI images has been compared with
-the established methods for identifying zonal anatomy using T2 MR images.  In
-XX cases of prostates containing varying degrees of PCa and BPH, perpipheral
-zone volumes\ldots XXXXX and central gland volumes\ldots XXXXX.  Aspect ratios
-of the central glad agreed to within XX\% between MR and ARFI imaging datasets.
-Appreciable amounts of BPH made determining the transition between the central
-gland and peripheral zone difficult to discern and, these cases were not
-included in this analysis.  Additionally, large PCa lesions can also distort
-prostate zonal anatomy appreciably, making the distinction between peripheral
-zone and central gland challenging.  Overall, ARFI imaging is able to delineate
-central gland from peripheral zone in the prostate in the absence of excessive
-BPH or PCa tumor burden.
+the established methods for identifying zonal anatomy using MR T2W images.
+Both imaging modalities showed moderate correlations (0.39 $<$ R$^2 < $ 0.74)
+between estimated organ volume and gross pathologic weights ARFI and MR total
+prostate gland volumes were well-correlated (R$^2$ = 0.68), but ARFI images
+yielded prostate volumes that were, on average, larger (36\% $\pm$ 28\%) than
+MR images, primarily due to over-estimation of the anterior-to-posterior
+dimension of the prostate total gland (17.0 $\pm$ 12.1\%), while over-estimated
+of the other dimensions were less significant contributors (8.1 $\pm$ 18.4\%
+and 0.58 $\pm$ 12.9\%).  The central zone volumes of ARFI and MR images were
+also moderated correlated (R$^2$ = 0.41), with minimal volume bias between the
+imaging modalities, but significant variability case-to-case (2.1 $\pm$
+39.1\%).  Central zone volume differences were, again, strongly attributed to
+over-estimation of the anterior-to-posterior axis (14.8 $\pm$ 23.1\%), with a
+significant underestimation of the apex-to-base dimension (-10.8 $\pm$ 22.3\%)
+and no mean bias in the lateral-to-lateral measurements (0.006 $\pm$ 17.2\%).
+Strong variability in central gland volumes is believed to be related to the
+extent of benign prostatic hyperplasia (BPH) for select cases.  Overall, ARFI
+imaging of the prostate yielded prostate volumes and dimensions that were
+correlated with MR T2WI estimates, with biases in the anterior-to-posterior
+dimension, most likely related to poor displacement SNR in the anterior region
+of the prostate from greater distance from the rectal wall imaging surface.
+ARFI imaging is a promising low-cost, real-time imaging modality that can
+compliment MR imaging for diagnosis, treatment planning and management of PCa.

### 2013-09-22 fig_mr_arfi_path_axes.tex
updated colors to agree w/ slicer figure

@@
     apex-to-base (c).  The correlation between the MR and ARFI imaging axis
-    measurements was performed in each orientation for the total gland (black
-    crosses) and central gland (red circles).  The black dashed-line represents
+    measurements was performed in each orientation for the total gland (red
+    crosses) and central gland (blue circles).  The black dashed-line represents
     the projection of perfectly-correlated measurements between imaging and

### 2013-09-22 figs/mr_arfi_path_axes.m
changed colors to match slicer figure

@@ -92,8 +92,8 @@ function mr_arfi_path_axes
     for i=1:3,
         figure;
         hold on;
-        plot(MR_total_axes(:,i)/10,ARFI_total_axes(:,i)/10,'kx','MarkerSize',14,'LineWidth',3);
-        plot(MR_central_axes(:,i)/10,ARFI_central_axes(:,i)/10,'ro','MarkerSize',14,'LineWidth',3);
+        plot(MR_total_axes(:,i)/10,ARFI_total_axes(:,i)/10,'rx','MarkerSize',14,'LineWidth',3);
+        plot(MR_central_axes(:,i)/10,ARFI_central_axes(:,i)/10,'bo','MarkerSize',14,'LineWidth',3);
         xlabel('MR Measurement (cm)','FontSize',fs);
         ylabel('ARFI Measurement (cm)','FontSize',fs);
 
@@ -111,9 +111,9 @@ function mr_arfi_path_axes
             'FontSize'    , fs);
 
          [image_total_fit,image_total_Rsq]=compute_linreg_Rsq(MR_total_axes(:,i)/10,ARFI_total_axes(:,i)/10);
-         plot(MR_total_axes(:,i)/10,image_total_fit,'-k','LineWidth',3);
+         plot(MR_total_axes(:,i)/10,image_total_fit,'-r','LineWidth',3);
          [image_central_fit,image_central_Rsq]=compute_linreg_Rsq(MR_central_axes(:,i)/10,ARFI_central_axes(:,i)/10);
-         plot(MR_central_axes(:,i)/10,image_central_fit,'-r','LineWidth',3);
+         plot(MR_central_axes(:,i)/10,image_central_fit,'-b','LineWidth',3);
 
        axis([2 8 2 8]); 
 

### 2013-09-23 abstract.tex
Kirema typo corrections

@@
 than MR images, primarily due to over-estimation of the anterior-to-posterior
-dimension of the prostate total gland (17.0 $\pm$ 12.1\%), while over-estimated
+dimension of the prostate total gland (17.0 $\pm$ 12.1\%), while over-estimates
 of the other dimensions were less significant contributors (8.1 $\pm$ 18.4\%
 and 0.58 $\pm$ 12.9\%).  The central zone volumes of ARFI and MR images were
-also moderated correlated (R$^2$ = 0.41), with minimal volume bias between the
+also moderately correlated (R$^2$ = 0.41), with minimal volume bias between the
 imaging modalities, but significant variability case-to-case (2.1 $\pm$

### 2013-09-23 conclusions.tex
Kirema typo corrections

@@
 MR images, primarily due to over-estimation of the anterior-to-posterior
-dimension of the prostate total gland (17.0 $\pm$ 12.1\%), while over-estimated
+dimension of the prostate total gland (17.0 $\pm$ 12.1\%), while over-estimates
 of the other dimensions were less significant contributors (8.1 $\pm$ 18.4\%
 and 0.58 $\pm$ 12.9\%).  The central zone volumes of ARFI and MR images were
-also moderated correlated (R$^2$ = 0.41), with minimal volume bias between the
+also moderately correlated (R$^2$ = 0.41), with minimal volume bias between the
 imaging modalities, but significant variability case-to-case (2.1 $\pm$

### 2013-09-23 image_proc_methods.tex
more ultrasound segmentation details added

@@
 
-ARFI images were segmented using a similar procedure to the MR images.
+ARFI images were segmented using a similar procedure to the MR images.  B-mode
+images were used to segment the prostate capsules since contrast between the
+peripheral zone and the periprostatic fat can be low, and poor displacement SNR
+can be present along the anterior aspect of the prostate in ARFI images, making
+it difficult to delineate that boundary.  The central gland was segmented as a
+stiffer region, relative to the peripheral zone, in the center of the prostate.
 

### 2013-09-23 fig_mr_arfi_weight.tex
updated with all of the latest data

@@
     volumes (R$^2$ = 0.08 and 0.01, respectively) (c).  It should be noted in
-    these figures that Study Subject 4 had an excessively large prostate that
-    was difficult to fully capture in imaging, and was therefore grossly
-    underestimated in size.  We retrospectively excluded that study subject
-    from the analysis and re-calculated the coefficients of determination,
-    showing that MR and ARFI imaging improved to R$^2$ = 0.74 and 0.60,
-    respectively, for prostate weight (d), and R$^2$ = 0.39 for both imaging
-    modalities relative to pathology gross volume (e).}
+    these figures that study subject 4 had an excessively large prostate that
+    was difficult to fully capture in imaging, especially in its anterior
+    region, and was therefore grossly underestimated in size.  We
+    retrospectively excluded that study subject from the analysis and
+    re-calculated the coefficients of determination, showing that MR and ARFI
+    imaging improved to R$^2$ = 0.74 and 0.60, respectively, for prostate
+    weight (d), and R$^2$ = 0.39 for both imaging modalities relative to
+    pathology gross volume (e).}
 \label{fig:mr_arfi_weight}

### 2013-09-23 results.tex
updated figures for correlation analyses and removed image comparisons to path axes
updated with all of the latest data

@@
 \section{Results}\label{sect:results}
-
-\begin{itemize}
-    \item Images of select cases, demonstrating segmentation process
-\end{itemize}
 
@@
 
-The computed volumes of the prostates from the MR and ARFI imaging data are
-presented in Table~\ref{tab:mr_arfi_volumes}.  The subvolumes associated with
-the zonal anatomy in each imaging modality was measured
-(Figure~\ref{fig:mr_arfi_volumes}(a)), showing an mean overestimation of total
-prostate volume of 36.7 $\pm$ 27.9\% by ARFI imaging compared to MR volumes
-(Figure~\ref{fig:mr_arfi_volumes}(b)), and central gland volumes differed
-between ARFI and MR images by 2.1 $\pm$ 39.1\%
-(Figure~\ref{fig:mr_arfi_volumes}(c)).
-
-\input{tab_mr_arfi_volumes}
+The subvolumes associated with the zonal anatomy in each imaging modality were
+measured (Figure~\ref{fig:mr_arfi_volumes}(a)), with moderate correlation
+between the ARFI and MR total prostate gland volumes (R$^2$ = 0.68), with a
+mean overestimation of of 36.7 $\pm$ 27.9\% by ARFI imaging compared to MR
+volumes (Figure~\ref{fig:mr_arfi_volumes}(b)).  Central gland volumes were
+slightly less correlated between ARFI and MR images (R$^2$ = 0.41) with no
+significant mean over/under-estimation (2.1 $\pm$ 39.1\%), but significant
+variability between the cases (Figure~\ref{fig:mr_arfi_volumes}(c)).
+Table~\ref{tab:mr_arfi_volumes} has the individually-measured volumes for MR
+and ARFI imaging for each study subject.
 
@@
 apex-to-base), the prostate volume was approximated as a tri-axial ellipsoid,
-and its volume was estimated.
-
-\input{tab_path_data}
-
-MR and ARFI imaging total prostate volumes were also compared with the weights
-and prostate volumes, approximated to be tri-axial ellipsoids, and calculated
-from measurements of the prostate semi-principal axes.  Prostate weights were
-moderately correlated with estimated pathology ellipsoidal prostate volumes
-(Figure~\ref{fig:mr_arfi_weight}(a), R$^2$ = 0.68).  There was moderate
+and its volume was estimated (\ref{eqn:ellipsoid_volume}).  Prostate weights
+were moderately correlated with estimated pathology ellipsoidal prostate
+volumes (Figure~\ref{fig:mr_arfi_weight}(a), R$^2$ = 0.68).  There was moderate
 correlation between the prostate weight and the image-reconstructed prostate
@@
 (ARFI)), though there was weaker correlation with the ellipsoidal approximation
-of the measurement prostate volume and the image-recontructed volumens
-(Figure~\ref{fig:mr_arfi_weight}(c), R$^2$ = 0.08 (MR) and 0.01 (ARFI)).
+of the measurement prostate volume and the image-reconstructed volumes
+(Figure~\ref{fig:mr_arfi_weight}(c), R$^2$ = 0.08 (MR) and 0.01 (ARFI)).  Study
+subject \#4 had a very large prostate, which could not be completely captured
+in MR or ARFI imaging in the anterior region, so the same analyses were
+performed with that study subject excluded
+(Figure~\ref{fig:mr_arfi_weight}(d-e)), improving correlations for MR and ARFI
+to the prostate weights (R$^2$ = 0.74 and 0.60, respectively), and prostate
+volumes (R$^2$ = 0.39 for both imaging modalities).
 
@@
 
-Measurements of the prostate dimensions along the three standard anatomic axes
-(apex-to-base, lateral-to-lateral, and anterior-to-posterior) were made
-(Table~\ref{tab:mr_arfi_axes}), and the correlation between the imaging axis
-measurements and pathology was performed (Figure~\ref{fig:mr_arfi_path_axes}).
-To evaluate the relative shape of the prostate in each imaging modality, the
-relative ratios of the axes in imaging and pathology for the total prostate
-volume were compared (Figure~\ref{fig:mr_arfi_total_axes}), and the relative
-ratios of the central gland / zone anatomy were compared for MR and ARFI
-imaging (Figure~\ref{fig:mr_arfi_central_axes}).  The mean ratio differences
-between ARFI imaging, T2WI MR and pathology are compiled in
-Table~\ref{tab:axis_ratio_over_under}.
-
-\input{tab_mr_arfi_axes}
+Measurements of the prostate total and central gland dimensions along the three
+standard anatomic axes (apex-to-base, lateral-to-lateral, and
+anterior-to-posterior) were made (Table~\ref{tab:mr_arfi_axes}), and the
+correlation between the imaging axis measurements was analyzed
+(Figure~\ref{fig:mr_arfi_path_axes}).  ARFI was most correlated to MR in the
+lateral-to-lateral axis in both the total and central glands (R$^2$ = 0.47 and
+0.44, respectively), with mean overestimates of 8.1 $\pm$ 18.4\% and 0.06 $\pm$
+17.2\%, respectively (Table~\ref{tab:mr_arfi_axes_error}).  ARFI had very good
+correlation with the total prostate gland axis in the anterior-to-posterior
+dimension (R$^2$ = 0.58), but poor correlation in the central gland (R$^2$ =
+0.00), with overestimates of 17.0 $\pm$ 12.1 and 14.8 $\pm$ 23.1\%.  ARFI
+imaging had weak correlation with MR images in the apex-to-base dimension
+(R$^2$ = 0.15, total gland and R$^2$ = 0.20, central gland), with differences
+of 0.58 $\pm$ 12.9 and -10.8 $\pm$ 22.3\%, respectively.
 
@@
 
-\input{tab_mr_arfi_path_axes}
+\input{tab_mr_arfi_axes_error}
 
-\input{fig_mr_arfi_total_axes}
+%\input{fig_mr_arfi_total_axes}
 
-\input{fig_mr_arfi_central_axes}
+%\input{fig_mr_arfi_central_axes}
 
-\input{tab_axis_ratio_over_under}
+%\input{tab_axis_ratio_over_under}

### 2013-09-23 discussion.tex
near final draft version

--- 2013-09-13 discussion.tex
@@
 
-Discussion text will go here.
+This study was designed to compared the zonal anatomy of the prostate in
+patients with biopsy-proven PCa as seen in ARFI imaging and MR T2WI.  Overall,
+ARFI imaging performed well when compared with MR estimates of total and
+central gland volume (Figure~\ref{fig:mr_arfi_volumes}), though ARFI tended to
+overestimate prostate total gland volume, most commonly due to over-estimation
+of the anterior-to-posterior dimension of the prostate.  As seen in Figure XX,
+clear delineation of the anterior aspect of the prostate can be challenging in
+ARFI and B-mode images due to poor SNR, since this is the deepest aspect of the
+prostate relative to the rectal wall imaging surface.  MR T2WI can also have
+difficulties visualizing and diagnosing lesions in the anterior region of the
+prostate,~\cite{Gupta2013} and was evident in imaging study subject 4, with a
+$\sim$143 g prostate (Figure~\ref{fig:mr_arfi_weight}).  With study subject 4
+excluded, MR and ARFI imaging volumes and the ellipsoidal prostate volume
+estimates had similar correlation with prostate weights
+(Figure~\ref{fig:mr_arfi_weight}(d)).
+
+The most accurate measurements for both the total prostate gland and central
+gland were in the lateral-to-lateral dimension
+(Figure~\ref{fig:mr_arfi_path_axes}(a), Table~\ref{tab:mr_arfi_axes_error}),
+with ARFI imaging exhibiting a slight overestimate in both glands.  ARFI
+imaging had weaker correlations with MR imaging in the apex-to-base dimension
+(Figure~\ref{fig:mr_arfi_path_axes}(c)), though it should be noted that the MR
+images are limited in spatial resolution by slice thickness in this dimension
+($\sim$ 3 mm).  Finally, as expected from qualitative image evaluation, the
+anterior-to-posterior dimension had the weakest correlations, especially in
+the central gland, most likely related to the inability to clearly resolve an
+anterior border for both the prostate total and central glands.
+
+This study has several limitations that should be considered when interpreting
+these results.  Gross pathology weight and axis measurements both could be
+affected by the presence of periprostatic tissue that was excised during
+radical prostatectomy, especially in cases where more aggressive margins may
+have been necessary.  For this reason, unlike the image-to-image measurement
+comparisons, all of the image metrics presented relative to pathology metrics
+(Figure~\ref{fig:mr_arfi_weight}) were not characterized for absolute accuracy,
+but instead, relative correlations were evaluated.  Additionally, the volumes
+of the prostate from gross pathologic measurements were approximated as
+ellipsoids, which also introduced error, most likely an over-estimation of
+volume.  Interestingly, all pathology estimates were thought to have positive
+biases, but both imaging modalities tended to overestimate relative to the
+pathology measurements.
+
+It should also be noted that all of the prostates in this study contained
+varying amounts of PCa, BPH and atrophy, all of which can distort the zonal
+anatomy, especially in the case of BPH and central gland morphology.  While
+younger, healthier prostates could have been targeted, these healthy organs
+would not have been excised for pathology characterization, and the zonal
+anatomy of a health (young) prostate is expected to be different from the
+prostate of a middle-age man, who is the target demographic for PCa screening
+imaging and PCa characterization; therefore, it was felt that performing this
+analysis in the presence of these confounding factors was appropriate.

### 2013-09-23 discussion.tex
contains Kirema additions

@@
 difficulties visualizing and diagnosing lesions in the anterior region of the
-prostate,~\cite{Gupta2013} and was evident in imaging study subject 4, with a
-$\sim$143 g prostate (Figure~\ref{fig:mr_arfi_weight}).  With study subject 4
-excluded, MR and ARFI imaging volumes and the ellipsoidal prostate volume
-estimates had similar correlation with prostate weights
-(Figure~\ref{fig:mr_arfi_weight}(d)).
+prostate,~\cite{Gupta2013}, which is thought to be secondary to the
+heterogeneity of the gland in that area, especially in the presence of BPH.
+This was evident in imaging study subject 4, with a $\sim$143 g prostate
+(Figure~\ref{fig:mr_arfi_weight}), which was an outlier in terms of weight and
+volume.  With study subject 4 excluded, MR and ARFI imaging volumes and the
+ellipsoidal prostate volume estimates had similar correlation with prostate
+weights (Figure~\ref{fig:mr_arfi_weight}(d)).
 
@@
 (Figure~\ref{fig:mr_arfi_path_axes}(c)), though it should be noted that the MR
-images are limited in spatial resolution by slice thickness in this dimension
-($\sim$ 3 mm).  Finally, as expected from qualitative image evaluation, the
-anterior-to-posterior dimension had the weakest correlations, especially in
-the central gland, most likely related to the inability to clearly resolve an
-anterior border for both the prostate total and central glands.
+images are limited in spatial resolution by slice thickness and volume
+averaging in this dimension ($\sim$ 3 mm).  Finally, as expected from
+qualitative image evaluation, the anterior-to-posterior dimension had the
+weakest correlations, especially in the central gland, most likely related to
+the inability to clearly resolve an anterior border for both the prostate total
+and central glands. Of note, a slight overestimation of anatomic borders can
+result in a large relative difference from the true values (by histology) as the gland
+is small in size.
 
 This study has several limitations that should be considered when interpreting
-these results.  Gross pathology weight and axis measurements both could be
+these results.  Gross pathology weight and axis measurements could both be
 affected by the presence of periprostatic tissue that was excised during
@@
 analysis in the presence of these confounding factors was appropriate.
+
+Future directions for our research include deeper analysis of our findings,
+especially the lateral-to-lateral overestimation, in order to improve our
+anatomic delineation and assess further its clinical impact. Although work
+remains to be done, the fact that one can view the different prostatic zones on
+ARFI images in addition to MRI is encouraging and could lead to targeted
+diagnostic biopsies and therapies with real-time imaging for men with PCa.

### 2013-09-23 fig_arfi_segs.tex
new b-mode and arfi segs fig included

\begin{figure}[htb!]
\centering
\begin{tabular}{c}
\includegraphics[width=1.0\textwidth]{figs/Bmode_CapsuleSegs.eps} \\
\includegraphics[width=1.0\textwidth]{figs/ARFI_CGsegs.eps} \\
\end{tabular}
\caption{Representative B-mode (top row) and ARFI images (bottom row), with
    superimposed segmentation outlines of the capsule and CG, respectively, for
    the same study subject shown in Figure~\ref{fig:mr_segs_vol}.  The B-mode
    images were segmented in the axial view, using the hypoechoic border of the
    capsule.  Notice that for large prostates, like the one shown in this
    figure, that the edges of the organ were not always captured in the imaging
    field-of-view.  The CG in the ARFI images was typically segmented using the
    coronal view, which, especially in the presence of extensive BPH, can have
    a very characteristic, heterogeneous appearance.  It should be noted that
    the coronal views in these ultrasound images are flipped 180\degree relative
    to the same images in Figure~\ref{fig:mr_segs_vol}.}
\label{fig:arfi_segs} 
\end{figure}

### 2013-09-23 image_proc_methods.tex
updated with all of the latest data
new b-mode and arfi segs fig included

@@
 delineated based on its heterogeneous and lower signal intensity as well as its
-location. Although not readily visible on every case, the AFS was identified by
-its low T2 signal intensity and its location anterior to the central gland.
+location (Figure~\ref{fig:mr_anatomy}). Although not readily visible on every
+case, the AFS was identified by its low T2 signal intensity and its location
+anterior to the central gland.
+
+Figure~\ref{fig:mr_anatomy} shows the anatomic zones and some anatomic
+orientation labels in the prostate of a representative study subject MR T2WI.  The
+zonal anatomy of the prostate was manually segmented
+(Figure~\ref{fig:mr_segs_vol}, bottom left) across the image stack, and 3D
+models of the zonal anatomy volumes were rendered
+(Figure~\ref{fig:mr_segs_vol}, top).
+
+\input{fig_mr_anatomy}
+
+\input{fig_mr_segs_vol}
 
@@
 
-\emph{CO-AUTHOR QUESTION: CAN WE EASILY DELINEATE OUR SEGMENTATION CRITERIA?}
+\input{fig_arfi_segs}
 

### 2013-09-23 results.tex
new b-mode and arfi segs fig included

@@
 \section{Results}\label{sect:results}
-
-Figure~\ref{fig:mr_anatomy} shows the anatomic zones and some anatomic
-orientation labels in the prostate of a representative study subject.  The
-zonal anatomy of the prostate was manually segmented
-(Figure~\ref{fig:mr_segs_vol}, bottom left) across the image stack, and 3D
-models of the zonal anatomy volumes were rendered
-(Figure~\ref{fig:mr_segs_vol}, top).
-
-\input{fig_mr_anatomy}
-
-\input{fig_mr_segs_vol}
-
 The subvolumes associated with the zonal anatomy in each imaging modality were

### 2013-09-23 image_proc_methods.tex
cherry-picked corrected method ref and merged Kirema MS3 submission

@@
 independently according to their well-established anatomical characteristics on
-T2WI.~\cite{Verma2011,Jung2012,Poon1985,Hricak2007,Bonekamp2011} The PZ was
+T2WI.~\cite{Barentsz2012,Jung2012,Poon1985,Hricak2007,Bonekamp2011} The PZ was
 identified by its homogenous high signal intensity on T2WI, which is usually

### 2013-09-23 clinical_significance.tex
fixed background reference

@@
 PCa located outside the routine sampling sites can be missed and the extent
-of the cancer might be underestimated.~\cite{Gupta2013,Cornud1012} For example,
+of the cancer might be underestimated.~\cite{Gupta2013,Cornud2012} For example,
 in a study by Mufarrij \etal, 45.9--47.2\% of patients who were candidates for

### 2013-09-23 acknowledgements.tex

--- 2013-09-13 acknowledgements.tex
@@
-\section*{ACKNOWLEDGEMENTS}
-The authors would like to thank Siemens Medical Solution USA, Ultrasound
-Division for their in-kind technical support and Ned Danieley for system administration support.
-This work was supported by NIH R01CA142824 and the Duke Coulter Translational Grant Program.
+\section*{Acknowledgements} 
+The authors would like to thank Dr. Ned Rouze for his design of the ultrasound
+transducer rotation apparatus and experimental data acquisition, Siemens
+Medical Solution USA, Ultrasound Division for their in-kind technical support
+and Ned Danieley for computer system administration support.  This work was
+supported by NIH R01CA142824 and the Duke Coulter Translational Grant Program.

### 2013-10-07 abstract.tex
added period

@@
 volumes.  Both imaging modalities showed moderate correlations (0.39 $<$ R$^2 <
-$ 0.74) between estimated organ volume and gross pathologic weights ARFI and MR
+$ 0.74) between estimated organ volume and gross pathologic weights.  ARFI and MR
 total prostate gland volumes were well-correlated (R$^2$ = 0.68), but ARFI

### 2013-10-07 mr_background.tex
re-organized radical prostatectomy text

@@
 significantly improve the performance of MRI in cancer
-diagnosis.~\cite{Turkbey2012}  The European Society of Urogenital Radiology’s
+diagnosis.~\cite{Turkbey2012}  The European Society of Urogenital Radiology
 (ESUR) prostate MR guidelines recommend at least 2 functional imaging

### 2013-10-07 path_methods.tex
updated with all of the latest data
re-organized radical prostatectomy text

@@
 \subsection{Pathology Processing}
-After excision, the prostates were weighed and tri-axially measured, formalin
-fixed for at least 24 hours without being cut, and then processed for whole
-mount histology.
-While tri-axial measurements of the gross prostate were not necessarily anatomically aligned, rough estimates of the prostate volume were made using a tri-axial, ellipsoidal organ volume assumption, where the volume could be estimated as:
+Prostates were radically removed using a da Vinci Surgical System (Intuitive
+Surgical\textregistered, Inc., Sunnyvale, CA).  After excision, the prostates
+were weighed and tri-axially measured, formalin fixed for at least 24 hours
+without being cut, and then processed for whole mount histology.  While
+tri-axial measurements of the gross prostate were not necessarily anatomically
+aligned, rough estimates of the prostate volume were made using a tri-axial,
+ellipsoidal organ volume assumption, where the volume could be estimated as:
 
-\begin{equation}
+\begin{equation}\label{eqn:ellipsoid_volume}
 V = \frac{4}{3}\pi \sqrt{\det{A^{-1}}},

### 2013-10-07 conclusions.tex
added period

@@
 Both imaging modalities showed moderate correlations (0.39 $<$ R$^2 < $ 0.74)
-between estimated organ volume and gross pathologic weights ARFI and MR total
+between estimated organ volume and gross pathologic weights.  ARFI and MR total
 prostate gland volumes were well-correlated (R$^2$ = 0.68), but ARFI images

### 2013-10-07 abstract.tex
updated text to reflect new volume data (Fig 4)

@@
 $ 0.74) between estimated organ volume and gross pathologic weights.  ARFI and MR
-total prostate gland volumes were well-correlated (R$^2$ = 0.68), but ARFI
-images yielded prostate volumes that were, on average, larger (36\% $\pm$ 28\%)
+total prostate gland volumes were well-correlated (R$^2$ = 0.63), but ARFI
+images yielded prostate volumes that were, on average, larger (6.1\% $\pm$ 25\%)
 than MR images, primarily due to over-estimation of the anterior-to-posterior
@@
 and 0.58 $\pm$ 12.9\%).  The central zone volumes of ARFI and MR images were
-also moderately correlated (R$^2$ = 0.41), with minimal volume bias between the
-imaging modalities, but significant variability case-to-case (2.1 $\pm$
-39.1\%).  Central zone volume differences were, again, strongly attributed to
+also moderately correlated (R$^2$ = 0.38), with minimal volume bias between the
+imaging modalities, but significant variability case-to-case (-5.0 $\pm$
+39.5\%).  Central zone volume differences were, again, strongly attributed to
 over-estimation of the anterior-to-posterior axis (14.8 $\pm$ 23.1\%), with a

### 2013-10-07 conclusions.tex
updated text to reflect new volume data (Fig 4)

@@
 between estimated organ volume and gross pathologic weights.  ARFI and MR total
-prostate gland volumes were well-correlated (R$^2$ = 0.68), but ARFI images
-yielded prostate volumes that were, on average, larger (36\% $\pm$ 28\%) than
+prostate gland volumes were well-correlated (R$^2$ = 0.63), but ARFI images
+yielded prostate volumes that were, on average, larger (6.1\% $\pm$ 25\%) than
 MR images, primarily due to over-estimation of the anterior-to-posterior

### 2013-10-07 fig_mr_arfi_volumes.tex
updated text to reflect new volume data (Fig 4)

@@
 \caption{Comparison of MR and ARFI zonal anatomy volume estimates from
-    manually-segmented images.  Total prostate gland volumes ranged from 19.6--71.0
-    cm$^3$ based on MR image models (a), with a moderate correlation between MR
-    and ARFI imaging for both the total gland volume (R$^2$ = 0.68, (b)) and the
-    central gland (R$^2$ = 0.41, (c)).  The ARFI total gland volumes had a positive bias
-    compared with the MR volumes (36.7 $\pm$ 27.9\% (b)), while the ARFI
-    central zone volumes were on average more similar (2.1 $\pm$ 39.0\%, (c)), but
-    with some significant variability between study subjects.
-    Table~\ref{tab:mr_arfi_volumes} contains the individual volume estimates
-    for the total prostate and central glands.}
+    manually-segmented images.  Total prostate gland volumes ranged from
+    19.6--71.0 cm$^3$ based on MR image models (a), with a moderate correlation
+    between MR and ARFI imaging for both the total gland volume (R$^2$ = 0.63,
+    (b)) and the central gland (R$^2$ = 0.38, (c)).  The ARFI total gland
+    volumes had a positive bias compared with the MR volumes (6.1 $\pm$ 25.0\%
+    (b)), while the ARFI central zone volumes were on average more similar
+    (-5.0 $\pm$ 39.5\%, (c)), but with some significant variability between
+    study subjects.  Table~\ref{tab:mr_arfi_volumes} contains the individual
+    volume estimates for the total prostate and central glands.}
 \label{fig:mr_arfi_volumes} 

### 2013-10-07 results.tex
updated text to reflect new volume data (Fig 4)

@@
 measured (Figure~\ref{fig:mr_arfi_volumes}(a)), with moderate correlation
-between the ARFI and MR total prostate gland volumes (R$^2$ = 0.68), with a
-mean overestimation of of 36.7 $\pm$ 27.9\% by ARFI imaging compared to MR
+between the ARFI and MR total prostate gland volumes (R$^2$ = 0.63), with a
+mean overestimation of of 6.1 $\pm$ 25.0\% by ARFI imaging compared to MR
 volumes (Figure~\ref{fig:mr_arfi_volumes}(b)).  Central gland volumes were
-slightly less correlated between ARFI and MR images (R$^2$ = 0.41) with no
-significant mean over/under-estimation (2.1 $\pm$ 39.1\%), but significant
+slightly less correlated between ARFI and MR images (R$^2$ = 0.38) with no
+significant mean over/under-estimation (-5.0 $\pm$ 39.5\%), but significant
 variability between the cases (Figure~\ref{fig:mr_arfi_volumes}(c)).

### 2013-10-07 abstract.tex
updated for numbers and now under 300 words

--- 2013-10-07 abstract.tex
@@
 Prostate cancer (PCa) is the most common non-cutaneous malignancy among men in
-the United States and the second leading cause of cancer related death.
-Non-invasive imaging could lead to improved diagnosis, risk-stratification, and
-PCa management.  Magnetic resonance imaging (MRI) has been available for use in
-the workup of patients with PCa since the early 1980s, and recent advances with
-functional parameters has greatly improved its clinical diagnostic utility.
-Acoustic Radiation Force Impulse (ARFI) imaging is an ultrasound-based modality
-that evaluates the mechanical properties of soft tissues. ARFI imaging has the
-potential to aid in PCa diagnosis and management by evaluating the structural
-composition of prostate zones and tumors based on their stiffness.  In this
-study, MR and ARFI imaging datasets were compared to one another and with gross
-pathology measurements made immediately post radical prostatectomy.  Imaging
-datasets were manually segmented to delineate the central gland and prostate
-capsule, and 3D models were rendered to evaluate zonal anatomy dimensions and
-volumes.  Both imaging modalities showed moderate correlations (0.39 $<$ R$^2 <
-$ 0.74) between estimated organ volume and gross pathologic weights.  ARFI and MR
-total prostate gland volumes were well-correlated (R$^2$ = 0.68), but ARFI
-images yielded prostate volumes that were, on average, larger (36\% $\pm$ 28\%)
-than MR images, primarily due to over-estimation of the anterior-to-posterior
-dimension of the prostate total gland (17.0 $\pm$ 12.1\%), while over-estimates
-of the other dimensions were less significant contributors (8.1 $\pm$ 18.4\%
-and 0.58 $\pm$ 12.9\%).  The central zone volumes of ARFI and MR images were
-also moderately correlated (R$^2$ = 0.41), with minimal volume bias between the
-imaging modalities, but significant variability case-to-case (2.1 $\pm$
-39.1\%).  Central zone volume differences were, again, strongly attributed to
-over-estimation of the anterior-to-posterior axis (14.8 $\pm$ 23.1\%), with a
-significant underestimation of the apex-to-base dimension (-10.8 $\pm$ 22.3\%)
-and no mean bias in the lateral-to-lateral measurements (0.006 $\pm$ 17.2\%).
-Strong variability in central gland volumes is believed to be related to the
-extent of benign prostatic hyperplasia (BPH) for select cases.  Overall, ARFI
-imaging of the prostate yielded prostate volumes and dimensions that were
-correlated with MR T2WI estimates, with biases in the anterior-to-posterior
-dimension, most likely related to poor displacement SNR in the anterior region
-of the prostate from greater distance from the rectal wall imaging surface.
-ARFI imaging is a promising low-cost, real-time imaging modality that can
-compliment MR imaging for diagnosis, treatment planning and management of PCa.
+the United States and the second leading cause of cancer-related death.
+Multi-parametric Magnetic Resonance Imaging (mpMRI) has gained recent
+popularity to characterize PCa.  ARFI imaging has the potential to aid in PCa
+diagnosis and management by evaluating the structural composition of prostate
+zones and tumors based on their stiffness.  In this study, MR and ARFI imaging
+datasets were compared to one another and with gross pathology measurements
+made immediately post radical prostatectomy.  Images were manually segmented to
+delineate the central gland (CG) and prostate capsule, and 3D models were
+rendered to evaluate zonal anatomy dimensions and volumes.  Both imaging
+modalities showed moderate correlations between estimated organ volume and
+gross pathologic weights.  ARFI and MR total prostate gland volumes were
+well-correlated (R$^2$ = 0.63), but ARFI images yielded prostate volumes that
+were, on average, larger (6.1\% $\pm$ 25\%) than MR images, primarily due to
+over-estimation of the lateral dimension of the total gland (13.5 $\pm$
+11.0\%), while over-estimates of the other dimensions were less significant
+(5.7 $\pm$ 20.6\% and -2.4 $\pm$ 17.0\%).  The ARFI and MR CG volumes were also
+moderately correlated (R$^2$ = 0.38).  CG volume differences were attributed to
+under-estimation of the anterior-to-posterior axis (-8.4 $\pm$ 24.0\%), and
+over-estimation bias in the lateral dimension (11.5 $\pm$ 22.5\%).  ARFI
+prostate imaging yielded prostate volumes and dimensions that were correlated
+with MR T2WI estimates, with biases in the AP dimension, most likely related to
+poor displacement SNR in the anterior region of the prostate from greater
+distance from the rectal wall imaging surface.  Lateral dimension contrast was
+also challenging due to peri-prostatic fat.  ARFI imaging is a promising
+low-cost, real-time imaging modality that can compliment MR imaging for
+diagnosis, treatment planning and management of PCa.

### 2013-10-07 fig_mr_arfi_weight.tex
no longer removing outliers

--- 2013-09-15 fig_mr_arfi_weight.tex
@@
 \end{tabular}
+%\begin{tabular}{ll}
+%\includegraphics[width=0.3\linewidth]{figs/corr_weight_vol_no4} &
+%\includegraphics[width=0.3\linewidth]{figs/corr_pathVol_vol_no4} \\
+%(d) Image Volume : Prostate Weight (-4) & (e) Image Volume : Path Volume (-4) \\
+%\end{tabular}
 \caption{Tri-axial pathology measurements were used to make an ellipsoidal
@@
     ARFI image volumens and approaximated ellipsoidal prostate pathology
-    volumes (R$^2$ = 0.08 and 0.01, respectively) (c).}
+    volumes (R$^2$ = 0.08 and 0.01, respectively) (c).  It should be noted in
+    these figures that two study subjects had excessively large prostates that
+    were difficult to fully capture in imaging (volumes $>$ 60 cm$^3$),
+    especially in their anterior region, and they were, therefore, grossly
+    underestimated in size.}
 \label{fig:mr_arfi_weight}

### 2013-10-07 coverletter.tex
coverletter

\documentclass[10pt]{article}
\usepackage{graphicx}
\pagestyle{empty}
\topmargin=-0.75in
\oddsidemargin=0.0in
\textwidth=6.5in
\textheight=9.5in
\parindent=0.0in
\parskip=10pt

\begin{document}
\begin{minipage}[t]{0.5\linewidth}
\vspace{0pt}
\includegraphics[width=3.0in,trim=55 50 50 85,clip=true]{/home/mlp6/bme/Duke_BME_Logos/Pratt-HorizontalLogo-Vector-Black}
\end{minipage}\hfill
\begin{minipage}[t]{0.5\linewidth}
\vspace{0pt}
\hfill{\bf Biomedical Engineering}\\

\begin{small}
\hfill 136 Hudson Hall, Box 90281, Durham, NC 27708-0281

\hfill Tel: (919) 660 5158 $\bullet$ Fax: (919) 684 4488

\hfill mark.palmeri@duke.edu\\
\end{small}
\end{minipage}
\begin{Large}

\vspace{0.25in}
\today
\vspace{0.25in}

Re: Ultrasonic Imaging Manuscript Submission
\vspace{0.25in}

To whom it may concern:
\vspace{0.1in}

We thank you for taking the time and effort to review our manuscript, titled
{\it ``Acoustic Radiation Force Impulse (ARFI) Imaging Prostate Zonal Anatomy:
    Comparison with Endorectal T2-Weighted MR Imaging (T2WI)''}, for
consideration to be published in Ultrasonic Imaging.  All of the authors on
this manuscript have reviewed its content and approve of its submission.  This
work has not been published or submitted for review to any other journals.

Please do not hesitate to contact me if you have any questions regarding this
submission.

\vspace{0.25in}

%\hspace*{0.575\textwidth} Sincerely,
Sincerely,\\
\includegraphics{/home/mlp6/MarkLPalmeriSig}\\
%\vspace*{0.5in}
Mark L. Palmeri, M.D., Ph.D.\\
\end{Large}
Assistant Research Professor\\
Biomedical Engineering \& Anesthesiology
%\hspace*{0.575\textwidth} Mark L. Palmeri, M.D., Ph.D.\\
%\hspace*{0.575\textwidth} Assistant Research Professor\\
%\hspace*{0.575\textwidth} Biomedical Engineering \& Anesthesiology
\end{document}

### 2013-10-07 conclusions.tex
updated

@@
 the established methods for identifying zonal anatomy using MR T2W images.
-Both imaging modalities showed moderate correlations (0.39 $<$ R$^2 < $ 0.74)
-between estimated organ volume and gross pathologic weights.  ARFI and MR total
-prostate gland volumes were well-correlated (R$^2$ = 0.63), but ARFI images
-yielded prostate volumes that were, on average, larger (6.1\% $\pm$ 25\%) than
-MR images, primarily due to over-estimation of the anterior-to-posterior
-dimension of the prostate total gland (17.0 $\pm$ 12.1\%), while over-estimates
-of the other dimensions were less significant contributors (8.1 $\pm$ 18.4\%
-and 0.58 $\pm$ 12.9\%).  The central zone volumes of ARFI and MR images were
-also moderately correlated (R$^2$ = 0.38), with minimal volume bias between the
-imaging modalities, but significant variability case-to-case (-5.0 $\pm$
-39.5\%).  Central zone volume differences were, again, strongly attributed to
-over-estimation of the anterior-to-posterior axis (14.8 $\pm$ 23.1\%), with a
-significant underestimation of the apex-to-base dimension (-10.8 $\pm$ 22.3\%)
-and no mean bias in the lateral-to-lateral measurements (0.006 $\pm$ 17.2\%).
-Strong variability in central gland volumes is believed to be related to the
-extent of benign prostatic hyperplasia (BPH) for select cases.  Overall, ARFI
-imaging of the prostate yielded prostate volumes and dimensions that were
-correlated with MR T2WI estimates, with biases in the anterior-to-posterior
-dimension, most likely related to poor displacement SNR in the anterior region
-of the prostate from greater distance from the rectal wall imaging surface.
-ARFI imaging is a promising low-cost, real-time imaging modality that can
-compliment MR imaging for diagnosis, treatment planning and management of PCa.
+Both imaging modalities showed moderate correlations between estimated organ
+volume and gross pathologic weights, and ARFI:MR total prostate gland volumes
+were well-correlated (R$^2$ = 0.63), but ARFI images yielded prostate volumes
+that were, on average, larger (6.1\% $\pm$ 25\%) than MR images, primarily due
+to over-estimation of the lateral dimension of the prostate total gland (13.5
+$\pm$ 11.0\%), while over-estimates of the other dimensions were less
+significant contributors (5.7 $\pm$ 20.6\% and -2.4 $\pm$ 17.0\%).  The central
+zone volumes of ARFI and MR images were also moderately correlated (R$^2$ =
+0.38), with minimal volume bias between the imaging modalities, but significant
+variability case-to-case (-5.0 $\pm$ 39.5\%).  Central zone volume differences
+were, again, strongly attributed to over-estimation of the lateral dimension
+(11.5 $\pm$ 22.5\%), with a significant underestimation of the
+anterior-to-posterior dimension (-8.4 $\pm$ 24.0\%).  Strong variability in
+central gland volumes is believed to be related to the extent of benign
+prostatic hyperplasia (BPH) for select cases.  Overall, ARFI imaging of the
+prostate yielded prostate volumes and dimensions that were correlated with MR
+T2WI estimates, with biases in the anterior-to-posterior dimension, most likely
+related to poor displacement SNR in the anterior region of the prostate from
+greater distance from the rectal wall imaging surface, and the lateral
+dimension, where contrast between the PZ and peri-prostatic fat could be
+limited.  ARFI imaging is a promising low-cost, real-time imaging modality that
+can compliment MR imaging for diagnosis, treatment planning and management of
+PCa.

### 2013-10-07 discussion.tex
updated

@@
 \section{Discussion}\label{sect:discussion}
-
 This study was designed to compared the zonal anatomy of the prostate in
@@
 overestimate prostate total gland volume, most commonly due to over-estimation
-of the anterior-to-posterior dimension of the prostate.  As seen in Figure XX,
-clear delineation of the anterior aspect of the prostate can be challenging in
-ARFI and B-mode images due to poor SNR, since this is the deepest aspect of the
-prostate relative to the rectal wall imaging surface. MR T2WI can also have
-difficulties visualizing and diagnosing lesions in the anterior region of the
-prostate,~\cite{Gupta2013}, which is thought to be secondary to the
-heterogeneity of the gland in that area, especially in the presence of BPH.
-This was evident in imaging study subject 4, with a $\sim$143 g prostate
-(Figure~\ref{fig:mr_arfi_weight}), which was an outlier in terms of weight and
-volume.  With study subject 4 excluded, MR and ARFI imaging volumes and the
-ellipsoidal prostate volume estimates had similar correlation with prostate
-weights (Figure~\ref{fig:mr_arfi_weight}(d)).
-
-The most accurate measurements for both the total prostate gland and central
-gland were in the lateral-to-lateral dimension
-(Figure~\ref{fig:mr_arfi_path_axes}(a), Table~\ref{tab:mr_arfi_axes_error}),
-with ARFI imaging exhibiting a slight overestimate in both glands.  ARFI
-imaging had weaker correlations with MR imaging in the apex-to-base dimension
-(Figure~\ref{fig:mr_arfi_path_axes}(c)), though it should be noted that the MR
-images are limited in spatial resolution by slice thickness and volume
-averaging in this dimension ($\sim$ 3 mm).  Finally, as expected from
-qualitative image evaluation, the anterior-to-posterior dimension had the
-weakest correlations, especially in the central gland, most likely related to
-the inability to clearly resolve an anterior border for both the prostate total
-and central glands. Of note, a slight overestimation of anatomic borders can
-result in a large relative difference from the true values (by histology) as the gland
-is small in size.
+of the lateral dimension of the prostate.  This can be due to poor contrast
+between the PZ and peri-prostate fat (Figure~\ref{fig:arfi_segs}).  As seen in
+Figure~\ref{fig:arfi_segs}, clear delineation of the anterior aspect of the
+prostate can also be challenging in ARFI and B-mode images due to poor SNR,
+since this is the deepest aspect of the prostate relative to the rectal wall
+imaging surface. MR T2WI can also have difficulties visualizing and diagnosing
+lesions in the anterior region of the prostate,~\cite{Gupta2013}, which is
+thought to be secondary to the heterogeneity of the gland in that area,
+especially in the presence of BPH.  This was evident in imaging study subject
+4, with a $\sim$143 g prostate (Figure~\ref{fig:mr_arfi_weight}), which was an
+outlier in terms of weight and volume.  The most accurate measurements for both
+the total prostate gland and central gland were in the apex-to-base dimension
+(Figure~\ref{fig:mr_arfi_path_axes}(c), Table~\ref{tab:mr_arfi_axes_error}),
+with no mean bias in ARFI imaging estimates, but significant case-to-case
+variability, leading to poor linear regression coefficients of determination.
 
@@
 
+It should also be noted that this study did not evaluate user biases in image
+segmentation.  While MR zonal anatomy delineation has some establishment in the
+clinical literature, this study is the first attempt to define the criteria for
+ARFI imaging zonal anatomy characteristics, and it is expected that such
+delineations will continue to be refined as we acquire more cases and continue
+to compare with MRI and pathology data.  Given this limitation, no attempts
+were made to further quantify reader-to-reader variability in this work.
+
 Future directions for our research include deeper analysis of our findings,
-especially the lateral-to-lateral overestimation, in order to improve our
-anatomic delineation and assess further its clinical impact. Although work
+especially the lateral dimension over-estimation, to improve our anatomic
+delineation and assess further its clinical impact.  Technological improvements
+to improve anterior prostate boundary delineation in both B-mode and ARFI
+imaging are also being pursued and will hopefully allow for the currently
+elusive anterior PCa lesions in MR to be seen in ARFI imaging.  Although work
 remains to be done, the fact that one can view the different prostatic zones on

### 2013-10-07 results.tex
updated

@@
 correlation between the prostate weight and the image-reconstructed prostate
-volumes (Figure~\ref{fig:mr_arfi_weight}(b), R$^2$ = 0.44 (MR) and 0.21
+volumes (Figure~\ref{fig:mr_arfi_weight}(b), R$^2$ = 0.44 (MR) and 0.18
 (ARFI)), though there was weaker correlation with the ellipsoidal approximation
 of the measurement prostate volume and the image-reconstructed volumes
-(Figure~\ref{fig:mr_arfi_weight}(c), R$^2$ = 0.08 (MR) and 0.01 (ARFI)).  Study
-subject \#4 had a very large prostate, which could not be completely captured
-in MR or ARFI imaging in the anterior region, so the same analyses were
-performed with that study subject excluded
-(Figure~\ref{fig:mr_arfi_weight}(d-e)), improving correlations for MR and ARFI
-to the prostate weights (R$^2$ = 0.74 and 0.60, respectively), and prostate
-volumes (R$^2$ = 0.39 for both imaging modalities).
+(Figure~\ref{fig:mr_arfi_weight}(c), R$^2$ = 0.08 (MR) and 0.00 (ARFI)).
 
@@
 
-Measurements of the prostate total and central gland dimensions along the three
+Measurements of the prostate total and CG dimensions along the three
 standard anatomic axes (apex-to-base, lateral-to-lateral, and
@@
 (Figure~\ref{fig:mr_arfi_path_axes}).  ARFI was most correlated to MR in the
-lateral-to-lateral axis in both the total and central glands (R$^2$ = 0.47 and
-0.44, respectively), with mean overestimates of 8.1 $\pm$ 18.4\% and 0.06 $\pm$
-17.2\%, respectively (Table~\ref{tab:mr_arfi_axes_error}).  ARFI had very good
+lateral-to-lateral axis in both the total and CG (R$^2$ = 0.57 and
+0.32, respectively), with mean overestimates of 13.5 $\pm$ 11.0\% and 11.5 $\pm$
+22.5\%, respectively (Table~\ref{tab:mr_arfi_axes_error}).  ARFI had moderate
 correlation with the total prostate gland axis in the anterior-to-posterior
-dimension (R$^2$ = 0.58), but poor correlation in the central gland (R$^2$ =
-0.00), with overestimates of 17.0 $\pm$ 12.1 and 14.8 $\pm$ 23.1\%.  ARFI
+dimension (R$^2$ = 0.23), but poor correlation in the CG (R$^2$ =
+0.00), with overestimates of 5.7 $\pm$ 20.6 and -8.4 $\pm$ 24.0\%.  ARFI
 imaging had weak correlation with MR images in the apex-to-base dimension
-(R$^2$ = 0.15, total gland and R$^2$ = 0.20, central gland), with differences
-of 0.58 $\pm$ 12.9 and -10.8 $\pm$ 22.3\%, respectively.
+(R$^2$ = 0.19, total gland and R$^2$ = 0.09, CG), with differences
+of -2.4 $\pm$ 17.0 and 1.2 $\pm$ 19.8\%, respectively.
 

### 2013-10-07 abstract.tex
corrected Zach P75 dimension

@@
 11.0\%), while over-estimates of the other dimensions were less significant
-(5.7 $\pm$ 20.6\% and -2.4 $\pm$ 17.0\%).  The ARFI and MR CG volumes were also
+(5.7 $\pm$ 20.6\% and 2.4 $\pm$ 17.0\%).  The ARFI and MR CG volumes were also
 moderately correlated (R$^2$ = 0.38).  CG volume differences were attributed to
-under-estimation of the anterior-to-posterior axis (-8.4 $\pm$ 24.0\%), and
+under-estimation of the anterior-to-posterior axis (-6.1 $\pm$ 21.1\%), and
 over-estimation bias in the lateral dimension (11.5 $\pm$ 22.5\%).  ARFI

### 2013-10-07 conclusions.tex
corrected Zach P75 dimension

@@
 $\pm$ 11.0\%), while over-estimates of the other dimensions were less
-significant contributors (5.7 $\pm$ 20.6\% and -2.4 $\pm$ 17.0\%).  The central
+significant contributors (5.7 $\pm$ 20.6\% and 2.4 $\pm$ 17.0\%).  The central
 zone volumes of ARFI and MR images were also moderately correlated (R$^2$ =
@@
 (11.5 $\pm$ 22.5\%), with a significant underestimation of the
-anterior-to-posterior dimension (-8.4 $\pm$ 24.0\%).  Strong variability in
+anterior-to-posterior dimension (-6.1 $\pm$ 21.1\%).  Strong variability in
 central gland volumes is believed to be related to the extent of benign

### 2013-10-07 results.tex
corrected Zach P75 dimension

@@
 
-Measurements of the prostate total and CG dimensions along the three
-standard anatomic axes (apex-to-base, lateral-to-lateral, and
-anterior-to-posterior) were made (Table~\ref{tab:mr_arfi_axes}), and the
-correlation between the imaging axis measurements was analyzed
-(Figure~\ref{fig:mr_arfi_path_axes}).  ARFI was most correlated to MR in the
-lateral-to-lateral axis in both the total and CG (R$^2$ = 0.57 and
-0.32, respectively), with mean overestimates of 13.5 $\pm$ 11.0\% and 11.5 $\pm$
-22.5\%, respectively (Table~\ref{tab:mr_arfi_axes_error}).  ARFI had moderate
-correlation with the total prostate gland axis in the anterior-to-posterior
-dimension (R$^2$ = 0.23), but poor correlation in the CG (R$^2$ =
-0.00), with overestimates of 5.7 $\pm$ 20.6 and -8.4 $\pm$ 24.0\%.  ARFI
-imaging had weak correlation with MR images in the apex-to-base dimension
-(R$^2$ = 0.19, total gland and R$^2$ = 0.09, CG), with differences
-of -2.4 $\pm$ 17.0 and 1.2 $\pm$ 19.8\%, respectively.
+Measurements of the prostate total and CG dimensions along the three standard
+anatomic axes (apex-to-base, lateral-to-lateral, and anterior-to-posterior)
+were made (Table~\ref{tab:mr_arfi_axes}), and the correlation between the
+imaging axis measurements was analyzed (Figure~\ref{fig:mr_arfi_path_axes}).
+ARFI was most correlated to MR in the lateral-to-lateral axis in both the total
+and CG (R$^2$ = 0.57 and 0.32, respectively), with mean overestimates of 13.5
+$\pm$ 11.0\% and 11.5 $\pm$ 22.5\%, respectively
+(Table~\ref{tab:mr_arfi_axes_error}).  ARFI had moderate correlation with the
+total prostate gland and CG axes in the anterior-to-posterior dimension (R$^2$
+= 0.23 and R$^2$ = 0.34), respectively, with overestimates of 5.7 $\pm$ 20.6
+and -6.1 $\pm$ 21.1\%.  ARFI imaging had weak correlation with MR images in the
+apex-to-base dimension (R$^2$ = 0.19, total gland and R$^2$ = 0.09, CG), with
+differences of 2.4 $\pm$ 17.0 and 1.2 $\pm$ 19.8\%, respectively.
 

### 2013-10-14 image_proc_methods.tex
fixed table reference and added Kiremas new addendum
add text about including urethra in CG segs

--- 2013-09-23 image_proc_methods.tex
@@
 bladder, and subsequent images were segmented until the last slice with visible
-prostatic tissue was identified caudally. The CG, PZ and AFS were segmented
-independently according to their well-established anatomical characteristics on
-T2WI.~\cite{Verma2011,Jung2012,Poon1985,Hricak2007,Bonekamp2011} The PZ was
+prostatic tissue was identified caudally. The CG and urethra, PZ and AFS were
+segmented independently according to their well-established anatomical
+characteristics on
+T2WI.~\cite{Barentsz2012,Jung2012,Poon1985,Hricak2007,Bonekamp2011} The PZ was
 identified by its homogenous high signal intensity on T2WI, which is usually
@@
 delineated based on its heterogeneous and lower signal intensity as well as its
-location (Figure~\ref{fig:mr_anatomy}). Although not readily visible on every
-case, the AFS was identified by its low T2 signal intensity and its location
-anterior to the central gland.
+location (Figure~\ref{fig:mr_anatomy}). The urethra was included in the CG
+segmentation. Although not readily visible on every case, the AFS was
+identified by its low T2 signal intensity and its location anterior to the
+central gland.
 
@@
 \end{tabular}
+\label{tab:3dslicer}
 \end{table}

### 2013-12-08 figs/corr_weight_vol.m
change Prostate*.csv -> data.csv

@@ -8,7 +8,7 @@ function corr_weight_vol()
     % mlp6@duke.edu
     % 2013-09-13
 
-    vols = csvread('../data/Prostate_CZ_PZ_Volume_Axis_Measurements.csv');
+    vols = csvread('../data/data.csv');
 
     Pnum = num2cell(vols(:,1));
     mr_total_vol = vols(:,3);

### 2013-12-08 figs/corr_weight_vol_no4.m

@@ -8,7 +8,7 @@ function corr_weight_vol_no4()
     % mlp6@duke.edu
     % 2013-09-17
 
-    vols = csvread('../data/Prostate_CZ_PZ_Volume_Axis_Measurements.csv');
+    vols = csvread('../data/data.csv');
 
     mr_total_vol = vols([1:3,5:end],3);
     arfi_total_vol = vols([1:3,5:end],11);

### 2013-12-08 figs/mr_arfi_path_axes.m

@@ -7,7 +7,7 @@ function mr_arfi_path_axes
     % mlp6@duke.edu
     % 2013-09-14
 
-    data = csvread('../data/Prostate_CZ_PZ_Volume_Axis_Measurements.csv');
+    data = csvread('../data/data.csv');
 
     Pnum = num2cell(data(:,1));
 

### 2013-12-08 figs/mr_arfi_volumes.m

@@ -8,7 +8,7 @@
 % mlp6@duke.edu
 % 2013-09-13
 
-vols = csvread('../data/Prostate_CZ_PZ_Volume_Axis_Measurements.csv');
+vols = csvread('../data/data.csv');
 
 Pnum = num2cell(vols(:,1));
 mr_cg_vol = vols(:,2);

### 2013-12-08 figs/mr_arfi_path_axes.m
generate axis errors in the Matlab script now

@@ -31,6 +31,10 @@ function mr_arfi_path_axes
     % fontsize
     fs = 18;
 
+    % open up the table of axis errors for LaTeX
+    latex_fid = fopen('../tab_mr_arfi_axes_error.tex','w');
+    print_tab_mr_arfi_axes_error_tex_header(latex_fid);
+
     % lets plot some direct axis measurements of the total prostate and central gland
     for i=1:3,
         figure;
@@ -83,10 +87,10 @@ function mr_arfi_path_axes
         arfi_axis_OverUnder(:,i) = compute_over_under(PATH_total_axes(:,i),ARFI_total_axes(:,i)/10);
     end;
 
-    mean(mr_axis_OverUnder)
-    std(mr_axis_OverUnder)
-    mean(arfi_axis_OverUnder)
-    std(arfi_axis_OverUnder)
+    %mean(mr_axis_OverUnder)
+    %std(mr_axis_OverUnder)
+    %mean(arfi_axis_OverUnder)
+    %std(arfi_axis_OverUnder)
 
     % let's make plots of the correlation between MR:ARFI for total gland and central gland
     for i=1:3,
@@ -140,11 +144,12 @@ function mr_arfi_path_axes
         mr_arfi_central_OverUnder(:,i) = compute_over_under(MR_central_axes(:,i)/10,ARFI_central_axes(:,i)/10);
     end;
 
-    disp('MR:ARFI Total and Central');
-    mean(mr_arfi_total_OverUnder)
-    std(mr_arfi_total_OverUnder)
-    mean(mr_arfi_central_OverUnder)
-    std(mr_arfi_central_OverUnder)
+    %disp('MR:ARFI Total and Central');
+    %mean(mr_arfi_total_OverUnder)
+    %std(mr_arfi_total_OverUnder)
+    %mean(mr_arfi_central_OverUnder)
+    %std(mr_arfi_central_OverUnder)
+    print_tab_mr_arfi_axes_error_tex_data(latex_fid,mr_arfi_total_OverUnder,mr_arfi_central_OverUnder);
 
     % I will now compute ratios of AB:LL (1:2), AB:AP (1:3), and LL:AP (2:3) for
[…]

### 2013-12-08 results.tex
added matlab syntax to generate Rsq for axis fits to input into LaTeX doc

@@
 ARFI was most correlated to MR in the lateral-to-lateral axis in both the total
-and CG (R$^2$ = 0.57 and 0.32, respectively), with mean overestimates of 13.5
+and CG (R$^2$ = \totalLatLatRsq and \centralLatLatRsq, respectively), with mean overestimates of 13.5
 $\pm$ 11.0\% and 11.5 $\pm$ 22.5\%, respectively
@@
 total prostate gland and CG axes in the anterior-to-posterior dimension (R$^2$
-= 0.23 and R$^2$ = 0.34), respectively, with overestimates of 5.7 $\pm$ 20.6
+= \totalAntPostRsq and \centralAntPostRsq), respectively, with overestimates of 5.7 $\pm$ 20.6
 and -6.1 $\pm$ 21.1\%.  ARFI imaging had weak correlation with MR images in the
-apex-to-base dimension (R$^2$ = 0.19, total gland and R$^2$ = 0.09, CG), with
+apex-to-base dimension (R$^2$ = \totalApexBaseRsq, total gland and R$^2$ = \centralApexBaseRsq, central gland), with
 differences of 2.4 $\pm$ 17.0 and 1.2 $\pm$ 19.8\%, respectively.

### 2013-12-08 figs/mr_arfi_path_axes.m

@@ -35,6 +35,9 @@ function mr_arfi_path_axes
     latex_fid = fopen('../tab_mr_arfi_axes_error.tex','w');
     print_tab_mr_arfi_axes_error_tex_header(latex_fid);
 
+    % open up Rsq file for writing
+    rsq_fid = fopen('../data/mr_arfi_axes_Rsq.tex','w');
+
     % lets plot some direct axis measurements of the total prostate and central gland
     for i=1:3,
         figure;
@@ -87,11 +90,6 @@ function mr_arfi_path_axes
         arfi_axis_OverUnder(:,i) = compute_over_under(PATH_total_axes(:,i),ARFI_total_axes(:,i)/10);
     end;
 
-    %mean(mr_axis_OverUnder)
-    %std(mr_axis_OverUnder)
-    %mean(arfi_axis_OverUnder)
-    %std(arfi_axis_OverUnder)
-
     % let's make plots of the correlation between MR:ARFI for total gland and central gland
     for i=1:3,
         figure;
@@ -123,13 +121,20 @@ function mr_arfi_path_axes
 
        title(sprintf('%s Axis',axis_titles{i}),'FontSize',fs);
 
+       % print legends and Rsq data for inclusion in LaTeX doc
        switch i
        case 1
            legend(sprintf('Total Gland (R^2 = %.2f)',image_total_Rsq),sprintf('Central Gland (R^2 = %.2f)',image_central_Rsq),'Location','NorthWest');
+           fprintf(rsq_fid,'\\newcommand{\\totalApexBaseRsq}{%.2f}\n',image_total_Rsq);
+           fprintf(rsq_fid,'\\newcommand{\\centralApexBaseRsq}{%.2f}\n',image_central_Rsq);
         case 2
            legend(sprintf('Total Gland (R^2 = %.2f)',image_total_Rsq),sprintf('Central Gland (R^2 = %.2f)',image_central_Rsq),'Location','SouthEast');
+           fprintf(rsq_fid,'\\newcommand{\\totalLatLatRsq}{%.2f}\n',image_total_Rsq);
+           fprintf(rsq_fid,'\\newcommand{\\centralLatLatRsq}{%.2f}\n',image_central_Rsq);
         case 3
[…]

### 2013-12-08 figs/mr_arfi_path_axes.m
updated scripts to generate axis errors (pct) in matlab and input as LaTeX variables

@@ -32,8 +32,7 @@ function mr_arfi_path_axes
     fs = 18;
 
     % open up the table of axis errors for LaTeX
-    latex_fid = fopen('../tab_mr_arfi_axes_error.tex','w');
-    print_tab_mr_arfi_axes_error_tex_header(latex_fid);
+    latex_fid = fopen('../data/mr_arfi_axes_error_data.tex','w');
 
     % open up Rsq file for writing
     rsq_fid = fopen('../data/mr_arfi_axes_Rsq.tex','w');
@@ -156,7 +155,7 @@ function mr_arfi_path_axes
     %std(mr_arfi_total_OverUnder)
     %mean(mr_arfi_central_OverUnder)
     %std(mr_arfi_central_OverUnder)
-    print_tab_mr_arfi_axes_error_tex_data(latex_fid,mr_arfi_total_OverUnder,mr_arfi_central_OverUnder);
+    print_mr_arfi_axes_error_tex_data(latex_fid,mr_arfi_total_OverUnder,mr_arfi_central_OverUnder);
 
     % I will now compute ratios of AB:LL (1:2), AB:AP (1:3), and LL:AP (2:3) for
     % imaging central and total and path total volumes
@@ -321,8 +320,6 @@ function mr_arfi_path_axes
         fprintf(fid,'ARFI & MR & Central & %s & %.1f $\\pm$ %.1f \\\\ \n',axis_combos{i},mean(ARFI_MR_Central_OverUnder(:,i)),std(ARFI_MR_Central_OverUnder(:,i)));
     end;
 
-    % print LaTeX axis error footer and close file
-    print_tab_mr_arfi_axes_error_tex_footer(latex_fid);
     fclose(latex_fid)
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@ -339,32 +336,24 @@ function [over_under]=compute_over_under(mr,arfi)
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-function print_tab_mr_arfi_axes_error_tex_header(fid)
-    fprintf(fid,'\\begin{table}[h!]\n');
[…]

### 2013-12-08 figs/mr_arfi_path_axes.m
comments on data files and changed author order

@@ -31,11 +31,11 @@ function mr_arfi_path_axes
     % fontsize
     fs = 18;
 
-    % open up the table of axis errors for LaTeX
+    % open up the table of axis errors and Rsq data files for LaTeX doc inputs
     latex_fid = fopen('../data/mr_arfi_axes_error_data.tex','w');
-
-    % open up Rsq file for writing
+    print_data_header(latex_fid);
     rsq_fid = fopen('../data/mr_arfi_axes_Rsq.tex','w');
+    print_data_header(rsq_fid);
 
     % lets plot some direct axis measurements of the total prostate and central gland
     for i=1:3,
@@ -357,3 +357,6 @@ function print_mr_arfi_axes_error_tex_data(fid,total,central)
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
+function print_data_header(fid)
+    fprintf(fid,'%% DO NOT MANUALLY EDIT THIS FILE\n');
+    fprintf(fid,'%% Generated by figs/mr_arfi_path_axes.m\n');

### 2013-12-08 abstract.tex
added variables for error estimates

@@
 were, on average, larger (6.1\% $\pm$ 25\%) than MR images, primarily due to
-over-estimation of the lateral dimension of the total gland (13.5 $\pm$
-11.0\%), while over-estimates of the other dimensions were less significant
-(5.7 $\pm$ 20.6\% and 2.4 $\pm$ 17.0\%).  The ARFI and MR CG volumes were also
-moderately correlated (R$^2$ = 0.38).  CG volume differences were attributed to
-under-estimation of the anterior-to-posterior axis (-6.1 $\pm$ 21.1\%), and
-over-estimation bias in the lateral dimension (11.5 $\pm$ 22.5\%).  ARFI
-prostate imaging yielded prostate volumes and dimensions that were correlated
-with MR T2WI estimates, with biases in the AP dimension, most likely related to
-poor displacement SNR in the anterior region of the prostate from greater
-distance from the rectal wall imaging surface.  Lateral dimension contrast was
-also challenging due to peri-prostatic fat.  ARFI imaging is a promising
-low-cost, real-time imaging modality that can compliment MR imaging for
-diagnosis, treatment planning and management of PCa.
+over-estimation of the lateral dimension of the total gland
+(\ARFImrTotalLatLatMeanPct $\pm$ \ARFImrTotalLatLatStdPct\%), while
+over-estimates of the other dimensions were less significant
+(\ARFImrTotalAntPostMeanPct $\pm$ \ARFImrTotalAntPostStdPct\% and
+\ARFImrTotalApexBaseMeanPct $\pm$ \ARFImrTotalApexBaseStdPct\%).  The ARFI and
+MR CG volumes were also moderately correlated (R$^2$ = 0.38).  CG volume
+differences were attributed to under-estimation of the apex-to-base axis
+(\ARFImrCentralApexBaseMeanPct $\pm$ \ARFImrCentralApexBaseStdPct\%), and
+over-estimation bias in the lateral dimension (\ARFImrCentralLatLatMeanPct
+$\pm$ \ARFImrCentralLatLatStdPct\%).  ARFI prostate imaging yielded prostate
+volumes and dimensions that were correlated with MR T2WI estimates, with biases
+in the AP dimension, most likely related to poor displacement SNR in the
+anterior region of the prostate from greater distance from the rectal wall
+imaging surface.  Lateral dimension contrast was also challenging due to
+peri-prostatic fat.  ARFI imaging is a promising low-cost, real-time imaging
+modality that can compliment MR imaging for diagnosis, treatment planning and
+management of PCa.

### 2013-12-08 results.tex
added variables for error estimates

--- 2013-10-07 results.tex
@@
 ARFI was most correlated to MR in the lateral-to-lateral axis in both the total
-and CG (R$^2$ = 0.57 and 0.32, respectively), with mean overestimates of 13.5
-$\pm$ 11.0\% and 11.5 $\pm$ 22.5\%, respectively
+and CG (R$^2$ = \totalLatLatRsq~and \centralLatLatRsq, respectively), with mean
+overestimates of \ARFImrTotalLatLatMeanPct~$\pm$ \ARFImrTotalLatLatStdPct\%~and
+\ARFImrCentralLatLatMeanPct~$\pm$ \ARFImrCentralLatLatStdPct\%, respectively
 (Table~\ref{tab:mr_arfi_axes_error}).  ARFI had moderate correlation with the
 total prostate gland and CG axes in the anterior-to-posterior dimension (R$^2$
-= 0.23 and R$^2$ = 0.34), respectively, with overestimates of 5.7 $\pm$ 20.6
-and -6.1 $\pm$ 21.1\%.  ARFI imaging had weak correlation with MR images in the
-apex-to-base dimension (R$^2$ = 0.19, total gland and R$^2$ = 0.09, CG), with
-differences of 2.4 $\pm$ 17.0 and 1.2 $\pm$ 19.8\%, respectively.
+= \totalAntPostRsq~and \centralAntPostRsq), respectively, with overestimates of
+\ARFImrTotalAntPostMeanPct~$\pm$ \ARFImrTotalAntPostStdPct and
+\ARFImrCentralAntPostMeanPct~$\pm$ \ARFImrCentralAntPostStdPct\%.  ARFI imaging
+had weak correlation with MR images in the apex-to-base dimension (R$^2$ =
+\totalApexBaseRsq, total gland and R$^2$ = \centralApexBaseRsq, central gland),
+with differences of \ARFImrTotalApexBaseMeanPct~$\pm$
+\ARFImrTotalApexBaseStdPct~and \ARFImrCentralApexBaseMeanPct~$\pm$
+\ARFImrCentralApexBaseStdPct\%, respectively.
 

### 2013-12-09 conclusions.tex
more variable subs

@@
 that were, on average, larger (6.1\% $\pm$ 25\%) than MR images, primarily due
-to over-estimation of the lateral dimension of the prostate total gland (13.5
-$\pm$ 11.0\%), while over-estimates of the other dimensions were less
-significant contributors (5.7 $\pm$ 20.6\% and 2.4 $\pm$ 17.0\%).  The central
+to over-estimation of the lateral dimension of the prostate total gland
+(\ARFImrTotalLatLatMeanPct~ $\pm$ \ARFImrTotalLatLatStdPct\%), while
+over-estimates of the other dimensions were less significant contributors
+(\ARFImrTotalAntPostMeanPct~$\pm$ \ARFImrTotalAntPostStdPct\%~and
+\ARFImrTotalApexBaseMeanPct~$\pm$ \ARFImrTotalApexBaseStdPct\%).  The central
 zone volumes of ARFI and MR images were also moderately correlated (R$^2$ =
@@
 were, again, strongly attributed to over-estimation of the lateral dimension
-(11.5 $\pm$ 22.5\%), with a significant underestimation of the
-anterior-to-posterior dimension (-6.1 $\pm$ 21.1\%).  Strong variability in
-central gland volumes is believed to be related to the extent of benign
-prostatic hyperplasia (BPH) for select cases.  Overall, ARFI imaging of the
-prostate yielded prostate volumes and dimensions that were correlated with MR
-T2WI estimates, with biases in the anterior-to-posterior dimension, most likely
-related to poor displacement SNR in the anterior region of the prostate from
-greater distance from the rectal wall imaging surface, and the lateral
+(\ARFImrCentralLatLatMeanPct~$\pm$ \ARFImrCentralLatLatStdPct\%), with a
+significant underestimation of the anterior-to-posterior dimension
+(\ARFImrCentralAntPostMeanPct~$\pm$ \ARFImrCentralAntPostStdPct\%).  Strong
+variability in central gland volumes is believed to be related to the extent of
+benign prostatic hyperplasia (BPH) for select cases.  Overall, ARFI imaging of
+the prostate yielded prostate volumes and dimensions that were correlated with
+MR T2WI estimates, with biases in the anterior-to-posterior dimension, most
+likely related to poor displacement SNR in the anterior region of the prostate
+from greater distance from the rectal wall imaging surface, and the lateral
 dimension, where contrast between the PZ and peri-prostatic fat could be

### 2013-12-09 results.tex
added variables for pathology correlations

@@
 were moderately correlated with estimated pathology ellipsoidal prostate
-volumes (Figure~\ref{fig:mr_arfi_weight}(a), R$^2$ = 0.68).  There was moderate
-correlation between the prostate weight and the image-reconstructed prostate
-volumes (Figure~\ref{fig:mr_arfi_weight}(b), R$^2$ = 0.44 (MR) and 0.18
-(ARFI)), though there was weaker correlation with the ellipsoidal approximation
-of the measurement prostate volume and the image-reconstructed volumes
-(Figure~\ref{fig:mr_arfi_weight}(c), R$^2$ = 0.08 (MR) and 0.00 (ARFI)).
+volumes (Figure~\ref{fig:mr_arfi_weight}(a), R$^2$ = \pathVolWeightRsq).  There
+was moderate correlation between the prostate weight and the
+image-reconstructed prostate volumes (Figure~\ref{fig:mr_arfi_weight}(b), R$^2$
+= \weightMRrsq~(MR) and \weightARFIrsq~ (ARFI)), though there was weaker
+correlation with the ellipsoidal approximation of the measurement prostate
+volume and the image-reconstructed volumes (Figure~\ref{fig:mr_arfi_weight}(c),
+R$^2$ = \pathVolMRrsq~(MR) and \pathVolARFIrsq~(ARFI)).
 

### 2013-12-09 figs/corr_weight_vol.m

@@ -10,6 +10,11 @@ function corr_weight_vol()
 
     vols = csvread('../data/data.csv');
 
+    % open data file for Rsq metrics to be read into LaTeX doc
+    fid = fopen('../data/corr_weight_vol_data.tex','w');
+    fprintf(fid,'%% DO NOT MANUAL EDIT THIS FILE\n');
+    fprintf(fid,'%% Generated by figs/corr_weight_vol.m\n');
+
     Pnum = num2cell(vols(:,1));
     mr_total_vol = vols(:,3);
     arfi_total_vol = vols(:,11);
@@ -19,12 +24,15 @@ function corr_weight_vol()
     mr_total_vol = mr_total_vol/1e3;
     arfi_total_vol = arfi_total_vol/1e3;
 
-    gen_fig_n_fits(w,mr_total_vol,arfi_total_vol,'weight');
-    gen_fig_n_fits(path_vols,mr_total_vol,arfi_total_vol,'pathVol');
-    gen_fig_n_fits(w,path_vols,path_vols,'path_vol_weight');
+    gen_fig_n_fits(fid,w,mr_total_vol,arfi_total_vol,'weight');
+    gen_fig_n_fits(fid,path_vols,mr_total_vol,arfi_total_vol,'pathVol');
+    gen_fig_n_fits(fid,w,path_vols,path_vols,'path_vol_weight');
+
+    fclose(fid);
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-function gen_fig_n_fits(w,mr_total_vol,arfi_total_vol,metric)
+
+function gen_fig_n_fits(fid,w,mr_total_vol,arfi_total_vol,metric)
     % fontsize
     fs = 18;
     figure;
@@ -74,11 +82,21 @@ function gen_fig_n_fits(w,mr_total_vol,arfi_total_vol,metric)
     if(strcmp(metric,'path_vol_weight')),
         legend(sprintf('R^2 = %.2f',mr_Rsq),'Location','NorthWest');
         legend boxoff;
+        fprintf(fid,'\\newcommand{\\pathVolWeightRsq}{%.2f}\n',mr_Rsq);
     else,
         legend(sprintf('MR (R^2 = %.2f)',mr_Rsq),sprintf('ARFI (R^2 = %.2f)',arfi_Rsq),'Location','NorthWest');
[…]

### 2013-12-09 abstract.tex
updated volume Rsq and errors to Matlab vars

@@
 gross pathologic weights.  ARFI and MR total prostate gland volumes were
-well-correlated (R$^2$ = 0.63), but ARFI images yielded prostate volumes that
-were, on average, larger (6.1\% $\pm$ 25\%) than MR images, primarily due to
+well-correlated (R$^2$ = \MRarfiVolTotalRsq), but ARFI images yielded prostate volumes that
+were, on average, larger (\MRarfiVolTotalMeanDiff\% $\pm$ \MRarfiVolTotalStdDiff\%) than MR images, primarily due to
 over-estimation of the lateral dimension of the total gland
@@
 \ARFImrTotalApexBaseMeanPct $\pm$ \ARFImrTotalApexBaseStdPct\%).  The ARFI and
-MR CG volumes were also moderately correlated (R$^2$ = 0.38).  CG volume
+MR CG volumes were also well-correlated (R$^2$ = \MRarfiVolCentralRsq).  CG volume
 differences were attributed to under-estimation of the apex-to-base axis

### 2013-12-09 conclusions.tex
updated volume Rsq and errors to Matlab vars

--- 2013-10-07 conclusions.tex
@@
 volume and gross pathologic weights, and ARFI:MR total prostate gland volumes
-were well-correlated (R$^2$ = 0.63), but ARFI images yielded prostate volumes
-that were, on average, larger (6.1\% $\pm$ 25\%) than MR images, primarily due
-to over-estimation of the lateral dimension of the prostate total gland (13.5
-$\pm$ 11.0\%), while over-estimates of the other dimensions were less
-significant contributors (5.7 $\pm$ 20.6\% and 2.4 $\pm$ 17.0\%).  The central
-zone volumes of ARFI and MR images were also moderately correlated (R$^2$ =
-0.38), with minimal volume bias between the imaging modalities, but significant
-variability case-to-case (-5.0 $\pm$ 39.5\%).  Central zone volume differences
-were, again, strongly attributed to over-estimation of the lateral dimension
-(11.5 $\pm$ 22.5\%), with a significant underestimation of the
-anterior-to-posterior dimension (-6.1 $\pm$ 21.1\%).  Strong variability in
-central gland volumes is believed to be related to the extent of benign
-prostatic hyperplasia (BPH) for select cases.  Overall, ARFI imaging of the
-prostate yielded prostate volumes and dimensions that were correlated with MR
-T2WI estimates, with biases in the anterior-to-posterior dimension, most likely
-related to poor displacement SNR in the anterior region of the prostate from
-greater distance from the rectal wall imaging surface, and the lateral
-dimension, where contrast between the PZ and peri-prostatic fat could be
-limited.  ARFI imaging is a promising low-cost, real-time imaging modality that
-can compliment MR imaging for diagnosis, treatment planning and management of
-PCa.
+were well-correlated (R$^2$ = \MRarfiVolTotalRsq), but ARFI images yielded
+prostate volumes that were, on average, larger (\MRarfiVolTotalMeanDiff\%~$\pm$
+\MRarfiVolTotalStdDiff\%) than MR images, primarily due to over-estimation of
+the lateral dimension of the prostate total gland (\ARFImrTotalLatLatMeanPct~
+$\pm$ \ARFImrTotalLatLatStdPct\%), while over-estimates of the other dimensions
+were less significant contributors (\ARFImrTotalAntPostMeanPct~$\pm$
+\ARFImrTotalAntPostStdPct\%~and \ARFImrTotalApexBaseMeanPct~$\pm$
+\ARFImrTotalApexBaseStdPct\%).  The central zone volumes of ARFI and MR images
+were also well-correlated (R$^2$ = \MRarfiVolCentralRsq), with minimal volume
+bias between the imaging modalities, but significant variability case-to-case
+(\MRarfiVolCentralMeanDiff~$\pm$ MRarfiVolCentralStdDiff\%).  Central zone
+volume differences were, again, strongly attributed to over-estimation of the
+lateral dimension (\ARFImrCentralLatLatMeanPct~$\pm$
+\ARFImrCentralLatLatStdPct\%), with a significant underestimation of the
+anterior-to-posterior dimension (\ARFImrCentralAntPostMeanPct~$\pm$
+\ARFImrCentralAntPostStdPct\%).  Strong variability in central gland volumes is
+believed to be related to the extent of benign prostatic hyperplasia (BPH) for
+select cases.  Overall, ARFI imaging of the prostate yielded prostate volumes
+and dimensions that were correlated with MR T2WI estimates, with biases in the
+anterior-to-posterior dimension, most likely related to poor displacement SNR
+in the anterior region of the prostate from greater distance from the rectal
+wall imaging surface, and the lateral dimension, where contrast between the PZ
+and peri-prostatic fat could be limited.  ARFI imaging is a promising low-cost,
+real-time imaging modality that can compliment MR imaging for diagnosis,
+treatment planning and management of PCa.

### 2013-12-09 results.tex
updated volume Rsq and errors to Matlab vars

@@
 measured (Figure~\ref{fig:mr_arfi_volumes}(a)), with moderate correlation
-between the ARFI and MR total prostate gland volumes (R$^2$ = 0.63), with a
-mean overestimation of of 6.1 $\pm$ 25.0\% by ARFI imaging compared to MR
-volumes (Figure~\ref{fig:mr_arfi_volumes}(b)).  Central gland volumes were
-slightly less correlated between ARFI and MR images (R$^2$ = 0.38) with no
-significant mean over/under-estimation (-5.0 $\pm$ 39.5\%), but significant
-variability between the cases (Figure~\ref{fig:mr_arfi_volumes}(c)).
-Table~\ref{tab:mr_arfi_volumes} has the individually-measured volumes for MR
-and ARFI imaging for each study subject.
+between the ARFI and MR total prostate gland volumes (R$^2$ =
+\MRarfiVolTotalRsq), with a mean overestimation of of
+\MRarfiVolTotalMeanDiff~$\pm$ \MRarfiVolTotalStdDiff\% by ARFI imaging compared
+to MR volumes (Figure~\ref{fig:mr_arfi_volumes}(b)).  Central gland volumes
+were well-correlated between ARFI and MR images (R$^2$ = \MRarfiVolCentralRsq)
+with no significant mean over/under-estimation (\MRarfiVolCentralMeanDiff~$\pm$
+\MRarfiVolCentralStdDiff\%), but significant variability between the cases
+(Figure~\ref{fig:mr_arfi_volumes}(c)).  Table~\ref{tab:mr_arfi_volumes} has the
+individually-measured volumes for MR and ARFI imaging for each study subject.
 

### 2013-12-09 figs/mr_arfi_volumes.m

@@ -10,6 +10,11 @@
 
 vols = csvread('../data/data.csv');
 
+% open data file for LaTeX doc input
+fid = fopen('../data/mr_arfi_volume_data.tex','w');
+fprintf(fid,'%% DO NOT MANUALLY EDIT THIS FILE\n');
+fprintf(fid,'%% Generated by figs/mr_arfi_volumes.m\n');
+
 Pnum = num2cell(vols(:,1));
 mr_cg_vol = vols(:,2);
 mr_total_vol = vols(:,3);
@@ -124,7 +129,10 @@ figure;
     %a=axis;                            
     %axis(a);
     text(50,35,sprintf('R^2 = %.2f',mr_arfi_total_Rsq),'FontSize',fs);
+    fprintf(fid,'\\newcommand{\\MRarfiVolTotalRsq}{%.2f}\n',mr_arfi_total_Rsq);
     text(50,30,sprintf('Mean Diff = %.1f +/- %.1f',mean(arfi_mr_vol_diff),std(arfi_mr_vol_diff)),'FontSize',fs);
+    fprintf(fid,'\\newcommand{\\MRarfiVolTotalMeanDiff}{%.2f}\n',mean(arfi_mr_vol_diff));
+    fprintf(fid,'\\newcommand{\\MRarfiVolTotalStdDiff}{%.2f}\n',std(arfi_mr_vol_diff));
     % add ideal line
     line([15 80],[15 80],'LineStyle','--','Color','k','LineWidth',3);
     print('-depsc2','mr_arfi_total_linreg.eps');
@@ -142,10 +150,16 @@ figure;
     %a=axis;                            
     %axis(a);
     text(5,55,sprintf('R^2 = %.2f',mr_arfi_central_Rsq),'FontSize',fs);
+    fprintf(fid,'\\newcommand{\\MRarfiVolCentralRsq}{%.2f}\n',mr_arfi_central_Rsq);
     text(5,50,sprintf('Mean Diff = %.1f +/- %.1f',mean(arfi_mr_central_diff),std(arfi_mr_central_diff)),'FontSize',fs);
+    fprintf(fid,'\\newcommand{\\MRarfiVolCentralMeanDiff}{%.2f}\n',mean(arfi_mr_central_diff));
+    fprintf(fid,'\\newcommand{\\MRarfiVolCentralStdDiff}{%.2f}\n',std(arfi_mr_central_diff));
     % add ideal line
[…]

### 2013-12-09 fig_mr_arfi_volumes.tex
updated caption for variables

@@
     19.6--71.0 cm$^3$ based on MR image models (a), with a moderate correlation
-    between MR and ARFI imaging for both the total gland volume (R$^2$ = 0.63,
-    (b)) and the central gland (R$^2$ = 0.38, (c)).  The ARFI total gland
-    volumes had a positive bias compared with the MR volumes (6.1 $\pm$ 25.0\%
-    (b)), while the ARFI central zone volumes were on average more similar
-    (-5.0 $\pm$ 39.5\%, (c)), but with some significant variability between
-    study subjects.  Table~\ref{tab:mr_arfi_volumes} contains the individual
-    volume estimates for the total prostate and central glands.}
+    between MR and ARFI imaging for both the total gland volume (R$^2$ =
+    \MRarfiVolTotalRsq, (b)) and the central gland (R$^2$ =
+    \MRarfiVolCentralRsq, (c)).  The ARFI total gland volumes had a positive
+    bias compared with the MR volumes (\MRarfiVolTotalMeanDiff~$\pm$
+    \MRarfiVolTotalStdDiff\% (b)), while the ARFI central zone volumes were on
+    average more similar (\MRarfiVolCentralMeanDiff~$\pm$
+    \MRarfiVolCentralStdDiff\%, (c)), but with some significant variability
+    between study subjects.  Table~\ref{tab:mr_arfi_volumes} contains the
+    individual volume estimates for the total prostate and central glands.}
 \label{fig:mr_arfi_volumes} 

### 2013-12-09 abstract.tex
added Zach's methods and figures; still need to refine

@@
 gross pathologic weights.  ARFI and MR total prostate gland volumes were
-well-correlated (R$^2$ = \MRarfiVolTotalRsq), but ARFI images yielded prostate volumes that
-were, on average, larger (\MRarfiVolTotalMeanDiff\% $\pm$ \MRarfiVolTotalStdDiff\%) than MR images, primarily due to
-over-estimation of the lateral dimension of the total gland
-(\ARFImrTotalLatLatMeanPct $\pm$ \ARFImrTotalLatLatStdPct\%), while
-over-estimates of the other dimensions were less significant
-(\ARFImrTotalAntPostMeanPct $\pm$ \ARFImrTotalAntPostStdPct\% and
-\ARFImrTotalApexBaseMeanPct $\pm$ \ARFImrTotalApexBaseStdPct\%).  The ARFI and
-MR CG volumes were also well-correlated (R$^2$ = \MRarfiVolCentralRsq).  CG volume
-differences were attributed to under-estimation of the apex-to-base axis
+well-correlated (R$^2$ = \MRarfiVolTotalRsq), but ARFI images yielded prostate
+volumes that were, on average, larger (\MRarfiVolTotalMeanDiff\% $\pm$
+\MRarfiVolTotalStdDiff\%) than MR images, primarily due to over-estimation of
+the lateral dimension of the total gland (\ARFImrTotalLatLatMeanPct $\pm$
+\ARFImrTotalLatLatStdPct\%), while over-estimates of the other dimensions were
+less significant (\ARFImrTotalAntPostMeanPct $\pm$ \ARFImrTotalAntPostStdPct\%
+and \ARFImrTotalApexBaseMeanPct $\pm$ \ARFImrTotalApexBaseStdPct\%).  The ARFI
+and MR CG volumes were also well-correlated (R$^2$ = \MRarfiVolCentralRsq).  CG
+volume differences were attributed to under-estimation of the apex-to-base axis
 (\ARFImrCentralApexBaseMeanPct $\pm$ \ARFImrCentralApexBaseStdPct\%), and

### 2013-12-09 fig_arfi_segs.tex
added Zach's methods and figures; still need to refine

@@
 \begin{tabular}{c}
-\includegraphics[width=1.0\textwidth]{figs/Bmode_CapsuleSegs.eps} \\
-\includegraphics[width=1.0\textwidth]{figs/ARFI_CGsegs.eps} \\
+%\includegraphics[width=1.0\textwidth]{figs/Bmode_CapsuleSegs.eps} \\
+%\includegraphics[width=1.0\textwidth]{figs/ARFI_CGsegs.eps} \\
+\includegraphics[width=0.8\textwidth]{zach/us_capsule_seg_model.png} \\
+\includegraphics[width=0.8\textwidth]{zach/arfi_central_seg.png} \\
+\includegraphics[width=0.8\textwidth]{zach/us_central_seg_model.png} \\
 \end{tabular}
-\caption{Representative B-mode (top row) and ARFI images (bottom row), with
-    superimposed segmentation outlines of the capsule and CG, respectively, for
-    the same study subject shown in Figure~\ref{fig:mr_segs_vol}.  The B-mode
-    images were segmented in the axial view, using the hypoechoic border of the
-    capsule.  Notice that for large prostates, like the one shown in this
-    figure, that the edges of the organ were not always captured in the imaging
-    field-of-view.  The CG in the ARFI images was typically segmented using the
-    coronal view, which, especially in the presence of extensive BPH, can have
-    a very characteristic, heterogeneous appearance.  It should be noted that
-    the coronal views in these ultrasound images are flipped 180\degree relative
-    to the same images in Figure~\ref{fig:mr_segs_vol}.}
+\caption{TOP ROW: B-mode capsule segmentation using all three orthogonal
+    planes.  The primary segmentation plane is sagittal (second pane), however
+    the axial (first pane) and coronal (third pane) are used to guide this
+    segmentation because as the label map is built up, the intersections of
+    the label map in the axial and sagittal planes can be viewed as well. The
+    block-like character of the axial/coronal sections of the label map are a
+    direct result of downsampling.  MIDDLE ROW: ARFI central gland segmentation
+    (blue) with the capsule model (pink) overlaid for guidance during
+    segmentation.  BOTTOM ROW: Demonstrates the modeled capsule with slice
+    intersections on the axial, sagittal, and coronal planes respectively (all
+    pink). This model was built up from the label map created in the B-mode
+    capsule segmentation.}
 \label{fig:arfi_segs} 

### 2013-12-09 image_proc_methods.tex
added Zach's methods and figures; still need to refine

@@
 \subsection{Image Zonal Anatomy Segmentation and 3D Model Rendering}
+\subsubsection{MR Image Segmentation and Modeling}
 Axial MR T2WI images were manually segmented using the smooth polygon tool in
@@
 central gland.
-
-Figure~\ref{fig:mr_anatomy} shows the anatomic zones and some anatomic
-orientation labels in the prostate of a representative study subject MR T2WI.  The
-zonal anatomy of the prostate was manually segmented
-(Figure~\ref{fig:mr_segs_vol}, bottom left) across the image stack, and 3D
-models of the zonal anatomy volumes were rendered
-(Figure~\ref{fig:mr_segs_vol}, top).
-
-\input{fig_mr_anatomy}
-
-\input{fig_mr_segs_vol}
-
-ARFI images were segmented using a similar procedure to the MR images.  B-mode
-images were used to segment the prostate capsules since contrast between the
-peripheral zone and the periprostatic fat can be low, and poor displacement SNR
-can be present along the anterior aspect of the prostate in ARFI images, making
-it difficult to delineate that boundary.  The central gland was segmented as a
-stiffer region, relative to the peripheral zone, in the center of the prostate.
-
-\input{fig_arfi_segs}
 
@@
 
-The 3D slicer models were used to render volume estimates of the PZ and CG,
-with the sum of PZ and CG representing the total prostate gland volume.  AFS
-volume estimates in select MR cases were included in the total prostate gland
-volume estimates.  Orthogonal tri-axial measurements in the lateral-to-lateral,
-axex-to-base, and anterior-to-posterior dimensions were made in 3D slicer.
+The 3D models were used to render volume estimates of the PZ and CG, with the
+sum of PZ and CG representing the total prostate gland volume. AFS volume
+estimates in select MR cases were included in the total prostate gland volume
+estimates. Orthogonal tri-axial measurements in the lateral-to-lateral,
+apex-to-base, and anterior-to-posterior dimensions were made in slicer by
+placing an ROI cube in the model's center and expanding the ROI cube to fit the
+maximal model dimensions.
+
+Figure~\ref{fig:mr_anatomy} shows the anatomic zones and some anatomic
+orientation labels in the prostate of a representative study subject MR T2WI.  The
+zonal anatomy of the prostate was manually segmented
+(Figure~\ref{fig:mr_segs_vol}, bottom left) across the image stack, and 3D
+models of the zonal anatomy volumes were rendered
+(Figure~\ref{fig:mr_segs_vol}, top).
+
+\input{fig_mr_anatomy}
+
+\input{fig_mr_segs_vol}
+
+\subsubsection{Ultrasound Image Segmentation and Modeling}
+Sagittal B-mode and ARFI ultrasound image stacks were segmented in 3D Slicer
+using the \verb+draweffect+ tool to model and measure the prostate capsule and CG
+volume and tri-axial dimensions.  The sagittal imaging plane was chosen for
+segmentation because it has a greater number of easily identifiable anatomical
+structures:  seminal vesicles in the left and right posterior base, urethra in
+the mid apex, surrounding fibromuscular tissue in the anterior , and rectum in
+the posterior  than either the axial or coronal views. Prostate capsule was
+segmented in B-mode as this modality offers both high contrast between
+prostatic tissue and surrounding periprostatic fat, and better SNR in the
+anterior capsule region than ARFI images. ARFI was used to segment central gland as
+there is clear delineation in this modality between central gland and
+peripheral zone. This delineation is a result of the greater density of
+connective tissue within the central gland versus the peripheral zone meaning
+the central gland is stiffer than the surrounding peripheral zone.  Prior to
+segmentation, both the B-mode and ARFI image stacks were downsampled along the
+sagittal axis by a factor of 20 (i.e., from 360 slices per image stack to 18
+slices per image stack). This was done because segmenting 360 slices per image
+is inordinately time intensive, and does not improve the quality of the 3D
+rendered model any more than segmenting the more reasonable 18 slices per image
+stack. The parameters used for the resample scalar volume tool are shown below:
+
+WHAT GOES HERE?
+
+Although, segmentations were performed along the sagittal plane, label map
+intersections in the axial and coronal views were used during active creation
+of the sagittal segmentation to further improve segmentation accuracy.  This
+was done by using the resampled image stack, B-mode for capsule and ARFI for
+CG, as the background in the sagittal plane, and the original image stack as
+background in the axial and coronal planes. All three orthogonal views were set
+to the label map assigned to the resampled volume. The B-mode derived capsule
+was segmented and modeled before the central gland. Parameters used for 3D
+model rendering were identical to those used in table X for MR modeling. The
+modeled capsule  helprf guide the CG segmentation by enabling capsule model
+intersections in the three orthogonal planes during CG segmentation. The
+capsule provided an outer boundary for the CG segmentation. The capsule and CG
+models were then measured identically to the MR measurement procedure.
+
+\input{fig_arfi_segs}
+
+\input{fig_roi_tool}

### 2013-12-29 image_proc_methods.tex
new ROI figure and associated text

@@
 placing an ROI cube in the model's center and expanding the ROI cube to fit the
-maximal model dimensions.
+maximal model dimensions (Figure~\ref{fig:roi_tool}).
 
 Figure~\ref{fig:mr_anatomy} shows the anatomic zones and some anatomic
-orientation labels in the prostate of a representative study subject MR T2WI.  The
-zonal anatomy of the prostate was manually segmented
+orientation labels in the prostate of a representative study subject MR T2WI.
+The zonal anatomy of the prostate was manually segmented
 (Figure~\ref{fig:mr_segs_vol}, bottom left) across the image stack, and 3D

### 2013-12-29 compile-n-view.sh
render pdf directly instead of via dvi

@@ -1,1 +1,1 @@
-latexmk -bibtex -pdfdvi -pvc mr_arfi_anatomy.tex
+latexmk -bibtex -pdf -pvc mr_arfi_anatomy.tex

### 2013-12-29 fig_arfi_segs.tex
latest cap and CG model / seg images

--- 2013-09-23 fig_arfi_segs.tex
@@
 \centering
-\begin{tabular}{c}
-\includegraphics[width=1.0\textwidth]{figs/Bmode_CapsuleSegs.eps} \\
-\includegraphics[width=1.0\textwidth]{figs/ARFI_CGsegs.eps} \\
+\includegraphics[width=0.3\textwidth]{zach/Cap_Modeling_Images/3D_Capsule.png} \\
+Prostate Capsule \\
+\begin{tabular}{ccc}
+\includegraphics[width=0.3\textwidth]{zach/Cap_Modeling_Images/Axial_Cap.png} &
+\includegraphics[width=0.3\textwidth]{zach/Cap_Modeling_Images/Coronal_Cap.png} &
+\includegraphics[width=0.3\textwidth]{zach/Cap_Modeling_Images/Sagittal_Cap.png} \\
+Axial B-mode Outline & Coronal B-mode Outline & Sagital B-mode Outline \\
 \end{tabular}
-\caption{Representative B-mode (top row) and ARFI images (bottom row), with
-    superimposed segmentation outlines of the capsule and CG, respectively, for
-    the same study subject shown in Figure~\ref{fig:mr_segs_vol}.  The B-mode
-    images were segmented in the axial view, using the hypoechoic border of the
-    capsule.  Notice that for large prostates, like the one shown in this
-    figure, that the edges of the organ were not always captured in the imaging
-    field-of-view.  The CG in the ARFI images was typically segmented using the
-    coronal view, which, especially in the presence of extensive BPH, can have
-    a very characteristic, heterogeneous appearance.  It should be noted that
-    the coronal views in these ultrasound images are flipped 180\degree relative
-    to the same images in Figure~\ref{fig:mr_segs_vol}.}
+\includegraphics[width=0.3\textwidth]{zach/CG_Modeling_Images/3D_CG.png} \\
+Prostate Central Gland \\
+\begin{tabular}{ccc}
+\includegraphics[width=0.3\textwidth]{zach/CG_Modeling_Images/Axial_CG.png} &
+\includegraphics[width=0.3\textwidth]{zach/CG_Modeling_Images/Coronal_CG.png} &
+\includegraphics[width=0.3\textwidth]{zach/CG_Modeling_Images/Sagittal_CG.png} \\
+Axial ARFI Outline & Coronal ARFI Outline & Sagital ARFI Outline \\
+\end{tabular}
+\caption{TOP ROWS: Prostate capsule 3D model (magenta) rendered from manual
+    segmentation of the B-mode images, with the 3D model outlines superimposed
+    on the central axial, coronal and sagital B-mode images.  The primary
+    segmentation plane was sagittal; however the axial and coronal planes were
+    used to guide the segmentation to insure 3D continuity of the capsule
+    structure.  BOTTOM ROWS: Prostate central gland 3D model (blue) rendered
+    from manual segmentation of the ARFI images, with the 3D model outlines
+    superimposed on the central axial, coronal and sagital ARFI images.  The
+    coronal ARFI imaging plane was the primary orientation used for image
+    segmentation, but like the capsule, the orthogonal planes were used to
+    insure continuity of the central gland in all three dimensions.}
 \label{fig:arfi_segs} 

### 2013-12-29 image_proc_methods.tex
added ref to new model fig; need updates to text from Zach

@@
 
-WHAT GOES HERE?
+ZACH - WHAT GOES HERE?
 
@@
 was segmented and modeled before the central gland. Parameters used for 3D
-model rendering were identical to those used in table X for MR modeling. The
-modeled capsule  helprf guide the CG segmentation by enabling capsule model
-intersections in the three orthogonal planes during CG segmentation. The
-capsule provided an outer boundary for the CG segmentation. The capsule and CG
-models were then measured identically to the MR measurement procedure.
+model rendering were identical to those used in table X [ZACH] for MR modeling.
+The modeled capsule helped guide the CG segmentation by enabling capsule model
+intersections in the three orthogonal planes during CG segmentation
+(Figure~\ref{fig:arfi_seg}). The capsule provided an outer boundary for the CG
+segmentation. The capsule and CG models were then measured identically to the
+MR measurement procedure.
 

### 2013-12-29 intro.tex
updated intro text

@@
 available for use in the workup of patients with PCa since the early 1980s, but
-early studies on its diagnostic accuracy were heterogeneous.  However, the more
+early studies on its diagnostic accuracy were heterogeneous.  The more
 recent ability to include functional parameters in PCa MRI analysis has yielded
@@
 \etal demonstrated the feasibility of ARFI prostate imaging
-\invivo.~\cite{Zhai2012}; however, to the authors’ best knowledge, there have
+\invivo~\cite{Zhai2012}; however, to the authors’ best knowledge, there have
 been no studies to date that compare \invivo ARFI prostate imaging to other
@@
 The goal of this study is to evaluate the ability of ARFI imaging to delineate
-prostate zonal anatomy, specifically central gland and total prostate gland,
+prostate zonal anatomy, specifically central gland and total prostate gland
 \invivo as compared to endorectal MR T2WI.  Section~\ref{sect:background}

### 2013-12-29 prostate_anatomy.tex
minor edits

@@
 healthy prostate, each zone contains approximately 5\%, 20\% and 70--80\% of
-glandular tissue, respectively.~\cite{Bonekamp2011} The non-glandular
-components include the anterior fibromuscular stroma (AFS) and the urethra.  In
-some imaging modalities, the central and transition zones cannot be visualized
-as discrete entities, and they are collectively identified as the central gland
-(CG).  This CG was one of the prostate structures manually segmented in MR and
-ARFI images in this study.
+glandular tissue, respectively.~\cite{Bonekamp2011}, but these ratios can
+change significantly in the presence of benign prostatic hyperplasia (BPH). The
+non-glandular components include the anterior fibromuscular stroma (AFS) and
+the urethra.  In some imaging modalities, the central and transition zones
+cannot be visualized as discrete entities, and they are collectively identified
+as the central gland (CG).  This CG was one of the prostate structures manually
+segmented in MR and ARFI images in this study.
 

### 2013-12-29 study_methods.tex
new b-mode and arfi segs fig included
minor text updates

@@
 enrolled as study subjects in this IRB-approved (Duke IRB\# Pro00006458),
-HIPAA-compliant study.  A total of 16 patients were recruited and enrolled in
-this study.  Inclusion criteria were undergoing complete pelvic MRI with
-endorectal coil for detection of prostate cancer, including multiplanar
-T2-weighted anatomic imaging, as well as pre-operative ARFI imaging and radical
-prostatectomy.  Patients with previous treatments of PCa or benign prostatic
-hyperplasia (BPH), or anatomic anomalies of the rectum, were excluded from this
-study.  All patients enrolled in this study provided written informed consent.
-
-\emph{CO-AUTHOR QUESTION: DO WE WANT TO INCLUDE A TABLE OF PATIENT DEMOGRAPHIC INFORMATION?}
+HIPAA-compliant study, with all enrolled patients providing written informed
+consent.  A total of 16 patients were recruited and enrolled in this study.
+Inclusion criteria were undergoing complete pelvic MRI with endorectal coil for
+detection of prostate cancer, including multiplanar T2-weighted anatomic
+imaging, as well as pre-operative ARFI imaging and radical prostatectomy.
+Patients with previous treatments of PCa or benign prostatic hyperplasia (BPH),
+or anatomic anomalies of the rectum, were excluded from this study.

### 2013-12-29 mr_methods.tex
re-organized radical prostatectomy text
minor text edits

@@
 images in the coronal, axial and sagittal planes.  Diffusion weighted images
-were obtained using multiple b-values and calculation of ADC maps was also
-performed.  Dynamic contrast enhanced MR sequences were obtained after
-administration of a weight-based dose of extracellular MR contrast agent with
-4-5 second temporal resolution for 5-6 minutes. (If we need it, can put in
-table with full MR parameters) Prostates were radically removed using a da
-Vinci Surgical System (Intuitive Surgical\textregistered, Inc., Sunnyvale, CA).
+were obtained using multiple b-values and calculation of Apparent Diffusion
+Coefficient (ADC) maps was also performed.  Dynamic contrast enhanced MR
+sequences were obtained after administration of a weight-based dose of
+extracellular MR contrast agent with 4-5 second temporal resolution for 5-6
+minutes.

### 2013-12-30 conclusions.tex
fixed some quantitative data

@@
 the lateral dimension of the prostate total gland (\ARFImrTotalLatLatMeanPct~
-$\pm$ \ARFImrTotalLatLatStdPct\%), while over-estimates of the other dimensions
+$\pm$ \ARFImrTotalLatLatStdPct\%), while differences in the other dimensions
 were less significant contributors (\ARFImrTotalAntPostMeanPct~$\pm$
@@
 \ARFImrCentralLatLatStdPct\%), with a significant underestimation of the
-anterior-to-posterior dimension (\ARFImrCentralAntPostMeanPct~$\pm$
-\ARFImrCentralAntPostStdPct\%).  Strong variability in central gland volumes is
-believed to be related to the extent of benign prostatic hyperplasia (BPH) for
-select cases.  Overall, ARFI imaging of the prostate yielded prostate volumes
-and dimensions that were correlated with MR T2WI estimates, with biases in the
-anterior-to-posterior dimension, most likely related to poor displacement SNR
-in the anterior region of the prostate from greater distance from the rectal
-wall imaging surface, and the lateral dimension, where contrast between the PZ
-and peri-prostatic fat could be limited.  ARFI imaging is a promising low-cost,
-real-time imaging modality that can compliment MR imaging for diagnosis,
-treatment planning and management of PCa.
+apex-to-base dimension (\ARFImrCentralAntPostMeanPct~$\pm$
+\ARFImrCentralAntPostStdPct\%).  Strong variability in CG volumes is believed
+to be related to the extent of BPH for select cases.  Overall, ARFI imaging of
+the prostate yielded prostate volumes and dimensions that were correlated with
+MR T2WI estimates, with differences in the lateral and apex-to-base dimensions.
+Poor image contrast between the prostate and peri-prostatic fat may be one
+cause of these differences.  ARFI imaging is a promising low-cost, real-time
+imaging modality that can compliment MR imaging for diagnosis, treatment
+planning and management of PCa.

### 2013-12-30 discussion.tex
fixed some quantitative data

@@
 of the lateral dimension of the prostate.  This can be due to poor contrast
-between the PZ and peri-prostate fat (Figure~\ref{fig:arfi_segs}).  As seen in
-Figure~\ref{fig:arfi_segs}, clear delineation of the anterior aspect of the
-prostate can also be challenging in ARFI and B-mode images due to poor SNR,
+between the PZ and peri-prostatic fat (Figure~\ref{fig:arfi_segs}).
+Figure~\ref{fig:arfi_segs} also demonstrates challenges in delineating the
+anterior aspect of the prostate in both ARFI and B-mode images due to poor SNR
 since this is the deepest aspect of the prostate relative to the rectal wall
@@
 4, with a $\sim$143 g prostate (Figure~\ref{fig:mr_arfi_weight}), which was an
-outlier in terms of weight and volume.  The most accurate measurements for both
-the total prostate gland and central gland were in the apex-to-base dimension
-(Figure~\ref{fig:mr_arfi_path_axes}(c), Table~\ref{tab:mr_arfi_axes_error}),
-with no mean bias in ARFI imaging estimates, but significant case-to-case
-variability, leading to poor linear regression coefficients of determination.
+outlier in terms of weight and volume.  While the anterior aspect of the
+prostate can be challenging to visualize, it was the axis with the best
+agreement between ARFI and MR imaging (Figure~\ref{fig:mr_arfi_path_axes}(b),
+Table~\ref{tab:mr_arfi_axes_error}).  ARFI images consistently underestimated
+the extent of the prostate capsule and CG along the apex-to-base axis relative
+to MR T2WI (Figure~\ref{fig:mr_arfi_path_axes}(c),
+Table~\ref{tab:mr_arfi_axes_error}), though the reasons for this
+underestimation were not clear.
 
@@
 volume.  Interestingly, all pathology estimates were thought to have positive
-biases, but both imaging modalities tended to overestimate relative to the
+biases, but both imaging modalities tended to overestimate volume relative to the
 pathology measurements.
@@
 would not have been excised for pathology characterization, and the zonal
-anatomy of a health (young) prostate is expected to be different from the
+anatomy of a healthy (young) prostate is expected to be different from the
 prostate of a middle-age man, who is the target demographic for PCa screening
-imaging and PCa characterization; therefore, it was felt that performing this
-analysis in the presence of these confounding factors was appropriate.
+imaging and PCa characterization.
 

### 2013-12-30 fig_mr_arfi_weight.tex
fixed some quantitative data

@@
     which was moderately well-correlated with the excised prostated weights (a,
-    R$^2$ = 0.68).  T2WI MR (blue, X) showed a moderate correlation between the
-    reconstructed volumes and prostate weight (R$^2$ = 0.44), while volumes
-    reconstructed from ARFI images (green, O) showed weaker correlation (R$^2$
-    = 0.21) (b).  Even weaker correlations existed between both T2WI MR and
-    ARFI image volumens and approaximated ellipsoidal prostate pathology
-    volumes (R$^2$ = 0.08 and 0.01, respectively) (c).  It should be noted in
-    these figures that two study subjects had excessively large prostates that
-    were difficult to fully capture in imaging (volumes $>$ 60 cm$^3$),
-    especially in their anterior region, and they were, therefore, grossly
-    underestimated in size.}
+    R$^2$ = \pathVolWeightRsq).  T2WI MR (blue, X) showed a moderate
+    correlation between the reconstructed volumes and prostate weight (R$^2$ =
+    \weightMRrsq), while volumes reconstructed from ARFI images (green, O)
+    showed weaker correlation (R$^2$ = \weightARFIrsq) (b).  Even weaker
+    correlations existed between both T2WI MR and ARFI image volumens and
+    approaximated ellipsoidal prostate pathology volumes (R$^2$ = \pathVolMRrsq
+    and \pathVolARFIrsq, respectively) (c).  It should be noted in these
+    figures that two study subjects had excessively large prostates that were
+    difficult to fully capture in imaging (volumes $>$ 60 cm$^3$), especially
+    in their anterior region, and they were, therefore, grossly underestimated
+    in size.}
 \label{fig:mr_arfi_weight}

### 2013-12-30 image_proc_methods.tex
fixed some quantitative data

--- 2013-12-29 image_proc_methods.tex
@@
 bladder, and subsequent images were segmented until the last slice with visible
-prostatic tissue was identified caudally. The CG and urethra, PZ and AFS were
+prostatic tissue was identified caudally. The CG, urethra, PZ and AFS were
 segmented independently according to their well-established anatomical
@@
 T2WI.~\cite{Barentsz2012,Jung2012,Poon1985,Hricak2007,Bonekamp2011} The PZ was
-identified by its homogenous high signal intensity on T2WI, which is usually
+identified by its homogeneous high signal intensity on T2WI, which is usually
 similar to that of the nearby periprostatic fat. The CG was visualized and
-delineated based on its heterogeneous and lower signal intensity as well as its
-location (Figure~\ref{fig:mr_anatomy}). The urethra was included in the CG
-segmentation. Although not readily visible on every case, the AFS was
-identified by its low T2 signal intensity and its location anterior to the
-central gland.
+delineated based on its heterogeneous, lower signal intensity and its location
+(Figure~\ref{fig:mr_anatomy}). The urethra was included in the CG segmentation.
+When visible, the AFS was identified by its low T2 signal intensity and its
+location anterior to the CG.
+
+\input{fig_mr_anatomy}
 
@@
 
+\input{fig_roi_tool}
+
 Figure~\ref{fig:mr_anatomy} shows the anatomic zones and some anatomic
@@
 
-\input{fig_mr_anatomy}
-
 \input{fig_mr_segs_vol}
@@
 Sagittal B-mode and ARFI ultrasound image stacks were segmented in 3D Slicer
-using the \verb+draweffect+ tool to model and measure the prostate capsule and CG
-volume and tri-axial dimensions.  The sagittal imaging plane was chosen for
+using the \verb+draweffect+ tool to model and measure the prostate capsule and
+CG volume and tri-axial dimensions.  The sagittal imaging plane was chosen for
 segmentation because it has a greater number of easily identifiable anatomical
@@
 prostatic tissue and surrounding periprostatic fat, and better SNR in the
-anterior capsule region than ARFI images. ARFI was used to segment central gland as
-there is clear delineation in this modality between central gland and
-peripheral zone. This delineation is a result of the greater density of
-connective tissue within the central gland versus the peripheral zone meaning
-the central gland is stiffer than the surrounding peripheral zone.  Prior to
-segmentation, both the B-mode and ARFI image stacks were downsampled along the
-sagittal axis by a factor of 20 (i.e., from 360 slices per image stack to 18
-slices per image stack). This was done because segmenting 360 slices per image
-is inordinately time intensive, and does not improve the quality of the 3D
-rendered model any more than segmenting the more reasonable 18 slices per image
-stack. The parameters used for the resample scalar volume tool are shown below:
+anterior capsule region than ARFI images. ARFI images were used for CG
+segmentation.  Prior to segmentation, both the B-mode and ARFI image stacks
+were down-sampled along the sagittal axis by a factor of 20 (i.e., from 360
+slices per image stack to 18 slices per image stack) to reduce the number of
+manually-segmented frames.
 
-WHAT GOES HERE?
-
-Although, segmentations were performed along the sagittal plane, label map
+While segmentations were performed along the sagittal plane, label map
 intersections in the axial and coronal views were used during active creation
-of the sagittal segmentation to further improve segmentation accuracy.  This
-was done by using the resampled image stack, B-mode for capsule and ARFI for
-CG, as the background in the sagittal plane, and the original image stack as
-background in the axial and coronal planes. All three orthogonal views were set
-to the label map assigned to the resampled volume. The B-mode derived capsule
-was segmented and modeled before the central gland. Parameters used for 3D
-model rendering were identical to those used in table X for MR modeling. The
-modeled capsule  helprf guide the CG segmentation by enabling capsule model
-intersections in the three orthogonal planes during CG segmentation. The
-capsule provided an outer boundary for the CG segmentation. The capsule and CG
-models were then measured identically to the MR measurement procedure.
+of the sagittal segmentation to further improve segmentation accuracy.
+Parameters used for 3D model rendering were identical to those used for MR
+modeling (Table~\ref{tab:3dslicer}).  The modeled capsule helped guide the CG
+segmentation by enabling capsule model intersections in the three orthogonal
+planes during CG segmentation (Figure~\ref{fig:arfi_segs}). The capsule provided
+an outer boundary for the CG segmentation. The capsule and CG models were then
+measured using the same procedure as the MR images.
 
 \input{fig_arfi_segs}
-
-\input{fig_roi_tool}

### 2013-12-30 results.tex
fixed some quantitative data

@@
 total prostate gland and CG axes in the anterior-to-posterior dimension (R$^2$
-= \totalAntPostRsq~and \centralAntPostRsq), respectively, with overestimates of
-\ARFImrTotalAntPostMeanPct~$\pm$ \ARFImrTotalAntPostStdPct and
+= \totalAntPostRsq~and \centralAntPostRsq), respectively, with differences of
+\ARFImrTotalAntPostMeanPct~$\pm$ \ARFImrTotalAntPostStdPct\% and
 \ARFImrCentralAntPostMeanPct~$\pm$ \ARFImrCentralAntPostStdPct\%.  ARFI imaging
 had weak correlation with MR images in the apex-to-base dimension (R$^2$ =
-\totalApexBaseRsq, total gland and R$^2$ = \centralApexBaseRsq, central gland),
-with differences of \ARFImrTotalApexBaseMeanPct~$\pm$
-\ARFImrTotalApexBaseStdPct~and \ARFImrCentralApexBaseMeanPct~$\pm$
+\totalApexBaseRsq, total gland and R$^2$ = \centralApexBaseRsq, CG),
+with underestimates of \ARFImrTotalApexBaseMeanPct~$\pm$
+\ARFImrTotalApexBaseStdPct\% and \ARFImrCentralApexBaseMeanPct~$\pm$
 \ARFImrCentralApexBaseStdPct\%, respectively.

### 2013-12-30 discussion.tex
updated for more impact

@@
 \section{Discussion}\label{sect:discussion}
-This study was designed to compared the zonal anatomy of the prostate in
-patients with biopsy-proven PCa as seen in ARFI imaging and MR T2WI.  Overall,
-ARFI imaging performed well when compared with MR estimates of total and
-central gland volume (Figure~\ref{fig:mr_arfi_volumes}), though ARFI tended to
-overestimate prostate total gland volume, most commonly due to over-estimation
-of the lateral dimension of the prostate.  This can be due to poor contrast
-between the PZ and peri-prostatic fat (Figure~\ref{fig:arfi_segs}).
-Figure~\ref{fig:arfi_segs} also demonstrates challenges in delineating the
-anterior aspect of the prostate in both ARFI and B-mode images due to poor SNR
-since this is the deepest aspect of the prostate relative to the rectal wall
-imaging surface. MR T2WI can also have difficulties visualizing and diagnosing
-lesions in the anterior region of the prostate,~\cite{Gupta2013}, which is
-thought to be secondary to the heterogeneity of the gland in that area,
-especially in the presence of BPH.  This was evident in imaging study subject
-4, with a $\sim$143 g prostate (Figure~\ref{fig:mr_arfi_weight}), which was an
-outlier in terms of weight and volume.  While the anterior aspect of the
-prostate can be challenging to visualize, it was the axis with the best
-agreement between ARFI and MR imaging (Figure~\ref{fig:mr_arfi_path_axes}(b),
-Table~\ref{tab:mr_arfi_axes_error}).  ARFI images consistently underestimated
-the extent of the prostate capsule and CG along the apex-to-base axis relative
-to MR T2WI (Figure~\ref{fig:mr_arfi_path_axes}(c),
+This study demonstrated that ARFI imaging delineates zonal anatomy of the
+prostate well when compared to MR T2WI, with comparable estimates of total and
+central gland volume (Figure~\ref{fig:mr_arfi_volumes}).  Establishing accurate
+delineation of the prostate's zonal anatomy is important since PCa has a strong
+correlations with location in the prostate, specifically in the peripheral
+zone, and accurate capsule delineation will facilitate future efforts to
+register 3D imaging datasets between ARFI and MR imaging.
+
+This was the first quantitative comparison between ARFI imaging and a newly
+established clinical imaging modality to evaluate the prostate for PCa
+detection and characterization, and we have drawn some useful conclusions in
+how our zonal anatomy boundaries are determined during image segmentation.
+Poor contrast between the PZ and peri-prostatic fat in B-mode images can lead
+to an overestimation of lateral boundaries of the prostate in our ultrasound
+datasets compared to MR T2WI, leading to an overall overestimation of prostate
+total gland volume.  The anterior aspect of the prostate can also be
+challenging to delineate, especially in large prostates, where SNR decreases
+with increasing depth away from our rectal wall imaging surface and MR images
+suffer from gland heterogeneity in this region~\cite{Gupta2013}, though we had
+good agreement between ARFI and MR T2W images in the anterior-to-posterior
+dimension (Figure~\ref{fig:mr_arfi_path_axes}(b),
+Table~\ref{tab:mr_arfi_axes_error}).  Surprisingly, ARFI imaging consistently
+underestimated the extent of the prostate capsule and CG along the apex-to-base
+axis relative to MR T2WI (Figure~\ref{fig:mr_arfi_path_axes}(c),
 Table~\ref{tab:mr_arfi_axes_error}), though the reasons for this
-underestimation were not clear.
+underestimation are not clear.
 
@@
 these results.  Gross pathology weight and axis measurements could both be
-affected by the presence of periprostatic tissue that was excised during
+affected by the presence of peri-prostatic tissue that was excised during
 radical prostatectomy, especially in cases where more aggressive margins may
@@
 
-It should also be noted that this study did not evaluate user biases in image
-segmentation.  While MR zonal anatomy delineation has some establishment in the
-clinical literature, this study is the first attempt to define the criteria for
-ARFI imaging zonal anatomy characteristics, and it is expected that such
-delineations will continue to be refined as we acquire more cases and continue
-to compare with MRI and pathology data.  Given this limitation, no attempts
-were made to further quantify reader-to-reader variability in this work.
+This study did not evaluate user biases in image segmentation.  While MR zonal
+anatomy delineation has some establishment in the clinical literature, this
+study is the first attempt to define the criteria for ARFI imaging zonal
+anatomy characteristics, and it is expected that such delineations will
+continue to be refined as we acquire more cases and continue to compare with
+MRI and pathology data.  Given this limitation, no attempts were made to
+further quantify reader-to-reader variability in this work.
 
@@
 imaging are also being pursued and will hopefully allow for the currently
-elusive anterior PCa lesions in MR to be seen in ARFI imaging.  Although work
-remains to be done, the fact that one can view the different prostatic zones on
-ARFI images in addition to MRI is encouraging and could lead to targeted
-diagnostic biopsies and therapies with real-time imaging for men with PCa.
+elusive anterior PCa lesions in MR to be seen in ARFI imaging.  3D models of
+the prostate and central glands will be used across ARFI and MR imaging
+datasets to spatially register them to facilitate correlation of regions of PCa
+suspicion between the two modalities, and to correlate with whole-mount
+histology of the excised prostate specimens.  This effort will allow the
+sensitivity and specificity of each imaging modality for PCa detection and
+characterization to be quantified.  While ARFI imaging of the prostate in a
+constantly evolving technology, the fact that one can view the different
+prostatic zones on ARFI images in addition to MRI is encouraging and could lead
+to targeted diagnostic biopsies and therapies with real-time imaging for men
+with PCa.

### 2013-12-30 conclusions.tex
corrected variable error

@@
 bias between the imaging modalities, but significant variability case-to-case
-(\MRarfiVolCentralMeanDiff~$\pm$ MRarfiVolCentralStdDiff\%).  Central zone
+(\MRarfiVolCentralMeanDiff~$\pm$ \MRarfiVolCentralStdDiff\%).  Central zone
 volume differences were, again, strongly attributed to over-estimation of the

### 2013-12-30 fig_arfi_segs.tex
updated captions

@@
 \end{tabular}
-\caption{TOP ROWS: Prostate capsule 3D model (magenta) rendered from manual
-    segmentation of the B-mode images, with the 3D model outlines superimposed
-    on the central axial, coronal and sagital B-mode images.  The primary
-    segmentation plane was sagittal; however the axial and coronal planes were
-    used to guide the segmentation to insure 3D continuity of the capsule
-    structure.  BOTTOM ROWS: Prostate central gland 3D model (blue) rendered
-    from manual segmentation of the ARFI images, with the 3D model outlines
-    superimposed on the central axial, coronal and sagital ARFI images.  The
-    coronal ARFI imaging plane was the primary orientation used for image
-    segmentation, but like the capsule, the orthogonal planes were used to
-    insure continuity of the central gland in all three dimensions.}
+\caption{Representative ultrasound prostate models from this study.  TOP ROWS:
+    Prostate capsule 3D model (magenta) rendered from manual segmentation of
+    the B-mode images, with the 3D model outlines superimposed on the central
+    axial, coronal and sagital B-mode images.  The primary segmentation plane
+    was sagittal; however the axial and coronal planes were used to guide the
+    segmentation to insure 3D continuity of the capsule structure.  BOTTOM
+    ROWS: Prostate central gland 3D model (blue) rendered from manual
+    segmentation of the ARFI images, with the 3D model outlines superimposed on
+    the central axial, coronal and sagital ARFI images.  The coronal ARFI
+    imaging plane was the primary orientation used for image segmentation, but
+    like the capsule, the orthogonal planes were used to insure continuity of
+    the central gland in all three dimensions.}
 \label{fig:arfi_segs} 

### 2014-01-03 disc_arfi_mr.tex
added compression and decimation discussion

\subsection{ARFI and MR T2WI Agreement}
This study demonstrated that ARFI imaging delineates zonal anatomy of the
prostate well when compared to MR T2WI, with comparable estimates of total and
central gland volume (Figure~\ref{fig:mr_arfi_volumes}).  Establishing accurate
delineation of the prostate's zonal anatomy is important since PCa has strong
correlations with location in the prostate, specifically in the peripheral
zone, and accurate capsule delineation will facilitate future efforts to
register 3D imaging datasets between ARFI and MR imaging.

This was the first quantitative comparison between ARFI imaging and a newly
established clinical imaging modality to evaluate the prostate for PCa
detection and characterization, and we have drawn some useful conclusions in
how our zonal anatomy boundaries are determined during image segmentation.
Poor contrast between the PZ and peri-prostatic fat in B-mode images can make
delineation of the prostate boundary in the lateral dimension challenging,
emphasizing the need to use the orthogonal imaging plane views to help endure
organ continuity in all three dimensions.  Our study demonstrated an
overestimation of lateral boundaries of the prostate in our ultrasound datasets
compared to MR T2WI, leading to an overall overestimation of prostate total
gland volume.  We did not expect perfect agreement between MR and ARFI imaging
in this dimension since each modality applies a different amount of external
compression to the rectal wall that can deform the prostate different amounts.
Presure on the prostate during ARFI imaging is modulated by the urologist when
he locks the transducer rotation apparatus in place prior to imaging, and
deformation of the prostate can be easily visualized during this alignment
process.  It was our goal to apply as little pressure to the prostate during
imaging, while still maintaining uniform contact throughout all imaging angles,
which is critical since poor acoustic coupling to the prostate through the
rectal wall can greatly diminish ARFI image displacement SNR.

The anterior aspect of the prostate can also be challenging to delineate,
especially in large prostates, where SNR decreases with increasing depth away
from our rectal wall imaging surface and MR images suffer from gland
heterogeneity in this region~\cite{Gupta2013}, though we had good agreement
between ARFI and MR T2W images in the anterior-to-posterior dimension
(Figure~\ref{fig:mr_arfi_path_axes}(b), Table~\ref{tab:mr_arfi_axes_error}).

ARFI imaging consistently underestimated the extent of the prostate capsule and
CG along the apex-to-base axis relative to MR T2WI
(Figure~\ref{fig:mr_arfi_path_axes}(c), Table~\ref{tab:mr_arfi_axes_error}).
This was the dimension with the worst spatial resolution in MR imaging, with a
typical slice thickness in MR T2WI being $\sim$3 mm for these studies.  While
higher-resolution MR images could be obtained and analyzed for more precise
comparisons in this imaging dimension, such data was not available for this
study.  Additionally, it should also be noted that the precision in ARFI
imaging for this study is also constrained by the decimation used to reduce the
number of imaging planes manually segmented, introducing a lower-bound on the
model precision of, again, $\sim$3 mm.  While less decimation could improve the
precion of the prostate and GG axis measurements, this would have introduced
sigificant overhead in the manual segmentation process, with diminishing
returns given the spatial smoothing of the modeling algorithms that were
utilized (Table~\ref{tab:3dslicer}).

### 2014-01-03 discussion.tex
added compression and decimation discussion

--- 2013-10-07 discussion.tex
@@
 \section{Discussion}\label{sect:discussion}
-This study was designed to compared the zonal anatomy of the prostate in
-patients with biopsy-proven PCa as seen in ARFI imaging and MR T2WI.  Overall,
-ARFI imaging performed well when compared with MR estimates of total and
-central gland volume (Figure~\ref{fig:mr_arfi_volumes}), though ARFI tended to
-overestimate prostate total gland volume, most commonly due to over-estimation
-of the lateral dimension of the prostate.  This can be due to poor contrast
-between the PZ and peri-prostate fat (Figure~\ref{fig:arfi_segs}).  As seen in
-Figure~\ref{fig:arfi_segs}, clear delineation of the anterior aspect of the
-prostate can also be challenging in ARFI and B-mode images due to poor SNR,
-since this is the deepest aspect of the prostate relative to the rectal wall
-imaging surface. MR T2WI can also have difficulties visualizing and diagnosing
-lesions in the anterior region of the prostate,~\cite{Gupta2013}, which is
-thought to be secondary to the heterogeneity of the gland in that area,
-especially in the presence of BPH.  This was evident in imaging study subject
-4, with a $\sim$143 g prostate (Figure~\ref{fig:mr_arfi_weight}), which was an
-outlier in terms of weight and volume.  The most accurate measurements for both
-the total prostate gland and central gland were in the apex-to-base dimension
-(Figure~\ref{fig:mr_arfi_path_axes}(c), Table~\ref{tab:mr_arfi_axes_error}),
-with no mean bias in ARFI imaging estimates, but significant case-to-case
-variability, leading to poor linear regression coefficients of determination.
+\input{disc_arfi_mr}
 
-This study has several limitations that should be considered when interpreting
-these results.  Gross pathology weight and axis measurements could both be
-affected by the presence of periprostatic tissue that was excised during
-radical prostatectomy, especially in cases where more aggressive margins may
-have been necessary.  For this reason, unlike the image-to-image measurement
-comparisons, all of the image metrics presented relative to pathology metrics
-(Figure~\ref{fig:mr_arfi_weight}) were not characterized for absolute accuracy,
-but instead, relative correlations were evaluated.  Additionally, the volumes
-of the prostate from gross pathologic measurements were approximated as
-ellipsoids, which also introduced error, most likely an over-estimation of
-volume.  Interestingly, all pathology estimates were thought to have positive
-biases, but both imaging modalities tended to overestimate relative to the
-pathology measurements.
+\input{limitations}
 
-It should also be noted that all of the prostates in this study contained
-varying amounts of PCa, BPH and atrophy, all of which can distort the zonal
-anatomy, especially in the case of BPH and central gland morphology.  While
-younger, healthier prostates could have been targeted, these healthy organs
-would not have been excised for pathology characterization, and the zonal
-anatomy of a health (young) prostate is expected to be different from the
-prostate of a middle-age man, who is the target demographic for PCa screening
-imaging and PCa characterization; therefore, it was felt that performing this
-analysis in the presence of these confounding factors was appropriate.
-
-It should also be noted that this study did not evaluate user biases in image
-segmentation.  While MR zonal anatomy delineation has some establishment in the
-clinical literature, this study is the first attempt to define the criteria for
-ARFI imaging zonal anatomy characteristics, and it is expected that such
-delineations will continue to be refined as we acquire more cases and continue
-to compare with MRI and pathology data.  Given this limitation, no attempts
-were made to further quantify reader-to-reader variability in this work.
-
-Future directions for our research include deeper analysis of our findings,
-especially the lateral dimension over-estimation, to improve our anatomic
-delineation and assess further its clinical impact.  Technological improvements
-to improve anterior prostate boundary delineation in both B-mode and ARFI
-imaging are also being pursued and will hopefully allow for the currently
-elusive anterior PCa lesions in MR to be seen in ARFI imaging.  Although work
-remains to be done, the fact that one can view the different prostatic zones on
-ARFI images in addition to MRI is encouraging and could lead to targeted
-diagnostic biopsies and therapies with real-time imaging for men with PCa.
+\input{future}

### 2014-01-03 future.tex
added compression and decimation discussion

\subsection{Future Directions}
Future directions for our research include deeper analysis of our findings,
especially the lateral dimension over-estimation, to improve our anatomic
delineation and assess further its clinical impact.  Technological improvements
to improve anterior prostate boundary delineation in both B-mode and ARFI
imaging are also being pursued and will hopefully allow for the currently
elusive anterior PCa lesions in MR to be seen in ARFI imaging.  3D models of
the prostate and central glands will be used across ARFI and MR imaging
datasets to spatially register them to facilitate correlation of regions of PCa
suspicion between the two modalities, and to correlate with whole-mount
histology of the excised prostate specimens.  This effort will allow the
sensitivity and specificity of each imaging modality for PCa detection and
characterization to be quantified.  While ARFI imaging of the prostate in a
constantly evolving technology, the fact that one can view the different
prostatic zones on ARFI images in addition to MRI is encouraging and could lead
to targeted diagnostic biopsies and therapies with real-time imaging for men
with PCa.

### 2014-01-03 limitations.tex
added compression and decimation discussion

\subsection{Limitations}
This study has several limitations that should be considered when interpreting
these results.  Gross pathology weight and axis measurements could both be
affected by the presence of peri-prostatic tissue that was excised during
radical prostatectomy, especially in cases where more aggressive margins may
have been necessary.  For this reason, unlike the image-to-image measurement
comparisons, all of the image metrics presented relative to pathology metrics
(Figure~\ref{fig:mr_arfi_weight}) were not characterized for absolute accuracy,
but instead, relative correlations were evaluated.  Additionally, the volumes
of the prostate from gross pathologic measurements were approximated as
ellipsoids, which also introduced error, most likely an over-estimation of
volume.  Interestingly, all pathology estimates were thought to have positive
biases, but both imaging modalities tended to overestimate volume relative to
the pathology measurements.

It should also be noted that all of the prostates in this study contained
varying amounts of PCa, BPH and atrophy, all of which can distort the zonal
anatomy, especially in the case of BPH and central gland morphology.  While
younger, healthier prostates could have been targeted, these healthy organs
would not have been excised for pathology characterization, and the zonal
anatomy of a healthy (young) prostate is expected to be different from the
prostate of a middle-age man, who is the target demographic for PCa screening
imaging and PCa characterization.

This study did not evaluate user biases in image segmentation.  While MR zonal
anatomy delineation has some establishment in the clinical literature, this
study is the first attempt to define the criteria for ARFI imaging zonal
anatomy characteristics, and it is expected that such delineations will
continue to be refined as we acquire more cases and continue to compare with
MRI and pathology data.  Given this limitation, no attempts were made to
further quantify reader-to-reader variability in this work.

### 2014-01-03 arfi_methods.tex
reminder to include text and figure about rotation setup
more notes to myself in the TO DO
new text regarding prostate probe rotation, etc.

@@
 
+3D volumetric imaging data was acquired using mechanical rotation of the ER7B
+transducer between sequential imaging frames, sweeping across the lateral
+extent of the prostate.  This rotation setup utilized a CIVCO
+Micro-Touch\texttrademark stabalizer (CIVCO Medical Solutions, Kalona, IA USA)
+that allows for complete 6-axis degrees of freedom for manual positioning of
+the transducer to sweep through the entire prostate during imaging.  Sequential
+imaging frames were $\sim$ 1-2\degree apart, depending on the absolute size of the
+prostate.  A custom optical angular feedback transduction circuit utilizing a
+reflective linear strip with 500 lines-per-inch (LPI) resolution (US Digital,
+Vancouver, WA, USA) was coupled with a 141 oz-in torque stepper motor with
+planetary gearbox (Model \#11YPG202S-LW4-R27, Anaheim Automation, Anaheim, CA,
+USA) (Figure~\ref{fig:setup_annotated}) to achieve accurate spatial
+localization of the imaging frames in the 3D dataset.
+
+\input{fig_setup_annotated}
+
 Displacement estimation was performed using a phase-shift estimator on the

### 2014-01-03 fig_setup_annotated.tex
new annotated setup figure

\begin{figure}[htb!]
\centering
\includegraphics[width=0.75\textwidth]{figs/setup_annotated.png}
\caption{The experimental setup used for ARFI imaging in the operating suite,
    including the Siemens SC2000 scanner with the ER7B endorectal transducer
    sitting in a custom cradle of a CIVCO Micro-Touch\texttrademark stabalizer
    with optical feedback controlling a stepper motor to acquire 3D ultrasound
    imaging data \invivo.  Custom-written Python code was run on the scanner to
    trigger B-mode and ARFI imaging acquisition sequences relative to completed
    rotation increments of $\sim$ 1-2\degree.}
\label{fig:setup_annotated} 
\end{figure}

### 2014-01-06 disc_arfi_mr.tex
disc_arfi_mr.tex edited online with Bitbucket

@@
 This study demonstrated that ARFI imaging delineates zonal anatomy of the
-prostate well when compared to MR T2WI, with comparable estimates of total and
-central gland volume (Figure~\ref{fig:mr_arfi_volumes}).  Establishing accurate
-delineation of the prostate's zonal anatomy is important since PCa has strong
-correlations with location in the prostate, specifically in the peripheral
-zone, and accurate capsule delineation will facilitate future efforts to
-register 3D imaging datasets between ARFI and MR imaging.
+prostate well when compared to MR T2WI with only slight overestimation of 
+ARFI total and central gland volume (Figure~\ref{fig:mr_arfi_volumes}). 
+ Establishing accurate delineation of the prostate's zonal anatomy is 
+important because PCa has strong correlations with location in the prostate, 
+specifically in the peripheral zone, and accurate capsule delineation will f
+acilitate future efforts to register 3D imaging datasets between ARFI and MR imaging.
 
-This was the first quantitative comparison between ARFI imaging and a newly
-established clinical imaging modality to evaluate the prostate for PCa
-detection and characterization, and we have drawn some useful conclusions in
-how our zonal anatomy boundaries are determined during image segmentation.
-Poor contrast between the PZ and peri-prostatic fat in B-mode images can make
-delineation of the prostate boundary in the lateral dimension challenging,
-emphasizing the need to use the orthogonal imaging plane views to help endure
-organ continuity in all three dimensions.  Our study demonstrated an
-overestimation of lateral boundaries of the prostate in our ultrasound datasets
-compared to MR T2WI, leading to an overall overestimation of prostate total
-gland volume.  We did not expect perfect agreement between MR and ARFI imaging
-in this dimension since each modality applies a different amount of external
-compression to the rectal wall that can deform the prostate different amounts.
-Presure on the prostate during ARFI imaging is modulated by the urologist when
-he locks the transducer rotation apparatus in place prior to imaging, and
-deformation of the prostate can be easily visualized during this alignment
-process.  It was our goal to apply as little pressure to the prostate during
-imaging, while still maintaining uniform contact throughout all imaging angles,
-which is critical since poor acoustic coupling to the prostate through the
-rectal wall can greatly diminish ARFI image displacement SNR.
+We did not expect perfect agreement between MR and ARFI imaging by volume or tri-axial
+ dimensions since each modality applies a different amount of external compression to 
+the rectal wall that deforms the prostate different amounts. Pressure on the prostate 
+during ARFI imaging is modulated by the urologist when he locks the transducer rotation 
+apparatus in place prior to imaging, and deformation of the prostate can be easily visualized
+ during this alignment process.  It was our goal to apply as little pressure while still maintaining
+ uniform contact throughout the image acquisition. This is critical since poor acoustic coupling to 
+the prostate through the rectal wall can greatly diminish ARFI image displacement SNR.
+
+The lateral boundaries of the prostate in the U/S dataset were overestimated relative to MR T2WI, increasing 
+ARFI total volume relative to MR. The lateral overestimation was a result of poor contrast between the PZ and 
+peri-prostatic fat in the B-mode images, and image decimation along the lateral dimension in both ARFI and B-mode 
+images. Without image decimation, the elevation and axial orthogonal planes can be used to correct errors introduced
+ by the poor contrast between the PZ and peri-prostatic fat. However, with image decimation, segmentation changes in 
+the elevation and axial orthogonal planes were limited to the lower bound on the model precision of, again, $\sim$3 mm.
+ While reduced decimation could improve the precision of the lateral measurements, this would have introduced significant 
+overhead in the manual segmentation process, with diminishing returns given the spatial smoothing of the modeling algorithms
+ that were utilized.
 
@@
 ARFI imaging consistently underestimated the extent of the prostate capsule and 
-CG along the apex-to-base axis relative to MR T2WI
-(Figure~\ref{fig:mr_arfi_path_axes}(c), Table~\ref{tab:mr_arfi_axes_error}).
-This was the dimension with the worst spatial resolution in MR imaging, with a
-typical slice thickness in MR T2WI being $\sim$3 mm for these studies.  While
+central gland apex-to-base dimension relative to MR T2WI ((Figure~\ref{fig:mr_arfi_path_axes}(c), 
+Table~\ref{tab:mr_arfi_axes_error}).  In fact, it is likely the degree of ARFI estimation relative 
+to the actual apex-to-base prostate length is even greater than suggested by the MR comparison as this 
+was the dimension with the worst spatial resolution in MR imaging, with a
+typical MR T2WI slice thickness of $\sim$3 mm for these studies.  While
 higher-resolution MR images could be obtained and analyzed for more precise
 comparisons in this imaging dimension, such data was not available for this
-study.  Additionally, it should also be noted that the precision in ARFI
-imaging for this study is also constrained by the decimation used to reduce the
-number of imaging planes manually segmented, introducing a lower-bound on the
-model precision of, again, $\sim$3 mm.  While less decimation could improve the
-precion of the prostate and GG axis measurements, this would have introduced
-sigificant overhead in the manual segmentation process, with diminishing
-returns given the spatial smoothing of the modeling algorithms that were
-utilized (Table~\ref{tab:3dslicer}).
+study.

### 2014-01-07 figs/corr_weight_vol.m
utilizing set_axes()

@@ -54,18 +54,7 @@ function gen_fig_n_fits(fid,w,mr_total_vol,arfi_total_vol,metric)
         xlabel('Prostate Pathology Volume (cm^3)','FontSize',fs);
     end;
 
-    set(gca, ...
-        'Box'         , 'off'     , ...
-        'TickDir'     , 'out'     , ...
-        'TickLength'  , [.01 .01] , ...
-        'XMinorTick'  , 'off'      , ...
-        'YMinorTick'  , 'off'      , ...
-        'XGrid'       , 'off'      , ...
-        'XColor'      , [0 0 0], ...
-        'YColor'      , [0 0 0], ...
-        'XMinorGrid'  , 'off'      , ...
-        'LineWidth'   , 2, ...
-        'FontSize'    , fs);
+    set_axes(gca,fs);
 
     % lets do a linear regression and see what we get
     [mr_fit,mr_Rsq]=compute_linreg_Rsq(w,mr_total_vol);

### 2014-01-07 fig_mr_arfi_weight.tex
path : imaging linear regressions with >80g prostate indicated as an outlier and not included in the analysis

@@
     \weightMRrsq), while volumes reconstructed from ARFI images (green, O)
-    showed weaker correlation (R$^2$ = \weightARFIrsq) (b).  Even weaker
-    correlations existed between both T2WI MR and ARFI image volumens and
-    approaximated ellipsoidal prostate pathology volumes (R$^2$ = \pathVolMRrsq
-    and \pathVolARFIrsq, respectively) (c).  It should be noted in these
-    figures that two study subjects had excessively large prostates that were
-    difficult to fully capture in imaging (volumes $>$ 60 cm$^3$), especially
-    in their anterior region, and they were, therefore, grossly underestimated
-    in size.}
+    showed weaker correlation (R$^2$ = \weightARFIrsq) (b).  Weaker
+    correlations existed between both T2WI MR and ARFI image volumes and
+    approximated ellipsoidal prostate pathology volumes (R$^2$ = \pathVolMRrsq
+    and \pathVolARFIrsq, respectively) (c).  One study subject had a very large
+    prostate ($>$ 80 g) that was not well visualized by both MR and ARFI
+    imaging, and it was excluded from all of the linear regression.  This
+    outlier was indicated with magenta circles in the plots.}
 \label{fig:mr_arfi_weight}

### 2014-01-07 figs/corr_weight_vol.m

@@ -37,12 +37,24 @@ function gen_fig_n_fits(fid,w,mr_total_vol,arfi_total_vol,metric)
     fs = 18;
     figure;
     hold on;
+
+    % circle the outlier(s), > 80g path weight
+    gt80g = find(w > 80);
+
     if(strcmp(metric,'path_vol_weight')),
         plot(w,mr_total_vol,'bx','MarkerSize',14,'LineWidth',3);
+        % circle the outlier
+        plot(w(gt80g),mr_total_vol(gt80g),'o','MarkerSize',32,'LineWidth',2,'Color','m');
     else,
         plot(w,mr_total_vol,'bx','MarkerSize',14,'LineWidth',3);
         plot(w,arfi_total_vol,'go','MarkerSize',14,'LineWidth',3);
+        % circle the outliers
+        plot(w(gt80g),mr_total_vol(gt80g),'o','MarkerSize',32,'LineWidth',2,'Color','m');
+        plot(w(gt80g),arfi_total_vol(gt80g),'o','MarkerSize',32,'LineWidth',2,'Color','m');
     end;
+
+
+
     if(strcmp(metric,'path_vol_weight')),
         ylabel('Pathology Ellipsoidal Volume (cm^3)','FontSize',fs)
     else,
@@ -57,15 +69,17 @@ function gen_fig_n_fits(fid,w,mr_total_vol,arfi_total_vol,metric)
     set_axes(gca,fs);
 
     % lets do a linear regression and see what we get
-    [mr_fit,mr_Rsq]=compute_linreg_Rsq(w,mr_total_vol);
 
-    plot(w,mr_fit,'-b','LineWidth',3);
+    % first we need to remove the prostate > 80 g
+    less80g = find(w < 80);
+
+    [mr_fit,mr_Rsq]=compute_linreg_Rsq(w(less80g),mr_total_vol(less80g));
+
+    plot(w(less80g),mr_fit,'-b','LineWidth',3);
 
     if(~strcmp(metric,'path_vol_weight')),
-        [arfi_fit,arfi_Rsq]=compute_linreg_Rsq(w,arfi_total_vol);
-
-        plot(w,arfi_fit,'-g','LineWidth',3);
-
+        [arfi_fit,arfi_Rsq]=compute_linreg_Rsq(w(less80g),arfi_total_vol(less80g));
+        plot(w(less80g),arfi_fit,'-g','LineWidth',3);
[…]

### 2014-01-07 disc_arfi_mr.tex

@@
 
-We did not expect perfect agreement between MR and ARFI imaging by volume or tri-axial
+We did not expect perfect agreement between MR and ARFI imaging volume or tri-axial
  dimensions since each modality applies a different amount of external compression to 

### 2014-01-14 arfi_methods.tex
updated text formatting

@@
 the transducer to sweep through the entire prostate during imaging.  Sequential
-imaging frames were $\sim$ 1-2\degree apart, depending on the absolute size of the
+imaging frames were $\sim$1\degree apart, depending on the absolute size of the
 prostate.  A custom optical angular feedback transduction circuit utilizing a
@@
 low-pass filtered with a cutoff-frequency of 0.8 mm$^{-1}$.
+
+The normalized displacement data were scan-converted using a Delaunay
+triangulation method with XXXX commaind in Matlab (R2012b, XXXX).  The
+scan-converted ARFI imaging data had a final voxel size of XX x XX x XX mm.

### 2014-01-14 fig_setup_annotated.tex
updated text formatting

@@
     trigger B-mode and ARFI imaging acquisition sequences relative to completed
-    rotation increments of $\sim$ 1-2\degree.}
+    rotation increments of $\sim$1\degree.}
 \label{fig:setup_annotated} 

### 2014-01-20 arfi_methods.tex
draft edits and spelling

--- 2014-01-03 arfi_methods.tex
@@
 mm, 22.5 mm, and 15 mm, respectively, and finally the response of the tissue
-was tracked for up to 6ms at a PRF of 8kHz. This pushing strategy is similar to
+was tracked for up to 6 ms at a PRF of 8kHz. This pushing strategy is similar to
 what has been published by Bercoff \etal~\cite{Bercoff2004}. The 30 mm and 22.5
@@
 extent of the prostate.  This rotation setup utilized a CIVCO
-Micro-Touch\texttrademark stabalizer (CIVCO Medical Solutions, Kalona, IA USA)
+Micro-Touch\texttrademark stabilizer (CIVCO Medical Solutions, Kalona, IA USA)
 that allows for complete 6-axis degrees of freedom for manual positioning of
 the transducer to sweep through the entire prostate during imaging.  Sequential
-imaging frames were $\sim$ 1-2\degree apart, depending on the absolute size of the
+imaging frames were $\sim$1\degree apart, depending on the absolute size of the
 prostate.  A custom optical angular feedback transduction circuit utilizing a
@@
 low-pass filtered with a cutoff-frequency of 0.8 mm$^{-1}$.
+
+The normalized displacement data were scan-converted using a 3D Delaunay
+triangulation interpolation method in Matlab (R2012b, The MathWorks, Inc,
+Natick, MA, USA).  The scan-converted ARFI imaging data had a final isotropic
+voxel size of 0.15 $\times$ 0.15 $\times$ 0.15 mm$^3$.

### 2014-01-20 disc_arfi_mr.tex
updated text formatting
draft edits and spelling

@@
 This study demonstrated that ARFI imaging delineates zonal anatomy of the
-prostate well when compared to MR T2WI with only slight overestimation of 
-ARFI total and central gland volume (Figure~\ref{fig:mr_arfi_volumes}). 
- Establishing accurate delineation of the prostate's zonal anatomy is 
-important because PCa has strong correlations with location in the prostate, 
-specifically in the peripheral zone, and accurate capsule delineation will f
-acilitate future efforts to register 3D imaging datasets between ARFI and MR imaging.
+prostate well when compared to MR T2WI with only slight overestimation of ARFI
+total and central gland volume (Figure~\ref{fig:mr_arfi_volumes}).
+Establishing accurate delineation of the prostate's zonal anatomy is important
+because PCa has strong correlations with location in the prostate, specifically
+in the peripheral zone, and accurate capsule delineation will facilitate
+future efforts to register 3D imaging datasets between ARFI and MR imaging.
 
-We did not expect perfect agreement between MR and ARFI imaging volume or tri-axial
- dimensions since each modality applies a different amount of external compression to 
-the rectal wall that deforms the prostate different amounts. Pressure on the prostate 
-during ARFI imaging is modulated by the urologist when he locks the transducer rotation 
-apparatus in place prior to imaging, and deformation of the prostate can be easily visualized
- during this alignment process.  It was our goal to apply as little pressure while still maintaining
- uniform contact throughout the image acquisition. This is critical since poor acoustic coupling to 
-the prostate through the rectal wall can greatly diminish ARFI image displacement SNR.
+We did not expect perfect agreement between MR and ARFI imaging volume or
+tri-axial dimensions since each modality applies a different amount of external
+compression to the rectal wall, which then deforms the prostate.  Pressure on
+the prostate during ARFI imaging is modulated by the urologist when he locks
+the transducer rotation apparatus in place prior to imaging, and deformation of
+the prostate can be easily visualized during this alignment process.  It was
+our goal to apply as little pressure while still maintaining uniform contact
+throughout the image acquisition. This is critical since poor acoustic coupling
+to the prostate through the rectal wall can greatly diminish ARFI image
+displacement SNR.
 
-The lateral boundaries of the prostate in the U/S dataset were overestimated relative to MR T2WI, increasing 
-ARFI total volume relative to MR. The lateral overestimation was a result of poor contrast between the PZ and 
-peri-prostatic fat in the B-mode images, and image decimation along the lateral dimension in both ARFI and B-mode 
-images. Without image decimation, the elevation and axial orthogonal planes can be used to correct errors introduced
- by the poor contrast between the PZ and peri-prostatic fat. However, with image decimation, segmentation changes in 
-the elevation and axial orthogonal planes were limited to the lower bound on the model precision of, again, $\sim$3 mm.
- While reduced decimation could improve the precision of the lateral measurements, this would have introduced significant 
-overhead in the manual segmentation process, with diminishing returns given the spatial smoothing of the modeling algorithms
- that were utilized.
+The lateral boundaries of the prostate in the U/S dataset were overestimated
+relative to MR T2WI, increasing ARFI total volume relative to MR. The lateral
+overestimation was a result of poor contrast between the PZ and peri-prostatic
+fat in the B-mode images, and image decimation along the lateral dimension in
+both ARFI and B-mode images. Without image decimation, the elevation and axial
+orthogonal planes can be used to correct errors introduced by the poor contrast
+between the PZ and peri-prostatic fat. However, with image decimation,
+segmentation changes in the elevation and axial orthogonal planes were limited
+to the lower bound on the model precision of, again, $\sim$3 mm.  While reduced
+decimation could improve the precision of the lateral measurements, this would
+have introduced significant overhead in the manual segmentation process, with
+diminishing returns given the spatial smoothing of the modeling algorithms that
+were utilized.
 
@@
 ARFI imaging consistently underestimated the extent of the prostate capsule and
-central gland apex-to-base dimension relative to MR T2WI ((Figure~\ref{fig:mr_arfi_path_axes}(c), 
-Table~\ref{tab:mr_arfi_axes_error}).  In fact, it is likely the degree of ARFI estimation relative 
-to the actual apex-to-base prostate length is even greater than suggested by the MR comparison as this 
-was the dimension with the worst spatial resolution in MR imaging, with a
-typical MR T2WI slice thickness of $\sim$3 mm for these studies.  While
-higher-resolution MR images could be obtained and analyzed for more precise
-comparisons in this imaging dimension, such data was not available for this
-study.
+central gland apex-to-base dimension relative to MR T2WI
+((Figure~\ref{fig:mr_arfi_path_axes}(c), Table~\ref{tab:mr_arfi_axes_error}).
+In fact, it is likely the degree of ARFI estimation relative to the actual
+apex-to-base prostate length is even greater than suggested by the MR
+comparison as this was the dimension with the worst spatial resolution in MR
+imaging, with a typical MR T2WI slice thickness of $\sim$3 mm for these
+studies.  While higher-resolution MR images could be obtained and analyzed for
+more precise comparisons in this imaging dimension, such data were not available
+for this study.

### 2014-01-20 fig_arfi_segs.tex
draft edits and spelling

@@
 \includegraphics[width=0.3\textwidth]{zach/Cap_Modeling_Images/Sagittal_Cap.png} \\
-Axial B-mode Outline & Coronal B-mode Outline & Sagital B-mode Outline \\
+Axial B-mode Outline & Coronal B-mode Outline & Sagittal B-mode Outline \\
 \end{tabular}
@@
 \includegraphics[width=0.3\textwidth]{zach/CG_Modeling_Images/Sagittal_CG.png} \\
-Axial ARFI Outline & Coronal ARFI Outline & Sagital ARFI Outline \\
+Axial ARFI Outline & Coronal ARFI Outline & Sagittal ARFI Outline \\
 \end{tabular}
@@
     the B-mode images, with the 3D model outlines superimposed on the central
-    axial, coronal and sagital B-mode images.  The primary segmentation plane
+    axial, coronal and sagittal B-mode images.  The primary segmentation plane
     was sagittal; however the axial and coronal planes were used to guide the
@@
     segmentation of the ARFI images, with the 3D model outlines superimposed on
-    the central axial, coronal and sagital ARFI images.  The coronal ARFI
+    the central axial, coronal and sagittal ARFI images.  The coronal ARFI
     imaging plane was the primary orientation used for image segmentation, but

### 2014-01-20 fig_mr_anatomy.tex
draft edits and spelling

@@
     central zone and the transitional zone. The apex (A) is composed mostly of
-    PZ glandular tissue and the urethra is seen at the level of the midgland as
+    PZ glandular tissue and the urethra is seen at the level of the mid-gland as
     an inverted ``U'' (b). Note the area of hypointense signal in the
-    peripheral zone at the midgland and base (asterisk, b and c), which
+    peripheral zone at the mid-gland and base (asterisk, b and c), which
     represents a prostatic tumor.  The posterior (P) aspect of the prostate is
     adjacent to the endorectal coil, and the right (R)-to-left (L) extent of
-    the prostate is refered to as the lateral-to-lateral axis in the subsequent
+    the prostate is referred to as the lateral-to-lateral axis in the subsequent
     analysis.}

### 2014-01-20 fig_mr_arfi_weight.tex
draft edits and spelling

@@
     prostate volume approximation based on gross pathology axis measurements,
-    which was moderately well-correlated with the excised prostated weights (a,
+    which was moderately well-correlated with the excised prostate weights (a,
     R$^2$ = \pathVolWeightRsq).  T2WI MR (blue, X) showed a moderate
@@
     approximated ellipsoidal prostate pathology volumes (R$^2$ = \pathVolMRrsq
-    and \pathVolARFIrsq, respectively) (c).  One study subject had a very large
-    prostate ($>$ 80 g) that was not well visualized by both MR and ARFI
-    imaging, and it was excluded from all of the linear regression.  This
-    outlier was indicated with magenta circles in the plots.}
+    and \pathVolARFIrsq, respectively) (c).  There was on pathology specimen
+    that had an erroneous prostate volume and weight recorded, which has been
+    excluded from the linear regressions (magenta circles).}
 \label{fig:mr_arfi_weight}

### 2014-01-20 fig_setup_annotated.tex
draft edits and spelling

@@
     including the Siemens SC2000 scanner with the ER7B endorectal transducer
-    sitting in a custom cradle of a CIVCO Micro-Touch\texttrademark stabalizer
+    sitting in a custom cradle of a CIVCO Micro-Touch\texttrademark stabilizer
     with optical feedback controlling a stepper motor to acquire 3D ultrasound

### 2014-01-20 image_proc_methods.tex
draft edits and spelling

@@
 ITK-SNAP~\cite{Yushkevich2006}, using unique labels for the PZ, CG and AFS. The
-gland was segmented from base to apex.  The base was identified below the
+gland was segmented from base-to-apex.  The base was identified below the
 bladder, and subsequent images were segmented until the last slice with visible
@@
 identified by its homogeneous high signal intensity on T2WI, which is usually
-similar to that of the nearby periprostatic fat. The CG was visualized and
+similar to that of the nearby peri-prostatic fat. The CG was visualized and
 delineated based on its heterogeneous, lower signal intensity and its location
@@
 estimates. Orthogonal tri-axial measurements in the lateral-to-lateral,
-apex-to-base, and anterior-to-posterior dimensions were made in slicer by
+apex-to-base, and anterior-to-posterior dimensions were made in 3D Slicer by
 placing an ROI cube in the model's center and expanding the ROI cube to fit the
@@
 using the \verb+draweffect+ tool to model and measure the prostate capsule and
-CG volume and tri-axial dimensions.  The sagittal imaging plane was chosen for
-segmentation because it has a greater number of easily identifiable anatomical
-structures:  seminal vesicles in the left and right posterior base, urethra in
-the mid apex, surrounding fibromuscular tissue in the anterior , and rectum in
-the posterior  than either the axial or coronal views. Prostate capsule was
-segmented in B-mode as this modality offers both high contrast between
-prostatic tissue and surrounding periprostatic fat, and better SNR in the
-anterior capsule region than ARFI images. ARFI images were used for CG
-segmentation.  Prior to segmentation, both the B-mode and ARFI image stacks
-were down-sampled along the sagittal axis by a factor of 20 (i.e., from 360
-slices per image stack to 18 slices per image stack) to reduce the number of
-manually-segmented frames.
+CG volume and tri-axial dimensions (Figure~\ref{fig:arfi_segs}).  The sagittal
+imaging plane was chosen for segmentation because it has a greater number of
+easily identifiable anatomical structures:  seminal vesicles in the left and
+right posterior base, urethra in the mid apex, surrounding fibromuscular tissue
+in the anterior, and rectum in the posterior. Prostate capsule was segmented in
+B-mode as this modality offers both high contrast between prostatic tissue and
+surrounding peri-prostatic fat, and better SNR in the anterior capsule region
+than ARFI images. ARFI images were used for CG segmentation.  Prior to
+segmentation, both the B-mode and ARFI image stacks were down-sampled along the
+sagittal axis by a factor of 20 (i.e., from 360 slices per image stack to 18
+slices per image stack) to reduce the number of manually-segmented frames.
 

### 2014-01-20 intro.tex
draft edits and spelling

@@
 assessment.  Additionally, the random nature of transrectal ultrasound
-(TRUS)-guided biopsies can miss or underestimate the burden of
-cancer.~\cite{Gupta2013} Earlier disease diagnosis leads to challenges in
+(TRUS)-guided biopsies can miss or underestimate the cancer
+burden.~\cite{Gupta2013} Earlier disease diagnosis leads to challenges in
 deciding optimal management strategies for patients presenting with less
@@
 currently used in the study of PCa, it is well established that T2-Weighted
-Imaging (T2WI) offers the best assessment of prostate anatomy due to its
+Imaging (T2WI) offers the best assessment of prostate anatomy based on its
 ability to delineate prostatic margins, distinguish internal structures and

### 2014-01-20 prostate_anatomy.tex
draft edits and spelling

@@
 healthy prostate, each zone contains approximately 5\%, 20\% and 70--80\% of
-glandular tissue, respectively.~\cite{Bonekamp2011}, but these ratios can
+glandular tissue, respectively~\cite{Bonekamp2011}, but these ratios can
 change significantly in the presence of benign prostatic hyperplasia (BPH). The
@@
 cannot be visualized as discrete entities, and they are collectively identified
-as the central gland (CG).  This CG was one of the prostate structures manually
+as the central gland (CG).  The CG was one of the prostate structures manually
 segmented in MR and ARFI images in this study.

### 2014-01-20 results.tex
draft edits and spelling

@@
 
-Measurements of the prostate total and CG dimensions along the three standard
-anatomic axes (apex-to-base, lateral-to-lateral, and anterior-to-posterior)
-were made (Table~\ref{tab:mr_arfi_axes}), and the correlation between the
-imaging axis measurements was analyzed (Figure~\ref{fig:mr_arfi_path_axes}).
-ARFI was most correlated to MR in the lateral-to-lateral axis in both the total
-and CG (R$^2$ = \totalLatLatRsq~and \centralLatLatRsq, respectively), with mean
-overestimates of \ARFImrTotalLatLatMeanPct~$\pm$ \ARFImrTotalLatLatStdPct\%~and
+Measurements of the prostate total gland and CG dimensions along the three
+standard anatomic axes (apex-to-base, lateral-to-lateral, and
+anterior-to-posterior) were made (Table~\ref{tab:mr_arfi_axes}), and the
+correlation between the imaging axis measurements was analyzed
+(Figure~\ref{fig:mr_arfi_path_axes}).  ARFI was most correlated to MR in the
+lateral-to-lateral axis in both the total gland and CG (R$^2$ =
+\totalLatLatRsq~and \centralLatLatRsq, respectively), with mean overestimates
+of \ARFImrTotalLatLatMeanPct~$\pm$ \ARFImrTotalLatLatStdPct\%~and
 \ARFImrCentralLatLatMeanPct~$\pm$ \ARFImrCentralLatLatStdPct\%, respectively
@@
 had weak correlation with MR images in the apex-to-base dimension (R$^2$ =
-\totalApexBaseRsq, total gland and R$^2$ = \centralApexBaseRsq, CG),
-with underestimates of \ARFImrTotalApexBaseMeanPct~$\pm$
+\totalApexBaseRsq, total gland and R$^2$ = \centralApexBaseRsq, CG), with
+underestimates of \ARFImrTotalApexBaseMeanPct~$\pm$
 \ARFImrTotalApexBaseStdPct\% and \ARFImrCentralApexBaseMeanPct~$\pm$

### 2014-01-20 tab_axis_ratio_over_under.tex
draft edits and spelling

@@
     imaging modalities.  The three axis ratios analyzed were:
-    lateral-to-lateral : apex-to-base (LL:AB), anterior-to-posterio :
+    lateral-to-lateral : apex-to-base (LL:AB), anterior-to-posterior :
     apex-to-base (AP:AB), and lateral-to-lateral : anterior-to-posterior

### 2014-01-20 compile-n-view.sh

@@ -1,1 +1,1 @@
-latexmk -bibtex -pdf -pvc mr_arfi_anatomy.tex
+latexmk -bibtex -pdf -pvc -outdir=/tmp/latex mr_arfi_anatomy.tex

### 2014-01-21 abstract.tex
abstract.tex edited online with Bitbucket

--- 2013-12-08 abstract.tex
@@
 the United States and the second leading cause of cancer-related death.
-Multi-parametric Magnetic Resonance Imaging (mpMRI) has gained recent
-popularity to characterize PCa.  ARFI imaging has the potential to aid in PCa
+Two imaging modalities, Multi-parametric Magnetic Resonance Imaging (mpMRI) and
+Acoustic Radiation Force Imaging have the potential to aid in PCa
 diagnosis and management by evaluating the structural composition of prostate
-zones and tumors based on their stiffness.  In this study, MR and ARFI imaging
-datasets were compared to one another and with gross pathology measurements
-made immediately post radical prostatectomy.  Images were manually segmented to
-delineate the central gland (CG) and prostate capsule, and 3D models were
-rendered to evaluate zonal anatomy dimensions and volumes.  Both imaging
-modalities showed moderate correlations between estimated organ volume and
-gross pathologic weights.  ARFI and MR total prostate gland volumes were
-well-correlated (R$^2$ = 0.63), but ARFI images yielded prostate volumes that
-were, on average, larger (6.1\% $\pm$ 25\%) than MR images, primarily due to
-over-estimation of the lateral dimension of the total gland
-(\ARFImrTotalLatLatMeanPct $\pm$ \ARFImrTotalLatLatStdPct\%), while
-over-estimates of the other dimensions were less significant
-(\ARFImrTotalAntPostMeanPct $\pm$ \ARFImrTotalAntPostStdPct\% and
-\ARFImrTotalApexBaseMeanPct $\pm$ \ARFImrTotalApexBaseStdPct\%).  The ARFI and
-MR CG volumes were also moderately correlated (R$^2$ = 0.38).  CG volume
-differences were attributed to under-estimation of the apex-to-base axis
+zones and tumors based on water content and stiffness respectively.  In this study, 
+MR and ARFI imaging datasets were compared to one another and against gross pathology.  
+MR and ARFI images were manually segmented in 3D Slicer to delineate the central gland (CG) and prostate capsule. 3D
+models were then rendered in 3D Slicer to evaluate zonal anatomy tri-axial dimensions and
+volumes.  Both imaging modalities showed moderate correlations between
+estimated organ volume and gross pathologic weights.  In terms of intra-modality comparison,
+ARFI and MR total prostate gland volumes were well-correlated (R$^2$ = \MRarfiVolTotalRsq), but
+ARFI yielded prostate volumes that were, on average, larger
+(\MRarfiVolTotalMeanDiff\% $\pm$ \MRarfiVolTotalStdDiff\%) than MR. This was
+primarily due to total gland over-estimation along the lateral dimension.
+(\ARFImrTotalLatLatMeanPct $\pm$ \ARFImrTotalLatLatStdPct\%). Differences between MR and
+ARFI in the other dimensions were less significant (\ARFImrTotalAntPostMeanPct $\pm$
+\ARFImrTotalAntPostStdPct\% and \ARFImrTotalApexBaseMeanPct $\pm$
+\ARFImrTotalApexBaseStdPct\%). ARFI and MR CG volumes were also
+well-correlated (R$^2$ = \MRarfiVolCentralRsq).  CG volume differences were
+attributed to apex-to-base axis underestimation by MR
 (\ARFImrCentralApexBaseMeanPct $\pm$ \ARFImrCentralApexBaseStdPct\%), and
-over-estimation bias in the lateral dimension (\ARFImrCentralLatLatMeanPct
-$\pm$ \ARFImrCentralLatLatStdPct\%).  ARFI prostate imaging yielded prostate
+lateral dimension over-estimation by ARFI (\ARFImrCentralLatLatMeanPct
+$\pm$ \ARFImrCentralLatLatStdPct\%).  Overall, ARFI prostate imaging yielded prostate
 volumes and dimensions that were correlated with MR T2WI estimates, with biases
-in the AP dimension, most likely related to poor displacement SNR in the
+in the lateral dimension, most likely related to poor displacement SNR in the
 anterior region of the prostate from greater distance from the rectal wall
-imaging surface.  Lateral dimension contrast was also challenging due to
+imaging surface. (I don't agree with this statement, our anterior to potsterior dimensions correlations were good)  Lateral dimension contrast was also challenging due to
 peri-prostatic fat.  ARFI imaging is a promising low-cost, real-time imaging
-modality that can compliment MR imaging for diagnosis, treatment planning and
+modality that can compliment mpMRI for diagnosis, treatment planning and
 management of PCa.

### 2014-01-21 compile-n-view.sh
compile in /tmp

@@ -1,1 +1,7 @@
-latexmk -bibtex -pdf -pvc -outdir=/tmp/latex mr_arfi_anatomy.tex
+tmp_latex = "/tmp/latex"
+
+if [ ! -d $tmp_latex ]; then
+    mkdir $tmp_latex
+fi
+
+latexmk -bibtex -pdf -pvc -outdir=$tmp_latex mr_arfi_anatomy.tex

### 2014-01-21 abstract.tex
abstract.tex edited online with Bitbucket

@@
 zones and tumors based on water content and stiffness respectively.  In this study, 
-MR and ARFI imaging datasets were compared to one another and against gross pathology.  
-MR and ARFI images were manually segmented in 3D Slicer to delineate the central gland (CG) and prostate capsule. 3D
-models were then rendered in 3D Slicer to evaluate zonal anatomy tri-axial dimensions and
-volumes.  Both imaging modalities showed moderate correlations between
-estimated organ volume and gross pathologic weights.  In terms of intra-modality comparison,
+MR and ARFI imaging datasets were compared to one another and against gross pathology in order to evaluate their
+ability to delineate central gland (CG) and prostate capsule anatomy. 3D slicer, an open source image processing program,  was used for MR and ARFI manual
+ for manual segmentation, 3d modeling, and measurement of zonal anatomy tri-axial dimensions and volumes.   
+ Both imaging modalities showed moderate correlations between
+modeled organ volume and gross pathologic weights.  In the intra-modality comparison,
 ARFI and MR total prostate gland volumes were well-correlated (R$^2$ = \MRarfiVolTotalRsq), but
@@
 well-correlated (R$^2$ = \MRarfiVolCentralRsq).  CG volume differences were
-attributed to apex-to-base axis underestimation by MR
+attributed to MR apex-to-base axis underestimation 
 (\ARFImrCentralApexBaseMeanPct $\pm$ \ARFImrCentralApexBaseStdPct\%), and
-lateral dimension over-estimation by ARFI (\ARFImrCentralLatLatMeanPct
+ARFI lateral dimension over-estimation  (\ARFImrCentralLatLatMeanPct
 $\pm$ \ARFImrCentralLatLatStdPct\%).  Overall, ARFI prostate imaging yielded prostate
 volumes and dimensions that were correlated with MR T2WI estimates, with biases
-in the lateral dimension, most likely related to poor displacement SNR in the
-anterior region of the prostate from greater distance from the rectal wall
-imaging surface. (I don't agree with this statement, our anterior to potsterior dimensions correlations were good)  Lateral dimension contrast was also challenging due to
-peri-prostatic fat.  ARFI imaging is a promising low-cost, real-time imaging
+in the lateral dimension. ARFI imaging is a promising low-cost, real-time imaging
 modality that can compliment mpMRI for diagnosis, treatment planning and

### 2014-01-21 abstract.tex
Kathy's corrections

--- 2013-12-09 abstract.tex
@@
 diagnosis and management by evaluating the structural composition of prostate
-zones and tumors based on their stiffness.  In this study, MR and ARFI imaging
-datasets were compared to one another and with gross pathology measurements
-made immediately post radical prostatectomy.  Images were manually segmented to
-delineate the central gland (CG) and prostate capsule, and 3D models were
-rendered to evaluate zonal anatomy dimensions and volumes.  Both imaging
-modalities showed moderate correlations between estimated organ volume and
-gross pathologic weights.  ARFI and MR total prostate gland volumes were
-well-correlated (R$^2$ = \MRarfiVolTotalRsq), but ARFI images yielded prostate
-volumes that were, on average, larger (\MRarfiVolTotalMeanDiff\% $\pm$
-\MRarfiVolTotalStdDiff\%) than MR images, primarily due to over-estimation of
-the lateral dimension of the total gland (\ARFImrTotalLatLatMeanPct $\pm$
-\ARFImrTotalLatLatStdPct\%), while over-estimates of the other dimensions were
-less significant (\ARFImrTotalAntPostMeanPct $\pm$ \ARFImrTotalAntPostStdPct\%
-and \ARFImrTotalApexBaseMeanPct $\pm$ \ARFImrTotalApexBaseStdPct\%).  The ARFI
-and MR CG volumes were also well-correlated (R$^2$ = \MRarfiVolCentralRsq).  CG
-volume differences were attributed to under-estimation of the apex-to-base axis
-(\ARFImrCentralApexBaseMeanPct $\pm$ \ARFImrCentralApexBaseStdPct\%), and
-over-estimation bias in the lateral dimension (\ARFImrCentralLatLatMeanPct
-$\pm$ \ARFImrCentralLatLatStdPct\%).  ARFI prostate imaging yielded prostate
-volumes and dimensions that were correlated with MR T2WI estimates, with biases
-in the AP dimension, most likely related to poor displacement SNR in the
-anterior region of the prostate from greater distance from the rectal wall
-imaging surface.  Lateral dimension contrast was also challenging due to
-peri-prostatic fat.  ARFI imaging is a promising low-cost, real-time imaging
-modality that can compliment MR imaging for diagnosis, treatment planning and
-management of PCa.
+zones and tumors based on their stiffness.  In this study, \invivo MR and ARFI
+imaging datasets were compared to one another and with gross pathology
+measurements made immediately post radical prostatectomy.  Images were manually
+segmented in 3D Slicer to delineate the central gland (CG) and prostate
+capsule, and 3D models were rendered in 3D Slicer to evaluate zonal anatomy
+dimensions and volumes.  Both imaging modalities showed moderate correlations
+between estimated organ volume and gross pathologic weights.  ARFI and MR total
+prostate gland volumes were well-correlated (R$^2$ = \MRarfiVolTotalRsq), but
+ARFI images yielded prostate volumes that were, on average, larger
+(\MRarfiVolTotalMeanDiff\% $\pm$ \MRarfiVolTotalStdDiff\%) than MR images,
+primarily due to over-estimation of the lateral dimension of the total gland
+(\ARFImrTotalLatLatMeanPct $\pm$ \ARFImrTotalLatLatStdPct\%), while differences
+in the other dimensions were less significant (\ARFImrTotalAntPostMeanPct $\pm$
+\ARFImrTotalAntPostStdPct\%, anterior-to-posterior, and
+\ARFImrTotalApexBaseMeanPct $\pm$ \ARFImrTotalApexBaseStdPct\%, apex-to-base).
+The ARFI and MR CG volumes were also well-correlated (R$^2$ =
+\MRarfiVolCentralRsq).  CG volume differences were attributed to
+under-estimation of the apex-to-base axis (\ARFImrCentralApexBaseMeanPct $\pm$
+\ARFImrCentralApexBaseStdPct\%), and over-estimation bias in the lateral
+dimension (\ARFImrCentralLatLatMeanPct $\pm$ \ARFImrCentralLatLatStdPct\%).
+ARFI prostate imaging yielded prostate volumes and dimensions that were
+correlated with MR T2WI estimates, with biases in the AP dimension, most likely
+related to poor displacement SNR in the anterior region of the prostate due to
+greater distance from the rectal wall imaging surface.  Lateral dimension
+contrast was also challenging due to peri-prostatic fat.  ARFI imaging is a
+promising low-cost, real-time imaging modality that can compliment mpMRI for
+diagnosis, treatment planning and management of PCa.

### 2014-01-21 abstract.tex
abstract.tex edited online with Bitbucket

--- 2014-01-21 abstract.tex
@@
 MR and ARFI imaging datasets were compared to one another and against gross pathology in order to evaluate their
-ability to delineate central gland (CG) and prostate capsule anatomy. 3D slicer, an open source image processing program,  was used for MR and ARFI manual
- for manual segmentation, 3d modeling, and measurement of zonal anatomy tri-axial dimensions and volumes.   
+ability to delineate central gland (CG) and prostate capsule anatomy. 3D slicer, an open source image processing program,  was used
+ for MR and ARFI  manual segmentation, 3d modeling, and measurement of zonal anatomy tri-axial dimensions and volumes.   
  Both imaging modalities showed moderate correlations between
@@
 ARFI yielded prostate volumes that were, on average, larger
-(\MRarfiVolTotalMeanDiff\% $\pm$ \MRarfiVolTotalStdDiff\%) than MR. This was
-primarily due to total gland over-estimation along the lateral dimension.
+(\MRarfiVolTotalMeanDiff\% $\pm$ \MRarfiVolTotalStdDiff\%) than MR. This volume bias was
+primarily due to ARFI lateral dimension over-estimation.
 (\ARFImrTotalLatLatMeanPct $\pm$ \ARFImrTotalLatLatStdPct\%). Differences between MR and

### 2014-01-21 abstract.tex

--- 2014-01-21 abstract.tex
@@
 Prostate cancer (PCa) is the most common non-cutaneous malignancy among men in
-the United States and the second leading cause of cancer-related death.
-Two imaging modalities, Multi-parametric Magnetic Resonance Imaging (mpMRI) and
-Acoustic Radiation Force Imaging have the potential to aid in PCa
-diagnosis and management by evaluating the structural composition of prostate
-zones and tumors based on water content and stiffness respectively.  In this study, 
-MR and ARFI imaging datasets were compared to one another and against gross pathology in order to evaluate their
-ability to delineate central gland (CG) and prostate capsule anatomy. 3D slicer, an open source image processing program,  was used for MR and ARFI manual
- for manual segmentation, 3d modeling, and measurement of zonal anatomy tri-axial dimensions and volumes.   
- Both imaging modalities showed moderate correlations between
-modeled organ volume and gross pathologic weights.  In the intra-modality comparison,
-ARFI and MR total prostate gland volumes were well-correlated (R$^2$ = \MRarfiVolTotalRsq), but
-ARFI yielded prostate volumes that were, on average, larger
+the United States and the second leading cause of cancer-related death.  Two
+imaging modalities, Multi-parametric Magnetic Resonance Imaging (mpMRI) and
+Acoustic Radiation Force Imaging have the potential to aid in PCa diagnosis and
+management by evaluating the structural composition of prostate zones and
+tumors based on water content and stiffness respectively.  In this study,
+\invivo MR and ARFI imaging datasets were compared to one another and against
+gross pathology in order to evaluate their ability to delineate central gland
+(CG) and prostate capsule anatomy. 3D slicer, an open source image processing
+program,  was used for MR and ARFI manual for manual segmentation, 3d modeling,
+and measurement of zonal anatomy tri-axial dimensions and volumes.   Both
+imaging modalities showed moderate correlations between modeled organ volume
+and gross pathologic weights.  In the intra-modality comparison, ARFI and MR
+total prostate gland volumes were well-correlated (R$^2$ = \MRarfiVolTotalRsq),
+but ARFI yielded prostate volumes that were, on average, larger
 (\MRarfiVolTotalMeanDiff\% $\pm$ \MRarfiVolTotalStdDiff\%) than MR. This was
 primarily due to total gland over-estimation along the lateral dimension.
-(\ARFImrTotalLatLatMeanPct $\pm$ \ARFImrTotalLatLatStdPct\%). Differences between MR and
-ARFI in the other dimensions were less significant (\ARFImrTotalAntPostMeanPct $\pm$
-\ARFImrTotalAntPostStdPct\% and \ARFImrTotalApexBaseMeanPct $\pm$
-\ARFImrTotalApexBaseStdPct\%). ARFI and MR CG volumes were also
+(\ARFImrTotalLatLatMeanPct $\pm$ \ARFImrTotalLatLatStdPct\%). Differences
+between MR and ARFI in the other dimensions were less significant
+(\ARFImrTotalAntPostMeanPct $\pm$ \ARFImrTotalAntPostStdPct\%,
+anterior-to-posterior, and \ARFImrTotalApexBaseMeanPct $\pm$
+\ARFImrTotalApexBaseStdPct\%, apex-to-base). ARFI and MR CG volumes were also
 well-correlated (R$^2$ = \MRarfiVolCentralRsq).  CG volume differences were
 attributed to MR apex-to-base axis underestimation
-(\ARFImrCentralApexBaseMeanPct $\pm$ \ARFImrCentralApexBaseStdPct\%), and
-ARFI lateral dimension over-estimation  (\ARFImrCentralLatLatMeanPct
-$\pm$ \ARFImrCentralLatLatStdPct\%).  Overall, ARFI prostate imaging yielded prostate
+(\ARFImrCentralApexBaseMeanPct $\pm$ \ARFImrCentralApexBaseStdPct\%), and ARFI
+lateral dimension over-estimation  (\ARFImrCentralLatLatMeanPct $\pm$
+\ARFImrCentralLatLatStdPct\%).  Overall, ARFI prostate imaging yielded prostate
 volumes and dimensions that were correlated with MR T2WI estimates, with biases
-in the lateral dimension. ARFI imaging is a promising low-cost, real-time imaging
-modality that can compliment mpMRI for diagnosis, treatment planning and
-management of PCa.
+in the lateral dimension. ARFI imaging is a promising low-cost, real-time
+imaging modality that can compliment mpMRI for diagnosis, treatment planning
+and management of PCa.

### 2014-01-21 abstract.tex

@@
 \ARFImrTotalApexBaseStdPct\%, apex-to-base). ARFI and MR CG volumes were also
+the United States and the second leading cause of cancer-related death.
+Two imaging modalities, Multi-parametric Magnetic Resonance Imaging (mpMRI) and
+Acoustic Radiation Force Imaging have the potential to aid in PCa
+diagnosis and management by evaluating the structural composition of prostate
+zones and tumors based on water content and stiffness respectively.  In this study, 
+MR and ARFI imaging datasets were compared to one another and against gross pathology in order to evaluate their
+ability to delineate central gland (CG) and prostate capsule anatomy. 3D slicer, an open source image processing program,  was used
+ for MR and ARFI  manual segmentation, 3d modeling, and measurement of zonal anatomy tri-axial dimensions and volumes.   
+ Both imaging modalities showed moderate correlations between
+modeled organ volume and gross pathologic weights.  In the intra-modality comparison,
+ARFI and MR total prostate gland volumes were well-correlated (R$^2$ = \MRarfiVolTotalRsq), but
+ARFI yielded prostate volumes that were, on average, larger
+(\MRarfiVolTotalMeanDiff\% $\pm$ \MRarfiVolTotalStdDiff\%) than MR. This volume bias was
+primarily due to ARFI lateral dimension over-estimation.
+(\ARFImrTotalLatLatMeanPct $\pm$ \ARFImrTotalLatLatStdPct\%). Differences between MR and
+ARFI in the other dimensions were less significant (\ARFImrTotalAntPostMeanPct $\pm$
+\ARFImrTotalAntPostStdPct\% and \ARFImrTotalApexBaseMeanPct $\pm$
+\ARFImrTotalApexBaseStdPct\%). ARFI and MR CG volumes were also
 well-correlated (R$^2$ = \MRarfiVolCentralRsq).  CG volume differences were

### 2014-01-21 compile-n-view.sh
corrected syntax

@@ -1,4 +1,4 @@
-tmp_latex = "/tmp/latex"
+tmp_latex="/tmp/latex"
 
 if [ ! -d $tmp_latex ]; then
     mkdir $tmp_latex

### 2014-01-21 abstract.tex
reverted abstract to manually merge in changes

--- 2014-01-21 abstract.tex
@@
 the United States and the second leading cause of cancer-related death.
-Two imaging modalities, Multi-parametric Magnetic Resonance Imaging (mpMRI) and
-Acoustic Radiation Force Imaging have the potential to aid in PCa
+Multi-parametric Magnetic Resonance Imaging (mpMRI) has gained recent
+popularity to characterize PCa.  ARFI imaging has the potential to aid in PCa
 diagnosis and management by evaluating the structural composition of prostate
-zones and tumors based on water content and stiffness respectively.  In this study, 
-MR and ARFI imaging datasets were compared to one another and against gross pathology in order to evaluate their
-ability to delineate central gland (CG) and prostate capsule anatomy. 3D slicer, an open source image processing program,  was used
- for MR and ARFI  manual segmentation, 3d modeling, and measurement of zonal anatomy tri-axial dimensions and volumes.   
- Both imaging modalities showed moderate correlations between
-modeled organ volume and gross pathologic weights.  In the intra-modality comparison,
-ARFI and MR total prostate gland volumes were well-correlated (R$^2$ = \MRarfiVolTotalRsq), but
-ARFI yielded prostate volumes that were, on average, larger
-(\MRarfiVolTotalMeanDiff\% $\pm$ \MRarfiVolTotalStdDiff\%) than MR. This volume bias was
-primarily due to ARFI lateral dimension over-estimation.
-(\ARFImrTotalLatLatMeanPct $\pm$ \ARFImrTotalLatLatStdPct\%). Differences between MR and
-ARFI in the other dimensions were less significant (\ARFImrTotalAntPostMeanPct $\pm$
+zones and tumors based on their stiffness.  In this study, MR and ARFI imaging
+datasets were compared to one another and with gross pathology measurements
+made immediately post radical prostatectomy.  Images were manually segmented in
+3D Slicer to delineate the central gland (CG) and prostate capsule, and 3D
+models were rendered in 3D Slicer to evaluate zonal anatomy dimensions and
+volumes.  Both imaging modalities showed moderate correlations between
+estimated organ volume and gross pathologic weights.  ARFI and MR total
+prostate gland volumes were well-correlated (R$^2$ = \MRarfiVolTotalRsq), but
+ARFI images yielded prostate volumes that were, on average, larger
+(\MRarfiVolTotalMeanDiff\% $\pm$ \MRarfiVolTotalStdDiff\%) than MR images,
+primarily due to over-estimation of the lateral dimension of the total gland
+(\ARFImrTotalLatLatMeanPct $\pm$ \ARFImrTotalLatLatStdPct\%), while differences
+in the other dimensions were less significant (\ARFImrTotalAntPostMeanPct $\pm$
 \ARFImrTotalAntPostStdPct\% and \ARFImrTotalApexBaseMeanPct $\pm$
-\ARFImrTotalApexBaseStdPct\%). ARFI and MR CG volumes were also
+\ARFImrTotalApexBaseStdPct\%).  The ARFI and MR CG volumes were also
 well-correlated (R$^2$ = \MRarfiVolCentralRsq).  CG volume differences were
-attributed to MR apex-to-base axis underestimation 
+attributed to under-estimation of the apex-to-base axis
 (\ARFImrCentralApexBaseMeanPct $\pm$ \ARFImrCentralApexBaseStdPct\%), and
-ARFI lateral dimension over-estimation  (\ARFImrCentralLatLatMeanPct
-$\pm$ \ARFImrCentralLatLatStdPct\%).  Overall, ARFI prostate imaging yielded prostate
+over-estimation bias in the lateral dimension (\ARFImrCentralLatLatMeanPct
+$\pm$ \ARFImrCentralLatLatStdPct\%).  ARFI prostate imaging yielded prostate
 volumes and dimensions that were correlated with MR T2WI estimates, with biases
-in the lateral dimension. ARFI imaging is a promising low-cost, real-time imaging
+in the AP dimension, most likely related to poor displacement SNR in the
+anterior region of the prostate from greater distance from the rectal wall
+imaging surface.  Lateral dimension contrast was also challenging due to
+peri-prostatic fat.  ARFI imaging is a promising low-cost, real-time imaging
 modality that can compliment mpMRI for diagnosis, treatment planning and

### 2014-01-21 intro.tex
Kathy's changes

@@
 clinical motivation for prostate imaging.  Section~\ref{sect:methods} describes
-the methods used to experimentally-acquire our imaging data, process of gross
+the methods used to experimentally-acquire our imaging data, process the gross
 pathological specimens post radical prostatectomy, and image process our

### 2014-01-21 arfi_background.tex
implemented all of Kathy's changes and URLs / citations for ITK-SNAP and 3DSlicer

@@
 for investigation of mechanical properties of human soft tissues, including
-liver, arteries, heart, prostate, and
-breast~\cite{Bouchard2009,Dumont2009,Trahey2004,Dahl2009,proc_congdon04,Palmeri2011,Zhai2010a,hsu07,Palmeri2011a}.
-and it provides a unique tool to use to characterize the cervix during
-pregnancy.  The acoustic radiation force $\vec{F}$ generated in soft tissues by
+liver, arteries, heart, prostate, breast, and cervix~\cite{Bouchard2009,Dumont2009,Trahey2004,Dahl2009,proc_congdon04,Palmeri2011,Zhai2010a,hsu07,Palmeri2011a,Palmeri2013}.
+The acoustic radiation force $\vec{F}$ generated in soft tissues by
 focused ultrasound can be described by~\cite{nyborg65,torr84}:

### 2014-01-21 arfi_methods.tex
implemented all of Kathy's changes and URLs / citations for ITK-SNAP and 3DSlicer

@@
 with all 4 beams located inside the beamwidth of the pushing pulses such that
-the beam spacing was 0.17 mm.
+the tracking beam spacing was 0.17 mm.
 

### 2014-01-21 clinical_significance.tex
implemented all of Kathy's changes and URLs / citations for ITK-SNAP and 3DSlicer

@@
 risk of over-diagnosis and treatment of low-grade, and possibly clinically
-insignificant, disease exists.  Moreover, due to the random nature of biopsies,
-PCa located outside the routine sampling sites can be missed and the extent
-of the cancer might be underestimated.~\cite{Gupta2013,Cornud2012} For example,
-in a study by Mufarrij \etal, 45.9--47.2\% of patients who were candidates for
-active surveillance, but underwent radical prostatectomy, had a higher Gleason
-score on final histopathology than after TRUS biopsy.~\cite{Mufarrij2010} These
-inaccuracies may lead to inappropriate diagnosis, imprecise risk assessment and
-potentially avoidable morbidity.
+insignificant, disease exists.  Moreover, due to the random nature of
+TRUS-guided systematic biopsies, PCa located outside the routine sampling sites
+can be missed and the extent of the cancer might be
+underestimated.~\cite{Gupta2013,Cornud2012} For example, in a study by Mufarrij
+\etal, 45.9--47.2\% of patients who were candidates for active surveillance,
+but underwent radical prostatectomy, had a higher Gleason score on final
+histopathology than after TRUS biopsy.~\cite{Mufarrij2010} These inaccuracies
+may lead to inappropriate diagnosis, imprecise risk assessment and potentially
+avoidable morbidity.

### 2014-01-21 disc_arfi_mr.tex
implemented all of Kathy's changes and URLs / citations for ITK-SNAP and 3DSlicer

@@
 ((Figure~\ref{fig:mr_arfi_path_axes}(c), Table~\ref{tab:mr_arfi_axes_error}).
-In fact, it is likely the degree of ARFI estimation relative to the actual
+In fact, it is likely the degree of ARFI underestimation relative to the actual
 apex-to-base prostate length is even greater than suggested by the MR

### 2014-01-21 fig_mr_arfi_volumes.tex
implemented all of Kathy's changes and URLs / citations for ITK-SNAP and 3DSlicer

@@
     manually-segmented images.  Total prostate gland volumes ranged from
-    19.6--71.0 cm$^3$ based on MR image models (a), with a moderate correlation
-    between MR and ARFI imaging for both the total gland volume (R$^2$ =
+    19.6--71.0 cm$^3$ based on MR image models (a), with good correlation
+    between MR and ultrasound imaging for both the total gland volume (R$^2$ =
     \MRarfiVolTotalRsq, (b)) and the central gland (R$^2$ =
-    \MRarfiVolCentralRsq, (c)).  The ARFI total gland volumes had a positive
+    \MRarfiVolCentralRsq, (c)).  The ultrasound total gland volumes had a positive
     bias compared with the MR volumes (\MRarfiVolTotalMeanDiff~$\pm$

### 2014-01-21 fig_mr_arfi_weight.tex
implemented all of Kathy's changes and URLs / citations for ITK-SNAP and 3DSlicer

@@
     approximated ellipsoidal prostate pathology volumes (R$^2$ = \pathVolMRrsq
-    and \pathVolARFIrsq, respectively) (c).  There was on pathology specimen
+    and \pathVolARFIrsq, respectively) (c).  There was one pathology specimen
     that had an erroneous prostate volume and weight recorded, which has been

### 2014-01-21 future.tex
draft edits and spelling
implemented all of Kathy's changes and URLs / citations for ITK-SNAP and 3DSlicer

@@
 to improve anterior prostate boundary delineation in both B-mode and ARFI
-imaging are also being pursued and will hopefully allow for the currently
-elusive anterior PCa lesions in MR to be seen in ARFI imaging.  3D models of
-the prostate and central glands will be used across ARFI and MR imaging
-datasets to spatially register them to facilitate correlation of regions of PCa
-suspicion between the two modalities, and to correlate with whole-mount
-histology of the excised prostate specimens.  This effort will allow the
-sensitivity and specificity of each imaging modality for PCa detection and
-characterization to be quantified.  While ARFI imaging of the prostate in a
-constantly evolving technology, the fact that one can view the different
+imaging are also being pursued for better visualization in this region.  3D
+models of the prostate and central glands will be used across ARFI and MR
+imaging datasets to spatially register them to facilitate correlation of
+regions of PCa suspicion between the two modalities, and to correlate with
+whole-mount histology of the excised prostate specimens.  This effort will
+allow the sensitivity and specificity of each imaging modality for PCa
+detection and characterization to be quantified.  Additionally, we will be
+pursuing quantitative measurement of the pressure being applied during imaging
+to help reduce prostate deformation and reduce the impact of prostate stiffness
+nonlinearity in our displacement images.  While ARFI imaging of the prostate is
+a constantly evolving technology, the fact that one can view the different
 prostatic zones on ARFI images in addition to MRI is encouraging and could lead

### 2014-01-21 image_proc_methods.tex
implemented all of Kathy's changes and URLs / citations for ITK-SNAP and 3DSlicer

@@
 Axial MR T2WI images were manually segmented using the smooth polygon tool in
-ITK-SNAP~\cite{Yushkevich2006}, using unique labels for the PZ, CG and AFS. The
+ITK-SNAP (\url{http://www.itksnap.org})~\cite{Yushkevich2006}, using unique labels for the PZ, CG and AFS. The
 gland was segmented from base-to-apex.  The base was identified below the
@@
 
-Segmented image stacks were imported into 3D Slicer (v4.3.0) and 3D
+Segmented image stacks were imported into 3D Slicer (\url{http://www.slicer.org}, v4.3.0)~\cite{Fedorov2012} and 3D
 models were rendered using the following parameters (Table~\ref{tab:3dslicer}):

### 2014-01-21 mr_background.tex
removed unnecessary higher-order MR background
implemented all of Kathy's changes and URLs / citations for ITK-SNAP and 3DSlicer

@@
 dynamic parameters, into MR analysis via diffusion-weighted imaging (DWI),
-dynamic contrast-enhanced (DCE) imaging or MR spectroscopic imaging (MRSI) is
+dynamic contrast-enhanced (DCE) imaging or MR spectroscopic imaging (MRSI) are
 promising in the future diagnosis and management of PCa.
@@
 glandular components include the peripheral zone (PZ) and the central gland,
-which are typically easily distinguishable on T2WI. The central gland includes
-the central zone, transition zone and the periurethral glandular
-tissue.~\cite{Jung2012} Other anatomical markers such as the urethra,
-verumontanum and ejaculatory are also often seen on T2WI.
+which are typically easily distinguishable on T2WI.  Other anatomical markers
+such as the urethra, verumontanum and ejaculatory are also often seen on T2WI.
 
@@
 images, T2WI were used exclusively in this study.
-
-%DWI is based on the free movement of water particles in tissue and measures the
-%degree of motion restriction.~\cite{Koh2007} Normal prostatic tissue is very
-%glandular with plenty of water molecule movement. On the other hand, tumors
-%have high cellular density and restricted water movement, which leads to
-%decreased diffusion. Due to this restricted diffusion, tumors seem to be of
-%higher intensity on DWI.~\cite{Koh2007}
-%
-%DCE-MRI provides a measure of tumor vascularity and takes advantage of
-%tumor-induced angiogenesis in PCa.~\cite{Noworolski2005} Increased vascularity
-%in PCa results in earlier and higher peak contrast enhancement and a faster
-%washout when compared to normal prostatic tissue. These known characteristics
-%of PCa are helpful when trying to localize lesions. 
-%
-%Like DWI, MRSI can be helpful in lesion characterization as it provides
-%information on the presence of certain metabolites in
-%tissue.~\cite{Barentsz2012} Choline and citrate are especially useful in the
-%setting of PCa. Choline is critical in cell membrane synthesis and is thus
-%usually elevated in cells with cancerous behavior.~\cite{Gupta2013} Healthy
-%prostate epithelium synthesizes and secretes large quantities of citrate but
-%levels are decreased in PCa.~\cite{Costello1999,Kurhanewicz2002} Therefore, an
-%increase in the choline-to-citrate ratio on MRSI can be used as an indicator of
-%malignancy.~\cite{Bonekamp2011}

### 2014-01-21 results.tex
implemented all of Kathy's changes and URLs / citations for ITK-SNAP and 3DSlicer

@@
 The subvolumes associated with the zonal anatomy in each imaging modality were
-measured (Figure~\ref{fig:mr_arfi_volumes}(a)), with moderate correlation
-between the ARFI and MR total prostate gland volumes (R$^2$ =
+measured (Figure~\ref{fig:mr_arfi_volumes}(a)), with good correlation
+between the ultrasound and MR total prostate gland volumes (R$^2$ =
 \MRarfiVolTotalRsq), with a mean overestimation of of
-\MRarfiVolTotalMeanDiff~$\pm$ \MRarfiVolTotalStdDiff\% by ARFI imaging compared
+\MRarfiVolTotalMeanDiff~$\pm$ \MRarfiVolTotalStdDiff\% by ultrasound imaging compared
 to MR volumes (Figure~\ref{fig:mr_arfi_volumes}(b)).  Central gland volumes

### 2014-01-21 disc_arfi_mr.tex
integrtaed Ned's changes

@@
 compression to the rectal wall, which then deforms the prostate.  Pressure on
-the prostate during ARFI imaging is modulated by the urologist when he locks
-the transducer rotation apparatus in place prior to imaging, and deformation of
-the prostate can be easily visualized during this alignment process.  It was
-our goal to apply as little pressure while still maintaining uniform contact
-throughout the image acquisition. This is critical since poor acoustic coupling
-to the prostate through the rectal wall can greatly diminish ARFI image
-displacement SNR.
+the prostate during ARFI imaging is modulated by the urologist when the
+transducer rotation apparatus is locked into position prior to imaging, and
+deformation of the prostate can be easily visualized during this alignment
+process.  It was our goal to apply as little pressure while still maintaining
+uniform contact throughout the image acquisition. This is critical since poor
+acoustic coupling to the prostate through the rectal wall can greatly diminish
+ARFI image displacement SNR.
 

### 2014-01-21 intro.tex
integrtaed Ned's changes

@@
 evaluating the structural composition of prostate zones and tumors based on
-their stiffness contrast.  Zhai \etal were able to visualize prostatic anatomy
+their stiffness contrast.  Zhai \etal~were able to visualize prostatic anatomy
 by utilizing ARFI imaging in freshly-excised prostates. In a second study, Zhai
-\etal demonstrated the feasibility of ARFI prostate imaging
+\etal~demonstrated the feasibility of ARFI prostate imaging
 \invivo~\cite{Zhai2012}; however, to the authors’ best knowledge, there have

### 2014-01-21 abstract.tex
changed conclusion in abstract

@@
 diagnosis and management by evaluating the structural composition of prostate
-zones and tumors based on their stiffness.  In this study, MR and ARFI imaging
-datasets were compared to one another and with gross pathology measurements
-made immediately post radical prostatectomy.  Images were manually segmented in
-3D Slicer to delineate the central gland (CG) and prostate capsule, and 3D
-models were rendered in 3D Slicer to evaluate zonal anatomy dimensions and
-volumes.  Both imaging modalities showed moderate correlations between
-estimated organ volume and gross pathologic weights.  ARFI and MR total
+zones and tumors based on their stiffness.  In this study, \invivo MR and ARFI
+imaging datasets were compared to one another and with gross pathology
+measurements made immediately post radical prostatectomy.  Images were manually
+segmented in 3D Slicer to delineate the central gland (CG) and prostate
+capsule, and 3D models were rendered in 3D Slicer to evaluate zonal anatomy
+dimensions and volumes.  Both imaging modalities showed moderate correlations
+between estimated organ volume and gross pathologic weights.  ARFI and MR total
 prostate gland volumes were well-correlated (R$^2$ = \MRarfiVolTotalRsq), but
@@
 in the other dimensions were less significant (\ARFImrTotalAntPostMeanPct $\pm$
-\ARFImrTotalAntPostStdPct\% and \ARFImrTotalApexBaseMeanPct $\pm$
-\ARFImrTotalApexBaseStdPct\%).  The ARFI and MR CG volumes were also
-well-correlated (R$^2$ = \MRarfiVolCentralRsq).  CG volume differences were
-attributed to under-estimation of the apex-to-base axis
-(\ARFImrCentralApexBaseMeanPct $\pm$ \ARFImrCentralApexBaseStdPct\%), and
-over-estimation bias in the lateral dimension (\ARFImrCentralLatLatMeanPct
-$\pm$ \ARFImrCentralLatLatStdPct\%).  ARFI prostate imaging yielded prostate
-volumes and dimensions that were correlated with MR T2WI estimates, with biases
-in the AP dimension, most likely related to poor displacement SNR in the
-anterior region of the prostate from greater distance from the rectal wall
-imaging surface.  Lateral dimension contrast was also challenging due to
-peri-prostatic fat.  ARFI imaging is a promising low-cost, real-time imaging
-modality that can compliment mpMRI for diagnosis, treatment planning and
-management of PCa.
+\ARFImrTotalAntPostStdPct\%, anterior-to-posterior, and
+\ARFImrTotalApexBaseMeanPct $\pm$ \ARFImrTotalApexBaseStdPct\%, apex-to-base).
+The ARFI and MR CG volumes were also well-correlated (R$^2$ =
+\MRarfiVolCentralRsq).  CG volume differences were attributed to
+under-estimation of the apex-to-base axis (\ARFImrCentralApexBaseMeanPct $\pm$
+\ARFImrCentralApexBaseStdPct\%), and over-estimation bias in the lateral
+dimension (\ARFImrCentralLatLatMeanPct $\pm$ \ARFImrCentralLatLatStdPct\%).
+ARFI prostate imaging yielded prostate volumes and dimensions that were well
+correlated with MR T2WI estimates, with biases in the lateral dimension due to
+poor contrast caused by peri-prostatic fat.  ARFI imaging is a promising
+low-cost, real-time imaging modality that can compliment mpMRI for diagnosis,
+treatment planning and management of PCa.

### 2014-01-21 disc_arfi_mr.tex
integrated Raj's comments

--- 2014-01-21 disc_arfi_mr.tex
@@
 tri-axial dimensions since each modality applies a different amount of external
-compression to the rectal wall, which then deforms the prostate.  Pressure on
-the prostate during ARFI imaging is modulated by the urologist when he locks
-the transducer rotation apparatus in place prior to imaging, and deformation of
-the prostate can be easily visualized during this alignment process.  It was
-our goal to apply as little pressure while still maintaining uniform contact
-throughout the image acquisition. This is critical since poor acoustic coupling
-to the prostate through the rectal wall can greatly diminish ARFI image
+compression to the rectal wall, which then deforms the prostate.  The
+endorectal coil places pressure on the prostate that is dependent on the
+patient's anatomy and overall prostate size, while pressure on the prostate
+during ultrasound imaging is modulated by the urologist when the transducer
+rotation apparatus is locked into position prior to imaging.  The differential
+deformation of the prostate can be easily visualized during the transducer
+alignment process during ultrasound imaging.  It was our goal to apply as
+little pressure while still maintaining uniform contact throughout the
+ultrasound image acquisition. This is critical since poor acoustic coupling to
+the prostate through the rectal wall can greatly diminish ARFI image
 displacement SNR.

### 2014-01-21 intro.tex
integrated Raj's comments

@@
 the United States and the second leading cause of cancer related
-death.~\cite{Howlader2011} Screening with prostate specific antigen (PSA) and
-digital rectal examination (DRE) has led to earlier PCa detection, but
-performance of these measures is not optimal, leading to imprecise risk
-assessment.  Additionally, the random nature of transrectal ultrasound
-(TRUS)-guided biopsies can miss or underestimate the cancer
-burden.~\cite{Gupta2013} Earlier disease diagnosis leads to challenges in
-deciding optimal management strategies for patients presenting with less
-disease burden, whereas missed tumors on TRUS-guided biopsies can result in
-inappropriate diagnoses.
+death.~\cite{Howlader2011} Approximately 1 in every 6 men will develop PCa
+during their lifetime, with the median age of diagnosis at 67 years
+old.~\cite{Howlader2011} PCa is also the second leading cause of cancer-related
+death, with 1 in 36 men dying from the disease.  The National Cancer Institute
+(NCI) estimates that 238,590 men will be diagnosed with PCa in 2013 and 29,720
+will die from the disease.~\cite{Howlader2011}
+
+PCa diagnosis usually begins by screening with prostate specific antigen (PSA)
+and digital rectal examination (DRE).  More definitive diagnosis can be made by
+random transrectal ultrasound (TRUS)-guided biopsies, which are then used to
+provide the clinician with the proper Gleason score. The combination of these
+factors, as well as staging, determines the appropriate therapy and prognosis.
+
+PCa screening has led to earlier diagnosis of smaller tumors and more localized
+disease; however, it is well known that the sensitivity and specificity of PSA
+and DRE are not optimal. In addition, DRE has a low predictive value at lower
+PSA ranges, and PSA yields many false
+positives.~\cite{Gosselaar2007,Gupta2013,Hricak2007} As such, a theoretical
+risk of over-diagnosis and treatment of low-grade, and possibly clinically
+insignificant, disease exists.  Moreover, due to the random nature of
+TRUS-guided systematic biopsies, PCa located outside the routine sampling sites
+can be missed and the extent of the cancer might be
+underestimated.~\cite{Gupta2013,Cornud2012} For example, in a study by Mufarrij
+\etal, 45.9--47.2\% of patients who were candidates for active surveillance,
+but underwent radical prostatectomy, had a higher Gleason score on final
+histopathology than after TRUS biopsy.~\cite{Mufarrij2010} These inaccuracies
+may lead to inappropriate diagnosis, imprecise risk assessment and potentially
+avoidable morbidity.
 
@@
 recent ability to include functional parameters in PCa MRI analysis has yielded
-promising results.~\cite{Gupta2013,Hricak2007} Among the MRI modalities
+promising results.~\cite{Gupta2013,Hricak2007} Among the MRI sequences
 currently used in the study of PCa, it is well established that T2-Weighted
@@
 prostate zonal anatomy, specifically central gland and total prostate gland
-\invivo as compared to endorectal MR T2WI.  Section~\ref{sect:background}
-provides an overview of MR and ARFI imaging in the prostate and an overall
-clinical motivation for prostate imaging.  Section~\ref{sect:methods} describes
-the methods used to experimentally-acquire our imaging data, process the gross
-pathological specimens post radical prostatectomy, and image process our
-datasets.  The results of our analysis, including gross pathology and imaging
-prostate axis and volume estimates, are presented in
-Section~\ref{sect:results}, along with a statistical analysis of the bias and
-variability associated with each imaging modalities measurements, all of which
-is discussed in Section~\ref{sect:discussion}.
+\invivo as compared to T2WI MR at 3T using an endorectal coil.
+Section~\ref{sect:background} provides an overview of MR and ARFI imaging in
+the prostate and an overall clinical motivation for prostate imaging.
+Section~\ref{sect:methods} describes the methods used to experimentally-acquire
+our imaging data, process the gross pathological specimens post radical
+prostatectomy, and image process our datasets.  The results of our analysis,
+including gross pathology and imaging prostate axis and volume estimates, are
+presented in Section~\ref{sect:results}, along with a statistical analysis of
+the bias and variability associated with each imaging modalities measurements,
+all of which is discussed in Section~\ref{sect:discussion}.

### 2014-01-21 prostate_anatomy.tex
integrated Raj's comments

@@
 \subsection{Prostate Anatomy}
-The prostate gland sits below the urinary bladder and surrounds the urethra.
+The prostate gland sits caudal to the urinary bladder and surrounds the urethra.
 Its superior borders include the bladder and seminal vesicles, and the
@@
 
-Although not a true capsule, an outer band of fibromuscular tissue surrounds
-the prostate.~\cite{Bonekamp2011} This ``capsule'' is important when assessing
-the extraprostatic extension of cancer as tumor can spread by disrupting this
+An outer band of fibromuscular tissue surrounds the anterior aspect of the
+prostate.~\cite{Bonekamp2011} This boundary is important when assessing the
+extraprostatic extension of cancer as tumor can spread by disrupting this
 tissue. Two neurovascular bundles course posterior and lateral to the prostate,
 which can also be invaded by malignant cells.  The prostate capsule is another
-structure that was manually segmented in both MR and ARFI images in this study,
-and its overall extent was used to delineate the total prostate gland.
+structure that was manually segmented in both MR and ultrasound images in this
+study, and its overall extent was used to delineate the total prostate gland.

### 2014-01-23 fig_mr_segs_vol.tex
draft edits and spelling
modified fig path
updated filepath and caption for new mr methods fig

@@
 \centering
-\includegraphics[width=1.0\textwidth]{figs/MR_P79_SegsVol.eps}
+\includegraphics[width=1.0\textwidth]{tyler/MR_segs_methods.png}
 \caption{Representative MR T2WI segmentations and rendered volume performed in
@@
     middle and right images show the projections of the rendered model segment
-    outlines in the sagital and coronal views, respectively.}
+    outlines in the sagittal and coronal views, respectively. 3D view of the rendered volume shows anatomical markers for patient left (L), right (R), superior (S), inferior (I), and posterior (P).}
 \label{fig:mr_segs_vol} 

### 2014-01-23 fig_arfi_segs.tex
updated ARFI and bmode seg images and caption to reflect proper apex-base flip, included capsule and CG in lower prostate model

--- 2013-12-30 fig_arfi_segs.tex
@@
 \centering
-\includegraphics[width=0.3\textwidth]{zach/Cap_Modeling_Images/3D_Capsule.png} \\
+\includegraphics[width=0.3\textwidth]{tyler/3D_capsule.png} \\
 Prostate Capsule \\
+\vspace{3 mm}
 \begin{tabular}{ccc}
-\includegraphics[width=0.3\textwidth]{zach/Cap_Modeling_Images/Axial_Cap.png} &
-\includegraphics[width=0.3\textwidth]{zach/Cap_Modeling_Images/Coronal_Cap.png} &
-\includegraphics[width=0.3\textwidth]{zach/Cap_Modeling_Images/Sagittal_Cap.png} \\
-Axial B-mode Outline & Coronal B-mode Outline & Sagital B-mode Outline \\
+\includegraphics[width=0.3\textwidth]{tyler/axial_bmode.png} &
+\includegraphics[width=0.3\textwidth]{tyler/sagittal_bmode.png} &
+\includegraphics[width=0.3\textwidth]{tyler/coronal_bmode.png} \\
+Axial B-mode Outline & Sagittal B-mode Outline & Coronal B-mode Outline \\
+\vspace{5 mm}
 \end{tabular}
-\includegraphics[width=0.3\textwidth]{zach/CG_Modeling_Images/3D_CG.png} \\
+\includegraphics[width=0.3\textwidth]{tyler/3D_CG.png} \\
 Prostate Central Gland \\
+\vspace{3 mm}
 \begin{tabular}{ccc}
-\includegraphics[width=0.3\textwidth]{zach/CG_Modeling_Images/Axial_CG.png} &
-\includegraphics[width=0.3\textwidth]{zach/CG_Modeling_Images/Coronal_CG.png} &
-\includegraphics[width=0.3\textwidth]{zach/CG_Modeling_Images/Sagittal_CG.png} \\
-Axial ARFI Outline & Coronal ARFI Outline & Sagital ARFI Outline \\
+\includegraphics[width=0.3\textwidth]{tyler/axial_ARFI.png} &
+\includegraphics[width=0.3\textwidth]{tyler/sagittal_ARFI.png} &
+\includegraphics[width=0.3\textwidth]{tyler/coronal_ARFI.png} \\
+Axial ARFI Outline & Sagittal ARFI Outline & Coronal ARFI Outline \\
 \end{tabular}
@@
     the B-mode images, with the 3D model outlines superimposed on the central
-    axial, coronal and sagital B-mode images.  The primary segmentation plane
+    axial, sagittal and coronal B-mode images.  The primary segmentation plane
     was sagittal; however the axial and coronal planes were used to guide the
-    segmentation to insure 3D continuity of the capsule structure.  BOTTOM
-    ROWS: Prostate central gland 3D model (blue) rendered from manual
-    segmentation of the ARFI images, with the 3D model outlines superimposed on
-    the central axial, coronal and sagital ARFI images.  The coronal ARFI
+    segmentation to ensure 3D continuity of the capsule structure.  BOTTOM
+    ROWS: Prostate central gland 3D model (blue) inside prostate capsule 3D model (magenta) rendered from manual
+    segmentation of the ARFI images, with the prostate central gland 3D model outlines superimposed on
+    the central axial, sagittal and coronal ARFI images.  The coronal ARFI
     imaging plane was the primary orientation used for image segmentation, but
-    like the capsule, the orthogonal planes were used to insure continuity of
+    like capsule segmentation, the orthogonal planes were used to ensure continuity of
     the central gland in all three dimensions.}

### 2014-02-08 abstract.tex
Tyler's comments

@@
 imaging datasets were compared to one another and with gross pathology
-measurements made immediately post radical prostatectomy.  Images were manually
-segmented in 3D Slicer to delineate the central gland (CG) and prostate
-capsule, and 3D models were rendered in 3D Slicer to evaluate zonal anatomy
-dimensions and volumes.  Both imaging modalities showed moderate correlations
-between estimated organ volume and gross pathologic weights.  ARFI and MR total
-prostate gland volumes were well-correlated (R$^2$ = \MRarfiVolTotalRsq), but
-ARFI images yielded prostate volumes that were, on average, larger
+measurements made immediately after radical prostatectomy.  Images were
+manually segmented in 3D Slicer to delineate the central gland (CG) and
+prostate capsule, and 3D models were rendered to evaluate zonal anatomy
+dimensions and volumes.  Both imaging modalities showed good correlation
+between estimated organ volume and gross pathologic weights.  Ultrasound and MR
+total prostate gland volumes were well-correlated (R$^2$ = \MRarfiVolTotalRsq),
+but ARFI images yielded prostate volumes that were larger
 (\MRarfiVolTotalMeanDiff\% $\pm$ \MRarfiVolTotalStdDiff\%) than MR images,
-primarily due to over-estimation of the lateral dimension of the total gland
+primarily due to overestimation of the lateral dimension
 (\ARFImrTotalLatLatMeanPct $\pm$ \ARFImrTotalLatLatStdPct\%), while differences
@@
 \MRarfiVolCentralRsq).  CG volume differences were attributed to
-under-estimation of the apex-to-base axis (\ARFImrCentralApexBaseMeanPct $\pm$
-\ARFImrCentralApexBaseStdPct\%), and over-estimation bias in the lateral
+ARFI underestimation of the apex-to-base axis (\ARFImrCentralApexBaseMeanPct $\pm$
+\ARFImrCentralApexBaseStdPct\%), and ARFI overestimation bias in the lateral
 dimension (\ARFImrCentralLatLatMeanPct $\pm$ \ARFImrCentralLatLatStdPct\%).
-ARFI prostate imaging yielded prostate volumes and dimensions that were well
-correlated with MR T2WI estimates, with biases in the lateral dimension due to
-poor contrast caused by peri-prostatic fat.  ARFI imaging is a promising
-low-cost, real-time imaging modality that can compliment mpMRI for diagnosis,
-treatment planning and management of PCa.
+ARFI imaging yielded prostate volumes and dimensions that were well-correlated
+with MR T2WI estimates, with biases in the lateral dimension due to poor
+contrast caused by peri-prostatic fat.  ARFI imaging is a promising low-cost,
+real-time imaging modality that can compliment mpMRI for diagnosis, treatment
+planning and management of PCa.

### 2014-02-08 fig_mr_arfi_weight.tex
Tyler's comments

@@
 \centering
+\begin{small}
 \begin{tabular}{lll}
@@
 \includegraphics[width=0.3\linewidth]{figs/corr_pathVol_vol} \\
-(a) Path Weight : Path Volume & (b) Image Volume : Prostate Weight & (c) Image Volume : Path Volume \\
+(a) Prostate Weight : Prostate Volume & (b) Image Volume : Prostate Weight & (c) Image Volume : Prostate Volume \\
 \end{tabular}
+\end{small}
 %\begin{tabular}{ll}
@@
     correlations existed between both T2WI MR and ARFI image volumes and
-    approximated ellipsoidal prostate pathology volumes (R$^2$ = \pathVolMRrsq
-    and \pathVolARFIrsq, respectively) (c).  There was one pathology specimen
-    that had an erroneous prostate volume and weight recorded, which has been
-    excluded from the linear regressions (magenta circles).}
+    approximated ellipsoidal prostate pathology volumes (R$^2$ =
+    \pathVolMRrsq~and \pathVolARFIrsq, respectively) (c).  There was one
+    pathology specimen that had an erroneous prostate volume and weight
+    recorded, which has been excluded from the linear regressions (magenta
+    circles).}
 \label{fig:mr_arfi_weight}

### 2014-02-08 image_proc_methods.tex
Tyler's comments

@@
 segmentation, both the B-mode and ARFI image stacks were down-sampled along the
-sagittal axis by a factor of 20 (i.e., from 360 slices per image stack to 18
-slices per image stack) to reduce the number of manually-segmented frames.
+sagittal axis to a slice thickness of 3 mm to reduce the number of
+manually-segmented frames (18 vs. 360 slices / image stack) and match the MRI
+slice thickness.
 
@@
 segmentation by enabling capsule model intersections in the three orthogonal
-planes during CG segmentation (Figure~\ref{fig:arfi_segs}). The capsule provided
-an outer boundary for the CG segmentation. The capsule and CG models were then
-measured using the same procedure as the MR images.
+planes during CG segmentation (Figure~\ref{fig:arfi_segs}). The capsule
+provided an outer boundary for the CG segmentation. The capsule and CG models
+dimensions were then measured using the same procedure as the MR images.
 

### 2014-02-08 abstract.tex
Polascik's comments

@@
 The ARFI and MR CG volumes were also well-correlated (R$^2$ =
-\MRarfiVolCentralRsq).  CG volume differences were attributed to
-ARFI underestimation of the apex-to-base axis (\ARFImrCentralApexBaseMeanPct $\pm$
+\MRarfiVolCentralRsq).  CG volume differences were attributed to ARFI
+underestimation of the apex-to-base axis (\ARFImrCentralApexBaseMeanPct $\pm$
 \ARFImrCentralApexBaseStdPct\%), and ARFI overestimation bias in the lateral
@@
 contrast caused by peri-prostatic fat.  ARFI imaging is a promising low-cost,
-real-time imaging modality that can compliment mpMRI for diagnosis, treatment
-planning and management of PCa.
+portable, real-time imaging modality that can compliment mpMRI for diagnosis,
+treatment planning and management of PCa.

### 2014-02-08 intro.tex
Polascik's comments

@@
 histopathology than after TRUS biopsy.~\cite{Mufarrij2010} These inaccuracies
-may lead to inappropriate diagnosis, imprecise risk assessment and potentially
+may lead to missed diagnoses, imprecise risk assessment and potentially
 avoidable morbidity.

### 2014-02-08 limitations.tex
Polascik's comments

@@
 anatomy of a healthy (young) prostate is expected to be different from the
-prostate of a middle-age man, who is the target demographic for PCa screening
+prostate of a middle-age and older man, who is the target demographic for PCa screening
 imaging and PCa characterization.

### 2014-02-08 mr_background.tex
integrtaed Ned's changes
Polascik's comments

@@
 dynamic contrast-enhanced (DCE) imaging or MR spectroscopic imaging (MRSI) are
-promising in the future diagnosis and management of PCa.
+promising in the future diagnosis and characterization of PCa.
 
@@
 which are typically easily distinguishable on T2WI.  Other anatomical markers
-such as the urethra, verumontanum and ejaculatory are also often seen on T2WI.
+such as the urethra, verumontanum and ejaculatory ducts are also often seen on
+T2WI.
 

### 2014-02-08 arfi_methods.tex
Stephen's comments

@@
 For the tracking pulses, 16 parallel receive lines at 5.0 MHz were spaced to
-observe both the on and off-axis response of the tissue to the pushing pulses.
-Specifically, four lines were dedicated to tracking the on-axis displacement,
-with all 4 beams located inside the beamwidth of the pushing pulses such that
-the tracking beam spacing was 0.17 mm.
+observe the response of the tissue to the pushing pulses.  Specifically, four
+lines were dedicated to tracking the on-axis displacement, with all 4 beams
+located inside the beamwidth of the pushing pulses such that the tracking beam
+spacing was 0.17 mm.  The tracking beams were focused at 60 mm with F/3
+transmit and F/0.5 receive focal configurations.
 
@@
 prostate.  A custom optical angular feedback transduction circuit utilizing a
-reflective linear strip with 500 lines-per-inch (LPI) resolution (US Digital,
-Vancouver, WA, USA) was coupled with a 141 oz-in torque stepper motor with
-planetary gearbox (Model \#11YPG202S-LW4-R27, Anaheim Automation, Anaheim, CA,
-USA) (Figure~\ref{fig:setup_annotated}) to achieve accurate spatial
-localization of the imaging frames in the 3D dataset.
+reflective linear strip with 212 lines-per-inch (LPI) resolution (US Digital,
+Vancouver, WA, USA) with a QSB-S Quadrature-to-USB adapter to achieve 9 line /
+degree resolution.  Rotation was performed with a 141 oz-in torque stepper
+motor with planetary gearbox (Model \#11YPG202S-LW4-R27, Anaheim Automation,
+Anaheim, CA, USA) (Figure~\ref{fig:setup_annotated}) to achieve accurate
+spatial localization of the imaging frames in the 3D dataset.
 
@@
 focal gain effects.  This normalization was performed using a displacement
-profile measured in a homogeneous tissue-mimicking phantom, which was applied
-to all displacements in the entire data set at each time step, and then
-low-pass filtered with a cutoff-frequency of 0.8 mm$^{-1}$.
+profile measured in a homogeneous tissue-mimicking phantom that was low-pass
+filtered with a cutoff-frequency of 0.8 mm$^{-1}$.  This filtered normalization
+profile was applied to all displacements in the entire data set at each time
+step.
 
-The normalized displacement data were scan-converted using a 3D Delaunay
+The normalized displacement data were scan-converted using a 2D Delaunay
 triangulation interpolation method in Matlab (R2012b, The MathWorks, Inc,
-Natick, MA, USA).  The scan-converted ARFI imaging data had a final isotropic
-voxel size of 0.15 $\times$ 0.15 $\times$ 0.15 mm$^3$.
+Natick, MA, USA) for each lateral location in the dataset.  The scan-converted
+ARFI imaging data had a final isotropic voxel size of 0.15 $\times$ 0.15
+$\times$ 0.15 mm$^3$.

### 2014-02-08 conclusions.tex
Stephen's comments

--- 2013-12-30 conclusions.tex
@@
 \ARFImrTotalAntPostStdPct\%~and \ARFImrTotalApexBaseMeanPct~$\pm$
-\ARFImrTotalApexBaseStdPct\%).  The central zone volumes of ARFI and MR images
+\ARFImrTotalApexBaseStdPct\%).  The central gland volumes of ARFI and MR images
 were also well-correlated (R$^2$ = \MRarfiVolCentralRsq), with minimal volume
 bias between the imaging modalities, but significant variability case-to-case
-(\MRarfiVolCentralMeanDiff~$\pm$ MRarfiVolCentralStdDiff\%).  Central zone
+(\MRarfiVolCentralMeanDiff~$\pm$ \MRarfiVolCentralStdDiff\%).  Central gland
 volume differences were, again, strongly attributed to over-estimation of the

### 2014-02-08 image_proc_methods.tex
Stephen's comments

@@
 \subsubsection{MR Image Segmentation and Modeling}
-Axial MR T2WI images were manually segmented using the smooth polygon tool in
-ITK-SNAP (\url{http://www.itksnap.org})~\cite{Yushkevich2006}, using unique labels for the PZ, CG and AFS. The
-gland was segmented from base-to-apex.  The base was identified below the
-bladder, and subsequent images were segmented until the last slice with visible
-prostatic tissue was identified caudally. The CG, urethra, PZ and AFS were
-segmented independently according to their well-established anatomical
-characteristics on
+Axial MR T2WI images were manually segmented by an experienced radiologist
+using the smooth polygon tool in ITK-SNAP
+(\url{http://www.itksnap.org})~\cite{Yushkevich2006}, using unique labels for
+the PZ, CG and AFS. The gland was segmented from base-to-apex.  The base was
+identified below the bladder, and subsequent images were segmented until the
+last slice with visible prostatic tissue was identified caudally. The CG,
+urethra, PZ and AFS were segmented independently according to their
+well-established anatomical characteristics on
 T2WI.~\cite{Barentsz2012,Jung2012,Poon1985,Hricak2007,Bonekamp2011} The PZ was

### 2014-02-08 mr_background.tex
Stephen's comments

@@
 contrast resolution, which can be maximized by using thin sections of 3--4 mm
-and a small field of view of approximately 14 cm.~\cite{Gupta2013,Bonekamp2011}
-T2 sequences are the most helpful for tumor localization, as they can clearly
-show overall prostate morphology, internal structures and prostatic
-margins.~\cite{Gupta2013}
+and a small field of view of approximately 14
+cm$^2$.~\cite{Gupta2013,Bonekamp2011} T2 sequences are the most helpful for
+tumor localization, as they can clearly show overall prostate morphology,
+internal structures and prostatic margins.~\cite{Gupta2013}
 

### 2014-02-08 tab_mr_arfi_volumes.tex
updated figures for correlation analyses and removed image comparisons to path axes
Stephen's comments

@@
-\begin{table}
+\begin{table}[h!]
 \centering
@@
 \begin{tabular}{|l|l|l|l|l|} \hline
-{\bf Study Subject} & {\bf MR Central Gland} & {\bf MR Total} & {\bf ARFI Central Zone} & {\bf ARFI Total} \\ 
+{\bf Study Subject} & {\bf MR Central Gland} & {\bf MR Total} & {\bf ARFI Central Gland} & {\bf ARFI Total} \\ 
 & {\bf Volume (cm$^3$)} & {\bf Volume (cm$^3$)} & {\bf Volume (cm$^3$)} & {\bf Volume (cm$^3$)} \\ \hline

### 2014-02-08 abstract.tex
shortened abstract for TMI limits

--- 2014-02-08 abstract.tex
@@
 Prostate cancer (PCa) is the most common non-cutaneous malignancy among men in
-the United States and the second leading cause of cancer-related death.
+the United States and the second-leading cause of cancer-related death.
 Multi-parametric Magnetic Resonance Imaging (mpMRI) has gained recent
-popularity to characterize PCa.  ARFI imaging has the potential to aid in PCa
-diagnosis and management by evaluating the structural composition of prostate
-zones and tumors based on their stiffness.  In this study, \invivo MR and ARFI
-imaging datasets were compared to one another and with gross pathology
-measurements made immediately after radical prostatectomy.  Images were
-manually segmented in 3D Slicer to delineate the central gland (CG) and
-prostate capsule, and 3D models were rendered to evaluate zonal anatomy
-dimensions and volumes.  Both imaging modalities showed good correlation
-between estimated organ volume and gross pathologic weights.  Ultrasound and MR
-total prostate gland volumes were well-correlated (R$^2$ = \MRarfiVolTotalRsq),
-but ARFI images yielded prostate volumes that were larger
-(\MRarfiVolTotalMeanDiff\% $\pm$ \MRarfiVolTotalStdDiff\%) than MR images,
-primarily due to overestimation of the lateral dimension
-(\ARFImrTotalLatLatMeanPct $\pm$ \ARFImrTotalLatLatStdPct\%), while differences
-in the other dimensions were less significant (\ARFImrTotalAntPostMeanPct $\pm$
-\ARFImrTotalAntPostStdPct\%, anterior-to-posterior, and
-\ARFImrTotalApexBaseMeanPct $\pm$ \ARFImrTotalApexBaseStdPct\%, apex-to-base).
-The ARFI and MR CG volumes were also well-correlated (R$^2$ =
-\MRarfiVolCentralRsq).  CG volume differences were attributed to
-ARFI underestimation of the apex-to-base axis (\ARFImrCentralApexBaseMeanPct $\pm$
-\ARFImrCentralApexBaseStdPct\%), and ARFI overestimation bias in the lateral
-dimension (\ARFImrCentralLatLatMeanPct $\pm$ \ARFImrCentralLatLatStdPct\%).
-ARFI imaging yielded prostate volumes and dimensions that were well-correlated
-with MR T2WI estimates, with biases in the lateral dimension due to poor
-contrast caused by peri-prostatic fat.  ARFI imaging is a promising low-cost,
-real-time imaging modality that can compliment mpMRI for diagnosis, treatment
-planning and management of PCa.
+popularity to characterize PCa.  ARFI imaging has the potential to aid PCa
+diagnosis and management by using tissue stiffness to evaluate prostate zonal
+anatomy and lesions.  MR and ARFI \invivo imaging datasets were compared to one
+another and with gross pathology measurements made immediately after radical
+prostatectomy.  Images were manually segmented in 3D Slicer to delineate the
+central gland (CG) and prostate capsule, and 3D models were rendered to
+evaluate zonal anatomy dimensions and volumes.  Both imaging modalities showed
+good correlation between estimated organ volume and gross pathologic weights.
+Ultrasound and MR total prostate volumes were well-correlated (R$^2$ =
+\MRarfiVolTotalRsq), but ARFI images yielded prostate volumes that were larger
+(\MRarfiVolTotalMeanDiff\% $\pm$ \MRarfiVolTotalStdDiff\%) than MR images, due
+to overestimation of the lateral dimension (\ARFImrTotalLatLatMeanPct $\pm$
+\ARFImrTotalLatLatStdPct\%), with less significant differences in the other
+dimensions (\ARFImrTotalAntPostMeanPct $\pm$ \ARFImrTotalAntPostStdPct\%,
+anterior-to-posterior, and \ARFImrTotalApexBaseMeanPct $\pm$
+\ARFImrTotalApexBaseStdPct\%, apex-to-base).  ARFI and MR CG volumes were also
+well-correlated (R$^2$ = \MRarfiVolCentralRsq).  CG volume differences were
+attributed to ARFI underestimation of the apex-to-base axis
+(\ARFImrCentralApexBaseMeanPct $\pm$ \ARFImrCentralApexBaseStdPct\%), and ARFI
+overestimation of the lateral dimension (\ARFImrCentralLatLatMeanPct $\pm$
+\ARFImrCentralLatLatStdPct\%).  ARFI imaging yielded prostate volumes and
+dimensions that were well-correlated with MR T2WI estimates, with biases in the
+lateral dimension due to poor contrast caused by peri-prostatic fat.  ARFI
+imaging is a promising low-cost, portable, real-time modality that can
+compliment mpMRI for PCa diagnosis, treatment planning and management.

### 2014-02-08 acknowledgements.tex
moved Ned from acknow -> author list
added Duke PAs to acknowledgements
modified format and cover letter for Ultrasonic Imaging submission

--- 2013-09-15 acknowledgements.tex
@@
 The authors would like to thank Siemens Medical Solution USA, Ultrasound
-Division for their in-kind technical support and Ned Danieley for computer
-system administration support.  This work was supported by NIH R01CA142824 and
-the Duke Coulter Translational Grant Program.
+Division for their in-kind technical support.  Special thanks to Ned Danieley
+for computer system administration support, the pathology assistants and Dr.
+John Madden in the Duke University Medical Center for their gross prostate
+organ processing, and Matthew Brown for his assistance in designing the optical
+prostate rotation feedback system.  This work was supported by NIH R01CA142824
+and the Duke Coulter Translational Grant Program.

### 2014-02-08 coverletter.tex
modified format and cover letter for Ultrasonic Imaging submission

@@
 
+\newcommand{\journal}{Ultrasonic Imaging}
 \begin{document}
@@
 \vspace{0pt}
-\includegraphics[width=3.0in,trim=55 50 50 85,clip=true]{/home/mlp6/bme/Duke_BME_Logos/Pratt-HorizontalLogo-Vector-Black}
+\includegraphics[width=3.0in,trim=55 50 50 85,clip=true]{/home/mlp6/Documents/Duke_BME_Logos/Pratt-HorizontalLogo-Vector-Black}
 \end{minipage}\hfill
@@
 
-Re: Ultrasonic Imaging Manuscript Submission
+Re: \journal~Manuscript Submission
 \vspace{0.25in}
@@
 {\it ``Acoustic Radiation Force Impulse (ARFI) Imaging Prostate Zonal Anatomy:
-    Comparison with Endorectal T2-Weighted MR Imaging (T2WI)''}, for
-consideration to be published in Ultrasonic Imaging.  All of the authors on
-this manuscript have reviewed its content and approve of its submission.  This
-work has not been published or submitted for review to any other journals.
+    Comparison with 3T T2-Weighted MR Imaging (T2WI)''}, for consideration to
+be published in \journal.
+
+All of the authors on this manuscript have reviewed its content and approve of
+its submission.
+
+This work has not been published in or submitted for review to any other
+journals.
+
+The human subject data presented in this manuscript was acquired in a Duke
+University Medical Center IRB-approved study, with all study subjects providing
+written informed consent.
 
@@
 Sincerely,\\
-\includegraphics{/home/mlp6/MarkLPalmeriSig}\\
+\includegraphics{/home/mlp6/Documents/MarkLPalmeriSig}\\
 %\vspace*{0.5in}

### 2014-02-08 image_proc_methods.tex
modified format and cover letter for Ultrasonic Imaging submission

--- 2014-02-08 image_proc_methods.tex
@@
 \subsubsection{MR Image Segmentation and Modeling}
-Axial MR T2WI images were manually segmented using the smooth polygon tool in
-ITK-SNAP (\url{http://www.itksnap.org})~\cite{Yushkevich2006}, using unique labels for the PZ, CG and AFS. The
-gland was segmented from base-to-apex.  The base was identified below the
-bladder, and subsequent images were segmented until the last slice with visible
-prostatic tissue was identified caudally. The CG, urethra, PZ and AFS were
-segmented independently according to their well-established anatomical
-characteristics on
+Axial MR T2WI images were manually segmented by an experienced radiologist
+using the smooth polygon tool in ITK-SNAP
+(http://www.itksnap.org)~\cite{Yushkevich2006}, using unique labels for
+the PZ, CG and AFS. The gland was segmented from base-to-apex.  The base was
+identified below the bladder, and subsequent images were segmented until the
+last slice with visible prostatic tissue was identified caudally. The CG,
+urethra, PZ and AFS were segmented independently according to their
+well-established anatomical characteristics on
 T2WI.~\cite{Barentsz2012,Jung2012,Poon1985,Hricak2007,Bonekamp2011} The PZ was
@@
 
-Segmented image stacks were imported into 3D Slicer (\url{http://www.slicer.org}, v4.3.0)~\cite{Fedorov2012} and 3D
+Segmented image stacks were imported into 3D Slicer (http://www.slicer.org, v4.3.0)~\cite{Fedorov2012} and 3D
 models were rendered using the following parameters (Table~\ref{tab:3dslicer}):

### 2014-02-08 compile-n-view.sh

@@ -1,7 +1,8 @@
-tmp_latex="/tmp/latex"
+#tmp_latex="/tmp/mr_arfi_anatomy"
+tmp_latex="./"
 
 if [ ! -d $tmp_latex ]; then
     mkdir $tmp_latex
 fi
 
-latexmk -bibtex -pdf -pvc -outdir=$tmp_latex mr_arfi_anatomy.tex
+latexmk -f -bibtex -pdf -pvc -outdir=$tmp_latex mr_arfi_anatomy.tex

### 2014-02-08 parse.py

@@ -1,0 +1,49 @@
+#!/bin/env python
+""""
+parse single PDF of manuscript into individual PDF files for upload
+
+The page breaks are hard-coded below and will need to be adjusted
+
+"""
+
+import os
+
+parent_pdf = "mr_arfi_anatomy.pdf"
+parsed_pdf_dir = "parsed_pdfs"
+
+if not os.path.exists(parsed_pdf_dir):
+    os.mkdir(parsed_pdf_dir)
+
+# section title, number of pages
+sections = (('title',   1),
+            ('main',    18),
+            ('figure1', 1),
+            ('figure2', 1),
+            ('figure3', 1),
+            ('figure4', 1),
+            ('figure5', 1),
+            ('figure6', 1),
+            ('figure7', 1),
+            ('figure8', 1))
+
+print("\nWARNING: Number of pages for each section is HARD-CODED\n")
+
+start_page = 0
+stop_page = 0
+for i in range(len(sections)):
+    # iteratively generate each section PDF
+    start_page = stop_page + 1
+    stop_page = start_page + int(sections[i][1]) - 1
+    os.system("pdftk A=%s cat A%i-%i output %s/%s.pdf" %
+              (parent_pdf,
+               start_page,
+               stop_page,
+               parsed_pdf_dir,
+               sections[i][0]))
+
+    # test for and indicate new file existence
+    new_file = ("./%s/%s.pdf" % (parsed_pdf_dir, sections[i][0]))
+    if os.path.exists(new_file):
+        print("Successfully created: %s" % new_file)
+    else:
+        print("ERROR: Failed to create %s" % new_file)

### 2014-02-08 rm_bib_urls.sh

@@ -1,0 +1,2 @@
+echo "Stripping URLs from library.bib . . ."
+grep -v "url =" ./library.bib > ./library_no_urls.bib

### 2014-02-08 figs/corr_weight_vol.m
added PNG versions of figs for PPT

@@ -103,6 +103,7 @@ function gen_fig_n_fits(fid,w,mr_total_vol,arfi_total_vol,metric)
     end;
 
     print('-depsc2',sprintf('corr_%s_vol.eps',metric));
+    print('-dpng',sprintf('corr_%s_vol.png',metric));
     close;
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2014-02-08 figs/mr_arfi_path_axes.m

@@ -82,6 +82,7 @@ function mr_arfi_path_axes
         line([2 8],[2 8],'LineStyle','--','Color','k','LineWidth',3);
 
         print('-depsc2',sprintf('%s.eps',axis_titles{i}));
+        print('-dpng',sprintf('%s.png',axis_titles{i}));
         close;
 
         % compute over/unders for each modality
@@ -141,6 +142,7 @@ function mr_arfi_path_axes
         line([2 8],[2 8],'LineStyle','--','Color','k','LineWidth',3);
 
         print('-depsc2',sprintf('Imaging_%s.eps',axis_titles{i}));
+        print('-dpng',sprintf('Imaging_%s.png',axis_titles{i}));
         close;
 
         % compute over/unders for each modality
@@ -202,6 +204,7 @@ function mr_arfi_path_axes
             'FontSize'    , fs);
 
         print('-depsc2',sprintf('mr_arfi_total_axes%i.eps',i));
+        print('-dpng',sprintf('mr_arfi_total_axes%i.png',i));
         close;
     end;
 
@@ -237,6 +240,7 @@ function mr_arfi_path_axes
             'FontSize'    , fs);
 
         print('-depsc2',sprintf('mr_arfi_central_axes%i.eps',i));
+        print('-dpng',sprintf('mr_arfi_central_axes%i.png',i));
         close;
     end;
 
@@ -281,6 +285,7 @@ function mr_arfi_path_axes
             'FontSize'    , fs);
 
         print('-depsc2',sprintf('mr_arfi_total_over_under%i.eps',i));
+        print('-dpng',sprintf('mr_arfi_total_over_under%i.png',i));
         close;
 
         % while I'm here, lets compute some stats to include in the caption / text, and LaTeX format it!!
@@ -316,6 +321,7 @@ function mr_arfi_path_axes
             'FontSize'    , fs);
 
         print('-depsc2',sprintf('mr_arfi_central_over_under%i.eps',i));
+        print('-dpng',sprintf('mr_arfi_central_over_under%i.png',i));
         close;
         fprintf(fid,'ARFI & MR & Central & %s & %.1f $\\pm$ %.1f \\\\ \n',axis_combos{i},mean(ARFI_MR_Central_OverUnder(:,i)),std(ARFI_MR_Central_OverUnder(:,i)));
     end;

### 2014-02-08 figs/mr_arfi_volumes.m

@@ -54,6 +54,7 @@ legend boxoff;
 set_axes(gca,fs);
 
 print('-depsc2','mr_arfi_volumes.eps');
+print('-dpng','mr_arfi_volumes.png');
 close;
 
 % lets do a little more processing to figure out what is going on... first the
@@ -73,6 +74,7 @@ a(4) = 110;
 axis(a);
 text(0.5,100,sprintf('Mean Diff = %.1f +/- %.1f',mean(arfi_mr_vol_diff),std(arfi_mr_vol_diff)),'FontSize',fs);
 print('-depsc2','mr_arfi_volume_diff.eps');
+print('-dpng','mr_arfi_volume_diff.png');
 close;
 
 % now lets look at the relative ratios of CG:total volume for each modality
@@ -136,6 +138,7 @@ figure;
     % add ideal line
     line([15 80],[15 80],'LineStyle','--','Color','k','LineWidth',3);
     print('-depsc2','mr_arfi_total_linreg.eps');
+    print('-dpng','mr_arfi_total_linreg.png');
 close;
 
 [mr_arfi_central_fit,mr_arfi_central_Rsq]=compute_linreg_Rsq(mr_cg_vol/1e3,arfi_cz_vol/1e3);
@@ -157,6 +160,7 @@ figure;
     % add ideal line
     line([5 60],[5 60],'LineStyle','--','Color','k','LineWidth',3);
     print('-depsc2','mr_arfi_central_linreg.eps');
+    print('-dpng','mr_arfi_central_linreg.png');
     close;
 
 

### 2014-04-14 Response_to_R0_Reviews.tex
responding to reviews

\documentclass[10pt]{article}
\usepackage{graphicx}
\pagestyle{empty}
\topmargin=-0.75in
\oddsidemargin=0.0in
\textwidth=6.5in
\textheight=9.5in
\parindent=0.0in
\parskip=10pt

\begin{document}

\section*{Response to Reviews}

We thank both reviewers for their insight and comments on the original
submission of our manuscript.  Overall, well agree with all of the comments
that were provided, and we have made edits to our original manuscript that we
feel respond to all of this feedback.  We have provided a detailed response to
the review below, with the original comments italicized, and the responses in
normal text.

Thank you, again, for your time and effort in this review process.

\textit{Reviewer: 1}

\textit{Comments to the Author}

\textit{This manuscripts describes exciting work demonstrating the potential relevance
of acoustic radiation force impulse (ARFI) and B-Mode ultrasound for
delineating prostate zonal anatomy dimensions and volumes in application to
prostate cancer diagnosis and treatment.  The current state of the art involves
MRI imaging, but an ultrasound-based complement would offer the benefits of
being low-cost, real-time and portable.  This work should be of great interest
to both the technically and clinically oriented UI readership.  Overall, the
manuscript is very well written, and the presented in vivo, clinical data
support the conclusions drawn. There are a few minor questions/comments for the
authors to consider before publication, which is recommended:}

\begin{itemize}

    \item \textit{An artistic rendering or other illustration of prostate
            anatomy would facilitate the readers' understanding of the
            description of prostate anatomy in section II.A.}

    \item \textit{Pg 5, lines 109-112, are redundant with other text previously
            written in the background.}

    \item \textit{The methods section describes that diffusion weighted images
            (with ADC maps) and dynamic contrast enhanced MRI data were
            acquired in addition to the T2W images.  Were the the diffusion and
            contrast enhanced MRI data used in the MRI segmentation and/or
            analysis of MRI performance (as compared to ARFI).  It does not
            appear that they were.  Why not?  Also, what contrast agent was
            used?}

    \item \textit{Please describe what is meant by the "longitudinal array" (pg
            6 line 151) and specify that the ER7B is an endorectal transducer.
            This was stated in the associated figure caption but not in the
            text.  Also, please state more clearly in the text that ARFI
            imaging was performed endorectally.}

    \item \textit{Why were the tracking beams focused (@ 60 mm) twice as deep
            as the pushing beams (@ 30 mm and above)?  It would seem that the
            radiation force would not propagate this far below the pushing
            focal position. (pg 6 lines 163 and 154, respectively).}

    \item \textit{For the displacement normalization factor derived in phantom
            materials (pg 7 lines 176-177), what were the acoustic and
            mechanical properties of the phantom, and were these well matched
            to the expected properties of prostate tissue?  How much are these
            properties expected to vary from individual to individual, with or
            without prostate cancer?  What would the impact of using of an
            incorrect normalization factor be on the overall ARFI results?  The
            authors should consider adding a discussion of the potential impact
            of the assumed normalization factor to the limitations sections.}

    \item \textit{Pg 8 lines 205-214 describe that both B-Mode (for prostate
            capsule) and ARFI (for CG) data were used for segmentation.  First,
            how were these data combined?  Second, the use of B-Mode
            segmentation in addition to ARFI is not clearly specified in the
            titles, text and figures.  For example, in Table III, was the total
            volume calculation a result of both ARFI and B-Mode segmentations?
            In that case, the title and column labels describing "ARFI" total
            volume are confusing because they imply that the B-Mode
            segmentation data were not used.  The same is true for the other
            tables, figures, text, and the main Title of the manuscript.  If I
            understand correctly that B-Mode data were important to the overall
            total volume segmentation, then it would be appropriate to better
            clarify this point.  The point is currently confused by the main
            manuscript title, text and figures that generally refer to "ARFI"
            performance.  I believe the authors try to indicate that both
            B-Mode and ARFI data are used for the segmentation by referring to
            the "ultrasound" imaging performance, as in "… with a mean
            overestimation of of 16.82 ± 22.45\% by ultrasound imaging compared
            to MR volumes (Figure 6(b))." (pg 9 line 225), but this reader did
            not immediately appreciate that "ultrasound" referred to combined
            B-mode and ARFI segmentation as opposed to ARFI alone, and the
            "ARFI" labels throughout the rest of the text, tables, titles, and
            figures further confused this point.}

    \item \textit{How many readers performed the B-Mode and ARFI segmentation
            per image? per entire study data set? Were the readers trained, and
            if so, how?  Did different readers segment the B-Mode and ARFI data
            sets?  Did the same radiologist who segmented the MRI images also
            segment the B-Mode and ARFI data?}

    \item \textit{How would limits to ARFI resolution impact the results?}

    \item \textit{The authors explain that gross pathology weights and axis
            measurements were affected by the presence of peri-prostatic
            tissue.  Why could the peri-prostatic tissue not be differentiated
            from the prostatic tissue in histology after processing?  Wouldn't
            this enable better gold standard axis measurements?}

    \item \textit{Further to the point above, if the prostatic tissue axes
            could be measured after histological processing, could you perform
            an ROC analysis to assess ARFI/B-Mode versus MRI performance for
            axis measurement?}

    \item \textit{pg 12, line 329, change "compliment" to "complement".}

    \item \textit{Note that in my version of the manuscript, the figures and
            the captions written below them are blurry and difficult to read.
            This may just be a formatting issue with the automatic reviewer
            support system, but it is worth confirming that figures and their
            captions will be clear for publication.}

\end{itemize}

\textit{Reviewer: 2}

\textit{Comments to the Author}

\textit{This is a nice study of ARFI imaging of the prostate that compares ARFI images
to MRI. Three areas of the paper could be improved:}

\begin{itemize}
    \item \textit{the ellipsoidal prostate volume estimate from pathology could be
        improved by using the prostate segmentation on each of the whole mount
        pathology slides to align the data to the MRI prostate surface. The
        volume comparison is further complicated by the shrinkage of the
        prostate after fixation.}

    \item \textit{Image registration. The three orthogonal dimensions of the
            images do not mean much. The slice thickness in MRI causes
            significant errors along this measurement axis, while the
            lateral-to-lateral dimension in ARFI seems to be enlarged by the
            TRUS deformation (this is not mentioned in the discussion, but
            seems to be fairly obvious from Fig. 2 and 5, for example). A
            Jaccard index type of comparison would be preferable after rigid or
            possibly volume-preserving deformable registration.}

    \item \textit{discussion of prior work is inadequate. Volumetric
            acquisition and comparison of mechanical imaging with B-mode and
            MRI done before for vibro-elastography if not for ARFI and that
            work should be cited and discussed.}
\end{itemize}

\end{document}

### 2014-04-22 fig_mcneal_anatomy.tex
mcneal anatomy figure added to background

\begin{figure}
\centering
\includegraphics[width=0.75\linewidth]{figs/Mcneal_Zonal_Anatomy.jpg}
\caption{\textbf{Diagrams of McNeal’s zonal anatomy of the human
    prostate.~\cite{mcneal_path} It divides a human prostate into one
    nonglandular zone (the anterior fibromuscular stroma) and three glandular
    zones: central zone, peripheral zone and transition zone, denoted by CZ, PZ
    and TZ, respectively. (a) Coronal section of the prostate. V represents the
    verumontanum. (b) Sagittal section of the prostate. In the top part of the
    diagram, the seminar vesicles and ductus deferens enter the central zone
    and connect to the ejaculatory ducts, which merge with the urethra at the
    verumontanum. (Figure reprinted with permission.~\cite{Zhai2010a})}}
\label{fig:mcneal_anatomy} 
\end{figure}

### 2014-04-22 mr_background.tex
removed redundant mpMRI background

@@
 mpMRI had a PCa detection positive predictive value of 98\%.~\cite{Turkbey2012}
-Functional sequences include diffusion-weighted imaging (DWI), dynamic contrast
+\st{Functional sequences include diffusion-weighted imaging (DWI), dynamic contrast
 enhanced imaging (DCE-MRI) and MR spectroscopic imaging (MRSI).  While mpMRI is
@@
 anatomy, and for the purposes of comparing prostate anatomy between ARFI and MR
-images, T2WI were used exclusively in this study.
+images, T2WI were used exclusively in this study.}

### 2014-04-22 Response_to_R0_Reviews.tex
responded to anatomy figure
removed redundant mpMRI background
corrected complement grammar

@@
 
+    We agree with this suggestion, and we have added Figure 1 to the manuscript
+    that outlines the McNeals zonal anatomy of the prostaet in the coronal and
+    sagittal views.
+
     \item \textit{Pg 5, lines 109-112, are redundant with other text previously
             written in the background.}
+
+    We agree that this text is redundant, and we have removed it from this
+    section of the manuscript.
 
@@
 
-    \item \textit{pg 12, line 329, change "compliment" to "complement".}
+    \item \textit{pg 12, line 329, change ``compliment'' to ``complement''.}
+
+    Thank you for catching this grammatical error.  It actually existed in both
+    the abstract and conclusion sections, and it has been corrected in both
+    sections.
 

### 2014-04-22 conclusions.tex
corrected complement grammar

@@
 cause of these differences.  ARFI imaging is a promising low-cost, real-time
-imaging modality that can compliment MR imaging for diagnosis, treatment
+imaging modality that can complement MR imaging for diagnosis, treatment
 planning and management of PCa.

### 2014-04-22 Response_to_R0_Reviews.tex
indicated how manuscript revisions are delineated
ellipsoidal volume choice

@@
 the review below, with the original comments italicized, and the responses in
-normal text.
+normal text.  The revised manuscript has significant additions indicated with
+bold text and removed text has been struck through.
 
@@
 \begin{itemize}
-    \item \textit{the ellipsoidal prostate volume estimate from pathology could be
+    \item \textit{The ellipsoidal prostate volume estimate from pathology could be
         improved by using the prostate segmentation on each of the whole mount
@@
         prostate after fixation.}
+
+   Our ellipsoidal approximation of the prostate volume is a rough
+   approximation from the surgical pathology reports, but we felt it was our
+   best metric to approximate prostate size for the following reasons:
+
+   \begin{itemize}
+        \item The gross surgical pathology mesaurements used to make the
+            ellipsoidal volume approximation were done after peri-prostatic
+            tissue was dissected away, but before formalin fixation, avoiding
+            the complication of tissue shrinkage associated with formalin
+            fixation, which our pathologists estimated could be as much as 30\%
+            of the prostate volume.
+        \item Whole mount histology slides were explored for this purpose, but
+            they suffer from post-fixation contraction, $\sim$ 3 mm slice
+            thickiness, and slice tearing and folding that compromise accurate
+            cross-sectional area representations on a single slide.
+            Additionally, the whole mount histology slides require some degree
+            of image registration due to their arbitrary orientation on the
+            slides, adding another source of error.
+    \end{itemize}
+
+    Given the reasons above, we decided to use the tri-axial measurements in
+    gross pathologic evaluation as our closest approximation of the glands
+    total volume as it would have been during MR and ultrasound imaging.  We
+    have added text to Section II.B that details this choice for estimating our
+    prostate volumes:
+
+    \textbf{\input{path_volume_choice}}
 

### 2014-04-22 path_volume_choice.tex
ellipsoidal volume choice

An alternative approach to estimating the prostate volume was to reconstruct
the prostate capsule from serial whole mount histology slides of the prostate.
This approach, however, was not chosen because the formalin fixation process
can cause up to $\sim$30\% contraction of the tissue [BUCK CITATION], the
$\sim$3 mm slice thickness introduces error, and tearing / folding of tissue on
the whole mount slides introduce additional artifacts that need to be
compensated for in the volume estimation process.  Finally, differences in the
tissue orientation on each whole mount slide would require an image
registration process to align all of the slides, introducing another source of
error in the process.

### 2014-04-23 intro.tex
Stephen's comments
updated response about vibro-elastography

@@
 
-PCa diagnosis usually begins by screening with prostate specific antigen (PSA)
-and digital rectal examination (DRE).  More definitive diagnosis can be made by
+PCa diagnosis usually begins by screening with a prostate specific antigen (PSA)
+test and digital rectal examination (DRE).  More definitive diagnosis can be made by
 random transrectal ultrasound (TRUS)-guided biopsies, which are then used to
@@
 \invivo as compared to T2WI MR at 3T using an endorectal coil.
-Section~\ref{sect:background} provides an overview of MR and ARFI imaging in
-the prostate and an overall clinical motivation for prostate imaging.
-Section~\ref{sect:methods} describes the methods used to experimentally-acquire
-our imaging data, process the gross pathological specimens post radical
-prostatectomy, and image process our datasets.  The results of our analysis,
-including gross pathology and imaging prostate axis and volume estimates, are
-presented in Section~\ref{sect:results}, along with a statistical analysis of
-the bias and variability associated with each imaging modalities measurements,
-all of which is discussed in Section~\ref{sect:discussion}.
+\input{intro_vibroelastography}
+Section~\ref{sect:background} of this manuscript provides an overview of MR and
+ARFI imaging in the prostate and an overall clinical motivation for prostate
+imaging.  Section~\ref{sect:methods} describes the methods used to
+experimentally-acquire our imaging data, process the gross pathological
+specimens post radical prostatectomy, and image process our datasets.  The
+results of our analysis, including gross pathology and imaging prostate axis
+and volume estimates, are presented in Section~\ref{sect:results}, along with a
+statistical analysis of the bias and variability associated with each imaging
+modalities measurements, all of which is discussed in
+Section~\ref{sect:discussion}.

### 2014-04-23 Response_to_R0_Reviews.tex
updated response about vibro-elastography
clarified not using mp MR sequence for anatomic identification

@@
 \parskip=10pt
+
+\input{newcommands}
 
@@
             used?}
+
+    The Methods section of the manuscript describes the comprehensive
+    multiparametric MRI protocol that is used at our institution for imaging of
+    the prostate gland.  It does include T2-weighted (T2W) imaging for anatomy
+    as well as diffusion weighted images (DWI) and dynamic contrast enhanced
+    imaging (DCE-MRI) for functional information as well as lesion
+    characterization.  The T2W images are the highest resolution images in our
+    dataset and therefore, would give the most accurate information for
+    prostate size as compared with ARFI.  The DWI and DCE-MRI images are used
+    in conjunction with the T2W images for lesion characterization, but due to
+    their lower resolution, would likely not add much information to the T2W
+    images for the purposes of prostate size and segmentation.  Additional
+    research ongoing in our group does focus on lesion segmentation, and for
+    this application, information from all three sequences are used in
+    conjunction. All MRIs were performed using a standard extracellular MR
+    contrast agent (Magnevist, Bayer Pharma AG).
+
+    To clarify this point for the reader, we have added the following text to
+    Section III(c) of the manuscript:
+
+    \textbf{\input{mr_dwi_dce_methods}}
+
+    The contrast agent that was used has also been added to the manuscript.
 
@@
             work should be cited and discussed.}
+
+    We did mistakenly omit very relevant work that has been done with
+    vibro-elastography and MR imaging in the prostate, and we have included
+    this work in the manuscript's Introduction:
+
+    \textbf{\input{intro_vibroelastography}}
+
 \end{itemize}

### 2014-04-23 mr_methods.tex
clarified not using mp MR sequence for anatomic identification

@@
 images in the coronal, axial and sagittal planes.  Diffusion weighted images
-were obtained using multiple b-values and calculation of Apparent Diffusion
-Coefficient (ADC) maps was also performed.  Dynamic contrast enhanced MR
-sequences were obtained after administration of a weight-based dose of
-extracellular MR contrast agent with 4-5 second temporal resolution for 5-6
-minutes.
+(DWI) were obtained using multiple b-values and calculation of Apparent
+Diffusion Coefficient (ADC) maps was also performed.  Dynamic contrast enhanced
+MR (DCE-MRI) sequences were obtained after administration of a weight-based
+dose of extracellular MR contrast agent \textbf{(Magenvist\textregistered,
+Bayer Pharma AG)} with 4-5 second temporal resolution for 5-6 minutes.
+\textbf{\input{mr_dwi_dce_methods}}

### 2014-04-23 Response_to_R0_Reviews.tex
emphasized that the ER7B uses a side-fire array and is used endorectally

@@
             imaging was performed endorectally.}
+
+    We have modified the text describing the ER7B array as follows:
+
+    \textbf{\ldots side-fire array of an Acuson ER7B endorectal transducer.}
+
+    We have also emphasized that the imaging was performed endorectally by
+    modifying the following statement in the same section:
+
+    \textbf{3D volumetric imaging data was acquired endorectally\ldots}
 

### 2014-04-23 arfi_methods.tex
emphasized that the ER7B uses a side-fire array and is used endorectally

@@
 SC2000\texttrademark~ultrasound scanner (Siemens Healthcare, Ultrasound Business Unit,
-Mountain View, CA, USA) and the longitudinal array of an Acuson ER7B
-transducer.  The ARFI imaging sequence was comprised of
+Mountain View, CA, USA) and the \textbf{side-fire array of an Acuson ER7B
+endorectal transducer}.  The ARFI imaging sequence was comprised of
 standard B-mode ultrasonic imaging, or tracking beams, and pushing beams. For
@@
 
-3D volumetric imaging data was acquired using mechanical rotation of the ER7B
-transducer between sequential imaging frames, sweeping across the lateral
-extent of the prostate.  This rotation setup utilized a CIVCO
-Micro-Touch\texttrademark stabilizer (CIVCO Medical Solutions, Kalona, IA USA)
-that allows for complete 6-axis degrees of freedom for manual positioning of
-the transducer to sweep through the entire prostate during imaging.  Sequential
-imaging frames were $\sim$1\degree apart, depending on the absolute size of the
-prostate.  A custom optical angular feedback transduction circuit utilizing a
-reflective linear strip with 212 lines-per-inch (LPI) resolution (US Digital,
-Vancouver, WA, USA) with a QSB-S Quadrature-to-USB adapter to achieve 9 line /
-degree resolution.  Rotation was performed with a 141 oz-in torque stepper
-motor with planetary gearbox (Model \#11YPG202S-LW4-R27, Anaheim Automation,
-Anaheim, CA, USA) (Figure~\ref{fig:setup_annotated}) to achieve accurate
-spatial localization of the imaging frames in the 3D dataset.
+\textbf{3D volumetric imaging data was acquired endorectally} using mechanical
+rotation of the ER7B transducer between sequential imaging frames, sweeping
+across the lateral extent of the prostate.  This rotation setup utilized a
+CIVCO Micro-Touch\texttrademark stabilizer (CIVCO Medical Solutions, Kalona, IA
+USA) that allows for complete 6-axis degrees of freedom for manual positioning
+of the transducer to sweep through the entire prostate during imaging.
+Sequential imaging frames were $\sim$1\degree apart, depending on the absolute
+size of the prostate.  A custom optical angular feedback transduction circuit
+utilizing a reflective linear strip with 212 lines-per-inch (LPI) resolution
+(US Digital, Vancouver, WA, USA) with a QSB-S Quadrature-to-USB adapter to
+achieve 9 line / degree resolution.  Rotation was performed with a 141 oz-in
+torque stepper motor with planetary gearbox (Model \#11YPG202S-LW4-R27, Anaheim
+Automation, Anaheim, CA, USA) (Figure~\ref{fig:setup_annotated}) to achieve
+accurate spatial localization of the imaging frames in the 3D dataset.
 

### 2014-04-23 Response_to_R0_Reviews.tex
responded to track Tx focus

@@
             focal position. (pg 6 lines 163 and 154, respectively).}
+
+    The tracking beam transmit was focused at 60 mm to insonify a large lateral
+    field of view and observe the off-axis shear wave propagation with the
+    remaining 12 parallel receive beams that were not utilized in this work
+    (lines 160-161). For each track beam, dynamic receive and receive aperture
+    growth with a F/0.5 configuration were used to minimize the effect of the
+    deep transmit focus.  Since only the ARFI imaging vectors were presented in
+    this work--not the shear wave data--we have not ammended the manuscript
+    text to avoid confusing the reader with details not relevant to the data
+    presented in this manuscript.
 

### 2014-04-23 Response_to_R0_Reviews.tex
disp normalization still in progress

@@
 
+    This is a great comment, and one that we have considered internally during
+    the development of our methodology for this study.  The normalization
+    profile for ARFI images is similar in concept to time gain compensation
+    (TGC) for B-mode images. It is used to apply gain throughout the image to
+    provide a more uniform appearance of the image through depth, which is
+    necessary to account for tissue attenuation and focal gain. The
+    normalization does not change the raw image data since it is applying
+    digital gain. Additionally, when a reader views the images, gain can be
+    added or removed as desired to further improve the image quality.
+    Therefore, on a patient-by-patient basis, the normalization profile can be
+    adjusted to account for differences in each individual to reduce the impact
+    of having an incorrect normalization profile.
+
+    Specifically, for this work, two calibrated CIRS (Norfolk, VA) elasticity
+    phantoms were used to determine the normalization factor; the background
+    stiffness of both materials was nominally 10 kPa Young's modulus, and the
+    acoustic attenuation of the two phantoms were 0.49 and 0.72 dB/cm/MHz,
+    respectively. The displacement-through-time profiles from each phantom were
+    averaged together to approximate a 10 kPa Young's modulus with a 0.6
+    dB/cm/MHz attenuation. The phantom stiffness is slightly lower than the
+    12.3 kPa Young's modulus reported as healthy peripheral zone tissue [1].
+    The attenuation coefficient for prostatic tissue has been reported as
+    0.54-1.04 dB/cm/MHz [2], with additional support that the attenuation is at
+    the lower end of that range [3]. Thus, the normalization profile closely
+    approximates the properties of healthy peripheral zone prostatic tissue,
+    but the transition and central zones as well as pathology such as cancer,
+    BPH, or atrophy can all potentially have different mechanical and
+    acoustical properties.
+
+    [1] Zhai, Liang, et al. ``Correlation between SWEI and ARFI image findings
+    in ex vivo human prostates.'' Ultrasonics Symposium (IUS), 2009 IEEE
+    International. IEEE, 2009.
+
+    [2] K.J Parker, S.R Huang, R.M Lerner et al. Elastic and ultrasonic
+    properties of the prostate IEEE Ultrason Sympos Proc, 2 (1993), pp.
+    1035–1038
+
+    [3] A.E Worthington, J Trachtenberg, M.D Sherar, Ultrasound properties of
+    human prostate tissue during heating, Ultrasound in Medicine \& Biology,
+    Volume 28, Issue 10, October 2002, Pages 1311-1318.
+
     \item \textit{Pg 8 lines 205-214 describe that both B-Mode (for prostate

### 2014-04-24 Response_to_R0_Reviews.tex
responded about number of segmenters for each imaging modality

@@
 
+    All of the B-mode and ARFI images were segemented by two individuals (ZAM,
+    TJG), and reviewed by a single, experienced ARFI imaging research (KRN).
+    There was no formal training in reading the ARFI images since this work is
+    our first effort to analyze our ability to evaluate our accuracy in
+    identifying structures in ARFI prostate images.  The same individuals
+    segmented both the B-mode images (for the capsules) and ARFI images (for
+    the central glands).  All of the MR images were segmented by an experienced
+    radiology fellow (CK) and reviewed by an experienced prostate radiologist
+    (RG).  We have added the following text to the manuscript to delineate the
+    readers / image segmenters for each part of the study (Section III(E.1)):
+
+    Axial MR T2WI images were manually segmented by a \textbf{single
+        radiologist (CK)} using the smooth polygon tool in ITK-SNAP\ldots
+
+    \textbf{All MR image segmentations were ultimately reviewed by an
+        experienced prostate radiologist (RTG).}
+
+    And Section III(E.2):
+
+    Sagittal B-mode and ARFI ultrasound image stacks were segmented in 3D
+    Slicer \textbf{by two researchers (TJG \& ZAM)}\ldots
+
+    \textbf{All B-mode and ARFI image segmentations were reviewed by an
+        experienced ARFI imaging researcher (KRN).}
+
     \item \textit{How would limits to ARFI resolution impact the results?}

### 2014-04-24 image_proc_methods.tex
responded about number of segmenters for each imaging modality

@@
 \subsubsection{MR Image Segmentation and Modeling}
-Axial MR T2WI images were manually segmented by an experienced radiologist
-using the smooth polygon tool in ITK-SNAP
-(http://www.itksnap.org)~\cite{Yushkevich2006}, using unique labels for
-the PZ, CG and AFS. The gland was segmented from base-to-apex.  The base was
-identified below the bladder, and subsequent images were segmented until the
-last slice with visible prostatic tissue was identified caudally. The CG,
-urethra, PZ and AFS were segmented independently according to their
-well-established anatomical characteristics on
+Axial MR T2WI images were manually segmented by a \textbf{single radiologist
+    (CK)} using the smooth polygon tool in ITK-SNAP
+(http://www.itksnap.org)~\cite{Yushkevich2006}, using unique labels for the PZ,
+CG and AFS. The gland was segmented from base-to-apex.  The base was identified
+below the bladder, and subsequent images were segmented until the last slice
+with visible prostatic tissue was identified caudally. The CG, urethra, PZ and
+AFS were segmented independently according to their well-established anatomical
+characteristics on
 T2WI.~\cite{Barentsz2012,Jung2012,Poon1985,Hricak2007,Bonekamp2011} The PZ was
@@
 When visible, the AFS was identified by its low T2 signal intensity and its
-location anterior to the CG.
+location anterior to the CG.  \textbf{All MR image segmentations were
+    ultimately reviewed by an experienced prostate radiologist (RTG).}
 
@@
 \subsubsection{Ultrasound Image Segmentation and Modeling}
-Sagittal B-mode and ARFI ultrasound image stacks were segmented in 3D Slicer
-using the \verb+draweffect+ tool to model and measure the prostate capsule and
+\Sagittal B-mode and ARFI ultrasound image stacks were segmented in 3D Slicer
+\textbf{by two researchers (TJG \& ZAM)} using the \verb+draweffect+ tool to model and measure the prostate capsule and
 CG volume and tri-axial dimensions (Figure~\ref{fig:arfi_segs}).  The sagittal
@@
 dimensions were then measured using the same procedure as the MR images.
+\textbf{All B-mode and ARFI image segmentations were reviewed by an experienced
+    ARFI imaging researcher (KRN).}
 

### 2014-04-24 path_volume_choice.tex
updated path contraction text and image segmentation

@@
 This approach, however, was not chosen because the formalin fixation process
-can cause up to $\sim$30\% contraction of the tissue [BUCK CITATION], the
-$\sim$3 mm slice thickness introduces error, and tearing / folding of tissue on
-the whole mount slides introduce additional artifacts that need to be
-compensated for in the volume estimation process.  Finally, differences in the
-tissue orientation on each whole mount slide would require an image
-registration process to align all of the slides, introducing another source of
-error in the process.
+can cause 0-25\% contraction of the tissue, the $\sim$3 mm slice thickness
+introduces error, and tearing / folding of tissue on the whole mount slides
+introduce additional artifacts that need to be compensated for in the volume
+estimation process.  Finally, differences in the tissue orientation on each
+whole mount slide would require an image registration process to align all of
+the slides, introducing another source of error in the process.

### 2014-04-24 Response_to_R0_Reviews.tex
added TO DO holders and response to image quality
more ARFI normalization info

@@
 
+    The following text has been added to Section III(C):
+
+    \input{arfi_normalization}
+
     \item \textit{Pg 8 lines 205-214 describe that both B-Mode (for prostate
@@
             figures further confused this point.}
+
+    TO DO
 
@@
 
+    TO DO
+
     \item \textit{The authors explain that gross pathology weights and axis
@@
 
+    TO DO
+
     \item \textit{Further to the point above, if the prostatic tissue axes
@@
             axis measurement?}
+
+    TO DO
 
@@
             captions will be clear for publication.}
+
+    High resolution images were uploaded for the review process, but their
+    quality was definitely degraded during the rendering of the review PDF.  We
+    will make sure that all images are propoerly reproduced at full resolution
+    in new PDFs that are rendered of this manuscript.
 
@@
 
+    TO DO
+
     \item \textit{discussion of prior work is inadequate. Volumetric

### 2014-04-24 arfi_methods.tex
more ARFI normalization info

@@
 Displacement estimation was performed using a phase-shift estimator on the
-beamformed in-phase and quadrature (IQ) data~\cite{Loupas95,pinton06}. The ARFI
-data were then normalized as a function of depth to account for attenuation and
-focal gain effects.  This normalization was performed using a displacement
-profile measured in a homogeneous tissue-mimicking phantom that was low-pass
-filtered with a cutoff-frequency of 0.8 mm$^{-1}$.  This filtered normalization
-profile was applied to all displacements in the entire data set at each time
-step.
+beamformed in-phase and quadrature (IQ) data~\cite{Loupas95,pinton06}.
+\input{arfi_normalization}
 

### 2014-04-24 arfi_normalization.tex
more ARFI normalization info

\textbf{The ARFI data were normalized as a function of depth to account for   
attenuation and focal gain effects to ensure that ARFI image brightness
differences are more directly related to stiffness differences.  This                
normalization was performed using a displacement profiles measured in two            
homogeneous, elastic tissue-mimicking phantoms (CIRS, Norfolk, VA) with a          
stiffness of 10.0 kPa and acoustic attenuations of 0.49 and 0.72 dB/cm/MHz.        
The mean displacement profiles from each phantom were averaged together to
achieve an effective attenuation of 0.6 dB/cm/MHz, which is within the measured
range of attenuation in healthy peripheral zone prostate
tissue~\cite{Parker1993}.  This mean displacement profile was low-pass filtered
with a cutoff-frequency of 0.8 mm$^{-1}$ and applied to all displacements in
the entire dataset at each time step.}

### 2014-04-24 Response_to_R0_Reviews.tex
more responses in progress
updated text to explicitely state that B-mode images were used with ARFI images for all of the total gland / capsule approximations

@@
 
-    TO DO
+    This comment is greatly appreciated since the use of both B-mode and ARFI
+    imaging for each segmentation was an important part of our methodology, and
+    we (1) did not do enough to emphasize this point, and (2) confused matters
+    by refering to `ARFI' exclusively in the results.  With respect to how the
+    data were combined, we have added the following text to the manuscript
+    (Section III(E.2)):
+
+    \input{bmode_arfi_coregistered}
+
+    We also agree that our exclusive use of 'ARFI' for many of the tables,
+    figures, manuscript title, and text was confusing, so we have updated all
+    references to appropriate include B-mode when appropriate (not delineated
+    here in the response due to the number of instances that were edited, but
+    all manuscript changes are indicated in \textbf{bold} typeface).  The title
+    of the manuscript has also been updated to reflect the importance of B-mode
+    imaging in this process, and the abstract and conclusions have also been
+    explicitely updated to include the fact that B-mode and ARFI imaging data
+    were used in this study.
 

### 2014-04-24 abstract.tex
corrected complement grammar
updated text to explicitely state that B-mode images were used with ARFI images for all of the total gland / capsule approximations

@@
 diagnosis and management by using tissue stiffness to evaluate prostate zonal
-anatomy and lesions.  MR and ARFI \invivo imaging datasets were compared to one
-another and with gross pathology measurements made immediately after radical
-prostatectomy.  Images were manually segmented in 3D Slicer to delineate the
-central gland (CG) and prostate capsule, and 3D models were rendered to
-evaluate zonal anatomy dimensions and volumes.  Both imaging modalities showed
-good correlation between estimated organ volume and gross pathologic weights.
-Ultrasound and MR total prostate volumes were well-correlated (R$^2$ =
-\MRarfiVolTotalRsq), but ARFI images yielded prostate volumes that were larger
-(\MRarfiVolTotalMeanDiff\% $\pm$ \MRarfiVolTotalStdDiff\%) than MR images, due
-to overestimation of the lateral dimension (\ARFImrTotalLatLatMeanPct $\pm$
-\ARFImrTotalLatLatStdPct\%), with less significant differences in the other
-dimensions (\ARFImrTotalAntPostMeanPct $\pm$ \ARFImrTotalAntPostStdPct\%,
+anatomy and lesions.  MR and \textbf{B-mode/}ARFI \invivo imaging datasets were
+compared to one another and with gross pathology measurements made immediately
+after radical prostatectomy.  Images were manually segmented in 3D Slicer to
+delineate the central gland (CG) and prostate capsule, and 3D models were
+rendered to evaluate zonal anatomy dimensions and volumes.  Both imaging
+modalities showed good correlation between estimated organ volume and gross
+pathologic weights.  Ultrasound and MR total prostate volumes were
+well-correlated (R$^2$ = \MRarfiVolTotalRsq), but \textbf{B-mode} images
+yielded prostate volumes that were larger (\MRarfiVolTotalMeanDiff\% $\pm$
+\MRarfiVolTotalStdDiff\%) than MR images, due to overestimation of the lateral
+dimension (\ARFImrTotalLatLatMeanPct $\pm$ \ARFImrTotalLatLatStdPct\%), with
+less significant differences in the other dimensions
+(\ARFImrTotalAntPostMeanPct $\pm$ \ARFImrTotalAntPostStdPct\%,
 anterior-to-posterior, and \ARFImrTotalApexBaseMeanPct $\pm$
@@
 overestimation of the lateral dimension (\ARFImrCentralLatLatMeanPct $\pm$
-\ARFImrCentralLatLatStdPct\%).  ARFI imaging yielded prostate volumes and
-dimensions that were well-correlated with MR T2WI estimates, with biases in the
-lateral dimension due to poor contrast caused by peri-prostatic fat.  ARFI
-imaging is a promising low-cost, portable, real-time modality that can
-compliment mpMRI for PCa diagnosis, treatment planning and management.
+\ARFImrCentralLatLatStdPct\%).  B-mode/ARFI imaging yielded prostate volumes
+and dimensions that were well-correlated with MR T2WI estimates, with biases in
+the lateral dimension due to poor contrast caused by peri-prostatic fat.
+\textbf{B-mode combined with} ARFI imaging is a promising low-cost, portable,
+real-time modality that can complement mpMRI for PCa diagnosis, treatment
+planning and management.

### 2014-04-24 arfi_methods.tex
updated text to explicitely state that B-mode images were used with ARFI images for all of the total gland / capsule approximations

@@
-\subsection{ARFI Imaging}
-Experimental ARFI imaging data were acquired using a modified Siemens Acuson
+\subsection{B-mode/ARFI Imaging}
+Experimental \textbf{B-mode and} ARFI imaging data were acquired using a modified Siemens Acuson
 SC2000\texttrademark~ultrasound scanner (Siemens Healthcare, Ultrasound Business Unit,

### 2014-04-24 conclusions.tex
updated text to explicitely state that B-mode images were used with ARFI images for all of the total gland / capsule approximations

--- 2013-12-30 conclusions.tex
@@
 \section{Conclusions}
-The delineation of prostate zonal anatomy in ARFI images has been compared with
-the established methods for identifying zonal anatomy using MR T2W images.
-Both imaging modalities showed moderate correlations between estimated organ
-volume and gross pathologic weights, and ARFI:MR total prostate gland volumes
-were well-correlated (R$^2$ = \MRarfiVolTotalRsq), but ARFI images yielded
-prostate volumes that were, on average, larger (\MRarfiVolTotalMeanDiff\%~$\pm$
-\MRarfiVolTotalStdDiff\%) than MR images, primarily due to over-estimation of
-the lateral dimension of the prostate total gland (\ARFImrTotalLatLatMeanPct~
-$\pm$ \ARFImrTotalLatLatStdPct\%), while differences in the other dimensions
-were less significant contributors (\ARFImrTotalAntPostMeanPct~$\pm$
-\ARFImrTotalAntPostStdPct\%~and \ARFImrTotalApexBaseMeanPct~$\pm$
-\ARFImrTotalApexBaseStdPct\%).  The central zone volumes of ARFI and MR images
-were also well-correlated (R$^2$ = \MRarfiVolCentralRsq), with minimal volume
-bias between the imaging modalities, but significant variability case-to-case
-(\MRarfiVolCentralMeanDiff~$\pm$ \MRarfiVolCentralStdDiff\%).  Central zone
-volume differences were, again, strongly attributed to over-estimation of the
-lateral dimension (\ARFImrCentralLatLatMeanPct~$\pm$
-\ARFImrCentralLatLatStdPct\%), with a significant underestimation of the
-apex-to-base dimension (\ARFImrCentralAntPostMeanPct~$\pm$
-\ARFImrCentralAntPostStdPct\%).  Strong variability in CG volumes is believed
-to be related to the extent of BPH for select cases.  Overall, ARFI imaging of
-the prostate yielded prostate volumes and dimensions that were correlated with
-MR T2WI estimates, with differences in the lateral and apex-to-base dimensions.
-Poor image contrast between the prostate and peri-prostatic fat may be one
-cause of these differences.  ARFI imaging is a promising low-cost, real-time
-imaging modality that can compliment MR imaging for diagnosis, treatment
-planning and management of PCa.
+The delineation of prostate zonal anatomy in \textbf{B-mode/}ARFI images has
+been compared with the established methods for identifying zonal anatomy using
+MR T2W images.  Both imaging modalities showed moderate correlations between
+estimated organ volume and gross pathologic weights, and \textbf{B-mode}:MR
+total prostate gland volumes were well-correlated (R$^2$ = \MRarfiVolTotalRsq),
+but ARFI images yielded prostate volumes that were, on average, larger
+(\MRarfiVolTotalMeanDiff\%~$\pm$ \MRarfiVolTotalStdDiff\%) than MR images,
+primarily due to over-estimation of the lateral dimension of the prostate total
+gland (\ARFImrTotalLatLatMeanPct~ $\pm$ \ARFImrTotalLatLatStdPct\%), while
+differences in the other dimensions were less significant contributors
+(\ARFImrTotalAntPostMeanPct~$\pm$ \ARFImrTotalAntPostStdPct\%~and
+\ARFImrTotalApexBaseMeanPct~$\pm$ \ARFImrTotalApexBaseStdPct\%).  The central
+gland volumes of ARFI and MR images were also well-correlated (R$^2$ =
+\MRarfiVolCentralRsq), with minimal volume bias between the imaging modalities,
+but significant variability case-to-case (\MRarfiVolCentralMeanDiff~$\pm$
+\MRarfiVolCentralStdDiff\%).  Central gland volume differences were, again,
+strongly attributed to over-estimation of the lateral dimension
+(\ARFImrCentralLatLatMeanPct~$\pm$ \ARFImrCentralLatLatStdPct\%), with a
+significant underestimation of the apex-to-base dimension
+(\ARFImrCentralAntPostMeanPct~$\pm$ \ARFImrCentralAntPostStdPct\%).  Strong
+variability in CG volumes is believed to be related to the extent of BPH for
+select cases.  Overall, B-mode/ARFI imaging of the prostate yielded prostate
+volumes and dimensions that were correlated with MR T2WI estimates, with
+differences in the lateral and apex-to-base dimensions.  Poor image contrast
+between the prostate and peri-prostatic fat may be one cause of these
+differences.  \textbf{B-mode imaging combined with} ARFI imaging is a promising
+low-cost, real-time imaging modality that can complement MR imaging for
+diagnosis, treatment planning and management of PCa.

### 2014-04-24 disc_arfi_mr.tex
updated text to explicitely state that B-mode images were used with ARFI images for all of the total gland / capsule approximations

@@
-\subsection{ARFI and MR T2WI Agreement}
-This study demonstrated that ARFI imaging delineates zonal anatomy of the
-prostate well when compared to MR T2WI with only slight overestimation of ARFI
-total and central gland volume (Figure~\ref{fig:mr_arfi_volumes}).
-Establishing accurate delineation of the prostate's zonal anatomy is important
-because PCa has strong correlations with location in the prostate, specifically
-in the peripheral zone, and accurate capsule delineation will facilitate
-future efforts to register 3D imaging datasets between ARFI and MR imaging.
+\subsection{B-mode/ARFI and MR T2WI Agreement}
+This study demonstrated that \textbf{B-mode/}ARFI imaging delineates zonal
+anatomy of the prostate well when compared to MR T2WI with only slight
+overestimation of \textbf{B-mode} total and \textbf{ARFI} central gland volume
+(Figure~\ref{fig:mr_arfi_volumes}).  Establishing accurate delineation of the
+prostate's zonal anatomy is important because PCa has strong correlations with
+location in the prostate, specifically in the peripheral zone, and accurate
+capsule delineation will facilitate future efforts to register 3D imaging
+datasets between ARFI and MR imaging.
 
-We did not expect perfect agreement between MR and ARFI imaging volume or
-tri-axial dimensions since each modality applies a different amount of external
-compression to the rectal wall, which then deforms the prostate.  The
-endorectal coil places pressure on the prostate that is dependent on the
+We did not expect perfect agreement between MR and \textbf{ultrasound} imaging
+volumes or tri-axial dimensions since each modality applies a different amount
+of external compression to the rectal wall, which then deforms the prostate.
+The endorectal coil places pressure on the prostate that is dependent on the
 patient's anatomy and overall prostate size, while pressure on the prostate
@@
 
-The lateral boundaries of the prostate in the U/S dataset were overestimated
-relative to MR T2WI, increasing ARFI total volume relative to MR. The lateral
-overestimation was a result of poor contrast between the PZ and peri-prostatic
-fat in the B-mode images, and image decimation along the lateral dimension in
-both ARFI and B-mode images. Without image decimation, the elevation and axial
-orthogonal planes can be used to correct errors introduced by the poor contrast
-between the PZ and peri-prostatic fat. However, with image decimation,
-segmentation changes in the elevation and axial orthogonal planes were limited
-to the lower bound on the model precision of, again, $\sim$3 mm.  While reduced
-decimation could improve the precision of the lateral measurements, this would
-have introduced significant overhead in the manual segmentation process, with
-diminishing returns given the spatial smoothing of the modeling algorithms that
-were utilized.
+The lateral boundaries of the prostate in the \textbf{ultrasound} dataset were
+overestimated relative to MR T2WI, increasing \textbf{B-mode} total volume
+relative to MR. The lateral overestimation was a result of poor contrast
+between the PZ and peri-prostatic fat in the B-mode images, and image
+decimation along the lateral dimension in both ARFI and B-mode images. Without
+image decimation, the elevation and axial orthogonal planes can be used to
+correct errors introduced by the poor contrast between the PZ and
+peri-prostatic fat. However, with image decimation, segmentation changes in the
+elevation and axial orthogonal planes were limited to the lower bound on the
+model precision of, again, $\sim$3 mm.  While reduced decimation could improve
+the precision of the lateral measurements, this would have introduced
+significant overhead in the manual segmentation process, with diminishing
+returns given the spatial smoothing of the modeling algorithms that were
+utilized.
 
@@
 heterogeneity in this region~\cite{Gupta2013}, though we had good agreement
-between ARFI and MR T2W images in the anterior-to-posterior dimension
+between ultrasound and MR T2W images in the anterior-to-posterior dimension
 (Figure~\ref{fig:mr_arfi_path_axes}(b), Table~\ref{tab:mr_arfi_axes_error}).
 
-ARFI imaging consistently underestimated the extent of the prostate capsule and
-central gland apex-to-base dimension relative to MR T2WI
+\textbf{B-mode/}ARFI imaging consistently underestimated the extent of the
+prostate capsule and central gland apex-to-base dimension relative to MR T2WI
 ((Figure~\ref{fig:mr_arfi_path_axes}(c), Table~\ref{tab:mr_arfi_axes_error}).
-In fact, it is likely the degree of ARFI underestimation relative to the actual
-apex-to-base prostate length is even greater than suggested by the MR
-comparison as this was the dimension with the worst spatial resolution in MR
-imaging, with a typical MR T2WI slice thickness of $\sim$3 mm for these
-studies.  While higher-resolution MR images could be obtained and analyzed for
-more precise comparisons in this imaging dimension, such data were not available
-for this study.
+In fact, it is likely the degree of \textbf{B-mode/}ARFI underestimation
+relative to the actual apex-to-base prostate length is even greater than
+suggested by the MR comparison as this was the dimension with the worst spatial
+resolution in MR imaging, with a typical MR T2WI slice thickness of $\sim$3 mm
+for these studies.  While higher-resolution MR images could be obtained and
+analyzed for more precise comparisons in this imaging dimension, such data were
+not available for this study.

### 2014-04-24 fig_mr_arfi_path_axes.tex
draft edits and spelling
modified format and cover letter for Ultrasonic Imaging submission
updated text to explicitely state that B-mode images were used with ARFI images for all of the total gland / capsule approximations

@@
-\begin{figure}[htb!]
+\begin{figure}
 \centering
@@
     anatomic axes: lateral-to-lateral (a), anterior-to-posterior (b) and
-    apex-to-base (c).  The correlation between the MR and ARFI imaging axis
-    measurements was performed in each orientation for the total gland (red
-    crosses) and central gland (blue circles).  The black dashed-line represents
-    the projection of perfectly-correlated measurements between imaging and
-    patholoy.  The over-/under-estimation of each imaging modality relative to
+    apex-to-base (c).  The correlation between the MR and B-mode/ARFI imaging
+    axis measurements was performed in each orientation for the total gland
+    (red crosses) and central gland (blue circles).  \textbf{B-mode images were
+        used to estimate the total gland volume, and ARFI images were used to
+        estimate the central glands.} The black dashed-line represents the
+    projection of perfectly-correlated measurements between imaging and
+    pathology.  The over-/under-estimation of each imaging modality relative to
     gross pathology and each other is summarized in

### 2014-04-24 fig_mr_arfi_volumes.tex
Stephen's comments
modified format and cover letter for Ultrasonic Imaging submission
updated text to explicitely state that B-mode images were used with ARFI images for all of the total gland / capsule approximations

--- 2013-12-09 fig_mr_arfi_volumes.tex
@@
 \end{tabular}
-\caption{Comparison of MR and ARFI zonal anatomy volume estimates from
-    manually-segmented images.  Total prostate gland volumes ranged from
-    19.6--71.0 cm$^3$ based on MR image models (a), with a moderate correlation
-    between MR and ARFI imaging for both the total gland volume (R$^2$ =
-    \MRarfiVolTotalRsq, (b)) and the central gland (R$^2$ =
-    \MRarfiVolCentralRsq, (c)).  The ARFI total gland volumes had a positive
-    bias compared with the MR volumes (\MRarfiVolTotalMeanDiff~$\pm$
-    \MRarfiVolTotalStdDiff\% (b)), while the ARFI central zone volumes were on
+\caption{Comparison of MR and \textbf{B-mode/}ARFI zonal anatomy volume
+    estimates from manually-segmented images.  Total prostate gland volumes
+    ranged from 19.6--71.0 cm$^3$ based on MR image models (a), with good
+    correlation between MR and ultrasound imaging for both the total gland
+    volume (R$^2$ = \MRarfiVolTotalRsq, (b)) and the central gland (R$^2$ =
+    \MRarfiVolCentralRsq, (c)).  The \textbf{B-mode} total gland volumes had a
+    positive bias compared with the MR volumes (\MRarfiVolTotalMeanDiff~$\pm$
+    \MRarfiVolTotalStdDiff\% (b)), while the ARFI central gland volumes were on
     average more similar (\MRarfiVolCentralMeanDiff~$\pm$

### 2014-04-24 fig_mr_arfi_weight.tex
modified format and cover letter for Ultrasonic Imaging submission
updated text to explicitely state that B-mode images were used with ARFI images for all of the total gland / capsule approximations

@@
     correlation between the reconstructed volumes and prostate weight (R$^2$ =
-    \weightMRrsq), while volumes reconstructed from ARFI images (green, O)
-    showed weaker correlation (R$^2$ = \weightARFIrsq) (b).  Weaker
-    correlations existed between both T2WI MR and ARFI image volumes and
-    approximated ellipsoidal prostate pathology volumes (R$^2$ =
+    \weightMRrsq), while volumes reconstructed from \textbf{B-mode} images
+    (green, O) showed weaker correlation (R$^2$ = \weightARFIrsq) (b).  Weaker
+    correlations existed between both T2WI MR and \textbf{B-mode} image volumes
+    and approximated ellipsoidal prostate pathology volumes (R$^2$ =
     \pathVolMRrsq~and \pathVolARFIrsq, respectively) (c).  There was one

### 2014-04-24 fig_setup_annotated.tex
modified format and cover letter for Ultrasonic Imaging submission
updated text to explicitely state that B-mode images were used with ARFI images for all of the total gland / capsule approximations

@@
-\begin{figure}[htb!]
+\begin{figure}
 \centering
 \includegraphics[width=0.75\textwidth]{figs/setup_annotated.png}
-\caption{The experimental setup used for ARFI imaging in the operating suite,
-    including the Siemens SC2000 scanner with the ER7B endorectal transducer
-    sitting in a custom cradle of a CIVCO Micro-Touch\texttrademark stabilizer
-    with optical feedback controlling a stepper motor to acquire 3D ultrasound
-    imaging data \invivo.  Custom-written Python code was run on the scanner to
-    trigger B-mode and ARFI imaging acquisition sequences relative to completed
-    rotation increments of $\sim$1\degree.}
+\caption{The experimental setup used for \textbf{B-mode/}ARFI imaging in the
+    operating suite, including the Siemens SC2000 scanner with the ER7B
+    endorectal transducer sitting in a custom cradle of a CIVCO
+    Micro-Touch\texttrademark stabilizer with optical feedback controlling a
+    stepper motor to acquire 3D ultrasound imaging data \invivo.
+    Custom-written Python code was run on the scanner to trigger B-mode and
+    ARFI imaging acquisition sequences relative to completed rotation
+    increments of $\sim$1\degree.}
 \label{fig:setup_annotated} 

### 2014-04-24 image_proc_methods.tex
updated path contraction text and image segmentation
updated text to explicitely state that B-mode images were used with ARFI images for all of the total gland / capsule approximations

@@
 \subsubsection{Ultrasound Image Segmentation and Modeling}
-\Sagittal B-mode and ARFI ultrasound image stacks were segmented in 3D Slicer
-\textbf{by two researchers (TJG \& ZAM)} using the \verb+draweffect+ tool to model and measure the prostate capsule and
-CG volume and tri-axial dimensions (Figure~\ref{fig:arfi_segs}).  The sagittal
-imaging plane was chosen for segmentation because it has a greater number of
-easily identifiable anatomical structures:  seminal vesicles in the left and
-right posterior base, urethra in the mid apex, surrounding fibromuscular tissue
-in the anterior, and rectum in the posterior. Prostate capsule was segmented in
-B-mode as this modality offers both high contrast between prostatic tissue and
-surrounding peri-prostatic fat, and better SNR in the anterior capsule region
-than ARFI images. ARFI images were used for CG segmentation.  Prior to
-segmentation, both the B-mode and ARFI image stacks were down-sampled along the
-sagittal axis to a slice thickness of 3 mm to reduce the number of
-manually-segmented frames (18 vs. 360 slices / image stack) and match the MRI
-slice thickness.
+Sagittal B-mode and ARFI ultrasound image stacks were segmented in 3D Slicer
+\textbf{by two researchers (TJG \& ZAM)} using the \verb+draweffect+ tool to
+model and measure the prostate capsule and CG volume and tri-axial dimensions
+(Figure~\ref{fig:arfi_segs}).  \input{bmode_arfi_coregistered} The sagittal imaging plane was chosen for
+segmentation because it has a greater number of easily identifiable anatomical
+structures:  seminal vesicles in the left and right posterior base, urethra in
+the mid apex, surrounding fibromuscular tissue in the anterior, and rectum in
+the posterior. Prostate capsule was segmented in B-mode as this modality offers
+both high contrast between prostatic tissue and surrounding peri-prostatic fat,
+and better SNR in the anterior capsule region than ARFI images. ARFI images
+were used for CG segmentation.  Prior to segmentation, both the B-mode and ARFI
+image stacks were down-sampled along the sagittal axis to a slice thickness of
+3 mm to reduce the number of manually-segmented frames (18 vs. 360 slices /
+image stack) and match the MRI slice thickness.
 

### 2014-04-24 intro.tex
updated text to explicitely state that B-mode images were used with ARFI images for all of the total gland / capsule approximations

@@
 
-The goal of this study is to evaluate the ability of ARFI imaging to delineate
-prostate zonal anatomy, specifically central gland and total prostate gland
-\invivo as compared to T2WI MR at 3T using an endorectal coil.
-\input{intro_vibroelastography}
-Section~\ref{sect:background} of this manuscript provides an overview of MR and
-ARFI imaging in the prostate and an overall clinical motivation for prostate
-imaging.  Section~\ref{sect:methods} describes the methods used to
-experimentally-acquire our imaging data, process the gross pathological
-specimens post radical prostatectomy, and image process our datasets.  The
-results of our analysis, including gross pathology and imaging prostate axis
-and volume estimates, are presented in Section~\ref{sect:results}, along with a
-statistical analysis of the bias and variability associated with each imaging
-modalities measurements, all of which is discussed in
-Section~\ref{sect:discussion}.
+The goal of this study is to evaluate the ability of \textbf{B-mode/}ARFI
+imaging to delineate prostate zonal anatomy, specifically central gland and
+total prostate gland \invivo as compared to T2WI MR at 3T using an endorectal
+coil.  \input{intro_vibroelastography} Section~\ref{sect:background} of this
+manuscript provides an overview of MR and ARFI imaging in the prostate and an
+overall clinical motivation for prostate imaging.  Section~\ref{sect:methods}
+describes the methods used to experimentally-acquire our imaging data, process
+the gross pathological specimens post radical prostatectomy, and image process
+our datasets.  The results of our analysis, including gross pathology and
+imaging prostate axis and volume estimates, are presented in
+Section~\ref{sect:results}, along with a statistical analysis of the bias and
+variability associated with each imaging modalities measurements, all of which
+is discussed in Section~\ref{sect:discussion}.

### 2014-04-24 results.tex
updated text to explicitely state that B-mode images were used with ARFI images for all of the total gland / capsule approximations

@@
 The subvolumes associated with the zonal anatomy in each imaging modality were
-measured (Figure~\ref{fig:mr_arfi_volumes}(a)), with good correlation
-between the ultrasound and MR total prostate gland volumes (R$^2$ =
+measured (Figure~\ref{fig:mr_arfi_volumes}(a)), with good correlation between
+the ultrasound and MR total prostate gland volumes (R$^2$ =
 \MRarfiVolTotalRsq), with a mean overestimation of of
-\MRarfiVolTotalMeanDiff~$\pm$ \MRarfiVolTotalStdDiff\% by ultrasound imaging compared
-to MR volumes (Figure~\ref{fig:mr_arfi_volumes}(b)).  Central gland volumes
-were well-correlated between ARFI and MR images (R$^2$ = \MRarfiVolCentralRsq)
-with no significant mean over/under-estimation (\MRarfiVolCentralMeanDiff~$\pm$
-\MRarfiVolCentralStdDiff\%), but significant variability between the cases
-(Figure~\ref{fig:mr_arfi_volumes}(c)).  Table~\ref{tab:mr_arfi_volumes} has the
-individually-measured volumes for MR and ARFI imaging for each study subject.
+\MRarfiVolTotalMeanDiff~$\pm$ \MRarfiVolTotalStdDiff\% by ultrasound imaging
+compared to MR volumes (Figure~\ref{fig:mr_arfi_volumes}(b)).  Central gland
+volumes were well-correlated between ARFI and MR images (R$^2$ =
+\MRarfiVolCentralRsq) with no significant mean over/under-estimation
+(\MRarfiVolCentralMeanDiff~$\pm$ \MRarfiVolCentralStdDiff\%), but significant
+variability between the cases (Figure~\ref{fig:mr_arfi_volumes}(c)).
+Table~\ref{tab:mr_arfi_volumes} has the individually-measured volumes for MR
+and \textbf{B-mode/}ARFI imaging for each study subject.
 
@@
 image-reconstructed prostate volumes (Figure~\ref{fig:mr_arfi_weight}(b), R$^2$
-= \weightMRrsq~(MR) and \weightARFIrsq~ (ARFI)), though there was weaker
+= \weightMRrsq~(MR) and \weightARFIrsq~ (\textbf{B-mode})), though there was weaker
 correlation with the ellipsoidal approximation of the measurement prostate
 volume and the image-reconstructed volumes (Figure~\ref{fig:mr_arfi_weight}(c),
-R$^2$ = \pathVolMRrsq~(MR) and \pathVolARFIrsq~(ARFI)).
+R$^2$ = \pathVolMRrsq~(MR) and \pathVolARFIrsq~(\textbf{B-mode})).
 
@@
 correlation between the imaging axis measurements was analyzed
-(Figure~\ref{fig:mr_arfi_path_axes}).  ARFI was most correlated to MR in the
-lateral-to-lateral axis in both the total gland and CG (R$^2$ =
-\totalLatLatRsq~and \centralLatLatRsq, respectively), with mean overestimates
-of \ARFImrTotalLatLatMeanPct~$\pm$ \ARFImrTotalLatLatStdPct\%~and
+(Figure~\ref{fig:mr_arfi_path_axes}). \textbf{Ultrasound images were} most
+correlated to MR in the lateral-to-lateral axis in both the total gland
+\textbf{B-mode} and CG \textbf{(ARFI)} (R$^2$ = \totalLatLatRsq~and
+\centralLatLatRsq, respectively), with mean overestimates of
+\ARFImrTotalLatLatMeanPct~$\pm$ \ARFImrTotalLatLatStdPct\%~and
 \ARFImrCentralLatLatMeanPct~$\pm$ \ARFImrCentralLatLatStdPct\%, respectively
-(Table~\ref{tab:mr_arfi_axes_error}).  ARFI had moderate correlation with the
-total prostate gland and CG axes in the anterior-to-posterior dimension (R$^2$
-= \totalAntPostRsq~and \centralAntPostRsq), respectively, with differences of
+(Table~\ref{tab:mr_arfi_axes_error}).  \textbf{Ultrasound images} had moderate
+correlation with the total prostate gland \textbf{(B-mode)} and CG
+\textbf{ARFI}) axes in the anterior-to-posterior dimension (R$^2$ =
+\totalAntPostRsq~and \centralAntPostRsq), respectively, with differences of
 \ARFImrTotalAntPostMeanPct~$\pm$ \ARFImrTotalAntPostStdPct\% and
-\ARFImrCentralAntPostMeanPct~$\pm$ \ARFImrCentralAntPostStdPct\%.  ARFI imaging
-had weak correlation with MR images in the apex-to-base dimension (R$^2$ =
-\totalApexBaseRsq, total gland and R$^2$ = \centralApexBaseRsq, CG), with
-underestimates of \ARFImrTotalApexBaseMeanPct~$\pm$
-\ARFImrTotalApexBaseStdPct\% and \ARFImrCentralApexBaseMeanPct~$\pm$
-\ARFImrCentralApexBaseStdPct\%, respectively.
+\ARFImrCentralAntPostMeanPct~$\pm$ \ARFImrCentralAntPostStdPct\%.  \textbf{The
+ultrasound images} had weak correlation with MR images in the apex-to-base
+dimension (R$^2$ = \totalApexBaseRsq, total gland and R$^2$ =
+\centralApexBaseRsq, CG), with underestimates of
+\ARFImrTotalApexBaseMeanPct~$\pm$ \ARFImrTotalApexBaseStdPct\% and
+\ARFImrCentralApexBaseMeanPct~$\pm$ \ARFImrCentralApexBaseStdPct\%,
+respectively.
 

### 2014-04-24 figs/mr_arfi_volumes.m
updated mr_arfi_volumes script to properly indicate B-mode over ARFI for some labels

@@ -41,9 +41,9 @@ set(h(1),'FaceColor',[0 0 1]);
 set(h(2),'FaceColor',[1 0 0]);
 set(g(1),'FaceColor',[0 1 0]);
 set(g(2),'FaceColor',[1 1 0]);
-ylabel('Gland/Zone Volumes (cm^3)','FontSize',fs)
+ylabel('Volume (cm^3)','FontSize',fs)
 xlabel('Study Subject','FontSize',fs);
-title('MR (Red/Blue) and ARFI (Yellow/Green) Volumes','FontSize',fs);
+title('MR (Red/Blue) and B-mode/ARFI (Yellow/Green) Volumes','FontSize',fs);
 a=axis;
 a(2) = 17;
 axis(a);
@@ -124,9 +124,9 @@ figure;
     hold on
     plot(mr_total_vol/1e3,arfi_total_vol/1e3,'kx','MarkerSize',14,'LineWidth',3);
     plot(mr_total_vol/1e3,mr_arfi_total_fit,'-k','LineWidth',3);
-    title('ARFI:MR Volumes','FontSize',fs);
+    title('B-mode:MR Volumes','FontSize',fs);
     xlabel('MR Volume (cm^3)','FontSize',fs);
-    ylabel('ARFI Volume (cm^3)','FontSize',fs);
+    ylabel('B-mode Volume (cm^3)','FontSize',fs);
     set(gca,'FontSize',fs);            
     %a=axis;                            
     %axis(a);

### 2014-04-24 figs/corr_weight_vol.m
updated path correlation plots to now reference B-mode instead of ARFI

@@ -88,9 +88,9 @@ function gen_fig_n_fits(fid,w,mr_total_vol,arfi_total_vol,metric)
         fprintf(fid,'\\newcommand{\\pathVolWeightRsq}{%.2f}\n',mr_Rsq);
     else,
         if(strcmp(metric,'pathVol')),
-            legend(sprintf('MR (R^2 = %.2f)',mr_Rsq),sprintf('ARFI (R^2 = %.2f)',arfi_Rsq),'Location','NorthEast');
+            legend(sprintf('MR (R^2 = %.2f)',mr_Rsq),sprintf('B-mode (R^2 = %.2f)',arfi_Rsq),'Location','NorthEast');
         else,
-            legend(sprintf('MR (R^2 = %.2f)',mr_Rsq),sprintf('ARFI (R^2 = %.2f)',arfi_Rsq),'Location','NorthWest');
+            legend(sprintf('MR (R^2 = %.2f)',mr_Rsq),sprintf('B-mode (R^2 = %.2f)',arfi_Rsq),'Location','NorthWest');
         end;
         legend boxoff;
         if(strcmp(metric,'pathVol')),

### 2014-04-24 figs/mr_arfi_path_axes.m
updated axis dimension figs to reflect b-mode when appropriate

@@ -97,7 +97,7 @@ function mr_arfi_path_axes
         plot(MR_total_axes(:,i)/10,ARFI_total_axes(:,i)/10,'rx','MarkerSize',14,'LineWidth',3);
         plot(MR_central_axes(:,i)/10,ARFI_central_axes(:,i)/10,'bo','MarkerSize',14,'LineWidth',3);
         xlabel('MR Measurement (cm)','FontSize',fs);
-        ylabel('ARFI Measurement (cm)','FontSize',fs);
+        ylabel('B-mode/ARFI Measurement (cm)','FontSize',fs);
 
         set(gca, ...
             'Box'         , 'off'     , ...
@@ -289,8 +289,8 @@ function mr_arfi_path_axes
         close;
 
         % while I'm here, lets compute some stats to include in the caption / text, and LaTeX format it!!
-        fprintf(fid,'ARFI & MR & Total & %s & %.1f $\\pm$ %.1f \\\\ \n',axis_combos{i},mean(ARFI_MR_Total_OverUnder(:,i)),std(ARFI_MR_Total_OverUnder(:,i)));
-        fprintf(fid,'ARFI & PATH & Total & %s & %.1f $\\pm$  %.1f \\\\ \n',axis_combos{i},mean(ARFI_PATH_Total_OverUnder(:,i)),std(ARFI_PATH_Total_OverUnder(:,i)));
+        fprintf(fid,'B-mode & MR & Total & %s & %.1f $\\pm$ %.1f \\\\ \n',axis_combos{i},mean(ARFI_MR_Total_OverUnder(:,i)),std(ARFI_MR_Total_OverUnder(:,i)));
+        fprintf(fid,'B-mode & PATH & Total & %s & %.1f $\\pm$  %.1f \\\\ \n',axis_combos{i},mean(ARFI_PATH_Total_OverUnder(:,i)),std(ARFI_PATH_Total_OverUnder(:,i)));
         fprintf(fid,'MR & PATH & Total & %s & %.1f $\\pm$ %.1f \\\\ \n',axis_combos{i},mean(MR_PATH_Total_OverUnder(:,i)),std(MR_PATH_Total_OverUnder(:,i)));
     end;
 

### 2014-04-24 fig_mr_arfi_path_axes.tex
updated caption to reflect usage of B-mode for total gland

@@
     (red crosses) and central gland (blue circles).  \textbf{B-mode images were
-        used to estimate the total gland volume, and ARFI images were used to
-        estimate the central glands.} The black dashed-line represents the
-    projection of perfectly-correlated measurements between imaging and
-    pathology.  The over-/under-estimation of each imaging modality relative to
-    gross pathology and each other is summarized in
-    Table~\ref{tab:mr_arfi_axes_error}.} 
+        used for total gland measurements, and ARFI images were for central
+        gland measurements.} The black dashed-line represents the projection of
+    perfectly-correlated measurements between imaging and pathology.  The
+    over-/under-estimation of each imaging modality relative to gross pathology
+    and each other is summarized in Table~\ref{tab:mr_arfi_axes_error}.} 
 \label{fig:mr_arfi_path_axes}

### 2014-04-24 tab_mr_arfi_axes.tex
updated table headings to reflect B-mode

@@
 \centering
-\caption{Comparison of Central Gland / Zone (C) and Total (T) Prostate Axes in MR T2WI and ARFI Imaging.
-Axes are approximated in orientation to match those specified in gross pathology: lateral-to-lateral (LL), anterior-to-posterior (AP) and apex-to-base (AB).}
+\caption{Comparison of Central Gland / Zone (C) and Total (T) Prostate Axes in
+    MR T2WI and \textbf{B-mode/}ARFI Imaging.  Axes are approximated in orientation to
+    match those specified in gross pathology: lateral-to-lateral (LL),
+    anterior-to-posterior (AP) and apex-to-base (AB).}
 \begin{tabular}{|l|l|l|l|l|l|l|l|l|l|l|l|l|} \hline
-{\bf Study} & {\bf MR} & {\bf ARFI} & {\bf MR} & {\bf ARFI} & {\bf MR} & {\bf ARFI} & {\bf MR} & {\bf ARFI} & {\bf MR} & {\bf ARFI} & {\bf MR} & {\bf ARFI} \\ 
+{\bf Study} & {\bf MR} & {\bf ARFI} & {\bf MR} & {\bf ARFI} & {\bf MR} & {\bf ARFI} & {\bf MR} & {\bf B-mode} & {\bf MR} & {\bf B-mode} & {\bf MR} & {\bf B-mode} \\ 
 {\bf Subject} & {\bf C-AB} & {\bf C-AB} & {\bf C-LL} & {\bf C-LL} & {\bf C-AP} & {\bf C-AP} & {\bf T-AB} & {\bf T-AB} & {\bf T-LL} & {\bf T-LL} & {\bf T-AP} & {\bf T-AP} \\

### 2014-04-24 tab_mr_arfi_volumes.tex
updated table headings to reflect B-mode

@@
 \centering
-\caption{Comparison of Central Gland / Zone and Total Prostate Volumes in MR T2WI and ARFI Imaging}
+\caption{Comparison of Central Gland / Zone and Total Prostate Volumes in MR T2WI and \textbf{B-mode/}ARFI Imaging}
 \begin{tabular}{|l|l|l|l|l|} \hline
-{\bf Study Subject} & {\bf MR Central Gland} & {\bf MR Total} & {\bf ARFI Central Gland} & {\bf ARFI Total} \\ 
+{\bf Study Subject} & {\bf MR Central Gland} & {\bf MR Total} & {\bf ARFI Central Gland} & {\bf B-mode Total} \\ 
 & {\bf Volume (cm$^3$)} & {\bf Volume (cm$^3$)} & {\bf Volume (cm$^3$)} & {\bf Volume (cm$^3$)} \\ \hline

### 2014-04-24 Response_to_R0_Reviews.tex
responded to peri-prostatic tissue comment

@@
 
-    TO DO
+    We agree that very careful dissection of the gross organ from adjacent
+    peri-prostatic tissue could be achieved, but to not disrupt the clinical
+    pathologic processing of these samples, we could not specify specific
+    pathology assisstants to process these specimens, so there was some
+    inherent variability associated with the time and effort that each
+    assistant dedicated to removing this peri-prostatic tissue.  We felt that
+    it was important to delineate this as a source of variability among our
+    pathology specimens, though it is expected to be a higher-order source of
+    error.
 

### 2014-04-24 Response_to_R0_Reviews.tex
responded to ROC for image axis analysis

@@
 
-    TO DO
+    As is also detailed in the response to comment 2 by the second reviewer, we
+    are actively pursuing correlation of our B-mode/ARFI and MR images with
+    whole mount histology, which includes rigid and diffeomorphic image
+    registration processes.  Given the that the methods and validation
+    associated with the more spatially-refined whole mount histology analysis
+    and diffeomorphic image registration is nontrivial, we have decided to
+    present that material in a dedicated manuscript that will allow us to
+    present our validation of that methodology.
 

### 2014-04-24 Response_to_R0_Reviews.tex
final responses in place; still need to spell check and proof read

@@
 
-    TO DO
+    ARFI resolution within a native imaging frame is not considered a limiting
+    factor in our ability to delineate the prostate central glands since the
+    scan-conversion and slice decimation introduce resolution distortions that
+    are at least an order-of-magnitude larger than our native image plane
+    resolution.  ARFI image resolution would be a more realistic limiting
+    factor if we were using those primary imaging frames for exclusive gland
+    delineation, but our use of the 3 orthogonal planes for segmentation makes
+    this resolution of higher-order parameter.  We have added to the following
+    text to Section V to emphasize this point to the reader:
+
+    \input{arfi_resolution}
 
@@
 
-    TO DO
+    We completely agree with the author that a more direct comparison of
+    B-mode/ARFI imaging data with MR after ``unwrapping'' the deformation to
+    register the images with one another would be more complete and satisfying,
+    but two of the primary objectives of this work were (1) to compare prostate
+    appearance and size in each native imaging, as would be seen by a clinician
+    in each setting, and (2) to quantify the differences in the sizes of each
+    dimension without the image registration to allow a clinician to
+    cognitively register the images with known amounts of bias in each
+    dimension.
+
+    We are actively working on analyses that are focusing on comparing each
+    imaging modality with whole mount histology as a gold standard, and then
+    utilizing diffeomorphic image registration to spatially align each image
+    for fused image guidance.  Both of these efforts require dedicated
+    metholodology description, and more importantly, method validation and
+    error analysis that, in our opinions, warrant dedicated publications to be
+    thoroughly presented to readers.
+
+    In the context of the analysis that we have presented in this manuscript
+    and your comment about TRUS deformation, we have expanded our comments
+    about this deformation and its impact on the results, along with more
+    description of the future analysis efforts.
+
+    Section V(A):
+
+    \input{disc_deformation}
+
+    And Section V(C):
+
+    \input{future_diffeomorphic}
 

### 2014-04-24 arfi_resolution.tex
final responses in place; still need to spell check and proof read

\textbf{The first-order limiting parameters in our ultrasound spatial
resolution to segment capsules and central glands were: (1) the scan conversion
of our native radial ultrasound imaging planes to yield uniform 3D voxels of
B-mode and ARFI imaging data, and (2) the decimation of imaging volume datasets
for manual segmentation.  The scan-conversion interpolation artifact is worse
at deeper depths, where native image data is more sparsely available.  For
these reasons, neither the native imaging plane resolution of B-mode or ARFI
imaging is considered a limiting factor in the total and central gland
measurements presented in this work.}

### 2014-04-24 disc_arfi_mr.tex
final responses in place; still need to spell check and proof read

@@
 
-We did not expect perfect agreement between MR and \textbf{ultrasound} imaging
-volumes or tri-axial dimensions since each modality applies a different amount
-of external compression to the rectal wall, which then deforms the prostate.
-The endorectal coil places pressure on the prostate that is dependent on the
-patient's anatomy and overall prostate size, while pressure on the prostate
-during ultrasound imaging is modulated by the urologist when the transducer
-rotation apparatus is locked into position prior to imaging.  The differential
-deformation of the prostate can be easily visualized during the transducer
-alignment process during ultrasound imaging.  It was our goal to apply as
-little pressure while still maintaining uniform contact throughout the
-ultrasound image acquisition. This is critical since poor acoustic coupling to
-the prostate through the rectal wall can greatly diminish ARFI image
-displacement SNR.
+\input{disc_deformation}
 
@@
 utilized.
+
+\input{arfi_resolution}
 

### 2014-04-24 disc_deformation.tex
final responses in place; still need to spell check and proof read

\textbf{We did not expect perfect agreement between MR and ultrasound imaging
volumes or tri-axial dimensions since each modality applies a different amount
of external compression to the rectal wall, which results in different degrees
of prostate deformation The endorectal coil used in MRI places pressure on the
prostate that is dependent on the patient's anatomy and overall prostate size,
while the pressure on the prostate during transrectal ultrasound imaging is
modulated by the urologist when the transducer rotation apparatus is locked
into position prior to imaging.  The differential prosate deformation during
transrectal ultrasound imaging can be easily visualized during the transducer
alignment process.  It was our goal to apply as little pressure while still
maintaining uniform contact throughout the ultrasound image acquisition. This
is critical since poor acoustic coupling to the prostate through the rectal
wall can greatly diminish ARFI image displacement SNR.}

### 2014-04-24 future.tex
final responses in place; still need to spell check and proof read

@@
 regions of PCa suspicion between the two modalities, and to correlate with
-whole-mount histology of the excised prostate specimens.  This effort will
-allow the sensitivity and specificity of each imaging modality for PCa
-detection and characterization to be quantified.  Additionally, we will be
-pursuing quantitative measurement of the pressure being applied during imaging
-to help reduce prostate deformation and reduce the impact of prostate stiffness
-nonlinearity in our displacement images.  While ARFI imaging of the prostate is
-a constantly evolving technology, the fact that one can view the different
-prostatic zones on ARFI images in addition to MRI is encouraging and could lead
-to targeted diagnostic biopsies and therapies with real-time imaging for men
-with PCa.
+whole-mount histology of the excised prostate specimens.  
+\input{future_diffeomorphic}
+These
+efforts for improved imaging performance and multi-modality image registration
+will allow for the quantification of sensitivity and specificity of a combined
+MR/B-mode/ARFI imaging system for PCa detection and characterization.
+Additionally, we will be pursuing quantitative measurement of the pressure
+being applied during imaging to help reduce prostate deformation and reduce the
+impact of prostate stiffness nonlinearity in our displacement images.  While
+ARFI imaging of the prostate is a constantly evolving technology, the fact that
+one can view the different prostatic zones on ARFI images in addition to MRI is
+encouraging and could lead to targeted diagnostic biopsies and therapies with
+real-time imaging for men with PCa.

### 2014-04-24 Response_to_R0_Reviews.tex
spell-checked

@@
     We agree with this suggestion, and we have added Figure 1 to the manuscript
-    that outlines the McNeals zonal anatomy of the prostaet in the coronal and
+    that outlines the McNeals zonal anatomy of the prostate in the coronal and
     sagittal views.
@@
     deep transmit focus.  Since only the ARFI imaging vectors were presented in
-    this work--not the shear wave data--we have not ammended the manuscript
+    this work--not the shear wave data--we have not amended the manuscript
     text to avoid confusing the reader with details not relevant to the data
@@
     we (1) did not do enough to emphasize this point, and (2) confused matters
-    by refering to `ARFI' exclusively in the results.  With respect to how the
+    by referring to `ARFI' exclusively in the results.  With respect to how the
     data were combined, we have added the following text to the manuscript
@@
     imaging in this process, and the abstract and conclusions have also been
-    explicitely updated to include the fact that B-mode and ARFI imaging data
+    explicitly updated to include the fact that B-mode and ARFI imaging data
     were used in this study.
@@
 
-    All of the B-mode and ARFI images were segemented by two individuals (ZAM,
+    All of the B-mode and ARFI images were segmented by two individuals (ZAM,
     TJG), and reviewed by a single, experienced ARFI imaging research (KRN).
@@
     pathologic processing of these samples, we could not specify specific
-    pathology assisstants to process these specimens, so there was some
+    pathology assistants to process these specimens, so there was some
     inherent variability associated with the time and effort that each
@@
     quality was definitely degraded during the rendering of the review PDF.  We
-    will make sure that all images are propoerly reproduced at full resolution
+    will make sure that all images are properly reproduced at full resolution
     in new PDFs that are rendered of this manuscript.
@@
    \begin{itemize}
-        \item The gross surgical pathology mesaurements used to make the
+        \item The gross surgical pathology measurements used to make the
             ellipsoidal volume approximation were done after peri-prostatic
@@
             they suffer from post-fixation contraction, $\sim$ 3 mm slice
-            thickiness, and slice tearing and folding that compromise accurate
+            thickness, and slice tearing and folding that compromise accurate
             cross-sectional area representations on a single slide.
@@
     for fused image guidance.  Both of these efforts require dedicated
-    metholodology description, and more importantly, method validation and
+    methodology description, and more importantly, method validation and
     error analysis that, in our opinions, warrant dedicated publications to be

### 2014-04-24 fig_mcneal_anatomy.tex
spell-checked

@@
     prostate.~\cite{mcneal_path} It divides a human prostate into one
-    nonglandular zone (the anterior fibromuscular stroma) and three glandular
+    non-glandular zone (the anterior fibromuscular stroma) and three glandular
     zones: central zone, peripheral zone and transition zone, denoted by CZ, PZ

### 2014-04-24 path_volume_choice.tex
added Buck reference on formalin prostate contraction

@@
 This approach, however, was not chosen because the formalin fixation process
-can cause 0-25\% contraction of the tissue, the $\sim$3 mm slice thickness
-introduces error, and tearing / folding of tissue on the whole mount slides
-introduce additional artifacts that need to be compensated for in the volume
-estimation process.  Finally, differences in the tissue orientation on each
-whole mount slide would require an image registration process to align all of
-the slides, introducing another source of error in the process.
+can cause 0-25\% contraction of the tissue~\cite{Schned1996}, the $\sim$3 mm
+slice thickness introduces error, and tearing / folding of tissue on the whole
+mount slides introduce additional artifacts that need to be compensated for in
+the volume estimation process.  Finally, differences in the tissue orientation
+on each whole mount slide would require an image registration process to align
+all of the slides, introducing another source of error in the process.

### 2014-04-25 Response_to_R0_Reviews.tex
updated response on extraprostatic tissue based on Buck feedback

@@
 
-    We agree that very careful dissection of the gross organ from adjacent
-    peri-prostatic tissue could be achieved, but to not disrupt the clinical
-    pathologic processing of these samples, we could not specify specific
-    pathology assistants to process these specimens, so there was some
-    inherent variability associated with the time and effort that each
-    assistant dedicated to removing this peri-prostatic tissue.  We felt that
-    it was important to delineate this as a source of variability among our
-    pathology specimens, though it is expected to be a higher-order source of
-    error.
+    The extraprostatic tissue that is left on the prostate by the surgeon is
+    usually minimal, but can vary case-by-case and is left in place during
+    gross and histologic processing to adhere to AJCC guidelines that require
+    the reporting of potential extraprostatic invasion of malignany tumors.
+    This extraprostatic tissue is left in place during whole-mount histology
+    processing, but given the challenges in estimating prostate gland size in
+    histology (detailed in the response to the second reviewer below), our
+    pathologist (AB) believed that the most accurate estimations of \invivo
+    prostate dimensions would be taken from the gross pathology specimens.  In
+    consulting with our pathologist on this specific comment, his impression
+    was that the impact of extraprostatic tissue would be minimal,  so we have
+    removed this section from the discussion to focus on more relevant points.
 
@@
             the complication of tissue shrinkage associated with formalin
-            fixation, which our pathologists estimated could be as much as 30\%
-            of the prostate volume.
+            fixation, which our pathologists estimated could be as much as 25\%
+            of the prostate volume (citation provided in new text).
         \item Whole mount histology slides were explored for this purpose, but

### 2014-04-25 abstract.tex
updated response on extraprostatic tissue based on Buck feedback

@@
 and dimensions that were well-correlated with MR T2WI estimates, with biases in
-the lateral dimension due to poor contrast caused by peri-prostatic fat.
+the lateral dimension due to poor contrast caused by extraprostatic fat.
 \textbf{B-mode combined with} ARFI imaging is a promising low-cost, portable,

### 2014-04-25 disc_arfi_mr.tex
updated response on extraprostatic tissue based on Buck feedback

@@
 relative to MR. The lateral overestimation was a result of poor contrast
-between the PZ and peri-prostatic fat in the B-mode images, and image
+between the PZ and extraprostatic fat in the B-mode images, and image
 decimation along the lateral dimension in both ARFI and B-mode images. Without
@@
 correct errors introduced by the poor contrast between the PZ and
-peri-prostatic fat. However, with image decimation, segmentation changes in the
+extraprostatic fat. However, with image decimation, segmentation changes in the
 elevation and axial orthogonal planes were limited to the lower bound on the

### 2014-04-25 image_proc_methods.tex
updated response on extraprostatic tissue based on Buck feedback

@@
 identified by its homogeneous high signal intensity on T2WI, which is usually
-similar to that of the nearby peri-prostatic fat. The CG was visualized and
+similar to that of the nearby extraprostatic fat. The CG was visualized and
 delineated based on its heterogeneous, lower signal intensity and its location
@@
 the posterior. Prostate capsule was segmented in B-mode as this modality offers
-both high contrast between prostatic tissue and surrounding peri-prostatic fat,
+both high contrast between prostatic tissue and surrounding extraprostatic fat,
 and better SNR in the anterior capsule region than ARFI images. ARFI images

### 2014-04-25 limitations.tex
updated response on extraprostatic tissue based on Buck feedback

@@
 This study has several limitations that should be considered when interpreting
-these results.  Gross pathology weight and axis measurements could both be
-affected by the presence of peri-prostatic tissue that was excised during
+these results.  \st{Gross pathology weight and axis measurements could both be
+affected by the presence of extraprostatic tissue that was excised during
 radical prostatectomy, especially in cases where more aggressive margins may
@@
 (Figure~\ref{fig:mr_arfi_weight}) were not characterized for absolute accuracy,
-but instead, relative correlations were evaluated.  Additionally, the volumes
+but instead, relative correlations were evaluated.}  Additionally, the volumes
 of the prostate from gross pathologic measurements were approximated as

### 2014-04-25 Response_to_R0_Reviews.tex
updated extraprostatic text based on more recs from Andrew

@@
     This extraprostatic tissue is left in place during whole-mount histology
-    processing, but given the challenges in estimating prostate gland size in
-    histology (detailed in the response to the second reviewer below), our
-    pathologist (AB) believed that the most accurate estimations of \invivo
-    prostate dimensions would be taken from the gross pathology specimens.  In
-    consulting with our pathologist on this specific comment, his impression
-    was that the impact of extraprostatic tissue would be minimal,  so we have
-    removed this section from the discussion to focus on more relevant points.
+    processing, but given the challenges in estimating prostate gland size post
+    histologic processing (detailed in the response to the second reviewer
+    below), our pathologist (AB) believed that the most accurate estimations of
+    \invivo prostate dimensions would be taken from the gross pathology
+    specimens.
+
+    We have modified the text in the Limitations section as follows:
+
+    \textbf{Gross pathology weight and axis measurements could both be affected
+        by the presence of extraprostatic tissue that was excised during
+        radical prostatectomy, especially in cases where more aggressive
+        margins may have been necessary; however, we still believe that these
+        weights and measurements represent the best estimation of \invivo
+        characteristics.}
 

### 2014-04-25 limitations.tex
updated extraprostatic text based on more recs from Andrew

--- 2014-02-08 limitations.tex
@@
 This study has several limitations that should be considered when interpreting
-these results.  Gross pathology weight and axis measurements could both be
-affected by the presence of peri-prostatic tissue that was excised during
+these results.  \textbf{Gross pathology weight and axis measurements could both
+be affected by the presence of extraprostatic tissue that was excised during
 radical prostatectomy, especially in cases where more aggressive margins may
-have been necessary.  For this reason, unlike the image-to-image measurement
-comparisons, all of the image metrics presented relative to pathology metrics
+have been necessary; however, we still believe that these weights and
+measurements represent the best estimation of \invivo characteristics.}  For
+this reason, unlike the image-to-image measurement comparisons, all of the
+image metrics presented relative to pathology metrics
 (Figure~\ref{fig:mr_arfi_weight}) were not characterized for absolute accuracy,

### 2014-05-02 authors.tex
broke out authors

\author{
    \IEEEauthorblockN{
        Mark L. Palmeri, M.D., Ph.D.\IEEEauthorrefmark{1}, 
        Zachary A. Miller\IEEEauthorrefmark{1}, 
        Tyler J. Glass\IEEEauthorrefmark{1}, 
        Kirema Garcia-Reyes\IEEEauthorrefmark{2}, 
        Rajan T. Gupta, M.D.\IEEEauthorrefmark{3}, 
        Stephen J.  Rosenzweig\IEEEauthorrefmark{1}, 
        Christopher Kauffman, M.D.\IEEEauthorrefmark{3}, 
        Thomas J. Polascik, M.D.\IEEEauthorrefmark{4},
        Andrew Buck, M.D.\IEEEauthorrefmark{5},
        Samantha L. Lipman\IEEEauthorrefmark{1},
        Ned C. Rouze\IEEEauthorrefmark{1},
        Kathryn R. Nightingale, Ph.D.\IEEEauthorrefmark{1}}

    \IEEEauthorblockN{
        {\IEEEauthorrefmark{1}Department of Biomedical Engineering, Pratt School of Engineering, Duke University}
        {\IEEEauthorrefmark{2}Duke University School of Medicine}
        {\IEEEauthorrefmark{3}Department of Radiology, Duke University Medical Center}
        {\IEEEauthorrefmark{4}Department of Surgery (Urology), Duke University Medical Center}
        {\IEEEauthorrefmark{5}Department of Pathology, Duke University Medical Center}
    }
}

### 2014-05-02 disc_deformation.tex
spell-checked
missing punctuation

@@
 of external compression to the rectal wall, which results in different degrees
-of prostate deformation The endorectal coil used in MRI places pressure on the
+of prostate deformation. The endorectal coil used in MRI places pressure on the
 prostate that is dependent on the patient's anatomy and overall prostate size,
@@
 modulated by the urologist when the transducer rotation apparatus is locked
-into position prior to imaging.  The differential prosate deformation during
+into position prior to imaging.  The differential prostate deformation during
 transrectal ultrasound imaging can be easily visualized during the transducer

### 2014-05-02 acknowledgements.tex
moved John Madded to author list and added Evan as an author

@@
 Division for their in-kind technical support.  Special thanks to Ned Danieley
-for computer system administration support, the pathology assistants and Dr.
-John Madden in the Duke University Medical Center for their gross prostate
-organ processing, and Matthew Brown for his assistance in designing the optical
-prostate rotation feedback system.  This work was supported by NIH R01CA142824
-and the Duke Coulter Translational Grant Program.
+for computer system administration support, the pathology assistants in the
+Duke University Medical Center for their gross prostate organ processing, and
+Matthew Brown for his assistance in designing the optical prostate rotation
+feedback system.  This work was supported by NIH R01CA142824 and the Duke
+Coulter Translational Grant Program.

### 2014-05-02 abstract.tex
define ARFI in abstract

@@
 Multi-parametric Magnetic Resonance Imaging (mpMRI) has gained recent
-popularity to characterize PCa.  ARFI imaging has the potential to aid PCa
-diagnosis and management by using tissue stiffness to evaluate prostate zonal
-anatomy and lesions.  MR and \textbf{B-mode/}ARFI \invivo imaging datasets were
-compared to one another and with gross pathology measurements made immediately
-after radical prostatectomy.  Images were manually segmented in 3D Slicer to
-delineate the central gland (CG) and prostate capsule, and 3D models were
-rendered to evaluate zonal anatomy dimensions and volumes.  Both imaging
-modalities showed good correlation between estimated organ volume and gross
-pathologic weights.  Ultrasound and MR total prostate volumes were
-well-correlated (R$^2$ = \MRarfiVolTotalRsq), but \textbf{B-mode} images
-yielded prostate volumes that were larger (\MRarfiVolTotalMeanDiff\% $\pm$
-\MRarfiVolTotalStdDiff\%) than MR images, due to overestimation of the lateral
-dimension (\ARFImrTotalLatLatMeanPct $\pm$ \ARFImrTotalLatLatStdPct\%), with
-less significant differences in the other dimensions
-(\ARFImrTotalAntPostMeanPct $\pm$ \ARFImrTotalAntPostStdPct\%,
+popularity to characterize PCa.  Acoustic Radiation Force Impulse (ARFI)
+imaging has the potential to aid PCa diagnosis and management by using tissue
+stiffness to evaluate prostate zonal anatomy and lesions.  MR and
+\textbf{B-mode/}ARFI \invivo imaging datasets were compared to one another and
+with gross pathology measurements made immediately after radical prostatectomy.
+Images were manually segmented in 3D Slicer to delineate the central gland (CG)
+and prostate capsule, and 3D models were rendered to evaluate zonal anatomy
+dimensions and volumes.  Both imaging modalities showed good correlation
+between estimated organ volume and gross pathologic weights.  Ultrasound and MR
+total prostate volumes were well-correlated (R$^2$ = \MRarfiVolTotalRsq), but
+\textbf{B-mode} images yielded prostate volumes that were larger
+(\MRarfiVolTotalMeanDiff\% $\pm$ \MRarfiVolTotalStdDiff\%) than MR images, due
+to overestimation of the lateral dimension (\ARFImrTotalLatLatMeanPct $\pm$
+\ARFImrTotalLatLatStdPct\%), with less significant differences in the other
+dimensions (\ARFImrTotalAntPostMeanPct $\pm$ \ARFImrTotalAntPostStdPct\%,
 anterior-to-posterior, and \ARFImrTotalApexBaseMeanPct $\pm$

### 2014-05-02 abstract.tex
defined T2WI in abstract

@@
 \ARFImrCentralLatLatStdPct\%).  B-mode/ARFI imaging yielded prostate volumes
-and dimensions that were well-correlated with MR T2WI estimates, with biases in
-the lateral dimension due to poor contrast caused by extraprostatic fat.
-\textbf{B-mode combined with} ARFI imaging is a promising low-cost, portable,
-real-time modality that can complement mpMRI for PCa diagnosis, treatment
-planning and management.
+and dimensions that were well-correlated with MR T2-weighted image (T2WI)
+estimates, with biases in the lateral dimension due to poor contrast caused by
+extraprostatic fat.  \textbf{B-mode combined with} ARFI imaging is a promising
+low-cost, portable, real-time modality that can complement mpMRI for PCa
+diagnosis, treatment planning and management.

### 2014-05-02 mr_background.tex
small typo

@@
 the early 1980s, but the early work on its diagnostic accuracy is heterogeneous
-due to heavy reliance on morphology via T1 and T2-weighted imaging (T2WI). More
+due to heavy reliance on morphology via T1- and T2-weighted imaging (T2WI). More
 recent abilities to include not only anatomic, but also biologic and functional

### 2014-05-02 arfi_methods.tex
corrected small typos

@@
 mm, 22.5 mm, and 15 mm, respectively, and finally the response of the tissue
-was tracked for up to 6 ms at a PRF of 8kHz. This pushing strategy is similar to
+was tracked for up to 6 ms at a PRF of 8 kHz. This pushing strategy is similar to
 what has been published by Bercoff \etal~\cite{Bercoff2004}. The 30 mm and 22.5
-mm foci pushing pulses were transmitted at 4.6 MHz with a F/2 geometry and the
+mm foci pushing pulses were transmitted at 4.6 MHz with a F/2 geometry, and the
 15 mm focus pushing pulse was transmitted at 5.4 MHz with a F/2.35 geometry to
@@
 
-\textbf{3D volumetric imaging data was acquired endorectally} using mechanical
+\textbf{3D volumetric imaging data were acquired endorectally} using mechanical
 rotation of the ER7B transducer between sequential imaging frames, sweeping

### 2014-05-02 arfi_normalization.tex
corrected small typos

@@
 attenuation and focal gain effects to ensure that ARFI image brightness
-differences are more directly related to stiffness differences.  This                
+differences were more directly related to stiffness differences.  This                
 normalization was performed using a displacement profiles measured in two            
 homogeneous, elastic tissue-mimicking phantoms (CIRS, Norfolk, VA) with a          
-stiffness of 10.0 kPa and acoustic attenuations of 0.49 and 0.72 dB/cm/MHz.        
+Youngs modulus of 10.0 kPa and acoustic attenuations of 0.49 and 0.72 dB/cm/MHz.        
 The mean displacement profiles from each phantom were averaged together to
-achieve an effective attenuation of 0.6 dB/cm/MHz, which is within the measured
+achieve a mean attenuation of 0.6 dB/cm/MHz, which is within the measured
 range of attenuation in healthy peripheral zone prostate

### 2014-05-02 image_proc_methods.tex
corrected small typos

@@
 model and measure the prostate capsule and CG volume and tri-axial dimensions
-(Figure~\ref{fig:arfi_segs}).  \input{bmode_arfi_coregistered} The sagittal imaging plane was chosen for
-segmentation because it has a greater number of easily identifiable anatomical
-structures:  seminal vesicles in the left and right posterior base, urethra in
-the mid apex, surrounding fibromuscular tissue in the anterior, and rectum in
-the posterior. Prostate capsule was segmented in B-mode as this modality offers
-both high contrast between prostatic tissue and surrounding extraprostatic fat,
-and better SNR in the anterior capsule region than ARFI images. ARFI images
-were used for CG segmentation.  Prior to segmentation, both the B-mode and ARFI
-image stacks were down-sampled along the sagittal axis to a slice thickness of
-3 mm to reduce the number of manually-segmented frames (18 vs. 360 slices /
-image stack) and match the MRI slice thickness.
+(Figure~\ref{fig:arfi_segs}).  \input{bmode_arfi_coregistered} The sagittal
+imaging plane was chosen for segmentation because it has a greater number of
+easily identifiable anatomical structures:  seminal vesicles in the left and
+right posterior base, urethra in the mid apex, surrounding fibromuscular tissue
+in the anterior, and rectum in the posterior. Prostate capsules were segmented
+in B-mode images instead of ARFI images since the B-mode images provided high
+image contrast between prostatic tissue and surrounding extraprostatic fat, and
+better SNR in the anterior capsule region. ARFI images were used for CG
+segmentation due to their better contrast over B-mode images.  Prior to
+segmentation, both the B-mode and ARFI image stacks were down-sampled along the
+sagittal axis to a slice thickness of 3 mm to reduce the number of
+manually-segmented frames (18 vs. 360 slices / image stack) and match the MRI
+slice thickness.
 
@@
 dimensions were then measured using the same procedure as the MR images.
-\textbf{All B-mode and ARFI image segmentations were reviewed by an experienced
-    ARFI imaging researcher (KRN).}
+\textbf{All B-mode and ARFI image segmentations were reviewed by two experienced
+    ARFI imaging researchers (MLP and KRN).}
 

### 2014-05-02 results.tex
corrected small typos

@@
 correlation with the total prostate gland \textbf{(B-mode)} and CG
-\textbf{ARFI}) axes in the anterior-to-posterior dimension (R$^2$ =
+(\textbf{ARFI}) axes in the anterior-to-posterior dimension (R$^2$ =
 \totalAntPostRsq~and \centralAntPostRsq), respectively, with differences of

### 2014-05-02 Response_to_R0_Reviews.tex
updated responses (minor)

@@
 
-    We also agree that our exclusive use of 'ARFI' for many of the tables,
+    We also agree that our exclusive use of `ARFI' for many of the tables,
     figures, manuscript title, and text was confusing, so we have updated all
@@
     All of the B-mode and ARFI images were segmented by two individuals (ZAM,
-    TJG), and reviewed by a single, experienced ARFI imaging research (KRN).
-    There was no formal training in reading the ARFI images since this work is
-    our first effort to analyze our ability to evaluate our accuracy in
+    TJG), and reviewed by two experienced ARFI imaging researchers (MLP and
+    KRN).  There was no formal training in reading the ARFI images since this
+    work is our first effort to analyze our ability to evaluate our accuracy in
     identifying structures in ARFI prostate images.  The same individuals
@@
 
-    \textbf{All B-mode and ARFI image segmentations were reviewed by an
-        experienced ARFI imaging researcher (KRN).}
+    \textbf{All B-mode and ARFI image segmentations were reviewed by two
+        experienced ARFI imaging researcheres (MLP and KRN).}
 
@@
 
+\bibliographystyle{vancouver}
+\bibliography{library_no_urls,slicer3d}
+
 \end{document}

### 2014-05-02 coverletter.tex
updated coverletter for R1 submission

@@
 
-Re: \journal~Manuscript Submission
+Re: UIX-14-0012 R1 Revision
 \vspace{0.25in}
@@
 
-We thank you for taking the time and effort to review our manuscript, titled
-{\it ``Acoustic Radiation Force Impulse (ARFI) Imaging Prostate Zonal Anatomy:
-    Comparison with 3T T2-Weighted MR Imaging (T2WI)''}, for consideration to
-be published in \journal.
+We thank you for taking the time and effort to review our original manuscript,
+and we have submitted a revision with a detailed, itemized response to the R0
+comments.
 
